---
title: "Manuscript December 2024, Figure 5 - survival features. RNA data release 1.4k10 prerelease"
author: Maria Jakobsdottir (Maria.Jakobsdottir@manchester.ac.uk)
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2:
    toc: true 
    toc_float: true
    toc_depth: 6
---

Analysis for figure 5 in Hypoxia manuscript March 2025.

# Libraries used
```{r message=FALSE, warning=FALSE}
library(dplyr)
library(finalfit)
library(ggplot2)
library(ggpubr)
library(ggsci)
library(ggsignif)
library(ggsurvfit)
library(grid)
library(gridExtra)
library(lmtest)
library(splitstackshape) # for paired analysis
library(survival)
library(tidyr)
library(viridis)
library(xtable)

options(bitmapType='cairo')
'%!in%' <- function(x,y)!('%in%'(x,y))
```

```{r}
hypoxia_colours_q <- list("Q1" = "#14556e",
                          "Q2" = "#67889a",
                          "Q3" = "#c06842",
                          "Q4" = "#9c2f15")
hypoxia_colours_q2 <- list("Q1" = "#8995bc",
                          "Q2" = "#8c7893",
                          "Q3" = "#955569",
                          "Q4" = "#823c52")
hypoxia_colours_surv <- c("#14556e","#14556e","#67889a","#67889a",  
                          "#c06842","#c06842","#9c2f15","#9c2f15")
```
```{r}
hypoxia_colours_bin <- list("Low" = "#3E6F84",
                            "High" = "#AE4C2C")
hypoxia_colours_bin2 <- list("Low" = "#8B87A8",
                             "High" = "#8C495E")
```

# Data
Read in data

```{r}
params <- list(ppcg_colourlist = "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/GitHub/PPCG_colours/PPCG_colourlist.RData",
               combined_data = "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_data_private/combined_PPCG_WGS_RNA_summary_data_clinical_v7_2024-12-06.txt")
```

```{r}
load(params$ppcg_colourlist)
```

Parameters, here for simplicity, otherwise kept in submission script. 
```{r}
print(params)
```

## Combined data

```{r}
combined_data <- read.delim(params$combined_data, header = TRUE, sep = "\t")
```

## Adjustments
```{r}
combined_data$pga_total <- combined_data$pga_clonal + combined_data$pga_subclonal
```

```{r}
combined_data$donor_relapse_interval <- as.numeric(combined_data$donor_relapse_interval)
```

# Merge data and select samples
Use Tier 1 Buffa signature sum.

## Selecting tumour only and one sample per patient
```{r}
hypoxia_signatures_tum_data_single <- subset(combined_data, !is.na(Buffa_signature_Sum_within_quartile_one_sample_per_donor))
```
```{r}
dim(hypoxia_signatures_tum_data_single)
```

```{r}
summary(hypoxia_signatures_tum_data_single)
```

Here is a list of the IDs that were used for the single tumour sample per patient analysis:
```{r}
hypoxia_signatures_tum_data_single$PPCG_SampleID
```

# Preparing survival data

```{r}
hist(hypoxia_signatures_tum_data_single$pga_clonal, breaks = 25)
hist(hypoxia_signatures_tum_data_single$pga_subclonal, breaks = 25)
hist((hypoxia_signatures_tum_data_single$pga_clonal + hypoxia_signatures_tum_data_single$pga_subclonal), breaks = 25)
```

## Survival - relapse
```{r}
hypoxia_signatures_tum_data_single$status_relapse <- NA
for(i in 1:nrow(hypoxia_signatures_tum_data_single)){
  if(is.na(hypoxia_signatures_tum_data_single$relapse_ind[i])){
    
  }else if(hypoxia_signatures_tum_data_single$relapse_ind[i] == "relapsed"){
    hypoxia_signatures_tum_data_single$status_relapse[i] <- 1
  }else if(hypoxia_signatures_tum_data_single$relapse_ind[i] == "no relapse"){
    hypoxia_signatures_tum_data_single$status_relapse[i] <- 0
  }
}
hypoxia_signatures_tum_data_single$donor_relapse_interval <- as.numeric(hypoxia_signatures_tum_data_single$donor_relapse_interval)
```


## Survival - mets
```{r}
hypoxia_signatures_tum_data_single$status_mets <- NA
for(i in 1:nrow(hypoxia_signatures_tum_data_single)){
  if(is.na(hypoxia_signatures_tum_data_single$mets_ind[i])){
    
  }else if(hypoxia_signatures_tum_data_single$mets_ind[i] == "mets"){
    hypoxia_signatures_tum_data_single$status_mets[i] <- 1
  }else if(hypoxia_signatures_tum_data_single$mets_ind[i] == "no mets"){
    hypoxia_signatures_tum_data_single$status_mets[i] <- 0
  }
}
hypoxia_signatures_tum_data_single$donor_first_mets_interval_from_m0 <- as.numeric(hypoxia_signatures_tum_data_single$donor_first_mets_interval_from_m0)
```

# Cox PH
```{r}
hypoxia_signatures_tum_data_single$Buffa_signature_Sum_within_quartile_one_sample_per_donor_bin <- ifelse(hypoxia_signatures_tum_data_single$Buffa_signature_Sum_within_quartile_one_sample_per_donor == "Q4", yes = "Q4", n = "Q1-Q3")
```

```{r}
hypoxia_signatures_tum_data_single$pga_status <- NA
for(i in 1:nrow(hypoxia_signatures_tum_data_single)){
  if(is.na(hypoxia_signatures_tum_data_single$pga_total[i])){
    
  }else if(hypoxia_signatures_tum_data_single$pga_total[i] > median(hypoxia_signatures_tum_data_single$pga_total, na.rm = TRUE)){
    hypoxia_signatures_tum_data_single$pga_status[i] <- "PGA High"
  }else if(hypoxia_signatures_tum_data_single$pga_total[i] <= median(hypoxia_signatures_tum_data_single$pga_total, na.rm = TRUE)){
    hypoxia_signatures_tum_data_single$pga_status[i] <- "PGA Low"
  }
}
```

```{r}
median(hypoxia_signatures_tum_data_single$pga_total, na.rm = TRUE)
```

## Preparing data for FinalFit
```{r}
hypoxia_signatures_tum_data_single$pga_status <- as.factor(hypoxia_signatures_tum_data_single$pga_status)
hypoxia_signatures_tum_data_single$pga_status <- factor(hypoxia_signatures_tum_data_single$pga_status, levels = c("PGA Low","PGA High"))

labels_df <- extract_variable_label(hypoxia_signatures_tum_data_single)

labels_df["Buffa_signature_Sum_within"] <- c("Hypoxia Score")
labels_df["Buffa_signature_Sum_within_quartile_one_sample_per_donor"] <- c("Hypoxia Quartile")
labels_df["Buffa_signature_Sum_within_quartile_one_sample_per_donor_bin"] <- c("Hypoxia Quartiles")
labels_df["pga_status"] <- c("Total PGA")
labels_df["age_at_tumour_collection"] <- c("Age")
labels_df["ISUP_Grade"] <- c("Grade Group")

hypoxia_signatures_tum_data_single <- ff_relabel(hypoxia_signatures_tum_data_single, labels_df)

```

```{r}
dependent_relapse <- "Surv(donor_relapse_interval, status_relapse)"
dependent_mets <- "Surv(donor_first_mets_interval_from_m0, status_mets)"

explanatory_cel   <- c("Buffa_signature_Sum_within", "pga_total", "age_at_tumour_collection", "ISUP_Grade", "Cellularity")
explanatory_quart_bin_rest_cel   <- c("Buffa_signature_Sum_within_quartile_one_sample_per_donor_bin", "pga_status", "age_at_tumour_collection", "ISUP_Grade","Cellularity")

```

## Relapse free survival
Get sample numbers
```{r}
full_rfs <- coxph(Surv(donor_relapse_interval, status_relapse) ~ Buffa_signature_Sum_within + pga_total + age_at_tumour_collection + ISUP_Grade + Cellularity, 
      data = hypoxia_signatures_tum_data_single)
```

```{r}
full_rfs
```

```{r}
full_rfs <- coxph(Surv(donor_relapse_interval, status_relapse) ~ Buffa_signature_Sum_within_quartile_one_sample_per_donor_bin + pga_status + age_at_tumour_collection + ISUP_Grade + Cellularity, 
      data = hypoxia_signatures_tum_data_single)
```

```{r}
full_rfs
```

### Finalfit

```{r}
hypoxia_signatures_tum_data_single %>% 
    finalfit(dependent_relapse, explanatory_cel, add_dependent_label = FALSE) %>% 
    rename("Relapse free survival" = label) %>% 
    rename(" " = levels) %>% 
    rename("  " = all) %>%
  xtable()
```
```{r}
hypoxia_signatures_tum_data_single %>% 
  hr_plot(dependent_relapse, explanatory_cel,
          plot_opts=list(geom_point(aes(size = Total), shape = 22, fill = "white"), 
                         geom_errorbarh(height=0.2), 
                         geom_vline(xintercept = 1, linetype = 1, alpha = 0.5, colour = "grey")),
          column_space = c(-1,0.2,1.5))
```

```{r}
hypoxia_signatures_tum_data_single %>% 
    finalfit(dependent_relapse, explanatory_quart_bin_rest_cel, add_dependent_label = FALSE) %>% 
    rename("Relapse free survival" = label) %>% 
    rename(" " = levels) %>% 
    rename("  " = all) %>%
  xtable()
```
```{r}
rfs_q4_rest_cel <- hypoxia_signatures_tum_data_single %>% 
  hr_plot(dependent_relapse, explanatory_quart_bin_rest_cel,
          plot_opts=list(geom_point(aes(size = Total), shape = 22, fill = "white"), 
                         geom_errorbarh(height=0.2), 
                         geom_vline(xintercept = 1, linetype = 1, alpha = 0.5, colour = "grey")),
          column_space = c(-1,0.2,1.5), 
          dependent_label = "C. Relapse Free Survival")
plot(rfs_q4_rest_cel)
```

```{r}
hypoxia_signatures_tum_data_single %>% 
    coxphmulti(dependent_relapse, explanatory_cel) %>% 
    cox.zph() %>% 
    {zph_result <<- .} %>% 
    plot()
```

```{r}
zph_result
```

```{r}
hypoxia_signatures_tum_data_single %>% 
    coxphmulti(dependent_relapse, explanatory_quart_bin_rest_cel) %>% 
    cox.zph() %>% 
    {zph_result <<- .} %>% 
    plot()
```

```{r}
zph_result
```

## Metastasis free survival
Get sample numbers
```{r}
full_mfs <- coxph(Surv(donor_first_mets_interval_from_m0, status_mets) ~ Buffa_signature_Sum_within + pga_total + age_at_tumour_collection + ISUP_Grade + Cellularity, 
      data = hypoxia_signatures_tum_data_single)
```

```{r}
full_mfs
```

```{r}
full_mfs <- coxph(Surv(donor_first_mets_interval_from_m0, status_mets) ~ Buffa_signature_Sum_within_quartile_one_sample_per_donor_bin + pga_status + age_at_tumour_collection + ISUP_Grade + Cellularity, 
      data = hypoxia_signatures_tum_data_single)
```

```{r}
full_mfs
```

### Finalfit

```{r}
hypoxia_signatures_tum_data_single %>% 
    finalfit(dependent_mets, explanatory_cel, add_dependent_label = FALSE) %>% 
    rename("Metastasis free survival" = label) %>% 
    rename(" " = levels) %>% 
    rename("  " = all) %>%
  xtable()
```
```{r}
hypoxia_signatures_tum_data_single %>% 
  hr_plot(dependent_mets, explanatory_cel,
          plot_opts=list(geom_point(aes(size = Total), shape = 22, fill = "white"), 
                         geom_errorbarh(height=0.2), 
                         geom_vline(xintercept = 1, linetype = 1, alpha = 0.5, colour = "grey")),
          column_space = c(-1,0.2,1.5))
```

```{r}
hypoxia_signatures_tum_data_single %>% 
    finalfit(dependent_mets, explanatory_quart_bin_rest_cel, add_dependent_label = FALSE) %>% 
    rename("Metastasis free survival" = label) %>% 
    rename(" " = levels) %>% 
    rename("  " = all) %>%
  xtable()
```
```{r}
mfs_q4_rest_cel <- hypoxia_signatures_tum_data_single %>% 
  hr_plot(dependent_mets, explanatory_quart_bin_rest_cel,
          plot_opts=list(geom_point(aes(size = Total), shape = 22, fill = "white"), 
                         geom_errorbarh(height=0.2), 
                         geom_vline(xintercept = 1, linetype = 1, alpha = 0.5, colour = "grey")),
          column_space = c(-1,0.2,1.5),
          dependent_label = "D. Metastasis Free Survival")
plot(mfs_q4_rest_cel)
```

```{r}
hypoxia_signatures_tum_data_single %>% 
    coxphmulti(dependent_mets, explanatory_cel) %>% 
    cox.zph() %>% 
    {zph_result <<- .} %>% 
    plot()
```

```{r}
zph_result
```

```{r}
hypoxia_signatures_tum_data_single %>% 
    coxphmulti(dependent_mets, explanatory_quart_bin_rest_cel) %>% 
    cox.zph() %>% 
    {zph_result <<- .} %>% 
    plot()
```

```{r}
zph_result
```

# Kaplan-Meier curves
```{r}
rel_lr <- survdiff(Surv(donor_relapse_interval,
                                                                         status_relapse) ~ Buffa_signature_Sum_within_quartile_one_sample_per_donor_bin,
                                                                    data=hypoxia_signatures_tum_data_single)
rel_lr
```

```{r fig.height=10, fig.width=8}
surv_rel_hypoxia <- survfit2(Surv(donor_relapse_interval, status_relapse) ~ Buffa_signature_Sum_within_quartile_one_sample_per_donor_bin, 
                             data = hypoxia_signatures_tum_data_single)

hypoxia_rel_bin <- ggsurvfit(surv_rel_hypoxia) +
  labs(
    x = "Days",
    y = "Relapse free survival probability"
    ) + 
  # add_confidence_interval() +
  # add_risktable()+
  scale_colour_manual(values = c("#3E6F84","#AE4C2C"))+
  scale_fill_manual(values = c("#3E6F84","#AE4C2C"))+
  # guides(colour=guide_legend(nrow=2,byrow=TRUE))+
  labs(tag = "A. ")+ 
  annotate("text", y = 0.95, x = 3000, label = c(paste0("log-rank p = ", 
                                                     format(rel_lr[[6]], scientific = TRUE, digits = 3))))
hypoxia_rel_bin
```

```{r}
mfs_lr <- survdiff(Surv(donor_first_mets_interval_from_m0,status_mets) ~ Buffa_signature_Sum_within_quartile_one_sample_per_donor_bin,
                   data=hypoxia_signatures_tum_data_single)
mfs_lr
```


```{r fig.height=10, fig.width=8}
surv_met_hypoxia <- survfit2(Surv(donor_first_mets_interval_from_m0, status_mets) ~ Buffa_signature_Sum_within_quartile_one_sample_per_donor_bin, 
                             data = hypoxia_signatures_tum_data_single)

hypoxia_mfs_bin <- ggsurvfit(surv_met_hypoxia) +
  labs(
    x = "Days",
    y = "Metastasis free survival probability"
    ) + 
  # add_confidence_interval() +
  # add_risktable()+
  scale_colour_manual(values = c("#3E6F84","#AE4C2C"))+
  scale_fill_manual(values = c("#3E6F84","#AE4C2C"))+
  # guides(colour=guide_legend(nrow=2,byrow=TRUE))+
  labs(tag = "B. ")+ 
  annotate("text", y = 0.98, x = 3000, label = c(paste0("log-rank p = ", 
                                                     format(mfs_lr[[6]], scientific = TRUE, digits = 3))))
hypoxia_mfs_bin
```

# Combined CoxPH plot

```{r fig.height=10, fig.width=7}
grid.arrange(hypoxia_rel_bin, hypoxia_mfs_bin,
             rfs_q4_rest_cel, mfs_q4_rest_cel, 
             nrow = 3, ncol = 4, layout_matrix = rbind(c(1,2),
                                                       c(3,3),
                                                       c(4,4)))
```


# Session Info
```{r}
sessionInfo()
```

 