---
title: "Figure 5 - survival features. RNA data release 1.4k10 prerelease"
author: Maria Jakobsdottir (Maria.Jakobsdottir@manchester.ac.uk)
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2:
    toc: true 
    toc_float: true
    toc_depth: 6
---

Analysis for figure 5 in Hypoxia manuscript.

# Libraries used
```{r message=FALSE, warning=FALSE}
library(dplyr)
library(finalfit)
library(ggplot2)
library(ggpubr)
library(ggsci)
library(ggsignif)
library(ggsurvfit)
library(grid)
library(gridExtra)
library(lmtest)
library(splitstackshape) # for paired analysis
library(survival)
library(tidyr)
library(viridis)
library(xtable)

options(bitmapType='cairo')
'%!in%' <- function(x,y)!('%in%'(x,y))
```

```{r}
hypoxia_colours_q <- list("Q1" = "#14556e",
                          "Q2" = "#67889a",
                          "Q3" = "#c06842",
                          "Q4" = "#9c2f15")
hypoxia_colours_q2 <- list("Q1" = "#8995bc",
                          "Q2" = "#8c7893",
                          "Q3" = "#955569",
                          "Q4" = "#823c52")
hypoxia_colours_surv <- c("#14556e","#14556e","#67889a","#67889a",  
                          "#c06842","#c06842","#9c2f15","#9c2f15")
```
```{r}
hypoxia_colours_bin <- list("Low" = "#3E6F84",
                            "High" = "#AE4C2C")
hypoxia_colours_bin2 <- list("Low" = "#8B87A8",
                             "High" = "#8C495E")
```

# Data
Read in data

```{r}
params <- list(ppcg_colourlist = "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/GitHub/PPCG_colours/PPCG_colourlist.RData",
               combined_data = "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_data_private/combined_PPCG_WGS_RNA_summary_data_clinical_v9_2026-05-31.txt")
```

```{r}
load(params$ppcg_colourlist)
```

Parameters, here for simplicity, otherwise kept in submission script. 
```{r}
print(params)
```
## Outdirs
```{r}
fig_outdir <- "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_paper_final_results/Figures/"
```

```{r}
fig_outdir_supp <- "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_paper_final_results/Supplementary_Figures/"
```

```{r}
tab_outdir_supp <- "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_paper_final_results/Supplementary_Tables/"
```

```{r}
fig_data_outdir <- "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_paper_final_results/Figure_Data/"
```

## Combined data

```{r}
combined_data <- read.delim(params$combined_data, header = TRUE, sep = "\t")
```

## Adjustments
```{r}
combined_data$pga_total <- combined_data$pga_clonal + combined_data$pga_subclonal
```

```{r}
combined_data$donor_relapse_interval <- as.numeric(combined_data$donor_relapse_interval)
```

# Functions

Modified `hr_plot` function from `finalfit` R package to print scientific p-values
```{r}
## Modified hr plot function to plot scientific p-values
#' Produce a hazard ratio table and plot
#'
#' Produce hazard ratio table and plot from a Cox Proportional Hazards analysis, \code{survival::coxph()}.
#'
#' @param .data Dataframe.
#' @param dependent Character vector of length 1:  name of survival object in form \code{Surv(time, status)}.
#' @param explanatory Character vector of any length: name(s) of explanatory variables.
#' @param factorlist Option to provide output directly from \code{\link{summary_factorlist}()}.
#' @param coxfit Option to provide output directly from \code{coxphmulti()}.
#' @param remove_ref Logical. Remove reference level for factors.
#' @param breaks Manually specify x-axis breaks in format \code{c(0.1, 1, 10)}.
#' @param column_space Adjust table column spacing.
#' @param dependent_label Main label for plot.
#' @param prefix Plots are titled by default with the dependent variable. This adds text before that label.
#' @param suffix Plots are titled with the dependent variable. This adds text after that label.
#' @param table_text_size Alter font size of table text.
#' @param title_text_size Alter font size of title text.
#' @param plot_opts A list of arguments to be appended to the ggplot call by "+".
#' @param table_opts A list of arguments to be appended to the ggplot table call by "+".
#' @param ... Other parameters passed to \code{fit2df()}.
#' @return Returns a table and plot produced in \code{ggplot2}.
#'
#' @family finalfit plot functions
#' @export
#'
#' @examples
#' # HR plot
#' library(finalfit)
#' library(dplyr)
#' library(ggplot2)
#'
#' explanatory = c("age.factor", "sex.factor", "obstruct.factor", "perfor.factor")
#' dependent = "Surv(time, status)"
#' colon_s %>%
#'   hr_plot(dependent, explanatory, dependent_label = "Survival")
#'
#' colon_s %>%
#'   hr_plot(dependent, explanatory, dependent_label = "Survival",
#'     table_text_size=4, title_text_size=14,
#'     plot_opts=list(xlab("HR, 95% CI"), theme(axis.title = element_text(size=12))))
#'
#' @import ggplot2
#'

hr_plot_mod = function(.data, dependent, explanatory, factorlist=NULL, coxfit=NULL,
                   remove_ref = FALSE,
                   breaks=NULL, column_space=c(-0.5, 0, 0.5),
                   dependent_label = "Survival", 
                   prefix = "", suffix = ": HR (95% CI, p-value)",
                   table_text_size = 4,
                   title_text_size = 13,
                   plot_opts = NULL, table_opts = NULL, ...){
  
  requireNamespace("ggplot2")
  
  # Generate or format factorlist object
  if(!is.null(factorlist)){
    if(is.null(factorlist$fit_id)) stop("summary_factorlist function must include fit_id=TRUE")
  }
  
  if(is.null(factorlist)){
    factorlist = summary_factorlist(.data, dependent, explanatory, fit_id=TRUE)
  }
  
  if(remove_ref){
    factorlist = factorlist %>%  
      dplyr::mutate(label = ifelse(label == "", NA, label)) %>% 
      tidyr::fill(label) %>% 
      dplyr::group_by(label) %>%
      dplyr::filter(dplyr::row_number() != 1 | 
                      dplyr::n() > 2 |
                      levels %in% c("Mean (SD)", "Median (IQR)")
      )%>% 
      rm_duplicate_labels()
  }
  
  # Specify breaks if provided
  if(is.null(breaks)){
    breaks = scales::breaks_log()
  }
  
  # Extract totals (this is CPH specific due to how summary_factorlist works)
  factorlist$Total = as.numeric(stringr::str_extract(as.character(factorlist$all), "^[:digit:]*"))
  
  # Fill in total for continuous variables
  factorlist$Total[which(factorlist$levels %in% c("Mean (SD)", "Median (IQR)"))] = dim(.data)[1]
  
  # For continuous variables, remove level label
  drop = grepl("Mean \\(SD\\)|Median \\(IQR\\)", factorlist$levels)
  factorlist$levels[drop] = "-"
  
  factorlist$all = NULL
  
  # Generate or format glm
  if(is.null(coxfit)){
    coxfit = coxphmulti(.data, dependent, explanatory)
  }
  coxfit_df_c = fit2df(coxfit, condense = TRUE, estimate_suffix = " (multivariable)", 
                       estimate_name = "HR", exp = TRUE)
  coxfit_df = fit2df(coxfit, condense = FALSE, 
                     estimate_name = "HR", exp = TRUE)

  
  # Merge
  df.out = finalfit_merge(factorlist, coxfit_df_c)
  df.out = finalfit_merge(df.out, coxfit_df, ref_symbol = "1.0")
  
  ###
  # EDIT: recreate printed information with scientific p-value
  ###
  df.out$`HR (multivariable)` <-  paste0(gsub("p=.*","",df.out$`HR (multivariable)`),"p=",format(df.out$p, scientific = TRUE, digits = 3),")")
  df.out$`HR (multivariable)` <- gsub("p=      NA)","",df.out$`HR (multivariable)`)
  
  ### end of EDIT
  
  # Remove unwanted lines, where there are more variables in model than wish to display.
  # Note merge function in summarizer merge is now `all` rather than `all.x` as wish to preserve interactions
  # These not named in factorlist, creating this problem. Interactions don't show on plot.
  if (any(
    is.na(df.out$label)
  )
  ){
    remove_rows = which(is.na(df.out$label)) # This row doesn't work when is.na == FALSE, hence if()
    df.out = df.out[-remove_rows,]
  } else {
    df.out
  }
  
  # Fix order
  df.out$levels = as.character(df.out$levels)
  df.out$fit_id = factor(df.out$fit_id, levels = df.out$fit_id[order(-df.out$index)])
  
  # Plot
  g1 = ggplot(df.out, aes(x = as.numeric(HR), xmin = as.numeric(L95), xmax  = as.numeric(U95),
                          y = fit_id))+
    geom_errorbarh(height=0.2) +
    geom_vline(xintercept = 1, linetype = "longdash", colour = "black")+
    geom_point(aes(size = Total), shape=22, fill="darkblue")+
    scale_x_continuous(trans="log10", breaks= breaks)+
    xlab("Hazard ratio (95% CI, log scale)")+ 
    theme_classic(11)+
    theme(axis.title.x = element_text(),
          axis.title.y = element_blank(),
          axis.text.y = element_blank(),
          axis.line.y = element_blank(),
          axis.ticks.y = element_blank(),
          legend.position="none")
  
  t1 = ggplot(df.out, aes(x = as.numeric(HR), y = fit_id))+
    annotate("text", x = column_space[1], y =  df.out$fit_id, label=df.out[,2], hjust=0, size=table_text_size)+
    annotate("text", x = column_space[2], y =  df.out$fit_id, label=df.out[,3], hjust=1, size=table_text_size)+
    annotate("text", x = column_space[3], y =  df.out$fit_id, label=df.out[,6], hjust=1, size=table_text_size)+
    theme_classic(14)+
    theme(axis.title.x = element_text(colour = "white"),
          axis.text.x = element_text(colour = "white"),
          axis.title.y = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          line = element_blank())
  
  # Add optional arguments
  g1 = g1 + plot_opts
  t1 = t1 + table_opts
  
  # Add dependent name label
  title_text = plot_title(.data, dependent, dependent_label = dependent_label, prefix = prefix, suffix = suffix)
  
  title_grob = cowplot::ggdraw() + 
    cowplot::draw_label(title_text, x = 0, hjust = 0, size = title_text_size) +
    theme(plot.margin = margin(0, 0, 0, 7))
  
  cowplot::plot_grid(title_grob,
                     cowplot::plot_grid(t1, g1, align = "h", rel_widths = c(3,2)),
                     ncol = 1, rel_heights = c(0.1, 1))
}

```


# Merge data and select samples
Use Tier 1 Buffa signature sum.

## Selecting tumour only and one sample per patient
```{r}
hypoxia_signatures_tum_data_single <- subset(combined_data, !is.na(Buffa_signature_Sum_within_quartile_one_sample_per_donor))
```
```{r}
dim(hypoxia_signatures_tum_data_single)
```

```{r}
summary(hypoxia_signatures_tum_data_single)
```

Here is a list of the IDs that were used for the single tumour sample per patient analysis:
```{r}
hypoxia_signatures_tum_data_single$PPCG_SampleID
```

# Preparing survival data

```{r}
hist(hypoxia_signatures_tum_data_single$pga_clonal, breaks = 25)
hist(hypoxia_signatures_tum_data_single$pga_subclonal, breaks = 25)
hist((hypoxia_signatures_tum_data_single$pga_clonal + hypoxia_signatures_tum_data_single$pga_subclonal), breaks = 25)
```

## Survival - relapse
```{r}
hypoxia_signatures_tum_data_single$status_relapse <- NA
for(i in 1:nrow(hypoxia_signatures_tum_data_single)){
  if(is.na(hypoxia_signatures_tum_data_single$relapse_ind[i])){
    
  }else if(hypoxia_signatures_tum_data_single$relapse_ind[i] == "relapsed"){
    hypoxia_signatures_tum_data_single$status_relapse[i] <- 1
  }else if(hypoxia_signatures_tum_data_single$relapse_ind[i] == "no relapse"){
    hypoxia_signatures_tum_data_single$status_relapse[i] <- 0
  }
}
hypoxia_signatures_tum_data_single$donor_relapse_interval <- as.numeric(hypoxia_signatures_tum_data_single$donor_relapse_interval)
```


## Survival - mets
```{r}
hypoxia_signatures_tum_data_single$status_mets <- NA
for(i in 1:nrow(hypoxia_signatures_tum_data_single)){
  if(is.na(hypoxia_signatures_tum_data_single$mets_ind[i])){
    
  }else if(hypoxia_signatures_tum_data_single$mets_ind[i] == "mets"){
    hypoxia_signatures_tum_data_single$status_mets[i] <- 1
  }else if(hypoxia_signatures_tum_data_single$mets_ind[i] == "no mets"){
    hypoxia_signatures_tum_data_single$status_mets[i] <- 0
  }
}
hypoxia_signatures_tum_data_single$donor_first_mets_interval_from_m0 <- as.numeric(hypoxia_signatures_tum_data_single$donor_first_mets_interval_from_m0)
```

# Cox PH
```{r}
hypoxia_signatures_tum_data_single$Buffa_signature_Sum_within_quartile_one_sample_per_donor_bin <- ifelse(hypoxia_signatures_tum_data_single$Buffa_signature_Sum_within_quartile_one_sample_per_donor == "Q4", yes = "Q4", n = "Q1-Q3")
```

```{r}
hypoxia_signatures_tum_data_single$pga_status <- NA
for(i in 1:nrow(hypoxia_signatures_tum_data_single)){
  if(is.na(hypoxia_signatures_tum_data_single$pga_total[i])){
    
  }else if(hypoxia_signatures_tum_data_single$pga_total[i] > median(hypoxia_signatures_tum_data_single$pga_total, na.rm = TRUE)){
    hypoxia_signatures_tum_data_single$pga_status[i] <- "PGA High"
  }else if(hypoxia_signatures_tum_data_single$pga_total[i] <= median(hypoxia_signatures_tum_data_single$pga_total, na.rm = TRUE)){
    hypoxia_signatures_tum_data_single$pga_status[i] <- "PGA Low"
  }
}
```

```{r}
median(hypoxia_signatures_tum_data_single$pga_total, na.rm = TRUE)
```

## Preparing data for FinalFit
```{r}
hypoxia_signatures_tum_data_single$pga_status <- as.factor(hypoxia_signatures_tum_data_single$pga_status)
hypoxia_signatures_tum_data_single$pga_status <- factor(hypoxia_signatures_tum_data_single$pga_status, levels = c("PGA Low","PGA High"))

labels_df <- extract_variable_label(hypoxia_signatures_tum_data_single)

labels_df["Buffa_signature_Sum_within"] <- c("Hypoxia Score")
labels_df["Buffa_signature_Sum_within_quartile_one_sample_per_donor"] <- c("Hypoxia Quartile")
labels_df["Buffa_signature_Sum_within_quartile_one_sample_per_donor_bin"] <- c("Hypoxia Quartiles")
labels_df["pga_status"] <- c("Total PGA")
labels_df["age_at_tumour_collection"] <- c("Age")
labels_df["ISUP_Grade"] <- c("Grade Group")

hypoxia_signatures_tum_data_single <- ff_relabel(hypoxia_signatures_tum_data_single, labels_df)

```

```{r}
dependent_relapse <- "Surv(donor_relapse_interval, status_relapse)"
dependent_mets <- "Surv(donor_first_mets_interval_from_m0, status_mets)"

explanatory_cel   <- c("Buffa_signature_Sum_within", "pga_total", "age_at_tumour_collection", "ISUP_Grade", "Cellularity")
explanatory_quart_bin_rest_cel   <- c("Buffa_signature_Sum_within_quartile_one_sample_per_donor_bin", "pga_status", "age_at_tumour_collection", "ISUP_Grade","Cellularity")

```

## Relapse free survival
Get sample numbers
```{r}
full_rfs <- coxph(Surv(donor_relapse_interval, status_relapse) ~ Buffa_signature_Sum_within + pga_total + age_at_tumour_collection + ISUP_Grade + Cellularity, 
      data = hypoxia_signatures_tum_data_single)
```

```{r}
full_rfs
```

```{r}
full_rfs <- coxph(Surv(donor_relapse_interval, status_relapse) ~ Buffa_signature_Sum_within_quartile_one_sample_per_donor_bin + pga_status + age_at_tumour_collection + ISUP_Grade + Cellularity, 
      data = hypoxia_signatures_tum_data_single)
```

```{r}
full_rfs
```

### Finalfit

```{r}
hypoxia_signatures_tum_data_single %>% 
    finalfit(dependent_relapse, explanatory_cel, add_dependent_label = FALSE, condense = FALSE) %>% 
    rename("Relapse free survival" = label) %>% 
    rename(" " = levels) %>% 
    rename("  " = all) %>%
  xtable(display = c("d","s","s","s","f","f","f","g","f","f","f","g"))
```

```{r}
print.xtable(hypoxia_signatures_tum_data_single %>% 
    finalfit(dependent_relapse, explanatory_cel, add_dependent_label = FALSE, condense = FALSE) %>% 
    rename("Relapse free survival" = label) %>% 
    rename(" " = levels) %>% 
    rename("  " = all) %>%
  xtable(display = c("d","s","s","s","f","f","f","g","f","f","f","g")), 
       type = "latex",
             file = paste0(tab_outdir_supp,"supplementary_table_8a_relapse_continuous_",Sys.Date(),".tex"))
```

```{r}
hypoxia_signatures_tum_data_single %>% 
  hr_plot_mod(dependent_relapse, explanatory_cel,
          plot_opts=list(geom_point(aes(size = Total), shape = 22, fill = "white"), 
                         geom_errorbarh(height=0.2), 
                         geom_vline(xintercept = 1, linetype = 1, alpha = 0.5, colour = "grey")),
          column_space = c(-1,0.2,1.5))
```

```{r}
hypoxia_signatures_tum_data_single %>% 
    finalfit(dependent_relapse, explanatory_quart_bin_rest_cel, add_dependent_label = FALSE, condense = FALSE) %>% 
    rename("Relapse free survival" = label) %>% 
    rename(" " = levels) %>% 
    rename("  " = all) %>%
  xtable(display = c("d","s","s","s","f","f","f","g","f","f","f","g"))
```

```{r}
print.xtable(hypoxia_signatures_tum_data_single %>% 
    finalfit(dependent_relapse, explanatory_quart_bin_rest_cel, add_dependent_label = FALSE, condense = FALSE) %>% 
    rename("Relapse free survival" = label) %>% 
    rename(" " = levels) %>% 
    rename("  " = all) %>%
  xtable(display = c("d","s","s","s","f","f","f","g","f","f","f","g")), 
       type = "latex",
             file = paste0(tab_outdir_supp,"supplementary_table_relapse_Q4_rest_",Sys.Date(),".tex"))
```

```{r}
rfs_q4_rest_cel <- hypoxia_signatures_tum_data_single %>% 
  hr_plot_mod(dependent_relapse, explanatory_quart_bin_rest_cel,
          plot_opts=list(geom_point(aes(size = Total), shape = 22, fill = "white"), 
                         geom_errorbarh(height=0.2), 
                         geom_vline(xintercept = 1, linetype = 1, alpha = 0.5, colour = "grey")),
          column_space = c(-1,0.2,1.5), 
          dependent_label = "C. Relapse Free Survival")
plot(rfs_q4_rest_cel)
```

```{r}
hypoxia_signatures_tum_data_single %>% 
    coxphmulti(dependent_relapse, explanatory_cel) %>% 
    cox.zph() %>% 
    {zph_result <<- .} %>% 
    plot()
```

```{r}
zph_result
```

```{r}
hypoxia_signatures_tum_data_single %>% 
    coxphmulti(dependent_relapse, explanatory_quart_bin_rest_cel) %>% 
    cox.zph() %>% 
    {zph_result <<- .} %>% 
    plot()
```

```{r}
zph_result
```

## Metastasis free survival
Get sample numbers
```{r}
full_mfs <- coxph(Surv(donor_first_mets_interval_from_m0, status_mets) ~ Buffa_signature_Sum_within + pga_total + age_at_tumour_collection + ISUP_Grade + Cellularity, 
      data = hypoxia_signatures_tum_data_single)
```

```{r}
full_mfs
```

```{r}
full_mfs <- coxph(Surv(donor_first_mets_interval_from_m0, status_mets) ~ Buffa_signature_Sum_within_quartile_one_sample_per_donor_bin + pga_status + age_at_tumour_collection + ISUP_Grade + Cellularity, 
      data = hypoxia_signatures_tum_data_single)
```

```{r}
full_mfs
```

### Finalfit

```{r}
hypoxia_signatures_tum_data_single %>% 
    finalfit(dependent_mets, explanatory_cel, add_dependent_label = FALSE, condense = FALSE) %>% 
    rename("Metastasis free survival" = label) %>% 
    rename(" " = levels) %>% 
    rename("  " = all) %>%
  xtable(display = c("d","s","s","s","f","f","f","g","f","f","f","g"))
```

```{r}
print.xtable(hypoxia_signatures_tum_data_single %>% 
    finalfit(dependent_mets, explanatory_cel, add_dependent_label = FALSE, condense = FALSE) %>% 
    rename("Metastasis free survival" = label) %>% 
    rename(" " = levels) %>% 
    rename("  " = all) %>%
  xtable(display = c("d","s","s","s","f","f","f","g","f","f","f","g")), 
       type = "latex",
             file = paste0(tab_outdir_supp,"supplementary_table_8b_mfs_continuous_",Sys.Date(),".tex"))
```

```{r}
hypoxia_signatures_tum_data_single %>% 
  hr_plot_mod(dependent_mets, explanatory_cel,
          plot_opts=list(geom_point(aes(size = Total), shape = 22, fill = "white"), 
                         geom_errorbarh(height=0.2), 
                         geom_vline(xintercept = 1, linetype = 1, alpha = 0.5, colour = "grey")),
          column_space = c(-1,0.2,1.5))
```

```{r}
hypoxia_signatures_tum_data_single %>% 
    finalfit(dependent_mets, explanatory_quart_bin_rest_cel, add_dependent_label = FALSE, condense = FALSE) %>% 
    rename("Metastasis free survival" = label) %>% 
    rename(" " = levels) %>% 
    rename("  " = all) %>%
  xtable(display = c("d","s","s","s","f","f","f","g","f","f","f","g"))
```

```{r}
print.xtable(hypoxia_signatures_tum_data_single %>% 
    finalfit(dependent_mets, explanatory_quart_bin_rest_cel, add_dependent_label = FALSE, condense = FALSE) %>% 
    rename("Metastasis free survival" = label) %>% 
    rename(" " = levels) %>% 
    rename("  " = all) %>%
  xtable(display = c("d","s","s","s","f","f","f","g","f","f","f","g")), 
       type = "latex",
             file = paste0(tab_outdir_supp,"supplementary_table_mfs_Q4_rest_",Sys.Date(),".tex"))
```

```{r}
mfs_q4_rest_cel <- hypoxia_signatures_tum_data_single %>% 
  hr_plot_mod(dependent_mets, explanatory_quart_bin_rest_cel,
          plot_opts=list(geom_point(aes(size = Total), shape = 22, fill = "white"), 
                         geom_errorbarh(height=0.2), 
                         geom_vline(xintercept = 1, linetype = 1, alpha = 0.5, colour = "grey")),
          column_space = c(-1,0.2,1.5),
          dependent_label = "D. Metastasis Free Survival")
plot(mfs_q4_rest_cel)
```

```{r}
hypoxia_signatures_tum_data_single %>% 
    coxphmulti(dependent_mets, explanatory_cel) %>% 
    cox.zph() %>% 
    {zph_result <<- .} %>% 
    plot()
```

```{r}
zph_result
```

```{r}
hypoxia_signatures_tum_data_single %>% 
    coxphmulti(dependent_mets, explanatory_quart_bin_rest_cel) %>% 
    cox.zph() %>% 
    {zph_result <<- .} %>% 
    plot()
```

```{r}
zph_result
```

#### Saving finalfit data

```{r}
# Relapse HR (Q4 vs rest) data save (variables used by explanatory_quart_bin_rest_cel + survival)
rfs_q4_data <- hypoxia_signatures_tum_data_single[, c("donor_relapse_interval",
                                                      "status_relapse",
                                                      "Buffa_signature_Sum_within_quartile_one_sample_per_donor_bin",
                                                      "pga_status",
                                                      "age_at_tumour_collection",
                                                      "ISUP_Grade",
                                                      "Cellularity")]
write.csv(rfs_q4_data, file = paste0(fig_data_outdir,"Figure_5_rfs_q4_data_",Sys.Date(),".csv"), row.names = FALSE)
```

```{r}
# recreate HR plot from saved data

rfs_q4_rest_cel <- rfs_q4_data %>% 
  hr_plot_mod(dependent_relapse, explanatory_quart_bin_rest_cel,
          plot_opts=list(geom_point(aes(size = Total), shape = 22, fill = "white"), 
                         geom_errorbarh(height=0.2), 
                         geom_vline(xintercept = 1, linetype = 1, alpha = 0.5, colour = "grey")),
          column_space = c(-1,0.2,1.5), 
          dependent_label = "C. Relapse Free Survival")
plot(rfs_q4_rest_cel)
```

Recreate table from saved data
```{r}
rfs_q4_data %>% 
    finalfit(dependent_relapse, explanatory_quart_bin_rest_cel, add_dependent_label = FALSE, condense = FALSE) %>% 
    rename("Relapse free survival" = label) %>% 
    rename(" " = levels) %>% 
    rename("  " = all) %>%
  xtable(display = c("d","s","s","s","f","f","f","g","f","f","f","g"))
```

```{r}
# Metastasis HR (Q4 vs rest) data save
mfs_q4_data <- hypoxia_signatures_tum_data_single[, c("donor_first_mets_interval_from_m0",
                                                      "status_mets",
                                                      "Buffa_signature_Sum_within_quartile_one_sample_per_donor_bin",
                                                      "pga_status",
                                                      "age_at_tumour_collection",
                                                      "ISUP_Grade",
                                                      "Cellularity")]
write.csv(mfs_q4_data, file = paste0(fig_data_outdir,"Figure_5_mfs_q4_data_",Sys.Date(),".csv"), row.names = FALSE)
```

```{r}
# recreate HR plot from saved data
mfs_q4_rest_cel <- mfs_q4_data %>% 
  hr_plot_mod(dependent_mets, explanatory_quart_bin_rest_cel,
          plot_opts=list(geom_point(aes(size = Total), shape = 22, fill = "white"), 
                         geom_errorbarh(height=0.2), 
                         geom_vline(xintercept = 1, linetype = 1, alpha = 0.5, colour = "grey")),
          column_space = c(-1,0.2,1.5),
          dependent_label = "D. Metastasis Free Survival")
plot(mfs_q4_rest_cel)
```

Recreate table from saved data
```{r}
mfs_q4_data %>% 
    finalfit(dependent_mets, explanatory_quart_bin_rest_cel, add_dependent_label = FALSE, condense = FALSE) %>% 
    rename("Metastasis free survival" = label) %>% 
    rename(" " = levels) %>% 
    rename("  " = all) %>%
  xtable(display = c("d","s","s","s","f","f","f","g","f","f","f","g"))
```

# Kaplan-Meier curves
```{r}
rel_lr <- survdiff(Surv(donor_relapse_interval,
                                                                         status_relapse) ~ Buffa_signature_Sum_within_quartile_one_sample_per_donor_bin,
                                                                    data=hypoxia_signatures_tum_data_single)
rel_lr
```

```{r fig.height=10, fig.width=8}
surv_rel_hypoxia <- survfit2(Surv(donor_relapse_interval, status_relapse) ~ Buffa_signature_Sum_within_quartile_one_sample_per_donor_bin, 
                             data = hypoxia_signatures_tum_data_single)
```

```{r}
relapse_km_data <- hypoxia_signatures_tum_data_single[, c("PPCG_Donor_ID",
                                                          "donor_relapse_interval",
                                                          "status_relapse",
                                                          "Buffa_signature_Sum_within",
                                                          "Buffa_signature_Sum_within_quartile_one_sample_per_donor",
                                                          "Buffa_signature_Sum_within_quartile_one_sample_per_donor_bin")]
write.csv(relapse_km_data, file = paste0(fig_data_outdir,"Figure_5_relapse_km_data_",Sys.Date(),".csv"), row.names = FALSE)
```

```{r}
surv_rel_hypoxia <- survfit2(Surv(donor_relapse_interval, status_relapse) ~ Buffa_signature_Sum_within_quartile_one_sample_per_donor_bin, 
                             data = relapse_km_data)

hypoxia_rel_bin <- ggsurvfit(surv_rel_hypoxia) +
  labs(
    x = "Days",
    y = "Relapse free survival probability"
    ) + 
  # add_confidence_interval() +
  # add_risktable()+
  scale_colour_manual(values = c("#3E6F84","#AE4C2C"))+
  scale_fill_manual(values = c("#3E6F84","#AE4C2C"))+
  # guides(colour=guide_legend(nrow=2,byrow=TRUE))+
  labs(tag = "A. ")+ 
  annotate("text", y = 0.95, x = 3000, label = c(paste0("log-rank p = ", 
                                                     format(rel_lr[[6]], scientific = TRUE, digits = 3))))
hypoxia_rel_bin
```

```{r}
mfs_lr <- survdiff(Surv(donor_first_mets_interval_from_m0,status_mets) ~ Buffa_signature_Sum_within_quartile_one_sample_per_donor_bin,
                   data=hypoxia_signatures_tum_data_single)
mfs_lr
```


```{r fig.height=10, fig.width=8}
surv_met_hypoxia <- survfit2(Surv(donor_first_mets_interval_from_m0, status_mets) ~ Buffa_signature_Sum_within_quartile_one_sample_per_donor_bin, 
                             data = hypoxia_signatures_tum_data_single)
```

```{r}
mets_km_data <- hypoxia_signatures_tum_data_single[, c("PPCG_Donor_ID",
                                                       "donor_first_mets_interval_from_m0",
                                                       "status_mets",
                                                       "Buffa_signature_Sum_within",
                                                       "Buffa_signature_Sum_within_quartile_one_sample_per_donor",
                                                       "Buffa_signature_Sum_within_quartile_one_sample_per_donor_bin")]
write.csv(mets_km_data, file = paste0(fig_data_outdir,"Figure_5_mets_km_data_",Sys.Date(),".csv"), row.names = FALSE)
```

```{r}
surv_met_hypoxia <- survfit2(Surv(donor_first_mets_interval_from_m0, status_mets) ~ Buffa_signature_Sum_within_quartile_one_sample_per_donor_bin, 
                             data = mets_km_data)

hypoxia_mfs_bin <- ggsurvfit(surv_met_hypoxia) +
  labs(
    x = "Days",
    y = "Metastasis free survival probability"
    ) + 
  # add_confidence_interval() +
  # add_risktable()+
  scale_colour_manual(values = c("#3E6F84","#AE4C2C"))+
  scale_fill_manual(values = c("#3E6F84","#AE4C2C"))+
  # guides(colour=guide_legend(nrow=2,byrow=TRUE))+
  labs(tag = "B. ")+ 
  annotate("text", y = 0.98, x = 3000, label = c(paste0("log-rank p = ", 
                                                     format(mfs_lr[[6]], scientific = TRUE, digits = 3))))
hypoxia_mfs_bin
```

# Combined CoxPH plot

```{r fig.height=10, fig.width=7}
grid.arrange(hypoxia_rel_bin, hypoxia_mfs_bin,
             rfs_q4_rest_cel, mfs_q4_rest_cel, 
             nrow = 3, ncol = 4, layout_matrix = rbind(c(1,2),
                                                       c(3,3),
                                                       c(4,4)))
```

```{r}
pdf(file = paste0(fig_outdir,"/","Figure_5_",Sys.Date(),".pdf"), width = 12, height = 10)
grid.arrange(hypoxia_rel_bin, hypoxia_mfs_bin,
             rfs_q4_rest_cel, mfs_q4_rest_cel, 
             nrow = 3, ncol = 4, layout_matrix = rbind(c(1,2),
                                                       c(3,3),
                                                       c(4,4)))
dev.off()
```


# Session Info
```{r}
sessionInfo()
```