---
title: "Manuscript Dec 2024, Figure 1. RNA data release 1.4k10prerelease"
author: Maria Jakobsdottir (Maria.Jakobsdottir@manchester.ac.uk)
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2:
    toc: true 
    toc_float: true
    toc_depth: 6
---

Analysis for figure 1 in Hypoxia manuscript December 2024.

# Libraries used
```{r message=FALSE, warning=FALSE}
library(circlize)
library(ComplexHeatmap)
library(data.table)
library(ggplot2)
library(ggsci)
library(ggsignif)
library(ggthemes)
library(gridExtra)
library(RColorBrewer)
library(scales)
library(xtable)

options(bitmapType='cairo')
'%!in%' <- function(x,y)!('%in%'(x,y))
```

Heatmap parameters
```{r}
ht_opt$heatmap_row_names_gp = gpar(fontsize = 12)
```

# Data
Read in data

```{r}
params <- list(ppcg_colourlist = "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/GitHub/PPCG_colours/PPCG_colourlist.RData",
               combined_data = "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_data_private/combined_PPCG_WGS_RNA_summary_data_clinical_v7_2024-12-06.txt",
               drivers = "/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_Drivers/Somatic_drivers/DRIVER_ANNOTATIONS__2023-09-25_release4/DRIVER_ANNOTATIONS__2023-09-25_release4__SNV+indel__filteredVariants__allGenes.txt",
               evofeatures = "/mnt/bmh01-rds/UoOxford_David_W/shared/ICGC_UK_RNA/ppcg_subtyping_weagles_nadiah_15112021.csv",
               rel_cna = "/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_Drivers/Somatic_drivers/CNA_ANNOTATIONS__2023-09-13_release4/relative_cna_table_by_gene_pam.tsv",
               total_model_coefs = "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_collaboration/manuscript_dec_2024_outputs/robustlmm_outputs/total_coefficient_tables_k10_rearranged_20241204.RData",
               nonzero_model_coefs = "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_collaboration/manuscript_dec_2024_outputs/robustlmm_outputs/nonzero_coefficient_tables_k10_rearranged_20241204.RData")
```

```{r}
load(params$ppcg_colourlist)
```
```{r}
hypoxia_colours_q <- list("Q1" = "#14556e",
                          "Q2" = "#67889a",
                          "Q3" = "#c06842",
                          "Q4" = "#9c2f15")
```

Parameters, here for simplicity, otherwise kept in submission script. 
```{r}
print(params)
```

## Combined data

```{r}
combined_data <- read.delim(params$combined_data, header = TRUE, sep = "\t")
```

## Driver data
Data release version: 4

Date: 2023-09-25

Source: Juri Reimand - FTP box

/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_Drivers/Somatic_drivers/DRIVER_ANNOTATIONS__2023-09-25_release4/DRIVER_ANNOTATIONS__2023-09-25_release4__SNV+indel__filteredVariants__allGenes.txt
```{r}
driver_data <- read.delim(params$drivers, row.names = 1, sep = "\t", header = TRUE)
```

## Relative CN in GISTIC peaks
Data release version: 4

Date: 2023-09-13

Source: Juri Reimand - FTP box

/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_Drivers/Somatic_drivers/Driver_annotations_2023-01-18/DRIVER_ANNOTATIONS__2023-01-18_release2__SNV+indel__filteredVariants__driverGenes.txt
```{r}
rel_cna <- read.delim(params$rel_cna, sep = "\t", header = TRUE)
```


## Other data
### Gene lists

```{r}
drg_genes_36 <- c("ATM","BAP1","BLM","BRCA1","BRCA2",
               "BRIP1","CDH1","CDK4","CDKN2A","CHEK2",
               "FANCA","FANCD2","FANCI","FANCL","FANCM",
               "MLH1","MSH2","MSH6","NBN","NTHL1",
               "PALB2","PMS2","POLD1","POLE","POT1",
               "PTCH1","PTEN","RAD51C","RAD51D","RB1",
               "SMAD4","SMARCA4","STK11","TP53","XPA",
               "XPC")
```

```{r}
drg_genes_128 <- c("AKT1","ALKBH3","APEX1","APLF","ATM",
                   "ATR","ATRIP","BAP1","BARD1","BLM",
                   "BRCA1","BRCA2","BRIP1","CCNH","CDC25A",
                   "CDC25C","CDH1","CDK1","CDK2","CDK4",
                   "CDKN2A","CHEK1","CHEK2","CLK2","DCLRE1A",
                   "DDB1","EME1","EME2","EPCAM","ERCC2",
                   "ERCC3","ERCC5","ERCC6","ESR2","EXO1",
                   "FAM175A","FANCA","FANCD2","FANCI","FANCL",
                   "FANCM","GADD45A","GEN1","GTF2H2","GTF2H3",
                   "GTF2H4","H2AFX","HUS1","LIG1","LIG3",
                   "LIG4","MLH1","MLH3","MMS19","MNAT1",
                   "MPG","MRE11","MRE11A","MSH2","MSH3",
                   "MSH5","MSH6","MSR1","NABP2","NBN",
                   "NEIL1","NEIL2","NHEJ1","NTHL1","OGG1",
                   "PALB2","PARP2","PER1","PER2","PMS1",
                   "PMS2","PNKP","POLD1","POLE","POLH",
                   "POLK","POLL","POLM","POLN","POLQ",
                   "POT1","PRSS1","PTCH1","PTEN","RAD1",
                   "RAD50","RAD51","RAD51B","RAD51C","RAD51D",
                   "RAD52","RAD54B","RAD54L","RB1","RDM1",
                   "RECQL","RECQL4","RECQL5","RINT1","RNF8",
                   "RPA1","RPA3","SETMAR","SLX4","SMAD4",
                   "SMARCA4","SMUG1","STK11","TDG","TDP1",
                   "TOP2A","TOP2B","TOP3A","TP53","TP53BP1",
                   "WRN","XAB2","XPA","XPC","XRCC1",
                   "XRCC2","XRCC4","XRCC5")
```

### Robustlmm coefficients 
```{r}
load(params$total_model_coefs)
```

```{r}
load(params$nonzero_model_coefs)
```

### IDC group renaming 
```{r}
combined_data$Cribriform_Intraductal_Y_N[combined_data$Cribriform_Intraductal_Y_N == "Y"] <- "IDC"
combined_data$Cribriform_Intraductal_Y_N[combined_data$Cribriform_Intraductal_Y_N == "N"] <- "No IDC"
combined_data$Cribriform_Intraductal_Y_N <- as.factor(combined_data$Cribriform_Intraductal_Y_N)
combined_data$Cribriform_Intraductal_Y_N <- factor(combined_data$Cribriform_Intraductal_Y_N, levels = c("No IDC","IDC"))
```


# Merge data and select samples
Use Tier 1 Buffa signature sum.

## Selecting tumour only and one sample per patient
Select only tumour samples. Then only one sample per individual donor. 

```{r}
hypoxia_signatures_tum_data_single <- subset(combined_data, !is.na(Buffa_signature_Sum_within_quartile_one_sample_per_donor))
```

```{r}
dim(hypoxia_signatures_tum_data_single)
```

```{r}
summary(hypoxia_signatures_tum_data_single)
```

## Tidy data levels and labels

```{r}
hypoxia_signatures_tum_data_single$gleason_total <- as.factor(hypoxia_signatures_tum_data_single$gleason_total)
hypoxia_signatures_tum_data_single$gleason_total <- factor(hypoxia_signatures_tum_data_single$gleason_total, levels = c("3+3","3+4","4+3","4+4","3+5","4+5","5+4","5+5"))
```

```{r}
hypoxia_signatures_tum_data_single$psa_bracket <- as.factor(hypoxia_signatures_tum_data_single$psa_bracket)
hypoxia_signatures_tum_data_single$psa_bracket <- factor(hypoxia_signatures_tum_data_single$psa_bracket, levels = c("0-9.9","10-19.9","≥20"))
```

```{r}
hypoxia_signatures_tum_data_single$age_bracket <- as.factor(hypoxia_signatures_tum_data_single$age_bracket)
hypoxia_signatures_tum_data_single$age_bracket <- factor(hypoxia_signatures_tum_data_single$age_bracket, levels = c("< 40","40-50","50-65","65-70","≥ 70"))
```

Here is a list of the IDs that were used for the single tumour sample per patient analysis:
```{r}
hypoxia_signatures_tum_data_single$PPCG_SampleID
```

## CN signature numbers
```{r}
cna_colnames <- c("CX1",  "CX2",  "CX3",  "CX4",  "CX5",  "CX7",  "CX8",  "CX11")
cna_colnames_counts <- paste0(cna_colnames,"_numbers")
```

### Multiply CNA signature proportions by number of CNAs
```{r}
hypoxia_signatures_tum_data_single[,cna_colnames_counts] <- hypoxia_signatures_tum_data_single[,cna_colnames]*hypoxia_signatures_tum_data_single$num_cnvs
```

# Clinical data associations table

```{r}
hypoxia_isup <- ggplot(subset(hypoxia_signatures_tum_data_single, !is.na(ISUP_Grade)), aes(x = as.factor(ISUP_Grade), y = Buffa_signature_Sum_within))+
  geom_jitter(width = 0.3, alpha = 0.3, aes(colour = as.factor(ISUP_Grade)), height = 0)+
  geom_boxplot(alpha = 0.5, outlier.colour = NA)+ # no outlier plotting because we have geom_jitter to plot the points
  theme_light()+
  scale_colour_manual(values = PPCG_colourlist$ISUP_colourlist)+
  xlab("Gleason Grade Group")+
  ylab("Hypoxia Score")+
  # ggtitle("ISUP Grade")+
  # geom_signif(map_signif_level = TRUE, tip_length = 0, comparisons = list(c("1","2"),
  #                                                                         c("2","3"),
  #                                                                         c("3","4"),
  #                                                                         c("4","5")))+
  theme(text = element_text(size = 16), legend.position = "none")

# hypoxia_survival
hypoxia_isup
```
```{r}
summary(hypoxia_signatures_tum_data_single$ISUP_Grade)
length(hypoxia_signatures_tum_data_single$ISUP_Grade)
```

```{r}
kruskal.test(hypoxia_signatures_tum_data_single$Buffa_signature_Sum_within,
             hypoxia_signatures_tum_data_single$ISUP_Grade)
```

```{r}
wilcox_isup_grade_stats <- as.data.frame(matrix(nrow = 10, ncol = 6))
colnames(wilcox_isup_grade_stats) <- c("Status_1","Status_2","n_Status_1","n_Status_2","wilcox_p","wilcox_p_adjusted")
wilcox_isup_grade_stats$Status_1 <- c(rep("1",4),rep("2",3),rep("3",2),"4")
wilcox_isup_grade_stats$Status_2 <- c("2","3","4","5",
                                      "3","4","5",
                                      "4","5",
                                      "5")

for(i in 1:nrow(wilcox_isup_grade_stats)){
  
  wilcox_isup_grade_stats$n_Status_1[i] <- nrow(subset(hypoxia_signatures_tum_data_single, ISUP_Grade == wilcox_isup_grade_stats$Status_1[i]))
  wilcox_isup_grade_stats$n_Status_2[i] <- nrow(subset(hypoxia_signatures_tum_data_single, ISUP_Grade == wilcox_isup_grade_stats$Status_2[i]))
  
  wilcox_isup_grade_stats$wilcox_p[i] <- wilcox.test(subset(hypoxia_signatures_tum_data_single, 
                                                            ISUP_Grade == wilcox_isup_grade_stats$Status_1[i])$Buffa_signature_Sum_within,
                                                     subset(hypoxia_signatures_tum_data_single, 
                                                            ISUP_Grade == wilcox_isup_grade_stats$Status_2[i])$Buffa_signature_Sum_within)[[3]]
}

wilcox_isup_grade_stats$wilcox_p_adjusted <- p.adjust(wilcox_isup_grade_stats$wilcox_p, method = "BH")
```

```{r}
wilcox_isup_grade_stats
```

```{r}
hypoxia_age <- ggplot(hypoxia_signatures_tum_data_single, aes(x = age_at_tumour_collection, y = Buffa_signature_Sum_within))+
  geom_point(alpha = 0.5)+ # no outlier plotting because we have geom_jitter to plot the points
  geom_smooth(method = "lm")+
  theme_light()+
  xlab("Age at tumour collection")+
  ylab("Hypoxia Score")+
  ggtitle("Age")+
  theme(plot.title = element_text(hjust = 0.5))
hypoxia_age
```
```{r}
cor.test(hypoxia_signatures_tum_data_single$age_at_tumour_collection,
         hypoxia_signatures_tum_data_single$Buffa_signature_Sum_within,
         method = "spearman")
```
```{r}
summary(hypoxia_signatures_tum_data_single$age_at_tumour_collection)
length(hypoxia_signatures_tum_data_single$age_at_tumour_collection)
```

```{r}
hypoxia_psa <- ggplot(hypoxia_signatures_tum_data_single, aes(x = psa_at_tumour_collection, y = Buffa_signature_Sum_within))+
  geom_point(alpha = 0.5)+ 
  geom_smooth(method = "lm")+
  theme_light()+
  xlab("PSA at tumour collection")+
  ylab("Hypoxia Score")+
  ggtitle("PSA")+
  scale_x_log10()+
  theme(plot.title = element_text(hjust = 0.5))
hypoxia_psa
```
```{r}
cor.test(hypoxia_signatures_tum_data_single$psa_at_tumour_collection,
         hypoxia_signatures_tum_data_single$Buffa_signature_Sum_within,
         method = "spearman")
```
```{r}
summary(hypoxia_signatures_tum_data_single$psa_at_tumour_collection)
length(hypoxia_signatures_tum_data_single$psa_at_tumour_collection)
```

```{r}
hypoxia_isup_cor <- ggplot(hypoxia_signatures_tum_data_single, aes(x = ISUP_Grade, y = Buffa_signature_Sum_within))+
  geom_point(alpha = 0.5)+ 
  geom_smooth(method = "lm")+
  theme_light()+
  xlab("Gleason Grade Group")+
  ylab("Hypoxia Score")+
  # ggtitle("ISUP Grade")+
  theme(plot.title = element_text(hjust = 0.5), legend.position = "none")+
  scale_colour_manual(values = PPCG_colourlist$ISUP_colourlist)
```


## IDC
```{r}
IDC_box <- ggplot(subset(hypoxia_signatures_tum_data_single, !is.na(Cribriform_Intraductal_Y_N)), 
                  aes(x = Cribriform_Intraductal_Y_N, y = Buffa_signature_Sum_within))+
  geom_jitter(alpha = 0.5, height = 0, aes(colour = Cribriform_Intraductal_Y_N))+
  geom_boxplot(alpha = 0.5, outlier.colour = NA)+ 
  theme_light()+
  xlab("IDC")+
  ylab("Hypoxia Score (Buffa Sum)")+
  geom_signif(map_signif_level = TRUE, comparisons = list(c("No IDC","IDC")),
              test = "wilcox.test", tip_length = 0, y_position = c(35))+
  scale_colour_manual(name = "IDC Presence", values = c("grey","tomato3"))+
  theme(text = element_text(size = 16),
        legend.position = "bottom")
IDC_box
```

```{r}
IDC_bar <- ggplot(subset(hypoxia_signatures_tum_data_single, !is.na(Cribriform_Intraductal_Y_N)), aes(fill = Cribriform_Intraductal_Y_N, x = Buffa_signature_Sum_within_quartile_one_sample_per_donor))+
  geom_bar(stat = "count", position = "fill")+ 
  geom_text(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], label=percent(..count../tapply(..count.., ..x.. ,sum)[..x..]) ),
              stat="count", position=position_fill(0.5), vjust=0.5, colour = "white")+
  theme_light()+
  xlab("Hypoxia Score Quartile")+
  ylab("IDC")+
  scale_fill_manual(name = "IDC Presence", values = c("grey","tomato3"))+
  theme(text = element_text(size = 16),
        legend.position = "bottom")
IDC_bar
```

```{r}
wilcox.test(subset(hypoxia_signatures_tum_data_single, Cribriform_Intraductal_Y_N == "IDC")$Buffa_signature_Sum_within,
            subset(hypoxia_signatures_tum_data_single, Cribriform_Intraductal_Y_N == "No IDC")$Buffa_signature_Sum_within)
```

## Output table
```{r}
patient_stat_table <- as.data.frame(matrix(nrow = 4, ncol = 5, data = NA))
colnames(patient_stat_table) <- c("Comparison","Test","n","p_val","Further_info")
patient_stat_table$Comparison <- c("Hypoxia by GG","Hypoxia vs Age","Hypoxia vs PSA","Hypoxia by IDC status")
patient_stat_table$Test <- c("Kruskal-Wallis Rank Sum Test", "Spearman Correlation","Spearman Correlation","Wilcoxon Rank Sum Test")

patient_stat_table$n <- c(sum(!is.na(hypoxia_signatures_tum_data_single$ISUP_Grade)),
                          sum(!is.na(hypoxia_signatures_tum_data_single$age_at_tumour_collection)),
                          sum(!is.na(hypoxia_signatures_tum_data_single$psa_at_tumour_collection)),
                          sum(!is.na(hypoxia_signatures_tum_data_single$Cribriform_Intraductal_Y_N)))

patient_stat_table$p_val <- c(kruskal.test(hypoxia_signatures_tum_data_single$Buffa_signature_Sum_within,
                                           hypoxia_signatures_tum_data_single$ISUP_Grade)$p.value,
                              cor.test(hypoxia_signatures_tum_data_single$age_at_tumour_collection,
                                       hypoxia_signatures_tum_data_single$Buffa_signature_Sum_within,
                                       method = "spearman")$p.value,
                              cor.test(hypoxia_signatures_tum_data_single$psa_at_tumour_collection,
                                       hypoxia_signatures_tum_data_single$Buffa_signature_Sum_within,
                                       method = "spearman")$p.value,
                              wilcox.test(subset(hypoxia_signatures_tum_data_single, Cribriform_Intraductal_Y_N == "IDC")$Buffa_signature_Sum_within,
                                          subset(hypoxia_signatures_tum_data_single, Cribriform_Intraductal_Y_N == "No IDC")$Buffa_signature_Sum_within)$p.value)
patient_stat_table$Further_info <- c("",
                                     paste0("Spearman $\\rho$ = ",round(cor.test(hypoxia_signatures_tum_data_single$age_at_tumour_collection,
                                       hypoxia_signatures_tum_data_single$Buffa_signature_Sum_within,
                                       method = "spearman")$estimate[[1]], digits = 3)),
                                     paste0("Spearman $\\rho$ = ",round(cor.test(hypoxia_signatures_tum_data_single$psa_at_tumour_collection,
                                       hypoxia_signatures_tum_data_single$Buffa_signature_Sum_within,
                                       method = "spearman")$estimate[[1]], digits = 3)),
                                     paste0("Median Hypoxia Score: IDC = ",
                                            round(median(subset(hypoxia_signatures_tum_data_single, Cribriform_Intraductal_Y_N == "IDC")$Buffa_signature_Sum_within),digits = 3),
                                            "; No IDC = ",
                                            round(median(subset(hypoxia_signatures_tum_data_single, Cribriform_Intraductal_Y_N == "No IDC")$Buffa_signature_Sum_within), digits = 3)))
```

Print latex style table
```{r}
xtable(patient_stat_table,
       display = c("s","s","s","d","e","s"))
```


# Drivers

Preparing analysis and plotting of driver data as a heatmap/oncoprint. 

```{r}
driver_mat <- as.matrix(subset(driver_data, gene_source != "."))
driver_mat = driver_mat[, -c(1,2)]
driver_mat <- gsub("\\.","", driver_mat)
```

Reorder signatures for oncoprint.  Only plot samples for which we have data from WGS/drivers. 
```{r}
hypoxia_op <- subset(hypoxia_signatures_tum_data_single, !is.na(WGS_AssayID) &
                       WGS_AssayID %!in% c("ADD",""))

rownames(hypoxia_op) <- hypoxia_op$WGS_AssayID

# order data by hypoxia scores
# hypoxia_op <- hypoxia_op[rownames(hypoxia_op) %in% colnames(driver_mat),]
hypoxia_op <- hypoxia_op[order(hypoxia_op$Buffa_signature_Sum_within, decreasing = FALSE),]

# order driver samples by hypoxia scores
driver_mat <- driver_mat[,rownames(hypoxia_op)[rownames(hypoxia_op) %in% colnames(driver_mat)]]

```

```{r}
driver_mat[driver_mat == "nsSNV"] <- "Mutation"
driver_mat[driver_mat == "splicing"] <- "Mutation"
driver_mat[driver_mat == "stopgain"] <- "Mutation"
driver_mat[driver_mat == "stoploss"] <- "Mutation"
driver_mat[driver_mat == "startloss"] <- "Mutation"
driver_mat[driver_mat == "stopgain+nsSNV"] <- "Mutation"
driver_mat[driver_mat == "nsSNV+splicing"] <- "Mutation"
driver_mat[driver_mat == "nsSNV+stopgain"] <- "Mutation"
driver_mat[driver_mat == "splicing+nsSNV"] <- "Mutation"

driver_mat[driver_mat == ""] <- "No Mutation"
```

Test for differences in scores
```{r}
driver_hypoxia_scores <- as.data.frame(matrix(nrow = nrow(driver_mat), ncol = 5))
colnames(driver_hypoxia_scores) <- c("Gene","n_mutated","n_wt","wilcox_p_val","wilcox_p_adj")
driver_hypoxia_scores$Gene <- rownames(driver_mat)

for(i in 1:nrow(driver_hypoxia_scores)){
  gene <- driver_hypoxia_scores$Gene[i]
  test_gene <- as.data.frame(t(driver_mat)[,gene])
  colnames(test_gene) <-"Test_gene"
  test_gene$sample <- rownames(test_gene)
  
  test_data <- merge(test_gene, hypoxia_op[,c("Buffa_signature_Sum_within","WGS_AssayID")],
                     by.x = "sample", by.y = "WGS_AssayID")
  
  table_data <- as.data.frame(table(test_data$Test_gene))
  if(nrow(table_data) > 1){
    driver_hypoxia_scores$wilcox_p_val[i] <- wilcox.test(subset(test_data, Test_gene == "Mutation")[,"Buffa_signature_Sum_within"],
                                                         subset(test_data, Test_gene == "No Mutation")[,"Buffa_signature_Sum_within"])[[3]]
  
    driver_hypoxia_scores$n_mutated[i] <- nrow(subset(test_data, Test_gene == "Mutation"))
    driver_hypoxia_scores$n_wt[i] <- nrow(subset(test_data, Test_gene == "No Mutation"))
  }
  
}

driver_hypoxia_scores$wilcox_p_adj <- p.adjust(driver_hypoxia_scores$wilcox_p_val, method = "BH")
```

```{r}
driver_hypoxia_scores$Significance <- ifelse(driver_hypoxia_scores$wilcox_p_adj < 0.05, yes = "Significant", no = "N.S.")
```

```{r}
rownames(driver_hypoxia_scores) <- driver_hypoxia_scores$Gene
```

Order data by p-value
```{r}
driver_hypoxia_scores <- driver_hypoxia_scores[order(driver_hypoxia_scores$wilcox_p_val),]
rownames(driver_hypoxia_scores) <- driver_hypoxia_scores$Gene
```
```{r}
head(driver_hypoxia_scores)
```

## Q1 Q4 difference
```{r}
driver_hypoxia_enrichment_quart <- as.data.frame(matrix(nrow = nrow(driver_mat), ncol = 8))
colnames(driver_hypoxia_enrichment_quart) <- c("Gene","n_q4_mutated","n_q1_mutated","n_q4_wt","n_q1_wt","fisher_p_val","fisher_p_adj","odds_ratio")
driver_hypoxia_enrichment_quart$Gene <- rownames(driver_mat)

for(i in 1:nrow(driver_hypoxia_enrichment_quart)){
  gene <- driver_hypoxia_enrichment_quart$Gene[i]
  test_gene <- as.data.frame(t(driver_mat)[,gene])
  colnames(test_gene) <-"Test_gene"
  test_gene$sample <- rownames(test_gene)
  
  test_data <- merge(test_gene, hypoxia_op[,c("Buffa_signature_Sum_within_quartile_one_sample_per_donor","WGS_AssayID")],
                     by.x = "sample", by.y = "WGS_AssayID")
  test_data <- subset(test_data, Buffa_signature_Sum_within_quartile_one_sample_per_donor %in% c("Q1","Q4"))
  
  table_data <- as.data.frame(table(test_data$Test_gene, test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
  if(nrow(table_data) > 2){
    driver_hypoxia_enrichment_quart$fisher_p_val[i] <- fisher.test(table(test_data$Test_gene, test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))[[1]]
    driver_hypoxia_enrichment_quart$odds_ratio[i] <- fisher.test(table(test_data$Test_gene, test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))[[3]]
  
  
  
    driver_hypoxia_enrichment_quart$n_q4_mutated[i] <- subset(table_data, (Var1 == "Mutation") & (Var2 == "Q4"))[,"Freq"]
    driver_hypoxia_enrichment_quart$n_q1_mutated[i] <- subset(table_data, (Var1 == "Mutation") & (Var2 == "Q1"))[,"Freq"]
    driver_hypoxia_enrichment_quart$n_q4_wt[i] <- subset(table_data, (Var1 == "No Mutation") & (Var2 == "Q4"))[,"Freq"]
    driver_hypoxia_enrichment_quart$n_q1_wt[i] <- subset(table_data, (Var1 == "No Mutation") & (Var2 == "Q1"))[,"Freq"]
  }
  
}

driver_hypoxia_enrichment_quart$fisher_p_adj <- p.adjust(driver_hypoxia_enrichment_quart$fisher_p_val, method = "BH")
```

```{r}
subset(driver_hypoxia_enrichment_quart, fisher_p_adj < 0.05)
```

```{r}
driver_hypoxia_enrichment_quart$Significance <- ifelse(driver_hypoxia_enrichment_quart$fisher_p_adj < 0.05, yes = "Significant", no = "N.S.")
```

```{r}
rownames(driver_hypoxia_enrichment_quart) <- driver_hypoxia_enrichment_quart$Gene
```

Order data by p-value
```{r}
driver_hypoxia_enrichment_quart <- driver_hypoxia_enrichment_quart[order(driver_hypoxia_enrichment_quart$fisher_p_val),]
rownames(driver_hypoxia_enrichment_quart) <- driver_hypoxia_enrichment_quart$Gene
```
```{r}
head(driver_hypoxia_enrichment_quart)
```

Select 10 genes to plot
```{r}
driver_top10_genes <- driver_hypoxia_enrichment_quart$Gene[1:10]
```

## Check mutation in group vs not
Check if mutation in any associated gene has differences in hypoxia score
```{r}
driver_mat_bin <- driver_mat
driver_mat_bin[driver_mat_bin == "Mutation"] <- as.numeric(1)
driver_mat_bin[driver_mat_bin == "No Mutation"] <- as.numeric(0)
```

```{r}
class(driver_mat_bin) <- "numeric"
```

```{r}
summary_somatic_driver <- as.data.frame(colSums(driver_mat_bin))
colnames(summary_somatic_driver) <- c("Muts_in_driver")
summary_somatic_driver$WGS_AssayID <- rownames(summary_somatic_driver)
summary_somatic_driver$bin_muts_in_driver <- ifelse(summary_somatic_driver$Muts_in_driver == 0, yes = "No Mutation", no = "Mutation")
```

```{r}
summary_somatic_driver <- merge(summary_somatic_driver, 
                               hypoxia_op[,c("Buffa_signature_Sum_within","WGS_AssayID",
                                             "Buffa_signature_Sum_within_bin_one_sample_per_donor",
                                             "Buffa_signature_Sum_within_quartile_one_sample_per_donor")],
                               by = "WGS_AssayID")
```

```{r}
wilcox.test(subset(summary_somatic_driver, bin_muts_in_driver == "Mutation")[,"Buffa_signature_Sum_within"], 
            subset(summary_somatic_driver, bin_muts_in_driver == "No Mutation")[,"Buffa_signature_Sum_within"])
```

```{r}
summary_somatic_driver_q1q4 <- subset(summary_somatic_driver, Buffa_signature_Sum_within_quartile_one_sample_per_donor %in% c("Q1","Q4"))
```

```{r}
fisher.test(table(summary_somatic_driver_q1q4$bin_muts_in_driver, summary_somatic_driver_q1q4$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
```
```{r}
table(summary_somatic_driver_q1q4$bin_muts_in_driver, 
      summary_somatic_driver_q1q4$Buffa_signature_Sum_within_quartile_one_sample_per_donor)
```

```{r}
wilcox.test(subset(summary_somatic_driver, bin_muts_in_driver == "Mutation")[,"Buffa_signature_Sum_within"], 
            subset(summary_somatic_driver, bin_muts_in_driver == "No Mutation")[,"Buffa_signature_Sum_within"])
```

# DRG 36 Somatic
DRG genes with somatic mutations identified by both callers or a tier 1 indel (filtered data)
```{r}
drg_genes_36[drg_genes_36 %in% rownames(driver_data)]
```

DRG genes that do not have somatic mutations identified by both callers or a tier 1 indel (filtered data)
```{r}
drg_genes_36[drg_genes_36 %!in% rownames(driver_data)]
```

```{r}
drg36_mat <- as.matrix(driver_data[rownames(driver_data) %in% drg_genes_36,])
drg36_mat = drg36_mat[, -c(1,2)]
drg36_mat <- gsub("\\.","", drg36_mat)
```
Reorder signatures for oncoprint.  Only plot samples for which we have data from WGS/drivers. 
```{r}
# order driver samples by hypoxia scores
drg36_mat <- drg36_mat[,rownames(hypoxia_op)[rownames(hypoxia_op) %in% colnames(drg36_mat)]]

```

```{r}
drg36_mat[drg36_mat == "nsSNV"] <- "Mutation"
drg36_mat[drg36_mat == "splicing"] <- "Mutation"
drg36_mat[drg36_mat == "stopgain"] <- "Mutation"
drg36_mat[drg36_mat == "stoploss"] <- "Mutation"
drg36_mat[drg36_mat == "startloss"] <- "Mutation"
drg36_mat[drg36_mat == "stopgain+nsSNV"] <- "Mutation"
drg36_mat[drg36_mat == "nsSNV+splicing"] <- "Mutation"
drg36_mat[drg36_mat == "nsSNV+stopgain"] <- "Mutation"
drg36_mat[drg36_mat == "splicing+nsSNV"] <- "Mutation"

drg36_mat[drg36_mat == ""] <- "No Mutation"
```

## Test for differences in scores
```{r}
drg36_hypoxia_scores <- as.data.frame(matrix(nrow = nrow(drg36_mat), ncol = 5))
colnames(drg36_hypoxia_scores) <- c("Gene","n_mutated","n_wt","wilcox_p_val","wilcox_p_adj")
drg36_hypoxia_scores$Gene <- rownames(drg36_mat)

for(i in 1:nrow(drg36_hypoxia_scores)){
  gene <- drg36_hypoxia_scores$Gene[i]
  test_gene <- as.data.frame(t(drg36_mat)[,gene])
  colnames(test_gene) <-"Test_gene"
  test_gene$sample <- rownames(test_gene)
  
  test_data <- merge(test_gene, hypoxia_op[,c("Buffa_signature_Sum_within","WGS_AssayID")],
                     by.x = "sample", by.y = "WGS_AssayID")
  
  table_data <- as.data.frame(table(test_data$Test_gene))
  if(nrow(table_data) > 1){
    drg36_hypoxia_scores$wilcox_p_val[i] <- wilcox.test(subset(test_data, Test_gene == "Mutation")[,"Buffa_signature_Sum_within"],
                                                         subset(test_data, Test_gene == "No Mutation")[,"Buffa_signature_Sum_within"])[[3]]
  
    drg36_hypoxia_scores$n_mutated[i] <- nrow(subset(test_data, Test_gene == "Mutation"))
    drg36_hypoxia_scores$n_wt[i] <- nrow(subset(test_data, Test_gene == "No Mutation"))
  }
  
}

drg36_hypoxia_scores$wilcox_p_adj <- p.adjust(drg36_hypoxia_scores$wilcox_p_val, method = "BH")
```

```{r}
drg36_hypoxia_scores$Significance <- ifelse(drg36_hypoxia_scores$wilcox_p_adj < 0.05, yes = "Significant", no = "N.S.")
```

```{r}
rownames(drg36_hypoxia_scores) <- drg36_hypoxia_scores$Gene
```

Order data by p-value
```{r}
drg36_hypoxia_scores <- drg36_hypoxia_scores[order(drg36_hypoxia_scores$wilcox_p_val),]
rownames(drg36_hypoxia_scores) <- drg36_hypoxia_scores$Gene
```
```{r}
head(drg36_hypoxia_scores)
```

## Q1 Q4 difference
```{r}
drg36_hypoxia_enrichment_quart <- as.data.frame(matrix(nrow = nrow(drg36_mat), ncol = 8))
colnames(drg36_hypoxia_enrichment_quart) <- c("Gene","n_q4_mutated","n_q1_mutated","n_q4_wt","n_q1_wt","fisher_p_val","fisher_p_adj","odds_ratio")
drg36_hypoxia_enrichment_quart$Gene <- rownames(drg36_mat)

for(i in 1:nrow(drg36_hypoxia_enrichment_quart)){
  gene <- drg36_hypoxia_enrichment_quart$Gene[i]
  test_gene <- as.data.frame(t(drg36_mat)[,gene])
  colnames(test_gene) <-"Test_gene"
  test_gene$sample <- rownames(test_gene)
  
  test_data <- merge(test_gene, hypoxia_op[,c("Buffa_signature_Sum_within_quartile_one_sample_per_donor","WGS_AssayID")],
                     by.x = "sample", by.y = "WGS_AssayID")
  test_data <- subset(test_data, Buffa_signature_Sum_within_quartile_one_sample_per_donor %in% c("Q1","Q4"))
  
  table_data <- as.data.frame(table(test_data$Test_gene, test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
  if(nrow(table_data) > 2){
    drg36_hypoxia_enrichment_quart$fisher_p_val[i] <- fisher.test(table(test_data$Test_gene, test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))[[1]]
    drg36_hypoxia_enrichment_quart$odds_ratio[i] <- fisher.test(table(test_data$Test_gene, test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))[[3]]
  
  
  
    drg36_hypoxia_enrichment_quart$n_q4_mutated[i] <- subset(table_data, (Var1 == "Mutation") & (Var2 == "Q4"))[,"Freq"]
    drg36_hypoxia_enrichment_quart$n_q1_mutated[i] <- subset(table_data, (Var1 == "Mutation") & (Var2 == "Q1"))[,"Freq"]
    drg36_hypoxia_enrichment_quart$n_q4_wt[i] <- subset(table_data, (Var1 == "No Mutation") & (Var2 == "Q4"))[,"Freq"]
    drg36_hypoxia_enrichment_quart$n_q1_wt[i] <- subset(table_data, (Var1 == "No Mutation") & (Var2 == "Q1"))[,"Freq"]
  }
  
}

drg36_hypoxia_enrichment_quart$fisher_p_adj <- p.adjust(drg36_hypoxia_enrichment_quart$fisher_p_val, method = "BH")
```

```{r}
subset(drg36_hypoxia_enrichment_quart, fisher_p_adj < 0.05)
```

```{r}
drg36_hypoxia_enrichment_quart$Significance <- ifelse(drg36_hypoxia_enrichment_quart$fisher_p_adj < 0.05, yes = "Significant", no = "N.S.")
```

```{r}
rownames(drg36_hypoxia_enrichment_quart) <- drg36_hypoxia_enrichment_quart$Gene
```

## Check mutation in group vs not
Check if mutation in any associated gene has differences in hypoxia score
```{r}
drg36_mat_bin <- drg36_mat
drg36_mat_bin[drg36_mat_bin == "Mutation"] <- as.numeric(1)
drg36_mat_bin[drg36_mat_bin == "No Mutation"] <- as.numeric(0)
```

```{r}
class(drg36_mat_bin) <- "numeric"
```

```{r}
summary_somatic_drg36 <- as.data.frame(colSums(drg36_mat_bin))
colnames(summary_somatic_drg36) <- c("Muts_in_drg36")
summary_somatic_drg36$WGS_AssayID <- rownames(summary_somatic_drg36)
summary_somatic_drg36$bin_muts_in_drg36 <- ifelse(summary_somatic_drg36$Muts_in_drg36 == 0, yes = "No Mutation", no = "Mutation")
```

```{r}
summary_somatic_drg36 <- merge(summary_somatic_drg36, 
                               hypoxia_op[,c("Buffa_signature_Sum_within","WGS_AssayID",
                                             "Buffa_signature_Sum_within_bin_one_sample_per_donor",
                                             "Buffa_signature_Sum_within_quartile_one_sample_per_donor")],
                               by = "WGS_AssayID")
```

```{r}
wilcox.test(subset(summary_somatic_drg36, bin_muts_in_drg36 == "Mutation")[,"Buffa_signature_Sum_within"], 
            subset(summary_somatic_drg36, bin_muts_in_drg36 == "No Mutation")[,"Buffa_signature_Sum_within"])
```

```{r}
summary_somatic_drg36_q1q4 <- subset(summary_somatic_drg36, Buffa_signature_Sum_within_quartile_one_sample_per_donor %in% c("Q1","Q4"))
```

```{r}
fisher.test(table(summary_somatic_drg36_q1q4$bin_muts_in_drg36, summary_somatic_drg36_q1q4$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
```
```{r}
table(summary_somatic_drg36_q1q4$bin_muts_in_drg36, 
      summary_somatic_drg36_q1q4$Buffa_signature_Sum_within_quartile_one_sample_per_donor)
```



# DRG 128 Somatic
DRG genes with somatic mutations identified by both callers or a tier 1 indel (filtered data)
```{r}
drg_genes_128[drg_genes_128 %in% rownames(driver_data)]
```
```{r}
length(drg_genes_128[drg_genes_128 %in% rownames(driver_data)])
```

DRG genes that do not have somatic mutations identified by both callers or a tier 1 indel (filtered data)
```{r}
drg_genes_128[drg_genes_128 %!in% rownames(driver_data)]
```
```{r}
length(drg_genes_128[drg_genes_128 %!in% rownames(driver_data)])
```

```{r}
drg128_mat <- as.matrix(driver_data[rownames(driver_data) %in% drg_genes_128,])
drg128_mat = drg128_mat[, -c(1,2)]
drg128_mat <- gsub("\\.","", drg128_mat)
```
Reorder signatures for oncoprint.  Only plot samples for which we have data from WGS/drivers. 
```{r}
# order driver samples by hypoxia scores
drg128_mat <- drg128_mat[,rownames(hypoxia_op)[rownames(hypoxia_op) %in% colnames(drg128_mat)]]

```

```{r}
drg128_mat[drg128_mat == "nsSNV"] <- "Mutation"
drg128_mat[drg128_mat == "splicing"] <- "Mutation"
drg128_mat[drg128_mat == "stopgain"] <- "Mutation"
drg128_mat[drg128_mat == "stoploss"] <- "Mutation"
drg128_mat[drg128_mat == "startloss"] <- "Mutation"
drg128_mat[drg128_mat == "stopgain+nsSNV"] <- "Mutation"
drg128_mat[drg128_mat == "nsSNV+splicing"] <- "Mutation"
drg128_mat[drg128_mat == "nsSNV+stopgain"] <- "Mutation"
drg128_mat[drg128_mat == "splicing+nsSNV"] <- "Mutation"

drg128_mat[drg128_mat == ""] <- "No Mutation"
```

## Test for differences in scores
```{r}
drg128_hypoxia_scores <- as.data.frame(matrix(nrow = nrow(drg128_mat), ncol = 5))
colnames(drg128_hypoxia_scores) <- c("Gene","n_mutated","n_wt","wilcox_p_val","wilcox_p_adj")
drg128_hypoxia_scores$Gene <- rownames(drg128_mat)

for(i in 1:nrow(drg128_hypoxia_scores)){
  gene <- drg128_hypoxia_scores$Gene[i]
  test_gene <- as.data.frame(t(drg128_mat)[,gene])
  colnames(test_gene) <-"Test_gene"
  test_gene$sample <- rownames(test_gene)
  
  test_data <- merge(test_gene, hypoxia_op[,c("Buffa_signature_Sum_within","WGS_AssayID")],
                     by.x = "sample", by.y = "WGS_AssayID")
  
  table_data <- as.data.frame(table(test_data$Test_gene))
  if(nrow(table_data) > 1){
    drg128_hypoxia_scores$wilcox_p_val[i] <- wilcox.test(subset(test_data, Test_gene == "Mutation")[,"Buffa_signature_Sum_within"],
                                                         subset(test_data, Test_gene == "No Mutation")[,"Buffa_signature_Sum_within"])[[3]]
  
    drg128_hypoxia_scores$n_mutated[i] <- nrow(subset(test_data, Test_gene == "Mutation"))
    drg128_hypoxia_scores$n_wt[i] <- nrow(subset(test_data, Test_gene == "No Mutation"))
  }
  
}

drg128_hypoxia_scores$wilcox_p_adj <- p.adjust(drg128_hypoxia_scores$wilcox_p_val, method = "BH")
```

```{r}
drg128_hypoxia_scores$Significance <- ifelse(drg128_hypoxia_scores$wilcox_p_adj < 0.05, yes = "Significant", no = "N.S.")
```

```{r}
rownames(drg128_hypoxia_scores) <- drg128_hypoxia_scores$Gene
```

Order data by p-value
```{r}
drg128_hypoxia_scores <- drg128_hypoxia_scores[order(drg128_hypoxia_scores$wilcox_p_val),]
rownames(drg128_hypoxia_scores) <- drg128_hypoxia_scores$Gene
```
```{r}
head(drg128_hypoxia_scores)
```

## Q1 Q4 difference
```{r}
drg128_hypoxia_enrichment_quart <- as.data.frame(matrix(nrow = nrow(drg128_mat), ncol = 8))
colnames(drg128_hypoxia_enrichment_quart) <- c("Gene","n_q4_mutated","n_q1_mutated","n_q4_wt","n_q1_wt","fisher_p_val","fisher_p_adj","odds_ratio")
drg128_hypoxia_enrichment_quart$Gene <- rownames(drg128_mat)

for(i in 1:nrow(drg128_hypoxia_enrichment_quart)){
  gene <- drg128_hypoxia_enrichment_quart$Gene[i]
  test_gene <- as.data.frame(t(drg128_mat)[,gene])
  colnames(test_gene) <-"Test_gene"
  test_gene$sample <- rownames(test_gene)
  
  test_data <- merge(test_gene, hypoxia_op[,c("Buffa_signature_Sum_within_quartile_one_sample_per_donor","WGS_AssayID")],
                     by.x = "sample", by.y = "WGS_AssayID")
  test_data <- subset(test_data, Buffa_signature_Sum_within_quartile_one_sample_per_donor %in% c("Q1","Q4"))
  
  table_data <- as.data.frame(table(test_data$Test_gene, test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
  if(nrow(table_data) > 2){
    drg128_hypoxia_enrichment_quart$fisher_p_val[i] <- fisher.test(table(test_data$Test_gene, test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))[[1]]
    drg128_hypoxia_enrichment_quart$odds_ratio[i] <- fisher.test(table(test_data$Test_gene, test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))[[3]]
  
  
  
    drg128_hypoxia_enrichment_quart$n_q4_mutated[i] <- subset(table_data, (Var1 == "Mutation") & (Var2 == "Q4"))[,"Freq"]
    drg128_hypoxia_enrichment_quart$n_q1_mutated[i] <- subset(table_data, (Var1 == "Mutation") & (Var2 == "Q1"))[,"Freq"]
    drg128_hypoxia_enrichment_quart$n_q4_wt[i] <- subset(table_data, (Var1 == "No Mutation") & (Var2 == "Q4"))[,"Freq"]
    drg128_hypoxia_enrichment_quart$n_q1_wt[i] <- subset(table_data, (Var1 == "No Mutation") & (Var2 == "Q1"))[,"Freq"]
  }
  
}

drg128_hypoxia_enrichment_quart$fisher_p_adj <- p.adjust(drg128_hypoxia_enrichment_quart$fisher_p_val, method = "BH")
```

```{r}
subset(drg128_hypoxia_enrichment_quart, fisher_p_adj < 0.05)
```

```{r}
drg128_hypoxia_enrichment_quart$Significance <- ifelse(drg128_hypoxia_enrichment_quart$fisher_p_adj < 0.05, yes = "Significant", no = "N.S.")
```

```{r}
rownames(drg128_hypoxia_enrichment_quart) <- drg128_hypoxia_enrichment_quart$Gene
```

```{r}
head(drg128_hypoxia_enrichment_quart)
```

## Check mutation in group vs not
Check if mutation in any associated gene has differences in hypoxia score
```{r}
drg128_mat_bin <- drg128_mat
drg128_mat_bin[drg128_mat_bin == "Mutation"] <- as.numeric(1)
drg128_mat_bin[drg128_mat_bin == "No Mutation"] <- as.numeric(0)
```

```{r}
class(drg128_mat_bin) <- "numeric"
```

```{r}
summary_somatic_drg128 <- as.data.frame(colSums(drg128_mat_bin))
colnames(summary_somatic_drg128) <- c("Muts_in_drg128")
summary_somatic_drg128$WGS_AssayID <- rownames(summary_somatic_drg128)
summary_somatic_drg128$bin_muts_in_drg128 <- ifelse(summary_somatic_drg128$Muts_in_drg128 == 0, yes = "No Mutation", no = "Mutation")
```

```{r}
summary_somatic_drg128 <- merge(summary_somatic_drg128, 
                               hypoxia_op[,c("Buffa_signature_Sum_within","WGS_AssayID",
                                             "Buffa_signature_Sum_within_bin_one_sample_per_donor",
                                             "Buffa_signature_Sum_within_quartile_one_sample_per_donor")],
                               by = "WGS_AssayID")
```

```{r}
wilcox.test(subset(summary_somatic_drg128, bin_muts_in_drg128 == "Mutation")[,"Buffa_signature_Sum_within"], 
            subset(summary_somatic_drg128, bin_muts_in_drg128 == "No Mutation")[,"Buffa_signature_Sum_within"])
```

```{r}
summary_somatic_drg128_q1q4 <- subset(summary_somatic_drg128, Buffa_signature_Sum_within_quartile_one_sample_per_donor %in% c("Q1","Q4"))
```

```{r}
fisher.test(table(summary_somatic_drg128_q1q4$bin_muts_in_drg128, summary_somatic_drg128_q1q4$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
```
```{r}
table(summary_somatic_drg128_q1q4$bin_muts_in_drg128, 
      summary_somatic_drg128_q1q4$Buffa_signature_Sum_within_quartile_one_sample_per_donor)
```


# CNAs
```{r}
rownames(rel_cna) <- paste0(rel_cna$gene_name,"_",rel_cna$gene_id)
```

Subset for genes in driver gene list and samples with data
```{r}
cna_mat <- as.matrix((rel_cna[(gsub("_.*","",rownames(rel_cna)) %in% driver_hypoxia_scores$Gene),colnames(rel_cna) %in% hypoxia_signatures_tum_data_single$WGS_AssayID]))
```


```{r}
dim(cna_mat)
```

Remove enormous cna data frame
```{r}
rm(rel_cna)
```

```{r}
cna_mat <- gsub("\\.","", cna_mat)
cna_mat[as.numeric(cna_mat) <0] <- "Loss"
cna_mat[as.numeric(cna_mat) >0] <- "Gain"
cna_mat[as.numeric(cna_mat) == 0] <- "Neutral"
```


## Q1 vs Q4 enrichment
```{r}
cna_hypoxia_enrichment_quart <- as.data.frame(matrix(nrow = nrow(cna_mat), ncol = 9))
colnames(cna_hypoxia_enrichment_quart) <- c("Gene","n_q4_gain","n_q1_gain","n_q4_neutral","n_q1_neutral","n_q4_loss","n_q1_loss","fisher_p_val","fisher_p_adj")
cna_hypoxia_enrichment_quart$Gene <- rownames(cna_mat)

for(i in 1:nrow(cna_hypoxia_enrichment_quart)){ #nrow(cna_hypoxia_enrichment)
  gene <- cna_hypoxia_enrichment_quart$Gene[i]
  test_gene <- as.data.frame(t(cna_mat)[,gene])
  colnames(test_gene) <-"Test_gene"
  test_gene$sample <- rownames(test_gene)
  
  test_data <- merge(test_gene, hypoxia_op[,c("Buffa_signature_Sum_within_quartile_one_sample_per_donor","WGS_AssayID")],
                     by.x = "sample", by.y = "WGS_AssayID")
  test_data <- subset(test_data, Buffa_signature_Sum_within_quartile_one_sample_per_donor %in% c("Q1","Q4"))
  
  table_data <- as.data.frame(table(test_data$Test_gene, test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
  if(nrow(table_data) > 2){
    cna_hypoxia_enrichment_quart$fisher_p_val[i] <- fisher.test(table(test_data$Test_gene, test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))[[1]]
    if("Gain" %in% table_data$Var1){
      cna_hypoxia_enrichment_quart$n_q4_gain[i] <- subset(table_data, (Var1 == "Gain") & (Var2 == "Q4"))[,"Freq"]
      cna_hypoxia_enrichment_quart$n_q1_gain[i] <- subset(table_data, (Var1 == "Gain") & (Var2 == "Q1"))[,"Freq"]
    }
    if("Neutral" %in% table_data$Var1){
      cna_hypoxia_enrichment_quart$n_q4_neutral[i] <- subset(table_data, (Var1 == "Neutral") & (Var2 == "Q4"))[,"Freq"]
      cna_hypoxia_enrichment_quart$n_q1_neutral[i] <- subset(table_data, (Var1 == "Neutral") & (Var2 == "Q1"))[,"Freq"]
    }
    if("Loss" %in% table_data$Var1){
      cna_hypoxia_enrichment_quart$n_q4_loss[i] <- subset(table_data, (Var1 == "Loss") & (Var2 == "Q4"))[,"Freq"]
      cna_hypoxia_enrichment_quart$n_q1_loss[i] <- subset(table_data, (Var1 == "Loss") & (Var2 == "Q1"))[,"Freq"]
    }
  }
}

cna_hypoxia_enrichment_quart$fisher_p_adj <- p.adjust(cna_hypoxia_enrichment_quart$fisher_p_val, method = "BH")
```

```{r}
rownames(cna_hypoxia_enrichment_quart) <- cna_hypoxia_enrichment_quart$Gene
```

```{r}
cna_hypoxia_enrichment_quart$Significance <- ifelse(cna_hypoxia_enrichment_quart$fisher_p_adj < 0.05, yes = "Significant", no = "N.S.")
```

Order data by p-value
```{r}
cna_hypoxia_enrichment_quart <- cna_hypoxia_enrichment_quart[order(cna_hypoxia_enrichment_quart$fisher_p_val),]
rownames(cna_hypoxia_enrichment_quart) <- cna_hypoxia_enrichment_quart$Gene
```
```{r}
cna_hypoxia_enrichment_quart
```

Order cna driver data by hypoxia score
```{r}
# order cna samples by hypoxia scores
cna_mat <- cna_mat[,rownames(hypoxia_op)[rownames(hypoxia_op) %in% colnames(cna_mat)]]

```

```{r}
colnames(driver_mat) == colnames(cna_mat)
```

For plotting purposes, only keep genes with some sort of significance
```{r}
sig_cna_genes <- subset(cna_hypoxia_enrichment_quart, Significance == "Significant")$Gene
cna_hypoxia_enrichment_quart_sub <- subset(cna_hypoxia_enrichment_quart, Gene %in% sig_cna_genes)
```

```{r}
cna_mat_plot <- cna_mat[sig_cna_genes,]
```

Change rownames so that it's just the gene name and not the ENSG name
```{r}
rownames(cna_hypoxia_enrichment_quart_sub) <- gsub("_.*","", rownames(cna_hypoxia_enrichment_quart_sub))
rownames(cna_mat_plot) <- gsub("_.*","", rownames(cna_mat_plot))
```


```{r}
dim(cna_mat_plot)
```

```{r}
colnames(driver_mat) == colnames(cna_mat_plot)
```

# Combined heatmap
## Preparing Signature Data
```{r}
sbs_sig_columns <- c("SBS1", "SBS2", "SBS3",	"SBS5",	"SBS8",	
                     "SBS13",	"SBS15",	"SBS17a",	"SBS17b", "SBS18",
                     "SBS21", "SBS22b",	"SBS28", "SBS34",	"SBS37",
                     "SBS38", "SBS40a",	"SBS40b",	"SBS44", "SBS52",
                     "SBS54",	"SBS96D")

id_sig_columns <- c("ID1","ID2","ID3","ID4","ID5",
                     "ID6","ID8","ID9","ID12","ID83C")

cx_sig_columns <- c("CX1_numbers","CX2_numbers","CX3_numbers","CX4_numbers","CX5_numbers",
                     "CX7_numbers","CX8_numbers","CX11_numbers")

mut_sig_columns <- c(sbs_sig_columns, id_sig_columns, cx_sig_columns)
mut_sig_split <- as.data.frame(mut_sig_columns)
colnames(mut_sig_split) <- c("sig_name")
mut_sig_split$Signature_type <- c(rep("SBS",22),
                                  rep("ID",10),
                                  rep("CX",8))
rownames(mut_sig_split) <- mut_sig_split$sig_name
```

```{r}
mut_sig_mat <- t(as.matrix(hypoxia_op[colnames(cna_mat),mut_sig_columns]))

sbs_sig_mat <- t(as.matrix(hypoxia_op[colnames(cna_mat),sbs_sig_columns]))
id_sig_mat <- t(as.matrix(hypoxia_op[colnames(cna_mat),id_sig_columns]))
cx_sig_mat <- t(as.matrix(hypoxia_op[colnames(cna_mat),cx_sig_columns]))
```

## Preparing signature annotation information
### Coefficient colours
```{r}
all_coefficients_total <- c(rlmer_CX_outputs$Coefficient,
                      rlmer_GI_outputs$Coefficient,
                      rlmer_ID_outputs$Coefficient,
                      rlmer_SBS_outputs$Coefficient)
all_coefficients_nonzero <- c(rlmer_CX_outputs_nonzero$Coefficient,
                      rlmer_ID_outputs_nonzero$Coefficient,
                      rlmer_SBS_outputs_nonzero$Coefficient)
```
```{r}
coefficient_colours_total <- colorRamp2(c(min(all_coefficients_total, na.rm = TRUE),
                                    -0.0000000001,
                                    0,
                                    0.0000000001,
                                    0.5,
                                    1,
                                    max(all_coefficients_total, na.rm = TRUE)),
                                  c("purple1","lavender","white","lightyellow","gold","orange1","red"))
coefficient_colours_nonzero <- colorRamp2(c(min(all_coefficients_nonzero, na.rm = TRUE),
                                    0,
                                    max(all_coefficients_nonzero, na.rm = TRUE)),
                                  c("royalblue","white","firebrick"))
```

### SBS annotations
```{r}
rlmer_SBS_outputs$Significance <- ifelse(rlmer_SBS_outputs$measure_p_adj < 0.05, yes = "Significant", no = "Not Significant")
rlmer_SBS_outputs_nonzero$Significance <- ifelse(rlmer_SBS_outputs_nonzero$measure_p_adj < 0.05, yes = "Significant", no = "Not Significant")
```


Make it clear which data is which
```{r}
colnames(rlmer_SBS_outputs) <- paste0(colnames(rlmer_SBS_outputs),"_total")
colnames(rlmer_SBS_outputs_nonzero) <- paste0(colnames(rlmer_SBS_outputs_nonzero),"_nonzero")
```

Add a measure name column so we can match to raw data
```{r}
rlmer_SBS_outputs$Measure_name <- gsub("rlmer_","",rlmer_SBS_outputs$Model_total)
rlmer_SBS_outputs_nonzero$Measure_name <- gsub("rlmer_","",rlmer_SBS_outputs_nonzero$Model_nonzero)
rlmer_SBS_outputs_nonzero$Measure_name <- gsub("_nonzero","",rlmer_SBS_outputs_nonzero$Measure_name)
```

Make a list of all the signatures present (not all have models but all are represented in the heatmap)
```{r}
rlmer_SBS_annotations <- as.data.frame(sbs_sig_columns)
colnames(rlmer_SBS_annotations) <- "SBS_list"
```

Merge SBS annotations
```{r}
rlmer_SBS_annotations <- merge(rlmer_SBS_annotations,
                              rlmer_SBS_outputs,
                              by.x = "SBS_list",
                              by.y = "Measure_name",
                              all = TRUE)
rlmer_SBS_annotations <- merge(rlmer_SBS_annotations,
                              rlmer_SBS_outputs_nonzero,
                              by.x = "SBS_list",
                              by.y = "Measure_name",
                              all = TRUE)
rownames(rlmer_SBS_annotations) <- rlmer_SBS_annotations$SBS_list
```
```{r}
xtable(rlmer_SBS_annotations[,c("SBS_list","Coefficient_total","CI_low_total","CI_high_total","p_total","measure_p_adj_total", 
                         "Coefficient_nonzero","CI_low_nonzero","CI_high_nonzero","p_nonzero","measure_p_adj_nonzero")], display = c("s","s","e","e","e","e","e","e","e","e","e","e"))
```

Rearrange
```{r}
rlmer_SBS_annotations <- rlmer_SBS_annotations[sbs_sig_columns,]
```

```{r}
sbs_row_annotation_right <- rowAnnotation("robustlmm" = rlmer_SBS_annotations$Significance_total,
                                         "Coefficient" = rlmer_SBS_annotations$Coefficient_total,
                                         "Non-zero robustlmm" = rlmer_SBS_annotations$Significance_nonzero,
                                         "Non-zero Coefficient" = rlmer_SBS_annotations$Coefficient_nonzero,
                                         col = list("robustlmm" = c("Significant" = "indianred", "Not Significant" = "grey"),
                                                    "Coefficient" = coefficient_colours_total,
                                                    "Non-zero robustlmm" = c("Significant" = "coral", "Not Significant" = "grey"),
                                                    "Non-zero Coefficient" = coefficient_colours_nonzero),
                           show_annotation_name = FALSE)
```

### ID annotations
```{r}
rlmer_ID_outputs$Significance <- ifelse(rlmer_ID_outputs$measure_p_adj < 0.05, yes = "Significant", no = "Not Significant")
rlmer_ID_outputs_nonzero$Significance <- ifelse(rlmer_ID_outputs_nonzero$measure_p_adj < 0.05, yes = "Significant", no = "Not Significant")
```

Make it clear which data is which
```{r}
colnames(rlmer_ID_outputs) <- paste0(colnames(rlmer_ID_outputs),"_total")
colnames(rlmer_ID_outputs_nonzero) <- paste0(colnames(rlmer_ID_outputs_nonzero),"_nonzero")
```

Add a measure name column so we can match to raw data
```{r}
rlmer_ID_outputs$Measure_name <- gsub("rlmer_","",rlmer_ID_outputs$Model_total)
rlmer_ID_outputs_nonzero$Measure_name <- gsub("rlmer_","",rlmer_ID_outputs_nonzero$Model_nonzero)
rlmer_ID_outputs_nonzero$Measure_name <- gsub("_nonzero","",rlmer_ID_outputs_nonzero$Measure_name)
```

Make a list of all the signatures present (not all have models but all are represented in the heatmap)
```{r}
rlmer_ID_annotations <- as.data.frame(id_sig_columns)
colnames(rlmer_ID_annotations) <- "ID_list"
```

Merge ID annotations
```{r}
rlmer_ID_annotations <- merge(rlmer_ID_annotations,
                              rlmer_ID_outputs,
                              by.x = "ID_list",
                              by.y = "Measure_name",
                              all = TRUE)
rlmer_ID_annotations <- merge(rlmer_ID_annotations,
                              rlmer_ID_outputs_nonzero,
                              by.x = "ID_list",
                              by.y = "Measure_name",
                              all = TRUE)
rownames(rlmer_ID_annotations) <- rlmer_ID_annotations$ID_list
```
```{r}
xtable(rlmer_ID_annotations[,c("ID_list","Coefficient_total","CI_low_total","CI_high_total","p_total","measure_p_adj_total", 
                         "Coefficient_nonzero","CI_low_nonzero","CI_high_nonzero","p_nonzero","measure_p_adj_nonzero")], display = c("s","s","e","e","e","e","e","e","e","e","e","e"))
```

Rearrange
```{r}
rlmer_ID_annotations <- rlmer_ID_annotations[id_sig_columns,]
```

```{r}
id_row_annotation_right <- rowAnnotation("robustlmm" = rlmer_ID_annotations$Significance_total,
                                         "Coefficient" = rlmer_ID_annotations$Coefficient_total,
                                         "Non-zero robustlmm" = rlmer_ID_annotations$Significance_nonzero,
                                         "Non-zero Coefficient" = rlmer_ID_annotations$Coefficient_nonzero,
                                         col = list("robustlmm" = c("Significant" = "indianred", "Not Significant" = "grey"),
                                                    "Coefficient" = coefficient_colours_total,
                                                    "Non-zero robustlmm" = c("Significant" = "coral", "Not Significant" = "grey"),
                                                    "Non-zero Coefficient" = coefficient_colours_nonzero),
                           show_annotation_name = FALSE)
```

### CX annotations
```{r}
rlmer_CX_outputs$Significance <- ifelse(rlmer_CX_outputs$measure_p_adj < 0.05, yes = "Significant", no = "Not Significant")
rlmer_CX_outputs_nonzero$Significance <- ifelse(rlmer_CX_outputs_nonzero$measure_p_adj < 0.05, yes = "Significant", no = "Not Significant")
```

Make it clear which data is which
```{r}
colnames(rlmer_CX_outputs) <- paste0(colnames(rlmer_CX_outputs),"_total")
colnames(rlmer_CX_outputs_nonzero) <- paste0(colnames(rlmer_CX_outputs_nonzero),"_nonzero")
```

Add a measure name column so we can match to raw data
```{r}
rlmer_CX_outputs$Measure_name <- gsub("rlmer_","",rlmer_CX_outputs$Parameter_total)
rlmer_CX_outputs_nonzero$Measure_name <- gsub("rlmer_","",rlmer_CX_outputs_nonzero$Parameter_nonzero)
rlmer_CX_outputs_nonzero$Measure_name <- gsub("_nonzero","",rlmer_CX_outputs_nonzero$Measure_name)
```

Make a list of all the signatures present (not all have models but all are represented in the heatmap)
```{r}
rlmer_CX_annotations <- as.data.frame(cx_sig_columns)
colnames(rlmer_CX_annotations) <- "CX_list"
```

Merge CX annotations
```{r}
rlmer_CX_annotations <- merge(rlmer_CX_annotations,
                              rlmer_CX_outputs,
                              by.x = "CX_list",
                              by.y = "Variable_total",
                              all = TRUE)
rlmer_CX_annotations <- merge(rlmer_CX_annotations,
                              rlmer_CX_outputs_nonzero,
                              by.x = "CX_list",
                              by.y = "Variable_nonzero",
                              all = TRUE)
rownames(rlmer_CX_annotations) <- rlmer_CX_annotations$CX_list
```
```{r}
xtable(rlmer_CX_annotations[,c("CX_list","Coefficient_total","CI_low_total","CI_high_total","p_total","measure_p_adj_total", 
                         "Coefficient_nonzero","CI_low_nonzero","CI_high_nonzero","p_nonzero","measure_p_adj_nonzero")],
       , display = c("s","s","e","e","e","e","e","e","e","e","e","e"))
```

Rearrange
```{r}
rlmer_CX_annotations <- rlmer_CX_annotations[cx_sig_columns,]
```

```{r}
cx_row_annotation_right <- rowAnnotation("robustlmm" = rlmer_CX_annotations$Significance_total,
                                         "Coefficient" = rlmer_CX_annotations$Coefficient_total,
                                         "Non-zero robustlmm" = rlmer_CX_annotations$Significance_nonzero,
                                         "Non-zero Coefficient" = rlmer_CX_annotations$Coefficient_nonzero,
                                         col = list("robustlmm" = c("Significant" = "indianred", "Not Significant" = "grey"),
                                                    "Coefficient" = coefficient_colours_total,
                                                    "Non-zero robustlmm" = c("Significant" = "coral", "Not Significant" = "grey"),
                                                    "Non-zero Coefficient" = coefficient_colours_nonzero),
                                         annotation_name_rot = 45)
```


## Preparing GI heatmap data
```{r}
GI_columns <- c("pga_clonal","pga_subclonal","num.SNVs","Tier_1.counts",
                "num_cnvs","num.svs","Kataegis_count","mtCN")

GI_columns_rlmer <- c("pga_clonal","pga_subclonal","num.SNVs_div100","Tier_1.counts_div100",
                    "num_cnvs","num.svs_div100","Kataegis_count","mtCN")

GI_columns_name <- c("PGA Clonal","PGA Subclonal","SNVs","INDELs","CNAs","SVs","Kataegis","mtCN")
```

```{r}
GI_columns_combined <- cbind(GI_columns, GI_columns_rlmer, GI_columns_name)
```

```{r}
gi_mat <- t(as.matrix(hypoxia_op[,GI_columns]))
rownames(gi_mat) <- c("PGA Clonal","PGA Subclonal","SNVs","INDELs","CNAs","SVs","Kataegis","mtCN")
```

### GI annotations

Merge GI annotations
```{r}
rlmer_GI_annotations <- merge(GI_columns_combined,
                              rlmer_GI_outputs,
                              by.x = "GI_columns_rlmer",
                              by.y = "Variable",
                              all = TRUE)

rownames(rlmer_GI_annotations) <- rlmer_GI_annotations$GI_columns_name
```

Rearrange
```{r}
rlmer_GI_annotations <- rlmer_GI_annotations[GI_columns_name,]
```

```{r}
rlmer_GI_annotations$Significance <- ifelse(rlmer_GI_annotations$measure_p_adj < 0.05, yes = "Significant", no = "Not Significant")
```


```{r}
gi_row_annotation_right <- rowAnnotation("robustlmm" = rlmer_GI_annotations$Significance,
                                         "Coefficient" = rlmer_GI_annotations$Coefficient,
                                         col = list("robustlmm" = c("Significant" = "indianred", "Not Significant" = "grey"),
                                                    "Coefficient" = coefficient_colours_total),
                                          show_annotation_name = FALSE)
```

## Preparing total DRG and driver heatmap
```{r}
total_drg_driver <- merge(summary_somatic_driver,
                          summary_somatic_drg36, 
                          by = "WGS_AssayID")
total_drg_driver <- merge(total_drg_driver,
                          summary_somatic_drg128, 
                          by = "WGS_AssayID")
rownames(total_drg_driver) <- total_drg_driver$WGS_Assay_ID

total_drg_driver_mat <- t(as.matrix(total_drg_driver[,c("bin_muts_in_driver","bin_muts_in_drg36","bin_muts_in_drg128")]))
rownames(total_drg_driver_mat) <- c("Mutations in Drivers","Mutations in DRG36","Mutations in DRG128")
```

```{r}
xtable(rlmer_GI_annotations[,c("GI_columns_name","Coefficient","CI_low","CI_high","p","measure_p_adj")],
       , display = c("s","s","e","e","e","e","e"))
```


## Colours
### CNAs
```{r}
cna_colours_heatmap <- c("Loss" = "blue", "Neutral" = "white", "Gain" = "red")
```

### Mutations
```{r}
mutation_colours <- c("Mutation" = "lightseagreen", "No Mutation" = "white")
```

```{r}
mutation_group_colours <- c("Mutation" = "seagreen", "No Mutation" = "white")
```

### Coefficient
```{r}
# coefficient_colours <- colorRamp2(c(min(rlmer_hallmark_outputs$Coefficient, na.rm = TRUE),
#                                     0,
#                                     max(rlmer_hallmark_outputs$Coefficient, na.rm = TRUE)),
#                                   c("purple1","white","orange1"))
```

### Annotation colours
```{r}
pga_colours <- colorRamp2(c(0,100),c("white","darkred"))
snv_colours <- colorRamp2(c(0,max(hypoxia_op$num.SNVs, na.rm = TRUE)),c("white","darkred"))
indel_colours <- colorRamp2(c(0,max(hypoxia_op$Tier_1.counts, na.rm = TRUE)),c("white","darkred"))
sv_colours <- colorRamp2(c(0,max(hypoxia_op$num.svs, na.rm = TRUE)),c("white","darkred"))
cna_colours <- colorRamp2(c(0,max(hypoxia_op$num_cnvs, na.rm = TRUE)),c("white","darkred"))
kataegis_colours <- colorRamp2(c(0,max(hypoxia_op$Kataegis_count, na.rm = TRUE)),c("white","darkred"))
mtcn_colours <- colorRamp2(c(0,max(hypoxia_op$mtCN, na.rm = TRUE)),c("white","darkred"))
cellularity_colours <- colorRamp2(c(0,max(hypoxia_op$Cellularity, na.rm = TRUE)),c("white","navyblue"))

age_colours <- c("< 40" = "white","40-50" = "grey75","50-65" = "grey50","65-70" = "grey25","≥ 70" = "black")
psa_colours <- c("0-9.9" = "navajowhite", "10-19.9" = "orange", "≥20" = "indianred")
gleason_colours <- c("3+3" = "white", "3+4" = "yellow", "4+3" = "orange", "4+4" = "red", "3+5" = "maroon3", "4+5" = "orangered4", "5+4" = "firebrick4", "5+5" = "black")
clin_t_colours <- c("T1" = "olivedrab2", "T2" = "green3", "T3" = "green4", "T4" = "darkgreen")
erg_colours <- c("TRUE" = PPCG_colourlist$ERG_colourlist$`ERG positive`, "FALSE" = PPCG_colourlist$ERG_colourlist$`ERG negative`)
erg_exp_colours <- c("ERG positive" = PPCG_colourlist$ERG_colourlist$`ERG positive`, "ERG negative" = PPCG_colourlist$ERG_colourlist$`ERG negative`)
isup_colours <- c("1" = PPCG_colourlist$ISUP_colourlist$`1`, 
                  "2" = PPCG_colourlist$ISUP_colourlist$`2`, 
                  "3" = PPCG_colourlist$ISUP_colourlist$`3`, 
                  "4" = PPCG_colourlist$ISUP_colourlist$`4`, 
                  "5" = PPCG_colourlist$ISUP_colourlist$`5`)
```

### Signature colours
```{r}
mutsig_colours <- colorRamp2(c(0,max(log10(mut_sig_mat), na.rm = TRUE)),c("white","#355C7D"))

sbs_colours <- colorRamp2(c(0,max(log10(sbs_sig_mat), na.rm = TRUE)),c("white","#F67280"))
id_colours <- colorRamp2(c(0,max(log10(id_sig_mat), na.rm = TRUE)),c("white","#C06C84"))
cx_colours <- colorRamp2(c(0,max(cx_sig_mat, na.rm = TRUE)),c("white","#355C7D"))
```

### GI colours
```{r}
gi_colours <- colorRamp2(c(0,max(log10(gi_mat), na.rm = TRUE)),c("white","darkred"))
```

## Heatmap annotations
### Top annotation
```{r}
ha_row_annotation_hypoxia <- HeatmapAnnotation(Hypoxia = anno_barplot(hypoxia_op$Buffa_signature_Sum_within))
ha_row_annotation_top <- HeatmapAnnotation(Hypoxia = anno_barplot(hypoxia_op$Buffa_signature_Sum_within),
                                           "PGA Clonal" = hypoxia_op$pga_clonal,
                                           "PGA Subclonal" = hypoxia_op$pga_subclonal,
                                           SNVs = hypoxia_op$num.SNVs,
                                           INDELs = hypoxia_op$Tier_1.counts,
                                           SVs = hypoxia_op$num.svs,
                                           CNAs = hypoxia_op$num_cnvs,
                                           Kataegis = hypoxia_op$Kataegis_count,
                                           mtCN = hypoxia_op$mtCN,
                                           "Grade Group (GG)" = hypoxia_op$ISUP,
                                           PSA = hypoxia_op$psa_bracket,
                                           Age = hypoxia_op$age_bracket,
                                           Cellularity = hypoxia_op$Cellularity,
                                           "ETS Status" = hypoxia_op$ETS,
                                           col = list("PGA Clonal" = pga_colours,
                                                      "PGA Subclonal" = pga_colours,
                                                      SNVs = snv_colours,
                                                      INDELs = indel_colours,
                                                      SVs = sv_colours,
                                                      CNAs = cna_colours,
                                                      Kataegis = kataegis_colours,
                                                      mtCN = mtcn_colours,
                                                      "Grade Group (GG)" = isup_colours,
                                                      PSA = psa_colours,
                                                      Age = age_colours,
                                                      Cellularity = cellularity_colours,
                                                      "ETS Status" = erg_colours
                                           ))

```

### Significance annotations
```{r}
ha_driver_significance_annotation <- rowAnnotation("Q1 v Q4" = driver_hypoxia_enrichment_quart[c(driver_top10_genes),"Significance"],
                                                   "Mutated v Not" = driver_hypoxia_scores[c(driver_top10_genes),"Significance"],
                                                   col = list("Q1 v Q4" = c("Significant" = "goldenrod", "N.S." = "grey"),
                                                              "Mutated v Not" = c("Significant" = "gold", "N.S." = "grey")),
                                                   show_annotation_name = FALSE)

ha_drg36_significance_annotation <- rowAnnotation("Q1 v Q4" = drg36_hypoxia_enrichment_quart[,"Significance"],
                                                   "Mutated v Not" = drg36_hypoxia_scores[,"Significance"],
                                                   col = list("Q1 v Q4" = c("Significant" = "goldenrod", "N.S." = "grey"),
                                                              "Mutated v Not" = c("Significant" = "gold", "N.S." = "grey")))

ha_cna_significance_annotation <- rowAnnotation("Q1 v Q4" = cna_hypoxia_enrichment_quart_sub$Significance,
                                                   col = list("Q1 v Q4" = c("Significant" = "goldenrod", "N.S." = "grey")),
                                                   show_annotation_name = FALSE)
```

```{r}
# ha_mutsig_significance_annotation <- rowAnnotation(Q1vQ4 = ssgsea_wilcox_table$significance_q1_q4,
#                                          # robustlmm = rlmer_hallmark_outputs$significance,
#                                          # Coefficient = rlmer_hallmark_outputs$Coefficient,
#                                          col = list(Q1vQ4 = c("Significant" = "firebrick", "N.S." = "grey")#,
#                                                     # robustlmm = c("Significant" = "indianred", "N.S." = "grey"),
#                                                     # Coefficient = coefficient_colours
#                                                     ))
```

## Somatic mutation annotation
```{r}
mut_sets_annotation <- as.data.frame(c("Mutations in Drivers","Mutations in DRG36","Mutations in DRG128"))
colnames(mut_sets_annotation) <- c("mutation_set")
rownames(mut_sets_annotation) <- mut_sets_annotation$mutation_set
mut_sets_annotation$q1_v_q4 <- NA
mut_sets_annotation$mut_v_not <- NA

mut_sets_annotation["Mutations in Drivers","q1_v_q4"] <- fisher.test(table(summary_somatic_driver_q1q4$bin_muts_in_driver, 
                    summary_somatic_driver_q1q4$Buffa_signature_Sum_within_quartile_one_sample_per_donor))[[1]]
mut_sets_annotation["Mutations in DRG36","q1_v_q4"] <- fisher.test(table(summary_somatic_drg36_q1q4$bin_muts_in_drg36, 
                    summary_somatic_drg36_q1q4$Buffa_signature_Sum_within_quartile_one_sample_per_donor))[[1]]
mut_sets_annotation["Mutations in DRG128","q1_v_q4"] <- fisher.test(table(summary_somatic_drg128_q1q4$bin_muts_in_drg128, 
                    summary_somatic_drg128_q1q4$Buffa_signature_Sum_within_quartile_one_sample_per_donor))[[1]]

mut_sets_annotation["Mutations in Drivers","mut_v_not"] <- wilcox.test(subset(summary_somatic_driver, bin_muts_in_driver == "Mutation")[,"Buffa_signature_Sum_within"], 
                    subset(summary_somatic_driver, bin_muts_in_driver == "No Mutation")[,"Buffa_signature_Sum_within"])[[3]]
mut_sets_annotation["Mutations in DRG36","mut_v_not"] <- wilcox.test(subset(summary_somatic_drg36, bin_muts_in_drg36 == "Mutation")[,"Buffa_signature_Sum_within"], 
                    subset(summary_somatic_drg36, bin_muts_in_drg36 == "No Mutation")[,"Buffa_signature_Sum_within"])[[3]]
mut_sets_annotation["Mutations in DRG128","mut_v_not"] <- wilcox.test(subset(summary_somatic_drg128, bin_muts_in_drg128 == "Mutation")[,"Buffa_signature_Sum_within"], 
                    subset(summary_somatic_drg128, bin_muts_in_drg128 == "No Mutation")[,"Buffa_signature_Sum_within"])[[3]]
```

```{r}
mut_sets_annotation$q1_v_q4_significance <- ifelse(mut_sets_annotation$q1_v_q4 < 0.05, yes = "Significant", no = "N.S.")
mut_sets_annotation$mut_v_not_significance <- ifelse(mut_sets_annotation$mut_v_not < 0.05, yes = "Significant", no = "N.S.")
```

```{r}
ha_mut_sets_annotation <- rowAnnotation("Q1 v Q4" = mut_sets_annotation[,"q1_v_q4_significance"],
                                                   "Mutated v Not" = mut_sets_annotation[,"mut_v_not_significance"],
                                                   col = list("Q1 v Q4" = c("Significant" = "goldenrod", "N.S." = "grey"),
                                                              "Mutated v Not" = c("Significant" = "gold", "N.S." = "grey")),
                                                   show_annotation_name = TRUE,
                                        annotation_name_rot = 45)
```

## Genomic Instability heatmap
```{r}
gen_inst_heatmap <- Heatmap(log10(gi_mat),
                              name = "GI Measures (log10)",
                              right_annotation = gi_row_annotation_right,
                              show_column_names = FALSE,
                              column_split = hypoxia_op$Buffa_signature_Sum_within_quartile_one_sample_per_donor,
                             col = gi_colours,
                             cluster_rows = FALSE)
```

## Driver Mutation heatmap
```{r}
driver_mut_heatmap <- Heatmap(driver_mat[driver_top10_genes,],
                              name = "Driver Mutations",
                              # bottom_annotation = ha_row_annotation_bottom,
                              right_annotation = ha_driver_significance_annotation,
                              show_column_names = FALSE,
                              column_split = hypoxia_op$Buffa_signature_Sum_within_quartile_one_sample_per_donor,
                              col = mutation_colours,
                             cluster_rows = TRUE)
```

## DRG and Driver total heatmap
```{r}
drg_mut_heatmap <- Heatmap(total_drg_driver_mat,
                              name = "Mutations in Groups",
                              right_annotation = ha_mut_sets_annotation,
                              show_column_names = FALSE,
                              column_split = hypoxia_op$Buffa_signature_Sum_within_quartile_one_sample_per_donor,
                              col = mutation_group_colours,
                             cluster_rows = TRUE)
```


## Driver Copy Number Heatmap
```{r}
driver_cn_heatmap <- Heatmap(cna_mat_plot,
                             name = "Driver CNAs",
                             # top_annotation = ha_row_annotation_top,
                             # bottom_annotation = ha_row_annotation_bottom,
                             top_annotation = ha_row_annotation_top,
                             right_annotation = ha_cna_significance_annotation,
                             show_column_names = FALSE,
                             column_split = hypoxia_op$Buffa_signature_Sum_within_quartile_one_sample_per_donor[rownames(hypoxia_op) %in% colnames(cna_mat_plot)],
                             col = cna_colours_heatmap,
                             cluster_rows = TRUE)
```

## Mutational Signature Heatmap

```{r}
# mutsig_heatmap <- Heatmap(mut_sig_mat,
#                           # right_annotation = ha_mutsig_significance_annotation,
#                           show_column_names = FALSE,
#                           column_split = hypoxia_op$Buffa_signature_Sum_within_quartile_one_sample_per_donor,
#                           row_split = mut_sig_split$Signature_type,
#                           col = mutsig_colours,
#                           cluster_rows = FALSE)
```

```{r}
sbs_sig_heatmap <- Heatmap(log10(sbs_sig_mat+1),
                           name = "SBS Mutations (log10)",
                           right_annotation = sbs_row_annotation_right,
                           show_column_names = FALSE,
                           column_split = hypoxia_op$Buffa_signature_Sum_within_quartile_one_sample_per_donor[rownames(hypoxia_op) %in% colnames(sbs_sig_mat)],
                           col = sbs_colours,
                           cluster_rows = FALSE)
```

```{r}
id_sig_heatmap <- Heatmap(log10(id_sig_mat+1),
                          name = "ID INDELS (log10)",
                          right_annotation = id_row_annotation_right,
                          show_column_names = FALSE,
                          column_split = hypoxia_op$Buffa_signature_Sum_within_quartile_one_sample_per_donor[rownames(hypoxia_op) %in% colnames(id_sig_mat)],
                          col = id_colours,
                          cluster_rows = FALSE)
```

```{r}
cx_sig_heatmap <- Heatmap(cx_sig_mat,
                          name = "CX CNAs",
                          right_annotation = cx_row_annotation_right,
                          show_column_names = FALSE,
                          column_split = hypoxia_op$Buffa_signature_Sum_within_quartile_one_sample_per_donor[rownames(hypoxia_op) %in% colnames(cx_sig_mat)],
                          col = cx_colours,
                          cluster_rows = FALSE)
```

# Combine heatmaps
```{r}
ht_list = driver_cn_heatmap %v% driver_mut_heatmap %v% drg_mut_heatmap %v% gen_inst_heatmap %v% sbs_sig_heatmap %v% id_sig_heatmap %v% cx_sig_heatmap
```

```{r fig.height=20, fig.width=17}
draw(ht_list, merge_legend = TRUE, ht_gap = unit(c(2, 2, 24, 2, 2, 2), "mm"))
```

# Supplementary figures
## GI boxplots
```{r}
snv_plot_q <- ggplot(hypoxia_signatures_tum_data_single, aes(x = Buffa_signature_Sum_within_quartile_one_sample_per_donor, y = num.SNVs))+
  geom_jitter(alpha = 0.3, width = 0.3, height = 0, aes(colour = Buffa_signature_Sum_within_quartile_one_sample_per_donor))+
  geom_boxplot(outlier.colour = NA, alpha = 0.3)+
  # scale_colour_manual(values = c("dodgerblue","black"))+
  xlab("")+
  ylab("Number of SNVs (log10)")+
  ggtitle("SNVs")+
  theme_light()+
  # geom_signif(map_signif_level = TRUE, test = "wilcox.test",
  #             comparisons = list(c("Q4","Q1")), tip_length = 0)+
  theme(plot.title = element_text(hjust = 0.5, size = 14), 
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.position = "none")+
  scale_y_log10()+
  scale_colour_manual(values = hypoxia_colours_q)
snv_plot_q
```

```{r}
indel_plot_q <- ggplot(hypoxia_signatures_tum_data_single, aes(x = Buffa_signature_Sum_within_quartile_one_sample_per_donor, y = Tier_1.counts))+
  geom_jitter(alpha = 0.3, width = 0.3, height = 0, aes(colour = Buffa_signature_Sum_within_quartile_one_sample_per_donor))+
  geom_boxplot(outlier.colour = NA, alpha = 0.3)+
  # scale_colour_manual(values = c("dodgerblue","black"))+
  xlab("")+
  ylab("Number of INDELs")+
  ggtitle("INDELs")+
  theme_light()+
  # geom_signif(map_signif_level = TRUE, test = "wilcox.test",
  #             comparisons = list(c("Q4","Q1")), tip_length = 0)+
  theme(plot.title = element_text(hjust = 0.5, size = 14), 
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.position = "none")+
  scale_colour_manual(values = hypoxia_colours_q)
indel_plot_q
```

```{r}
indel_plot_log_q <- 
ggplot(hypoxia_signatures_tum_data_single, aes(x = Buffa_signature_Sum_within_quartile_one_sample_per_donor, y = Tier_1.counts))+
  geom_jitter(alpha = 0.3, width = 0.3, height = 0, aes(colour = Buffa_signature_Sum_within_quartile_one_sample_per_donor))+
  geom_boxplot(outlier.colour = NA, alpha = 0.3)+
  # scale_colour_manual(values = c("dodgerblue","black"))+
  xlab("")+
  ylab("Number of INDELs (log10)")+
  ggtitle("INDELs")+
  theme_light()+
  scale_y_log10()+
  # geom_signif(map_signif_level = TRUE, test = "wilcox.test",
  #             comparisons = list(c("Q4","Q1")), tip_length = 0)+
  theme(plot.title = element_text(hjust = 0.5, size = 14), 
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.position = "none")+
  scale_colour_manual(values = hypoxia_colours_q)
indel_plot_log_q
```

```{r}
sv_plot_q <- ggplot(hypoxia_signatures_tum_data_single, aes(x = Buffa_signature_Sum_within_quartile_one_sample_per_donor, y = num.svs))+
  geom_jitter(alpha = 0.3, width = 0.3, height = 0, aes(colour = Buffa_signature_Sum_within_quartile_one_sample_per_donor))+
  geom_boxplot(outlier.colour = NA, alpha = 0.3)+
  # scale_colour_manual(values = c("dodgerblue","black"))+
  xlab("")+
  ylab("Number of SVs (log10)")+
  ggtitle("SVs")+
  theme_light()+
  # geom_signif(map_signif_level = TRUE, test = "wilcox.test",
  #             comparisons = list(c("Q4","Q1")), tip_length = 0)+
  theme(plot.title = element_text(hjust = 0.5, size = 14), 
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.position = "none")+
  scale_y_log10()+
  scale_colour_manual(values = hypoxia_colours_q)
sv_plot_q
```

```{r}
cna_plot_q <- ggplot(hypoxia_signatures_tum_data_single, aes(x = Buffa_signature_Sum_within_quartile_one_sample_per_donor, y = num_cnvs))+
  geom_jitter(alpha = 0.3, width = 0.3, height = 0, aes(colour = Buffa_signature_Sum_within_quartile_one_sample_per_donor))+
  geom_boxplot(outlier.colour = NA, alpha = 0.3)+
  # scale_colour_manual(values = c("dodgerblue","black"))+
  xlab("")+
  ylab("Number of CNAs (log10)")+
  ggtitle("CNAs")+
  theme_light()+
  # geom_signif(map_signif_level = TRUE, test = "wilcox.test",
  #             comparisons = list(c("Q4","Q1")), tip_length = 0)+
  theme(plot.title = element_text(hjust = 0.5, size = 14), 
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.position = "none")+
  scale_y_log10()+
  scale_colour_manual(values = hypoxia_colours_q)
cna_plot_q
```
```{r}
hypoxia_signatures_tum_data_single$pga_total <- hypoxia_signatures_tum_data_single$pga_clonal + hypoxia_signatures_tum_data_single$pga_subclonal
```

```{r}
pga_tot_plot_q <- ggplot(hypoxia_signatures_tum_data_single, aes(x = Buffa_signature_Sum_within_quartile_one_sample_per_donor, y = pga_total))+
  geom_jitter(alpha = 0.3, width = 0.3, height = 0, aes(colour = Buffa_signature_Sum_within_quartile_one_sample_per_donor))+
  geom_boxplot(outlier.colour = NA, alpha = 0.3)+
  # scale_colour_manual(values = c("dodgerblue","black"))+
  xlab("")+
  ylab("Total PGA")+
  ggtitle("Total PGA")+
  theme_light()+
  # geom_signif(map_signif_level = TRUE, test = "wilcox.test",
  #             comparisons = list(c("Q4","Q1")), tip_length = 0)+
  theme(plot.title = element_text(hjust = 0.5, size = 14),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.position = "none")+
  scale_colour_manual(values = hypoxia_colours_q)
pga_tot_plot_q
```

```{r}
pga_clonal_plot_q <- ggplot(hypoxia_signatures_tum_data_single, aes(x = Buffa_signature_Sum_within_quartile_one_sample_per_donor, y = pga_clonal))+
  geom_jitter(alpha = 0.3, width = 0.3, height = 0, aes(colour = Buffa_signature_Sum_within_quartile_one_sample_per_donor))+
  geom_boxplot(outlier.colour = NA, alpha = 0.3)+
  # scale_colour_manual(values = c("dodgerblue","black"))+
  xlab("")+
  ylab("Clonal PGA")+
  ggtitle("PGA Clonal")+
  theme_light()+
  # geom_signif(map_signif_level = TRUE, test = "wilcox.test",
  #             comparisons = list(c("Q4","Q1")), tip_length = 0)+
  theme(plot.title = element_text(hjust = 0.5, size = 14), 
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.position = "none")+
  scale_colour_manual(values = hypoxia_colours_q)
pga_clonal_plot_q
```

```{r}
pga_subclonal_plot_q <- ggplot(hypoxia_signatures_tum_data_single, aes(x = Buffa_signature_Sum_within_quartile_one_sample_per_donor, y = pga_subclonal))+
  geom_jitter(alpha = 0.3, width = 0.3, height = 0, aes(colour = Buffa_signature_Sum_within_quartile_one_sample_per_donor))+
  # scale_colour_manual(values = c("dodgerblue","black"))+
  geom_boxplot(outlier.colour = NA, alpha = 0.3)+
  xlab("")+
  ylab("Subclonal PGA")+
  ggtitle("PGA Subclonal")+
  theme_light()+
  # geom_signif(map_signif_level = TRUE, test = "wilcox.test",
  #             comparisons = list(c("Q4","Q1")), tip_length = 0)+
  theme(plot.title = element_text(hjust = 0.5, size = 14), 
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.position = "none")+
  scale_colour_manual(values = hypoxia_colours_q)
pga_subclonal_plot_q
```

```{r}
kataegis_plot_q <- ggplot(hypoxia_signatures_tum_data_single, aes(x = Buffa_signature_Sum_within_quartile_one_sample_per_donor, y = Kataegis_count))+
  geom_jitter(alpha = 0.3, width = 0.3, height = 0, aes(colour = Buffa_signature_Sum_within_quartile_one_sample_per_donor))+
  geom_boxplot(outlier.colour = NA, alpha = 0.3)+
  # scale_colour_manual(values = c("dodgerblue","black"))+
  xlab("")+
  ylab("Kataegis Events")+
  ggtitle("Kataegis")+
  theme_light()+
  # geom_signif(map_signif_level = TRUE, test = "wilcox.test",
  #             comparisons = list(c("Q4","Q1")), tip_length = 0)+
  theme(plot.title = element_text(hjust = 0.5, size = 14), 
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.position = "none")+
  scale_colour_manual(values = hypoxia_colours_q)
kataegis_plot_q
```

```{r}
mtcn_plot_q <- ggplot(hypoxia_signatures_tum_data_single, aes(x = Buffa_signature_Sum_within_quartile_one_sample_per_donor, y = mtCN))+
  geom_jitter(alpha = 0.3, width = 0.3, height = 0, aes(colour = Buffa_signature_Sum_within_quartile_one_sample_per_donor))+
  geom_boxplot(outlier.colour = NA, alpha = 0.3)+
  # scale_colour_manual(values = c("dodgerblue","black"))+
  xlab("")+
  ylab("mtCN")+
  ggtitle("mtCN")+
  theme_light()+
  # geom_signif(map_signif_level = TRUE, test = "wilcox.test",
  #             comparisons = list(c("Q4","Q1")), tip_length = 0)+
  theme(plot.title = element_text(hjust = 0.5, size = 14), 
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.position = "none")+
  scale_colour_manual(values = hypoxia_colours_q)
mtcn_plot_q
```

### Composite boxplot
```{r fig.height=8, fig.width=14}
grid.arrange(pga_clonal_plot_q+labs(tag = "A"), pga_subclonal_plot_q+labs(tag = "B"),pga_tot_plot_q+labs(tag = "C"), kataegis_plot_q+labs(tag = "D"),mtcn_plot_q+labs(tag = "E"),
             snv_plot_q+labs(tag = "F"), indel_plot_log_q+labs(tag = "G"), sv_plot_q+labs(tag = "H"), cna_plot_q+labs(tag = "I"),
             nrow = 2, ncol = 10,
             layout_matrix = rbind(c(1,1,2,2,3,3,4,4,5,5),
                                   c(NA,6,6,7,7,8,8,9,9,NA)))
```
## GI scatter
```{r}
snv_plot_s <- ggplot(hypoxia_signatures_tum_data_single, aes(x = Buffa_signature_Sum_within, y = num.SNVs))+
  geom_point(alpha = 0.3, aes(colour = Buffa_signature_Sum_within_quartile_one_sample_per_donor))+
  # scale_colour_manual(values = c("dodgerblue","black"))+
  xlab("")+
  ylab("Number of SNVs (log10)")+
  ggtitle("SNVs")+
  theme_light()+
  theme(plot.title = element_text(hjust = 0.5, size = 14), 
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.position = "none")+
  scale_y_log10()+
  scale_colour_manual(values = hypoxia_colours_q)+
  geom_smooth(method = "lm", se = FALSE, colour = "grey30")
snv_plot_s
```

```{r}
indel_plot_s <- ggplot(hypoxia_signatures_tum_data_single, aes(x = Buffa_signature_Sum_within, y = Tier_1.counts))+
  geom_point(alpha = 0.3, aes(colour = Buffa_signature_Sum_within_quartile_one_sample_per_donor))+
  # scale_colour_manual(values = c("dodgerblue","black"))+
  xlab("")+
  ylab("Number of INDELs")+
  ggtitle("INDELs")+
  theme_light()+
  theme(plot.title = element_text(hjust = 0.5, size = 14), 
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.position = "none")+
  scale_colour_manual(values = hypoxia_colours_q)+
  geom_smooth(method = "lm", se = FALSE, colour = "grey30")
indel_plot_s
```

```{r}
indel_plot_log_s <- 
ggplot(hypoxia_signatures_tum_data_single, aes(x = Buffa_signature_Sum_within, y = Tier_1.counts))+
  geom_point(alpha = 0.3, aes(colour = Buffa_signature_Sum_within_quartile_one_sample_per_donor))+
  # scale_colour_manual(values = c("dodgerblue","black"))+
  xlab("")+
  ylab("Number of INDELs (log10)")+
  ggtitle("INDELs")+
  theme_light()+
  scale_y_log10()+
  theme(plot.title = element_text(hjust = 0.5, size = 14), 
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.position = "none")+
  scale_colour_manual(values = hypoxia_colours_q)+
  geom_smooth(method = "lm", se = FALSE, colour = "grey30")
indel_plot_log_s
```

```{r}
sv_plot_s <- ggplot(hypoxia_signatures_tum_data_single, aes(x = Buffa_signature_Sum_within, y = num.svs))+
  geom_point(alpha = 0.3, aes(colour = Buffa_signature_Sum_within_quartile_one_sample_per_donor))+
  # scale_colour_manual(values = c("dodgerblue","black"))+
  xlab("")+
  ylab("Number of SVs (log10)")+
  ggtitle("SVs")+
  theme_light()+
  theme(plot.title = element_text(hjust = 0.5, size = 14), 
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.position = "none")+
  scale_y_log10()+
  scale_colour_manual(values = hypoxia_colours_q)+
  geom_smooth(method = "lm", se = FALSE, colour = "grey30")
sv_plot_s
```

```{r}
cna_plot_s <- ggplot(hypoxia_signatures_tum_data_single, aes(x = Buffa_signature_Sum_within, y = num_cnvs))+
  geom_point(alpha = 0.3, aes(colour = Buffa_signature_Sum_within_quartile_one_sample_per_donor))+
  # scale_colour_manual(values = c("dodgerblue","black"))+
  xlab("")+
  ylab("Number of CNAs (log10)")+
  ggtitle("CNAs")+
  theme_light()+
  theme(plot.title = element_text(hjust = 0.5, size = 14), 
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.position = "none")+
  scale_y_log10()+
  scale_colour_manual(values = hypoxia_colours_q)+
  geom_smooth(method = "lm", se = FALSE, colour = "grey30")
cna_plot_s
```

```{r}
pga_tot_plot_s <- ggplot(hypoxia_signatures_tum_data_single, aes(x = Buffa_signature_Sum_within, y = pga_total))+
  geom_point(alpha = 0.3, aes(colour = Buffa_signature_Sum_within_quartile_one_sample_per_donor))+
  # scale_colour_manual(values = c("dodgerblue","black"))+
  xlab("")+
  ylab("Total PGA")+
  ggtitle("Total PGA")+
  theme_light()+
  theme(plot.title = element_text(hjust = 0.5, size = 14), 
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.position = "none")+
  scale_colour_manual(values = hypoxia_colours_q)+
  geom_smooth(method = "lm", se = FALSE, colour = "grey30")
pga_tot_plot_s
```

```{r}
pga_clonal_plot_s <- ggplot(hypoxia_signatures_tum_data_single, aes(x = Buffa_signature_Sum_within, y = pga_clonal))+
  geom_point(alpha = 0.3, aes(colour = Buffa_signature_Sum_within_quartile_one_sample_per_donor))+
  # scale_colour_manual(values = c("dodgerblue","black"))+
  xlab("")+
  ylab("Clonal PGA")+
  ggtitle("PGA Clonal")+
  theme_light()+
  theme(plot.title = element_text(hjust = 0.5, size = 14), 
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.position = "none")+
  scale_colour_manual(values = hypoxia_colours_q)+
  geom_smooth(method = "lm", se = FALSE, colour = "grey30")
pga_clonal_plot_s
```

```{r}
pga_subclonal_plot_s <- ggplot(hypoxia_signatures_tum_data_single, aes(x = Buffa_signature_Sum_within, y = pga_subclonal))+
  geom_point(alpha = 0.3, aes(colour = Buffa_signature_Sum_within_quartile_one_sample_per_donor))+
  # scale_colour_manual(values = c("dodgerblue","black"))+
  xlab("")+
  ylab("Subclonal PGA")+
  ggtitle("PGA Subclonal")+
  theme_light()+
  theme(plot.title = element_text(hjust = 0.5, size = 14), 
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.position = "none")+
  scale_colour_manual(values = hypoxia_colours_q)+
  geom_smooth(method = "lm", se = FALSE, colour = "grey30")
pga_subclonal_plot_s
```

```{r}
kataegis_plot_s <- ggplot(hypoxia_signatures_tum_data_single, aes(x = Buffa_signature_Sum_within, y = Kataegis_count))+
  geom_point(alpha = 0.3, aes(colour = Buffa_signature_Sum_within_quartile_one_sample_per_donor))+
  # scale_colour_manual(values = c("dodgerblue","black"))+
  xlab("")+
  ylab("Kataegis Events")+
  ggtitle("Kataegis")+
  theme_light()+
  theme(plot.title = element_text(hjust = 0.5, size = 14), 
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.position = "none")+
  scale_colour_manual(values = hypoxia_colours_q)+
  geom_smooth(method = "lm", se = FALSE, colour = "grey30")
kataegis_plot_s
```

```{r}
mtcn_plot_s <- ggplot(hypoxia_signatures_tum_data_single, aes(x = Buffa_signature_Sum_within, y = mtCN))+
  geom_point(alpha = 0.3, aes(colour = Buffa_signature_Sum_within_quartile_one_sample_per_donor))+
  # scale_colour_manual(values = c("dodgerblue","black"))+
  xlab("")+
  ylab("mtCN")+
  ggtitle("mtCN")+
  theme_light()+
  theme(plot.title = element_text(hjust = 0.5, size = 14), 
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.position = "none")+
  scale_colour_manual(values = hypoxia_colours_q)+
  geom_smooth(method = "lm", se = FALSE, colour = "grey30")
mtcn_plot_s
```

### Composite scatter plot
```{r fig.height=8, fig.width=14}
grid.arrange(pga_clonal_plot_s+labs(tag = "A"), pga_subclonal_plot_s+labs(tag = "B"),pga_tot_plot_s+labs(tag = "C"), kataegis_plot_s+labs(tag = "D"), mtcn_plot_s+labs(tag = "E"),
             snv_plot_s+labs(tag = "F"), indel_plot_log_s+labs(tag = "G"), sv_plot_s+labs(tag = "H"), cna_plot_s+labs(tag = "I"),
             nrow = 2, ncol = 10,
             layout_matrix = rbind(c(1,1,2,2,3,3,4,4,5,5),
                                   c(NA,6,6,7,7,8,8,9,9,NA)))
```

```{r fig.height=16, fig.width=14}
grid.arrange(pga_clonal_plot_q+labs(tag = "A"), pga_subclonal_plot_q, pga_tot_plot_q, kataegis_plot_q, mtcn_plot_q,
             snv_plot_q, indel_plot_log_q, sv_plot_q, cna_plot_q,
             pga_clonal_plot_s+labs(tag = "B"), pga_subclonal_plot_s, pga_tot_plot_s, kataegis_plot_s, mtcn_plot_s,
             snv_plot_s, indel_plot_log_s, sv_plot_s, cna_plot_s,
             nrow = 4, ncol = 10,
             layout_matrix = rbind(c(1,1,2,2,3,3,4,4,5,5),
                                   c(NA,6,6,7,7,8,8,9,9,NA),
                                   c(10,10,11,11,12,12,13,13,14,14),
                                   c(NA,15,15,16,16,17,17,18,18,NA)))
```

## SBS box and scatter plots
```{r}
sbs_data_melt <- melt(as.data.table(hypoxia_signatures_tum_data_single), 
                      id.vars = c("PPCG_SampleID",
                                  "Buffa_signature_Sum_within", "Buffa_signature_Sum_within_quartile_one_sample_per_donor"),
                      measure.vars = sbs_sig_columns,
                      variable.name = "Signature", value.name = "Number")
```

```{r}
sbs_plot_q <- ggplot(sbs_data_melt, aes(x = Buffa_signature_Sum_within_quartile_one_sample_per_donor, y = Number))+
  geom_jitter(alpha = 0.3, width = 0.3, height = 0, aes(colour = Buffa_signature_Sum_within_quartile_one_sample_per_donor))+
  geom_boxplot(outlier.colour = NA, alpha = 0.3)+
  # scale_colour_manual(values = c("dodgerblue","black"))+
  xlab("Hypoxia Quartile")+
  ylab("Number of SNVs")+
  ggtitle("")+
  theme_light()+
  # geom_signif(map_signif_level = TRUE, test = "wilcox.test",
  #             comparisons = list(c("Q4","Q1")), tip_length = 0)+
  theme(text = element_text(size = 16),
        legend.position = "none")+
  facet_wrap(~Signature, scales = "free_y")+
  scale_colour_manual(values = hypoxia_colours_q)+
  scale_y_log10()
sbs_plot_q
```

```{r}
sbs_plot_s <- ggplot(sbs_data_melt, aes(x = Buffa_signature_Sum_within, y = Number))+
  geom_point(alpha = 0.3, aes(colour = Buffa_signature_Sum_within_quartile_one_sample_per_donor))+
  # scale_colour_manual(values = c("dodgerblue","black"))+
  xlab("Hypoxia Score")+
  ylab("Number of SNVs")+
  ggtitle("")+
  theme_light()+
  theme(text = element_text(size = 16),
        legend.position = "none")+
  facet_wrap(~Signature, scales = "free_y")+
  scale_colour_manual(values = hypoxia_colours_q)+
  scale_y_log10()+
  geom_smooth(method = "lm", se = FALSE, colour = "grey30")
sbs_plot_s
```

```{r fig.height=18, fig.width=14}
grid.arrange(sbs_plot_q+labs(tag = "A"),
             sbs_plot_s+labs(tag = "B"),
             nrow = 2, ncol = 1)
```

## ID box and scatter plots
```{r}
id_data_melt <- melt(as.data.table(hypoxia_signatures_tum_data_single), 
                      id.vars = c("PPCG_SampleID",
                                  "Buffa_signature_Sum_within", "Buffa_signature_Sum_within_quartile_one_sample_per_donor"),
                      measure.vars = id_sig_columns,
                      variable.name = "Signature", value.name = "Number")
```

```{r}
id_plot_q <- ggplot(id_data_melt, aes(x = Buffa_signature_Sum_within_quartile_one_sample_per_donor, y = Number))+
  geom_jitter(alpha = 0.3, width = 0.3, height = 0, aes(colour = Buffa_signature_Sum_within_quartile_one_sample_per_donor))+
  geom_boxplot(outlier.colour = NA, alpha = 0.3)+
  # scale_colour_manual(values = c("dodgerblue","black"))+
  xlab("Hypoxia Quartile")+
  ylab("Number of INDELs")+
  ggtitle("")+
  theme_light()+
  # geom_signif(map_signif_level = TRUE, test = "wilcox.test",
  #             comparisons = list(c("Q4","Q1")), tip_length = 0)+
  theme(text = element_text(size = 16),
        legend.position = "none")+
  facet_wrap(~Signature, scales = "free_y")+
  scale_colour_manual(values = hypoxia_colours_q)+
  scale_y_log10()
id_plot_q
```

```{r}
id_plot_s <- ggplot(id_data_melt, aes(x = Buffa_signature_Sum_within, y = Number))+
  geom_point(alpha = 0.3, aes(colour = Buffa_signature_Sum_within_quartile_one_sample_per_donor))+
  # scale_colour_manual(values = c("dodgerblue","black"))+
  xlab("Hypoxia Score")+
  ylab("Number of INDELs")+
  ggtitle("")+
  theme_light()+
  theme(text = element_text(size = 16),
        legend.position = "none")+
  facet_wrap(~Signature, scales = "free_y")+
  scale_colour_manual(values = hypoxia_colours_q)+
  scale_y_log10()+
  geom_smooth(method = "lm", se = FALSE, colour = "grey30")
id_plot_s
```

```{r fig.height=16, fig.width=14}
grid.arrange(id_plot_q+labs(tag = "A"),
             id_plot_s+labs(tag = "B"),
             nrow = 2, ncol = 1)
```

## CX box and scatter plots
```{r}
cx_data_melt <- melt(as.data.table(hypoxia_signatures_tum_data_single), 
                      id.vars = c("PPCG_SampleID",
                                  "Buffa_signature_Sum_within", "Buffa_signature_Sum_within_quartile_one_sample_per_donor"),
                      measure.vars = cx_sig_columns,
                      variable.name = "Signature", value.name = "Number")
```

```{r}
cx_plot_q <- ggplot(cx_data_melt, aes(x = Buffa_signature_Sum_within_quartile_one_sample_per_donor, y = Number))+
  geom_jitter(alpha = 0.3, width = 0.3, height = 0, aes(colour = Buffa_signature_Sum_within_quartile_one_sample_per_donor))+
  geom_boxplot(outlier.colour = NA, alpha = 0.3)+
  # scale_colour_manual(values = c("dodgerblue","black"))+
  xlab("Hypoxia Quartile")+
  ylab("Number of CNAs")+
  ggtitle("")+
  theme_light()+
  # geom_signif(map_signif_level = TRUE, test = "wilcox.test",
  #             comparisons = list(c("Q4","Q1")), tip_length = 0)+
  theme(text = element_text(size = 16),
        legend.position = "none")+
  facet_wrap(~Signature, scales = "free_y")+
  scale_colour_manual(values = hypoxia_colours_q)+
  scale_y_log10()
cx_plot_q
```

```{r}
cx_plot_s <- ggplot(cx_data_melt, aes(x = Buffa_signature_Sum_within, y = Number))+
  geom_point(alpha = 0.3, aes(colour = Buffa_signature_Sum_within_quartile_one_sample_per_donor))+
  # scale_colour_manual(values = c("dodgerblue","black"))+
  xlab("Hypoxia Score")+
  ylab("Number of CNAs")+
  ggtitle("")+
  theme_light()+
  theme(text = element_text(size = 16),
        legend.position = "none")+
  facet_wrap(~Signature, scales = "free_y")+
  scale_colour_manual(values = hypoxia_colours_q)+
  scale_y_log10()+
  geom_smooth(method = "lm", se = FALSE, colour = "grey30")
cx_plot_s
```

```{r fig.height=16, fig.width=14}
grid.arrange(cx_plot_q+labs(tag = "A"),
             cx_plot_s+labs(tag = "B"),
             nrow = 2, ncol = 1)
```

# Session Info
```{r}
sessionInfo()
```

