---
title: "Figure 4 - metastasis features. RNA data release 1.4k10prerelease"
author: Maria Jakobsdottir (Maria.Jakobsdottir@manchester.ac.uk)
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2:
    toc: true 
    toc_float: true
    toc_depth: 6
---

Analysis for figure 4 in Hypoxia manuscript.

# Libraries used
```{r message=FALSE, warning=FALSE}
library(dplyr)
library(finalfit)
library(ggplot2)
library(ggpubr)
library(ggsci)
library(ggsignif)
library(ggsurvfit)
library(grid)
library(gridExtra)
library(splitstackshape) # for paired analysis
library(survival)
library(tidyr)
library(viridis)
library(xtable)

options(bitmapType='cairo')
'%!in%' <- function(x,y)!('%in%'(x,y))
```

# Data
Read in data

```{r}
params <- list(ppcg_colourlist = "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/GitHub/PPCG_colours/PPCG_colourlist.RData",
               combined_data = "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_data_private/combined_PPCG_WGS_RNA_summary_data_clinical_v9_2026-05-31.txt")
```

```{r}
load(params$ppcg_colourlist)
```

Parameters, here for simplicity, otherwise kept in submission script. 
```{r}
print(params)
```

## Outdirs
```{r}
fig_outdir <- "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_paper_final_results/Figures/"
```

```{r}
fig_outdir_supp <- "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_paper_final_results/Supplementary_Figures/"
```

```{r}
fig_data_outdir <- "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_paper_final_results/Figure_Data/"
```

```{r}
supp_fig_data_outdir <- "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_paper_final_results/Supplementary_Figure_Data/"
```

## Combined data

```{r}
combined_data <- read.delim(params$combined_data, header = TRUE, sep = "\t")
```



# Merge data and select samples
Use Tier 1 Buffa signature sum.

## Selecting tumour only and one sample per patient
```{r}
hypoxia_signatures_tum_data_single <- subset(combined_data, !is.na(Buffa_signature_Sum_within_quartile_one_sample_per_donor))
```
```{r}
dim(hypoxia_signatures_tum_data_single)
```

```{r}
summary(hypoxia_signatures_tum_data_single)
```



Here is a list of the IDs that were used for the single tumour sample per patient analysis:
```{r}
hypoxia_signatures_tum_data_single$PPCG_SampleID
```
 


# Paired primaries and mets

```{r}
combined_data$RNA_Sample_Type <- as.factor(combined_data$RNA_Sample_Type)
combined_data$MetastaticSite <- as.factor(combined_data$MetastaticSite)
summary(combined_data)
```

```{r}
combined_data_paired <- subset(combined_data, (PPCG_Donor_ID %in% unique(combined_data$PPCG_Donor_ID[duplicated(combined_data$PPCG_Donor_ID)])) & 
                                 (RNA_Sample_Type != "BEN"))
combined_data_paired$donor_and_type <- paste0(combined_data_paired$PPCG_Donor_ID, "_", combined_data_paired$RNA_Sample_Type)
```
```{r}
dim(combined_data_paired)
```


```{r}
combined_data_paired_long <- subset(combined_data_paired, RNA_Sample_Type %in% c("MET","TUM"))
combined_data_paired_long <- getanID(combined_data_paired_long[,c("PPCG_Donor_ID","Buffa_signature_Sum_within","RNA_Sample_Type","MetastaticSite","donor_and_type")], c(paste("donor_and_type")))

for(i in 1:nrow(combined_data_paired_long)){
  if(combined_data_paired_long$RNA_Sample_Type[i] == "MET" & combined_data_paired_long$.id[i] == 2){
    temp_sample_data <- subset(combined_data_paired_long, PPCG_Donor_ID == combined_data_paired_long$PPCG_Donor_ID[i])
    if("TUM" %in% temp_sample_data$RNA_Sample_Type){
      temp_row <- subset(temp_sample_data, RNA_Sample_Type == "TUM")
      temp_row$.id <- 2
      
      combined_data_paired_long <-rbind(combined_data_paired_long, temp_row)
    }
  }
  if(combined_data_paired_long$RNA_Sample_Type[i] == "MET" & combined_data_paired_long$.id[i] == 3){
    temp_sample_data <- subset(combined_data_paired_long, PPCG_Donor_ID == combined_data_paired_long$PPCG_Donor_ID[i])
    if("TUM" %in% temp_sample_data$RNA_Sample_Type){
      temp_row <- subset(temp_sample_data, RNA_Sample_Type == "TUM")
      temp_row <- temp_row[1]
      temp_row$.id <- 3
      
      combined_data_paired_long <-rbind(combined_data_paired_long, temp_row)
    }
  }
    
}

combined_data_paired_long <- reshape(combined_data_paired_long, 
        idvar = c("PPCG_Donor_ID",".id"), timevar = "RNA_Sample_Type", direction = "wide")
```

```{r}
combined_data_paired_long_only <- subset(combined_data_paired_long, !is.na(Buffa_signature_Sum_within.MET) & !is.na(Buffa_signature_Sum_within.TUM))
```

```{r}
combined_data_paired_long_only$line_colour <- ifelse(combined_data_paired_long_only$Buffa_signature_Sum_within.TUM > combined_data_paired_long_only$Buffa_signature_Sum_within.MET, "Lower_in_Met", "Higher_in_MET")
combined_data_paired_long_only$multisample <- ifelse(combined_data_paired_long_only$PPCG_Donor_ID %in%  combined_data_paired_long_only$PPCG_Donor_ID[duplicated(combined_data_paired_long_only$PPCG_Donor_ID)],
                                                          "Multiple pairs per patient","Single pair per patient")
combined_data_paired_long_only$MetastaticSite.MET <- as.character(combined_data_paired_long_only$MetastaticSite.MET)
combined_data_paired_long_only$MetastaticSite.MET[is.na(combined_data_paired_long_only$MetastaticSite.MET)] <- "Unknown"
combined_data_paired_long_only$MetastaticSite.MET[combined_data_paired_long_only$MetastaticSite.MET == "BONE"] <- "Bone"
combined_data_paired_long_only$MetastaticSite.MET[combined_data_paired_long_only$MetastaticSite.MET == "LN"] <- "Lymph Node"
combined_data_paired_long_only$MetastaticSite.MET <- as.factor(combined_data_paired_long_only$MetastaticSite.MET)
```

```{r}
combined_data_paired_long_only
```

Make a data frame with a single pair per patient
```{r}
combined_data_paired_long_only_single <- subset(combined_data_paired_long_only, .id == "1")
```

```{r}
paired_plot <- ggpaired(combined_data_paired_long_only_single, cond1 = "Buffa_signature_Sum_within.TUM", cond2 = "Buffa_signature_Sum_within.MET", line.color = "line_colour", palette = "npg",
         xlab = "Sample Type", ylab = "Hypoxia Score")+
  scale_x_discrete(labels = c("Primary","Metastasis"))+
  scale_colour_npg(name = "Line Colour:", labels = c("Higher in metastatic\n         sample", "Lower in metastastatic\n         sample"))+
  geom_signif(map_signif_level = TRUE, comparisons = list(c("Buffa_signature_Sum_within.TUM","Buffa_signature_Sum_within.MET")), test = "wilcox.test", tip_length = 0, y_position = 50)+
  theme_light()+
  theme(legend.text = element_text(size = 12), legend.position = "bottom")
paired_plot
```

```{r}
supp_fig_7_data <- combined_data_paired_long_only_single[,c("Buffa_signature_Sum_within.TUM", "Buffa_signature_Sum_within.MET", "line_colour", "MetastaticSite.MET")]
```

```{r}
ggpaired(supp_fig_7_data, cond1 = "Buffa_signature_Sum_within.TUM", cond2 = "Buffa_signature_Sum_within.MET", line.color = "line_colour", palette = "npg",
         xlab = "Sample Type", ylab = "Hypoxia Score")+
  scale_x_discrete(labels = c("Primary","Metastasis"))+
  scale_colour_npg(name = "Line Colour:", labels = c("Higher in metastatic\n         sample", "Lower in metastastatic\n         sample"))+
  geom_signif(map_signif_level = TRUE, comparisons = list(c("Buffa_signature_Sum_within.TUM","Buffa_signature_Sum_within.MET")), test = "wilcox.test", tip_length = 0, y_position = 50)+
  theme_light()+
  theme(legend.text = element_text(size = 12), legend.position = "bottom")
```


```{r}
paired_plot_facet <- ggpaired(combined_data_paired_long_only_single, cond1 = "Buffa_signature_Sum_within.TUM", cond2 = "Buffa_signature_Sum_within.MET", facet.by =  "MetastaticSite.MET", line.color = "line_colour", palette = "npg",
         xlab = "Sample Type", ylab = "Hypoxia Score")+
  scale_x_discrete(labels = c("Primary","Metastasis"))+
  theme_light()+
  theme(legend.position = "none")+
  geom_signif(map_signif_level = TRUE, comparisons = list(c("Buffa_signature_Sum_within.TUM","Buffa_signature_Sum_within.MET")), test = "wilcox.test", tip_length = 0, y_position = 50)
paired_plot_facet
```

```{r}
ggpaired(supp_fig_7_data, cond1 = "Buffa_signature_Sum_within.TUM", cond2 = "Buffa_signature_Sum_within.MET", facet.by =  "MetastaticSite.MET", line.color = "line_colour", palette = "npg",
         xlab = "Sample Type", ylab = "Hypoxia Score")+
  scale_x_discrete(labels = c("Primary","Metastasis"))+
  theme_light()+
  theme(legend.position = "none")+
  geom_signif(map_signif_level = TRUE, comparisons = list(c("Buffa_signature_Sum_within.TUM","Buffa_signature_Sum_within.MET")), test = "wilcox.test", tip_length = 0, y_position = 50)
```


```{r fig.height=8, fig.width=6}
grid.arrange(paired_plot+labs(tag = "A"), paired_plot_facet +labs(tag = "B"), nrow = 2, ncol = 1)
```

```{r}
write.csv(supp_fig_7_data, file = paste0(supp_fig_data_outdir,"Supplementary_Figure_7_data.csv"), row.names = FALSE)
```


```{r}
wilcox.test(combined_data_paired_long_only_single$Buffa_signature_Sum_within.TUM, 
            combined_data_paired_long_only_single$Buffa_signature_Sum_within.MET,
            paired = TRUE,
            conf.int = TRUE
            )
```

```{r}
wilcox.test(subset(combined_data_paired_long_only_single, MetastaticSite.MET == "Bone")$Buffa_signature_Sum_within.TUM,
            subset(combined_data_paired_long_only_single, MetastaticSite.MET == "Bone")$Buffa_signature_Sum_within.MET,
            paired = TRUE,
            conf.int = TRUE
            )
```

```{r}
wilcox.test(subset(combined_data_paired_long_only_single, MetastaticSite.MET == "Lymph Node")$Buffa_signature_Sum_within.TUM,
            subset(combined_data_paired_long_only_single, MetastaticSite.MET == "Lymph Node")$Buffa_signature_Sum_within.MET,
            paired = TRUE,
            conf.int = TRUE
            )
```
```{r}
# wilcox.test(subset(combined_data_paired_long_only, MetastaticSite.MET == "BONE")$Buffa_signature_Sum_within.TUM,
#             subset(combined_data_paired_long_only, MetastaticSite.MET == "LN")$Buffa_signature_Sum_within.TUM
#             )
```

```{r}
dim(combined_data_paired_long_only_single)
```

```{r}
length(unique(combined_data_paired_long_only_single$PPCG_Donor_ID))
```

```{r}
cor.test(combined_data_paired_long_only_single$Buffa_signature_Sum_within.TUM, 
            combined_data_paired_long_only_single$Buffa_signature_Sum_within.MET,
            method = "spearman"
            )
```

```{r}
cor_pm_plot <- ggplot(combined_data_paired_long_only_single, aes(x = Buffa_signature_Sum_within.TUM, y = Buffa_signature_Sum_within.MET))+
  geom_smooth(method = "lm", se = FALSE, colour = "grey30")+
  geom_point(alpha = 0.8, aes(colour = MetastaticSite.MET))+
  scale_colour_bmj(name = "Metastatic Site")+
  # scale_shape_manual(name = "Multisample", values = c(17,19))+
  theme_light()+
  xlab("Hypoxia Score - Primary Sample")+
  ylab("Hypoxia Score - Metastatic Sample")+
  annotate("text", label = paste0("Spearman's rho = ",
                                 round(cor.test(combined_data_paired_long_only_single$Buffa_signature_Sum_within.TUM, 
            combined_data_paired_long_only_single$Buffa_signature_Sum_within.MET,
            method = "spearman")[[4]], digits = 3),
                                 "; p = ",
            format(cor.test(combined_data_paired_long_only_single$Buffa_signature_Sum_within.TUM, 
            combined_data_paired_long_only_single$Buffa_signature_Sum_within.MET,
            method = "spearman")[[3]], scientific = TRUE, digits = 3)), x = -15, y = 90, colour = "black")+
  theme(legend.position = "bottom")+
  guides(shape = guide_legend(nrow = 2), colour = guide_legend(nrow = 2))
cor_pm_plot
```

```{r}
fig_4b_data <- combined_data_paired_long_only_single[,c("Buffa_signature_Sum_within.TUM", "Buffa_signature_Sum_within.MET", "MetastaticSite.MET"), drop = FALSE]
```

```{r}
ggplot(fig_4b_data, aes(x = Buffa_signature_Sum_within.TUM, y = Buffa_signature_Sum_within.MET))+
  geom_smooth(method = "lm", se = FALSE, colour = "grey30")+
  geom_point(alpha = 0.8, aes(colour = MetastaticSite.MET))+
  scale_colour_bmj(name = "Metastatic Site")+
  # scale_shape_manual(name = "Multisample", values = c(17,19))+
  theme_light()+
  xlab("Hypoxia Score - Primary Sample")+
  ylab("Hypoxia Score - Metastatic Sample")+
  annotate("text", label = paste0("Spearman's ρ = ",
                                 round(cor.test(combined_data_paired_long_only_single$Buffa_signature_Sum_within.TUM, 
            combined_data_paired_long_only_single$Buffa_signature_Sum_within.MET,
            method = "spearman")[[4]], digits = 3),
                                 "; p = ",
            format(cor.test(combined_data_paired_long_only_single$Buffa_signature_Sum_within.TUM, 
            combined_data_paired_long_only_single$Buffa_signature_Sum_within.MET,
            method = "spearman")[[3]], scientific = TRUE, digits = 3)), x = -15, y = 90, colour = "black")+
  theme(legend.position = "bottom")+
  guides(shape = guide_legend(nrow = 2), colour = guide_legend(nrow = 2))
```

```{r}
write.csv(fig_4b_data, file = paste0(fig_data_outdir,"Figure_4b_data.csv"), row.names = FALSE)
```


```{r fig.height=12, fig.width=6}
grid.arrange(paired_plot+labs(tag = "A"), paired_plot_facet +labs(tag = "B"), cor_pm_plot+labs(tag = "C"),
             nrow = 3, ncol = 1)
```


# Primaries with or without mets
```{r}
# single sample of each type
combined_data_single <- combined_data[!duplicated(combined_data[,c("RNA_Sample_Type","PPCG_Donor_ID")]),]
combined_data_single <- subset(combined_data_single, RNA_Sample_Type %in% c("TUM","MET"))
combined_data_single <- as.data.frame(combined_data_single)

sample_and_type <- as.data.frame(table(combined_data_single$RNA_Sample_Type, combined_data_single$PPCG_Donor_ID))

combined_data_single$met_status <- NA
for(i in 1:nrow(combined_data_single)){
  sample_and_type_sub_tum <- subset(sample_and_type, Var2 == combined_data_single$PPCG_Donor_ID[i] &
                                      Var1 == "TUM")
  sample_and_type_sub_met <- subset(sample_and_type, Var2 == combined_data_single$PPCG_Donor_ID[i] &
                                      Var1 == "MET")
  
  if(combined_data_single$RNA_Sample_Type[i] == "TUM" & sample_and_type_sub_met$Freq > 0){
    combined_data_single$met_status[i] <- "Primary with met"
  }else if(combined_data_single$RNA_Sample_Type[i] == "TUM" & sample_and_type_sub_met$Freq == 0){
    combined_data_single$met_status[i] <- "Primary without met"
  }
  if(combined_data_single$RNA_Sample_Type[i] == "MET"){
    combined_data_single$met_status[i] <- "Metastasis"
  }
}

combined_data_single$met_status <- as.factor(combined_data_single$met_status)
combined_data_single$met_status <- factor(combined_data_single$met_status,
                                               levels = c("Primary without met","Primary with met","Metastasis"))
```

```{r}
ggplot(combined_data_single, aes(x = met_status, y = Buffa_signature_Sum_within))+
  geom_jitter(alpha = 0.5, width = 0.3, colour = "grey", height = 0)+
  geom_boxplot(alpha = 0.5)+
  xlab("Metastasis status")+
  ylab("Hypoxia Score")+
  theme_light()+
  geom_signif(map_signif_level = TRUE, comparisons = list(c("Primary without met","Primary with met"),
                                                          c("Primary without met","Metastasis"),
                                                          c("Primary with met","Metastasis")),
              test = "wilcox.test", tip_length = 0,
              y_position = c(75,90,115))+
  scale_x_discrete(labels=c("Primary with met" = "Primary with \nmetastasis", 
                            "Primary without met" = "Primary without \nmetastasis",
                            "Metastasis" = "Metastasis"))
```

```{r}
wilcox.test(subset(combined_data_single, met_status == "Primary without met")$Buffa_signature_Sum_within,
            subset(combined_data_single, met_status == "Primary with met")$Buffa_signature_Sum_within,
            conf.int = TRUE)
```

```{r}
wilcox.test(subset(combined_data_single, met_status == "Primary without met")$Buffa_signature_Sum_within,
            subset(combined_data_single, met_status == "Metastasis")$Buffa_signature_Sum_within,
            conf.int = TRUE)
```

```{r}
wilcox.test(subset(combined_data_single, met_status == "Metastasis")$Buffa_signature_Sum_within,
            subset(combined_data_single, met_status == "Primary with met")$Buffa_signature_Sum_within,
            conf.int = TRUE)
```

```{r}
dim(combined_data_single)
```

# Figure 2 Mets etc
```{r}
names(PPCG_colourlist$new_metastatic_biology_indicator_colourlist) <- c("no_mets_biol","mets_biol")
PPCG_colourlist$new_metastatic_biology_indicator_colourlist[1] <- "grey25"
```

```{r}
mets_biol_plot <- ggplot(hypoxia_signatures_tum_data_single, aes(x = new_metastatic_biology_indicator, y = Buffa_signature_Sum_within))+
  geom_jitter(width = 0.3, alpha = 0.3, aes(colour = new_metastatic_biology_indicator), height = 0)+
  geom_boxplot(alpha = 0.5, outlier.colour = NA)+ # no outlier plotting because we have geom_jitter to plot the points
  theme_light()+
  xlab("New Metastatic Biology Indicator")+
  ylab("Hypoxia Score")+
  geom_signif(map_signif_level = TRUE, comparisons = list(c("no_mets_biol","mets_biol")),
              test = "wilcox.test", tip_length = 0,
              y_position = c(60))+
  scale_colour_manual(values = PPCG_colourlist$new_metastatic_biology_indicator_colourlist)+
  theme(legend.position = "none")+
  scale_x_discrete(labels=c("mets_biol" = "Metastatic \nBiology", 
                            "no_mets_biol" = "No Metastatic \nBiology"))

met_status_plot <- ggplot(combined_data_single, aes(x = met_status, y = Buffa_signature_Sum_within))+
  geom_jitter(alpha = 0.5, width = 0.3, height = 0)+
  geom_boxplot(outlier.colour = NULL, alpha = 0.5)+
  xlab("Metastasis status")+
  ylab("Hypoxia Score")+
  theme_light()+
  geom_signif(map_signif_level = TRUE, comparisons = list(c("Primary without met","Primary with met"),
                                                          c("Primary without met","Metastasis"),
                                                          c("Primary with met","Metastasis")),
              test = "wilcox.test", tip_length = 0,
              y_position = c(75,100,125))

# paired_prim_met_plot <- ggpaired(subset(combined_data_paired_tum_met, has_tum_and_met == "Yes"), 
#          x = "RNA_Sample_Type", y = "Buffa_signature_Sum_within",
#  line.color = "tum_met_direction", line.size = 0.4,
#  palette = c("royalblue","firebrick"), id = "PPCG_Donor_ID")+
#   theme(legend.position = "none")

```

```{r fig.height=6, fig.width=8}
grid.arrange(mets_biol_plot, met_status_plot,
             nrow = 1, ncol = 4,
             layout_matrix = rbind(c(1,1,2,2)))
```

```{r fig.height=4, fig.width=8}
grid.arrange(mets_biol_plot+labs(tag = "A"), met_status_plot+labs(tag = "B"),
             nrow = 1, ncol = 2,
             layout_matrix = rbind(c(1,2)))
```

```{r}
summary(as.factor(hypoxia_signatures_tum_data_single$new_metastatic_biology_indicator))
```

```{r}
summary(combined_data_single$met_status)
```

# Try plotting data as Clarissa suggests
"Fig. 7: I suggest to combine A and B into one figure with the following panels: benign, no mets biol, mets biol, mets. (all mets biol samples by definition are from patients with a met. It does not matter whether we have analyzed it or not).  "

Make a column for new label - use one sample per patient per tissue type
```{r}
combined_data_single_all <- combined_data[!duplicated(combined_data[,c("RNA_Sample_Type","PPCG_Donor_ID")]),]
# combined_data_single_all$met_category <- combined_data_single_all$RNA_Sample_Type

combined_data_single_all$met_category[!is.na(combined_data_single_all$new_metastatic_biology_indicator)] <- combined_data_single_all$new_metastatic_biology_indicator

combined_data_single_all$met_category[combined_data_single_all$RNA_Sample_Type == "MET"] <- "Metastasis"
combined_data_single_all$met_category[combined_data_single_all$RNA_Sample_Type == "BEN"] <- "Benign"
combined_data_single_all$met_category[combined_data_single_all$RNA_Sample_Type == "NOR"] <- "Normal"
combined_data_single_all$met_category[combined_data_single_all$RNA_Sample_Type == "DistantNormal"] <- "Normal"
combined_data_single_all$met_category[combined_data_single_all$RNA_Sample_Type == "LYMPH_NOR"] <- "Normal"
combined_data_single_all$met_category[combined_data_single_all$met_category == "not applicable"] <- NA
combined_data_single_all$met_category[combined_data_single_all$met_category == "mets_biol"] <- "Metastatic\n Biology"
combined_data_single_all$met_category[combined_data_single_all$met_category == "no_mets_biol"] <- "No Metastatic\n Biology"
```

Exclude NOR and rename "not applicable" to NA
```{r}
combined_data_single_all_sub <- subset(combined_data_single_all, met_category != "Normal")
combined_data_single_all_sub$met_category <- as.factor(combined_data_single_all_sub$met_category)
combined_data_single_all_sub$met_category <- factor(combined_data_single_all_sub$met_category, levels = c("Benign",
                                                                                                          "No Metastatic\n Biology",
                                                                                                          "Metastatic\n Biology",
                                                                                                          "Metastasis"))
```

```{r}
dim(subset(combined_data_single_all_sub, !is.na(Buffa_signature_Sum_within)))
table(subset(combined_data_single_all_sub, !is.na(Buffa_signature_Sum_within))$met_category)
```

```{r}
combined_data_single_all_sub_hypoxia <- subset(combined_data_single_all_sub, !is.na(Buffa_signature_Sum_within))
```

```{r}
wilcox_met_status_stats <- as.data.frame(matrix(nrow = 6, ncol = 9))
colnames(wilcox_met_status_stats) <- c("Status_1","Status_2","n_Status_1","n_Status_2","wilcox_p","wilcox_p_adjusted", "wilcox_w", "wilcox_L95", "wilcox_U95")
wilcox_met_status_stats$Status_1 <- c(rep("Benign",3),rep("No Metastatic\n Biology",2),"Metastatic\n Biology")
wilcox_met_status_stats$Status_2 <- c("No Metastatic\n Biology","Metastatic\n Biology","Metastasis",
                                      "Metastatic\n Biology","Metastasis","Metastasis")

for(i in 1:nrow(wilcox_met_status_stats)){
  
  wilcox_met_status_stats$n_Status_1[i] <- nrow(subset(combined_data_single_all_sub_hypoxia, met_category == wilcox_met_status_stats$Status_1[i]))
  wilcox_met_status_stats$n_Status_2[i] <- nrow(subset(combined_data_single_all_sub_hypoxia, met_category == wilcox_met_status_stats$Status_2[i]))
  
  wilcox_met_status_stats$wilcox_p[i] <- wilcox.test(subset(combined_data_single_all_sub_hypoxia, 
                                                            met_category == wilcox_met_status_stats$Status_1[i])$Buffa_signature_Sum_within,
                                                     subset(combined_data_single_all_sub_hypoxia, 
                                                            met_category == wilcox_met_status_stats$Status_2[i])$Buffa_signature_Sum_within, 
                                                     conf.int = TRUE)[[3]]
  
  wilcox_met_status_stats$wilcox_w[i] <- wilcox.test(subset(combined_data_single_all_sub_hypoxia, 
                                                            met_category == wilcox_met_status_stats$Status_1[i])$Buffa_signature_Sum_within,
                                                     subset(combined_data_single_all_sub_hypoxia, 
                                                            met_category == wilcox_met_status_stats$Status_2[i])$Buffa_signature_Sum_within, 
                                                     conf.int = TRUE)[[1]]
  wilcox_met_status_stats$wilcox_L95[i] <- wilcox.test(subset(combined_data_single_all_sub_hypoxia, 
                                                            met_category == wilcox_met_status_stats$Status_1[i])$Buffa_signature_Sum_within,
                                                     subset(combined_data_single_all_sub_hypoxia, 
                                                            met_category == wilcox_met_status_stats$Status_2[i])$Buffa_signature_Sum_within, 
                                                     conf.int = TRUE)[[8]][1]
  wilcox_met_status_stats$wilcox_U95[i] <- wilcox.test(subset(combined_data_single_all_sub_hypoxia, 
                                                            met_category == wilcox_met_status_stats$Status_1[i])$Buffa_signature_Sum_within,
                                                     subset(combined_data_single_all_sub_hypoxia, 
                                                            met_category == wilcox_met_status_stats$Status_2[i])$Buffa_signature_Sum_within, 
                                                     conf.int = TRUE)[[8]][2]
  
}

wilcox_met_status_stats$wilcox_p_adjusted <- p.adjust(wilcox_met_status_stats$wilcox_p, method = "BH")
```

```{r}
wilcox_met_status_stats
```

```{r}
combined_data_single_all_sub_met <- subset(combined_data_single_all_sub, !is.na(met_category))
```

```{r}
new_met_status_plot <- ggplot(combined_data_single_all_sub_met, aes(x = met_category, y = Buffa_signature_Sum_within))+
  geom_boxplot(outlier.colour = NA)+ # Because we show individual data points
  geom_jitter(alpha = 0.5, aes(colour = met_category), height = 0)+
  theme_light()+
  xlab("Sample and Patient Status")+
  ylab("Hypoxia Score")+
  scale_colour_manual(values = c("Benign" = "skyblue",
                                 "No Metastatic\n Biology" = PPCG_colourlist$new_metastatic_biology_indicator_colourlist$no_mets_biology,
                                 "Metastatic\n Biology" = "#698B22",
                                 "Metastasis" = "forestgreen"))+
  theme(legend.position = "none")+
  geom_signif(map_signif_level = TRUE, test = "wilcox.test", tip_length = 0,
              comparisons = list(c("Benign","No Metastatic\n Biology"),
                                 c("Benign","Metastatic\n Biology"),
                                 c("Benign","Metastasis"),
                                 c("No Metastatic\n Biology","Metastatic\n Biology"),
                                 c("No Metastatic\n Biology","Metastasis"),
                                 c("Metastatic\n Biology","Metastasis")), 
              y_position = c(28,43,48,
                             33,38,28))+
  coord_cartesian(ylim = c(-35,55))
new_met_status_plot
```

```{r}
fig_4a_data <- combined_data_single_all_sub_met[,c("met_category","Buffa_signature_Sum_within")]
```

```{r}
ggplot(fig_4a_data, aes(x = met_category, y = Buffa_signature_Sum_within))+
  geom_boxplot(outlier.colour = NA)+ # Because we show individual data points
  geom_jitter(alpha = 0.5, aes(colour = met_category), height = 0)+
  theme_light()+
  xlab("Sample and Patient Status")+
  ylab("Hypoxia Score")+
  scale_colour_manual(values = c("Benign" = "skyblue",
                                 "No Metastatic\n Biology" = PPCG_colourlist$new_metastatic_biology_indicator_colourlist$no_mets_biology,
                                 "Metastatic\n Biology" = "#698B22",
                                 "Metastasis" = "forestgreen"))+
  theme(legend.position = "none")+
  geom_signif(map_signif_level = TRUE, test = "wilcox.test", tip_length = 0,
              comparisons = list(c("Benign","No Metastatic\n Biology"),
                                 c("Benign","Metastatic\n Biology"),
                                 c("Benign","Metastasis"),
                                 c("No Metastatic\n Biology","Metastatic\n Biology"),
                                 c("No Metastatic\n Biology","Metastasis"),
                                 c("Metastatic\n Biology","Metastasis")), 
              y_position = c(28,43,48,
                             33,38,28))+
  coord_cartesian(ylim = c(-35,55))
```

```{r}
write.csv(fig_4a_data, paste0(fig_data_outdir,"Figure_4a_data.csv"), row.names = FALSE)
```


# Combine figures for main and supplementary
## Main figure

```{r fig.width=8, fig.height=10}
grid.arrange(new_met_status_plot+labs(tag = "A"),
             cor_pm_plot+labs(tag = "B"),
             nrow = 2)
```

```{r}
pdf(file = paste0(fig_outdir,"/","Figure_4_",Sys.Date(),".pdf"), width = 8.5, height = 10)
grid.arrange(new_met_status_plot+labs(tag = "A"),
             cor_pm_plot+labs(tag = "B"),
             nrow = 2)
dev.off()
```

## Supp figure
```{r fig.height=8, fig.width=6}
grid.arrange(paired_plot+labs(tag = "A"), paired_plot_facet +labs(tag = "B"),
             nrow = 2, ncol = 1)
```

```{r}
png(filename = paste0(fig_outdir_supp,"/","supplementary_figure_7_paired_scores_",Sys.Date(),".png"), width = 6, height = 8, 
    units = "in", res = 300)
grid.arrange(paired_plot+labs(tag = "A"), paired_plot_facet +labs(tag = "B"),
             nrow = 2, ncol = 1)
dev.off()
```

# Session info
```{r}
sessionInfo()
```

