---
title: "Manuscript December 2024, Figure 3 - germline and patient level characteristics. RNA data release 1.4k10"
author: Maria Jakobsdottir (Maria.Jakobsdottir@manchester.ac.uk)
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2:
    toc: true 
    toc_float: true
    toc_depth: 6
---

Analysis for figure 3 in Hypoxia manuscript revisions.

# Libraries used
```{r message=FALSE, warning=FALSE}
library(circlize)
library(ComplexHeatmap)
library(ggplot2)
library(ggsci)
library(ggsignif)
library(ggthemes)
library(gridExtra)
library(rawr)
library(RColorBrewer)
library(scales)
library(xtable)

options(bitmapType='cairo')
'%!in%' <- function(x,y)!('%in%'(x,y))
```

Heatmap parameters
```{r}
ht_opt$heatmap_row_names_gp = gpar(fontsize = 12)
```

# Data
Read in data

```{r}
params <- list(ppcg_colourlist = "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/GitHub/PPCG_colours/PPCG_colourlist.RData",
               combined_data = "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_data_private/combined_PPCG_WGS_RNA_summary_data_clinical_v7_2024-12-06.txt",
               germline_tier_1_2_sample_list = "/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_Germline/20240716_PPCG_Germline_Data_Globus/20240716_PPCG_Germline_Data_Tier1_plus_2_samples_vs_gene_lists.txt",
               germline_tier_1_2_sample_gene = "/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_Germline/20240716_PPCG_Germline_Data_Globus/20240716_PPCG_Germline_Data_Tier1_plus_2_genes_vs_samples.txt",
               germline_tier_1_sample_list = "/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_Germline/20240716_PPCG_Germline_Data_Globus/20240716_PPCG_Germline_Data_Tier1_samples_vs_gene_lists.txt",
               germline_tier_1_sample_gene = "/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_Germline/20240716_PPCG_Germline_Data_Globus/20240716_PPCG_Germline_Data_Tier1_genes_vs_samples.txt",
               timing_trajectories = "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_data/Chris_trajectories/evolutionary_orderings_210824/PPCG_selected_primary_samples_sample_groups_G3.Rdata",
               drivers = "/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_Drivers/Somatic_drivers/DRIVER_ANNOTATIONS__2023-09-25_release4/DRIVER_ANNOTATIONS__2023-09-25_release4__SNV+indel__filteredVariants__allGenes.txt",
               hallmark_hypoxia_quartile_coefs = "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_paper_revisions/outputs/driver_models/hallmark_hypoxia_quartile_hypoxia_binomial_glm_summary_revisions.txt",
               hallmark_hypoxia_continuous_coefs = "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_paper_revisions/outputs/driver_models/hallmark_hypoxia_continuous_hypoxia_binomial_glm_summary_revisions.txt",
               drg36_quartile_coefs = "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_paper_revisions/outputs/driver_models/drg128s_quartile_hypoxia_binomial_glm_summary_revisions.txt",
               drg36_continuous_coefs = "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_paper_revisions/outputs/driver_models/drg36s_continuous_hypoxia_binomial_glm_summary_revisions.txt",
               drg128_quartile_coefs = "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_paper_revisions/outputs/driver_models/drg128s_quartile_hypoxia_binomial_glm_summary_revisions.txt",
               drg128_continuous_coefs = "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_paper_revisions/outputs/driver_models/drg128s_continuous_hypoxia_binomial_glm_summary_revisions.txt",
               gene_set_bionm_coefs = "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_paper_revisions/outputs/driver_models/gene_set_hypoxia_binomial_glm_summary_revisions.txt"
               )
```

```{r}
load(params$ppcg_colourlist)
```

Clone colours
```{r}
clone_colours <- brewer.pal(6, "BuPu")[2:5]
```


Parameters, here for simplicity, otherwise kept in submission script. 
```{r}
print(params)
```

## Combined data

```{r}
combined_data <- read.delim(params$combined_data, header = TRUE, sep = "\t")
```

## Driver data
Data release version: 4

Date: 2023-09-25

Source: Juri Reimand - FTP box

/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_Drivers/Somatic_drivers/DRIVER_ANNOTATIONS__2023-09-25_release4/DRIVER_ANNOTATIONS__2023-09-25_release4__SNV+indel__filteredVariants__allGenes.txt
```{r}
driver_data <- read.delim(params$drivers, row.names = 1, sep = "\t", header = TRUE)
```

## Germline data
Data release version: ?

Date: 2024-07-16

Source: Daniel Burns - FTP box

### germline_tier_1_sample_list
```{r}
germline_tier_1_sample_list <- read.delim(params$germline_tier_1_sample_list)
```
```{r}
summary(germline_tier_1_sample_list)
```

### germline_tier_1_sample_gene
```{r}
germline_tier_1_sample_gene <- read.delim(params$germline_tier_1_sample_gene)
```
```{r}
summary(germline_tier_1_sample_gene)
```

### germline_tier_1_2_sample_list
```{r}
germline_tier_1_2_sample_list <- read.delim(params$germline_tier_1_2_sample_list)
```
```{r}
summary(germline_tier_1_2_sample_list)
```

### germline_tier_1_2_sample_gene
```{r}
germline_tier_1_2_sample_gene <- read.delim(params$germline_tier_1_2_sample_gene)
```
```{r}
summary(germline_tier_1_2_sample_gene)
```

## Other data
### Timing trajectories
Source: Chris Wirth - Email 21.08.2024

/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_data/Chris_trajectories/
Object is named `sample_groups`
```{r}
load(params$timing_trajectories)
```

Turn this into a data frame
```{r}
trajectories <- as.data.frame(matrix(data = c(sample_groups[[1]], sample_groups[[2]], sample_groups[[3]],
                              rep("Trajectory 1", length(sample_groups[[1]])),
                              rep("Trajectory 2", length(sample_groups[[2]])),
                              rep("Trajectory 3", length(sample_groups[[3]]))),
                              ncol = 2, byrow = FALSE))
colnames(trajectories) <- c("WGS_AssayID_full","Trajectory")
trajectories$WGS_AssayID <- gsub("_vs_.*","",trajectories$WGS_AssayID_full)
```



### Gene lists

```{r}
drg_genes_36 <- c("ATM","BAP1","BLM","BRCA1","BRCA2",
               "BRIP1","CDH1","CDK4","CDKN2A","CHEK2",
               "FANCA","FANCD2","FANCI","FANCL","FANCM",
               "MLH1","MSH2","MSH6","NBN","NTHL1",
               "PALB2","PMS2","POLD1","POLE","POT1",
               "PTCH1","PTEN","RAD51C","RAD51D","RB1",
               "SMAD4","SMARCA4","STK11","TP53","XPA",
               "XPC")
```

```{r}
drg_genes_128 <- c("AKT1","ALKBH3","APEX1","APLF","ATM",
                   "ATR","ATRIP","BAP1","BARD1","BLM",
                   "BRCA1","BRCA2","BRIP1","CCNH","CDC25A",
                   "CDC25C","CDH1","CDK1","CDK2","CDK4",
                   "CDKN2A","CHEK1","CHEK2","CLK2","DCLRE1A",
                   "DDB1","EME1","EME2","EPCAM","ERCC2",
                   "ERCC3","ERCC5","ERCC6","ESR2","EXO1",
                   "FAM175A","FANCA","FANCD2","FANCI","FANCL",
                   "FANCM","GADD45A","GEN1","GTF2H2","GTF2H3",
                   "GTF2H4","H2AFX","HUS1","LIG1","LIG3",
                   "LIG4","MLH1","MLH3","MMS19","MNAT1",
                   "MPG","MRE11","MRE11A","MSH2","MSH3",
                   "MSH5","MSH6","MSR1","NABP2","NBN",
                   "NEIL1","NEIL2","NHEJ1","NTHL1","OGG1",
                   "PALB2","PARP2","PER1","PER2","PMS1",
                   "PMS2","PNKP","POLD1","POLE","POLH",
                   "POLK","POLL","POLM","POLN","POLQ",
                   "POT1","PRSS1","PTCH1","PTEN","RAD1",
                   "RAD50","RAD51","RAD51B","RAD51C","RAD51D",
                   "RAD52","RAD54B","RAD54L","RB1","RDM1",
                   "RECQL","RECQL4","RECQL5","RINT1","RNF8",
                   "RPA1","RPA3","SETMAR","SLX4","SMAD4",
                   "SMARCA4","SMUG1","STK11","TDG","TDP1",
                   "TOP2A","TOP2B","TOP3A","TP53","TP53BP1",
                   "WRN","XAB2","XPA","XPC","XRCC1",
                   "XRCC2","XRCC4","XRCC5")
```

```{r}
hallmark_hypoxia <- c("ADM",      "ADORA2B",  "AK4",      "AKAP12",   "ALDOA",    "ALDOB",   
                                     "ALDOC",    "AMPD3",    "ANGPTL4",  "ANKZF1",   "ANXA2",    "ATF3",
                                     "ATP7A",    "B3GALT6",  "B4GALNT2", "BCAN",     "BCL2",     "BGN",
                                     "BHLHE40",  "BNIP3L",   "BRS3",     "BTG1",     "CA12",     "CASP6",
                                     "CAV1",     "CCNG2",    "CCRN4L",   "CDKN1A",   "CDKN1B",   "CDKN1C",
                                     "CHST2",    "CHST3",    "CITED2",   "COL5A1",   "CP",       "CSRP2",
                                     "CTGF",     "CXCR4",    "CXCR7",    "CYR61",    "DCN",      "DDIT3",
                                     "DDIT4",    "DPYSL4",   "DTNA",     "DUSP1",    "EDN2",     "EFNA1",
                                     "EFNA3",    "EGFR",     "ENO1",     "ENO2",     "ENO3",     "ERO1L",
                                     "ERRFI1",   "ETS1",     "EXT1",     "F3",       "FAM162A",  "FBP1",
                                     "FOS",      "FOSL2",    "FOXO3",    "GAA",      "GALK1",    "GAPDH",
                                     "GAPDHS",   "GBE1",     "GCK",      "GCNT2",    "GLRX",     "GPC1",
                                     "GPC3",     "GPC4",     "GPI",      "GRHPR",    "GYS1",     "HAS1",
                                     "HDLBP",    "HEXA",     "HK1",      "HK2",      "HMOX1",    "HOXB9",
                                     "HS3ST1",   "HSPA5",    "IDS",      "IER3",     "IGFBP1",   "IGFBP3",
                                     "IL6",      "ILVBL",    "INHA",     "IRS2",     "ISG20",    "JMJD6",
                                     "JUN",      "KDELR3",   "KDM3A",    "KIF5A",    "KLF6",     "KLF7",
                                    "KLHL24",   "LALBA",    "LARGE",    "LDHA",     "LDHC",     "LOX",
                                    "LXN",      "MAFF",     "MAP3K1",   "MIF",      "MT1E",     "MT2A",
                                    "MXI1",     "MYH9",     "NAGK",     "NCAN",     "NDRG1",    "NDST1",
                                    "NDST2",    "NEDD4L",   "NFIL3",    "NR3C1",    "P4HA1",    "P4HA2",
                                    "PAM",      "PCK1",     "PDGFB",    "PDK1",     "PDK3",     "PFKFB3",
                                    "PFKL",     "PFKP",     "PGAM2",    "PGF",      "PGK1",     "PGM1",
                                    "PGM2",     "PHKG1",    "PIM1",     "PKLR",     "PKP1",     "PLAC8",
                                    "PLAUR",    "PLIN2",    "PNRC1",    "PPARGC1A", "PPFIA4",   "PPP1R15A",
                                    "PPP1R3C",  "PRDX5",    "PRKCA",    "PRKCDBP",  "PTRF",     "PYGM",
                                    "RBPJ",     "RORA",     "RRAGD",    "S100A4",   "SAP30",    "SCARB1",
                                    "SDC2",     "SDC3",     "SDC4",     "SELENBP1", "SERPINE1", "SIAH2",
                                    "SLC25A1",  "SLC2A1",   "SLC2A3",   "SLC2A5",   "SLC37A4",  "SLC6A6",
                                    "SRPX",     "STBD1",    "STC1",     "STC2",     "SULT2B1",  "TES",
                                    "TGFB3",    "TGFBI",    "TGM2",     "TIPARP",   "TKTL1",    "TMEM45A",
                                    "TNFAIP3",  "TPBG",     "TPD52",    "TPI1",     "TPST2",    "UGP2",
                                    "VEGFA",    "VHL",      "VLDLR",    "WISP2",    "WSB1",     "XPNPEP1",
                                    "ZFP36",    "ZNF292")
```

### Binomial model coefficients
Read in coefficients from binomial models that account for total mutation burden when looking at relationship between hypoxia and driver mutations

glm(as.numeric(gene) ~ Buffa_signature_Sum_within + log(num.SNVs), family = binomial, data = sub_data)

Hallmark hypoxia
```{r}
hallmark_hypoxia_continuous_coefficients <- read.delim(params$hallmark_hypoxia_continuous_coefs, header = TRUE)
rownames(hallmark_hypoxia_continuous_coefficients) <- hallmark_hypoxia_continuous_coefficients$Gene
hallmark_hypoxia_continuous_coefficients$Significance <- ifelse(hallmark_hypoxia_continuous_coefficients$Hypoxia_p_adj <= 0.05, yes = "Significant", no = "N.S.")
```

```{r}
hallmark_hypoxia_quartile_coefficients <- read.delim(params$hallmark_hypoxia_quartile_coefs, header = TRUE)
rownames(hallmark_hypoxia_quartile_coefficients) <- hallmark_hypoxia_quartile_coefficients$Gene
hallmark_hypoxia_quartile_coefficients$Significance <- ifelse(hallmark_hypoxia_quartile_coefficients$Hypoxia_p_adj <= 0.05, yes = "Significant", no = "N.S.")
```

DRG36
```{r}
drg36_continuous_coefficients <- read.delim(params$drg36_continuous_coefs, header = TRUE)
rownames(drg36_continuous_coefficients) <- drg36_continuous_coefficients$Gene
drg36_continuous_coefficients$Significance <- ifelse(drg36_continuous_coefficients$Hypoxia_p_adj <= 0.05, yes = "Significant", no = "N.S.")
```

```{r}
drg36_quartile_coefficients <- read.delim(params$drg36_quartile_coefs, header = TRUE)
rownames(drg36_quartile_coefficients) <- drg36_quartile_coefficients$Gene
drg36_quartile_coefficients$Significance <- ifelse(drg36_quartile_coefficients$Hypoxia_p_adj <= 0.05, yes = "Significant", no = "N.S.")
```

DRG128
```{r}
drg128_continuous_coefficients <- read.delim(params$drg128_continuous_coefs, header = TRUE)
rownames(drg128_continuous_coefficients) <- drg128_continuous_coefficients$Gene
drg128_continuous_coefficients$Significance <- ifelse(drg128_continuous_coefficients$Hypoxia_p_adj <= 0.05, yes = "Significant", no = "N.S.")
```

```{r}
drg128_quartile_coefficients <- read.delim(params$drg128_quartile_coefs, header = TRUE)
rownames(drg128_quartile_coefficients) <- drg128_quartile_coefficients$Gene
drg128_quartile_coefficients$Significance <- ifelse(drg128_quartile_coefficients$Hypoxia_p_adj <= 0.05, yes = "Significant", no = "N.S.")
```

Gene Sets
```{r}
gene_sets_binom_coefficients <- read.delim(params$gene_set_bionm_coefs, header = TRUE)
rownames(gene_sets_binom_coefficients) <- gene_sets_binom_coefficients$mutation_set
gene_sets_binom_coefficients$Score_Significance <- ifelse(gene_sets_binom_coefficients$Score_p <= 0.05, yes = "Significant", no = "N.S.")
gene_sets_binom_coefficients$Quart_Significance <- ifelse(gene_sets_binom_coefficients$Quart_p <= 0.05, yes = "Significant", no = "N.S.")

```

```{r}
gene_sets_binom_coefficients
```


# Re-arrange factors where necessary
```{r}
combined_data$age_bracket <- as.factor(combined_data$age_bracket)
combined_data$age_bracket <- factor(combined_data$age_bracket, levels = c("< 40","40-50","50-65","65-70","≥ 70"))

combined_data$psa_bracket <- as.factor(combined_data$psa_bracket)
combined_data$psa_bracket <- factor(combined_data$psa_bracket, levels = c("0-9.9","10-19.9","≥20"))

combined_data$gleason_total <- as.factor(combined_data$gleason_total)
combined_data$gleason_total <- factor(combined_data$gleason_total, levels = c("3+3","3+4","4+3","4+4","3+5","4+5","5+4","5+5"))
```

# Merge data and select samples
Use Tier 1 Buffa signature sum.

## Selecting tumour only and one sample per patient
```{r}
hypoxia_signatures_tum_data_single <- subset(combined_data, !is.na(Buffa_signature_Sum_within_quartile_one_sample_per_donor))
```
```{r}
dim(hypoxia_signatures_tum_data_single)
```

```{r}
summary(hypoxia_signatures_tum_data_single)
```



Here is a list of the IDs that were used for the single tumour sample per patient analysis:
```{r}
hypoxia_signatures_tum_data_single$PPCG_SampleID
```
 
# Trajectories

```{r}
hypoxia_signatures_tum_data_single_trajectory <- merge(hypoxia_signatures_tum_data_single,
                                                trajectories,
                                                by.x = "WGS_AssayID",
                                                by.y = "WGS_AssayID",
                                                all.x = TRUE)
```

Subset combined trajectory data so you only keep samples with available data:
```{r}
hypoxia_signatures_tum_data_single_trajectory_sub <- subset(hypoxia_signatures_tum_data_single_trajectory, 
                                                     !is.na(Trajectory))
```
```{r}
dim(hypoxia_signatures_tum_data_single_trajectory_sub)
```

## Bar plot
https://stackoverflow.com/questions/63653351/how-to-use-stat-count-to-label-a-bar-chart-with-counts-or-percentages-in-ggplo
```{r}
traj_hilo_bar <- ggplot(hypoxia_signatures_tum_data_single_trajectory_sub, aes(x = Buffa_signature_Sum_within_bin_one_sample_per_donor, fill = as.factor(Trajectory)))+
  geom_bar(stat = "count", position = "fill")+
  geom_text(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], label=percent(..count../tapply(..count.., ..x.. ,sum)[..x..]) ),
              stat="count", position=position_fill(0.5), vjust=0.5, colour = "white")+
  theme_light()+
  ylab("")+
  xlab("Hypoxia Signature")+
  scale_fill_manual(name = "Trajectory", values = c("grey","#1C1D3D","#E22B56"))+
  theme(text = element_text(size = 16),
        legend.position = "bottom")
traj_hilo_bar
```
```{r}
# fisher.test(table(hypoxia_signatures_tum_data_single_evo_sub$Buffa_signature_Sum_within_bin_one_sample_per_donor,
#                   hypoxia_signatures_tum_data_single_evo_sub$Trajectory))
```
```{r}
table(hypoxia_signatures_tum_data_single_trajectory_sub$Buffa_signature_Sum_within_bin_one_sample_per_donor,
                  hypoxia_signatures_tum_data_single_trajectory_sub$Trajectory)
```

### Quartiles
```{r}
traj_quart_bar <- ggplot(hypoxia_signatures_tum_data_single_trajectory_sub, aes(x = Buffa_signature_Sum_within_quartile_one_sample_per_donor, fill = as.factor(Trajectory)))+
  geom_bar(stat = "count", position = "fill")+
  geom_text(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], label=percent(..count../tapply(..count.., ..x.. ,sum)[..x..], accuracy = 0.1) ),
              stat="count", position=position_fill(0.5), vjust=0.5, colour = "white")+
  theme_light()+
  ylab("")+
  xlab("Hypoxia Signature")+
  # ggtitle("Trajectory")+
  scale_fill_manual(name = "Trajectory", values = c("grey","#1C1D3D","#E22B56"))+
  theme(text = element_text(size = 16),
        legend.position = "bottom")
traj_quart_bar
```

```{r}
hypoxia_signatures_tum_data_single_trajectory_sub_q1q4 <- subset(hypoxia_signatures_tum_data_single_trajectory_sub, Buffa_signature_Sum_within_quartile_one_sample_per_donor %in% c("Q1","Q4"))
```

```{r}
fisher.test(table(hypoxia_signatures_tum_data_single_trajectory_sub_q1q4$Buffa_signature_Sum_within_quartile_one_sample_per_donor,
                  hypoxia_signatures_tum_data_single_trajectory_sub_q1q4$Trajectory))
```

```{r}
table(hypoxia_signatures_tum_data_single_trajectory_sub$Buffa_signature_Sum_within_quartile_one_sample_per_donor,
                  hypoxia_signatures_tum_data_single_trajectory_sub$Trajectory)
```


## Box plot
```{r}
traj_score_diff <- ggplot(hypoxia_signatures_tum_data_single_trajectory_sub, aes(x = as.factor(Trajectory), y = Buffa_signature_Sum_within))+
  geom_jitter(width = 0.3, alpha = 0.5, aes(colour = Trajectory), height = 0)+
  geom_boxplot(alpha = 0.5, outlier.colour = NA)+ # no outlier plotting because we have geom_jitter to plot the points
  theme_light()+
  xlab("Trajectory")+
  ylab("Hypoxia Score")+
  # ggtitle("Trajectory")+
  theme(text = element_text(size = 16),
        legend.position = "bottom")+
  geom_signif(map_signif_level = TRUE, comparisons = list(c("Trajectory 1","Trajectory 2"),
                                                          c("Trajectory 1","Trajectory 3"),
                                                          c("Trajectory 2","Trajectory 3")),
              test = "wilcox.test", tip_length = 0, y_position = c(30,40,35))+
  scale_colour_manual(name = "Trajectory", values = c("grey","#1C1D3D","#E22B56"))

traj_score_diff
```
```{r}
wilcox.test(subset(hypoxia_signatures_tum_data_single_trajectory_sub, Trajectory == "Trajectory 1")[,"Buffa_signature_Sum_within"],
            subset(hypoxia_signatures_tum_data_single_trajectory_sub, Trajectory == "Trajectory 2")[,"Buffa_signature_Sum_within"])

wilcox.test(subset(hypoxia_signatures_tum_data_single_trajectory_sub, Trajectory == "Trajectory 1")[,"Buffa_signature_Sum_within"],
            subset(hypoxia_signatures_tum_data_single_trajectory_sub, Trajectory == "Trajectory 3")[,"Buffa_signature_Sum_within"])

wilcox.test(subset(hypoxia_signatures_tum_data_single_trajectory_sub, Trajectory == "Trajectory 2")[,"Buffa_signature_Sum_within"],
            subset(hypoxia_signatures_tum_data_single_trajectory_sub, Trajectory == "Trajectory 3")[,"Buffa_signature_Sum_within"])
```

# Subclones
Combine multiple sublcones into one category. 
```{r}
hypoxia_signatures_tum_data_single$Clone_count_summary <- NA
hypoxia_signatures_tum_data_single$Clone_count_summary[hypoxia_signatures_tum_data_single$Clone_count == 1] <- "1"
hypoxia_signatures_tum_data_single$Clone_count_summary[hypoxia_signatures_tum_data_single$Clone_count > 1] <- ">1"

hypoxia_signatures_tum_data_single$Clone_count_summary <- as.factor(hypoxia_signatures_tum_data_single$Clone_count_summary)
hypoxia_signatures_tum_data_single$Clone_count_summary <- factor(hypoxia_signatures_tum_data_single$Clone_count_summary, levels = c("1",">1"))

```


```{r}
ggplot(subset(hypoxia_signatures_tum_data_single, !is.na(Clone_count)), aes(y = as.factor(Clone_count), x = Buffa_signature_Sum_within_quartile_one_sample_per_donor))+
  geom_boxplot()+ 
  geom_jitter(alpha = 0.5, height = 0, aes(colour = as.factor(Clone_count)))+
  theme_light()+
  ylab("Clone count")+
  xlab("Hypoxia Score")+
  scale_colour_manual(values = clone_colours, name = "Number of Subclones")
```

```{r}
clone_boxplot <- ggplot(subset(hypoxia_signatures_tum_data_single, !is.na(Clone_count)), aes(x = as.factor(Clone_count), y = Buffa_signature_Sum_within))+
  geom_boxplot()+ 
  geom_jitter(alpha = 0.5, height = 0, aes(colour = as.factor(Clone_count)))+
  theme_light()+
  xlab("Clone count")+
  ylab("Hypoxia Score")+
  scale_colour_manual(values = clone_colours, name = "Number of Subclones")+
  theme(text = element_text(size = 16),
        legend.position = "bottom")+
  geom_signif(map_signif_level = TRUE, comparisons = list(c("1","2"),
                                                          c("1","3"),
                                                          c("2","3")),
              test = "wilcox.test", tip_length = 0, y_position = c(30,40,35))
clone_boxplot
```

```{r}
kruskal.test(hypoxia_signatures_tum_data_single$Buffa_signature_Sum_within, hypoxia_signatures_tum_data_single$Clone_count)
```

```{r}
cuzick.test(x = hypoxia_signatures_tum_data_single$Buffa_signature_Sum_within,
            g = hypoxia_signatures_tum_data_single$Clone_count)
```

```{r}
cuzick.test(x = subset(hypoxia_signatures_tum_data_single, Clone_count < 4)$Buffa_signature_Sum_within,
            g = subset(hypoxia_signatures_tum_data_single, Clone_count < 4)$Clone_count)
```

```{r}
dim(subset(hypoxia_signatures_tum_data_single, !is.na(Clone_count_summary)))
```

```{r}
clone_boxplot_sum <- ggplot(subset(hypoxia_signatures_tum_data_single, !is.na(Clone_count_summary)), aes(x = as.factor(Clone_count_summary), y = Buffa_signature_Sum_within))+
  geom_boxplot()+ 
  geom_jitter(alpha = 0.5, height = 0, aes(colour = as.factor(Clone_count_summary)))+
  theme_light()+
  xlab("Clone count")+
  ylab("Hypoxia Score")+
  scale_colour_manual(values = clone_colours[c(1,3)], name = "Number of Subclones")+
  theme(text = element_text(size = 16),
        legend.position = "bottom")+
  geom_signif(map_signif_level = TRUE, comparisons = list(c("1",">1")),
              test = "wilcox.test", tip_length = 0, y_position = c(30))
clone_boxplot_sum
```

```{r}
wilcox.test(subset(hypoxia_signatures_tum_data_single, Clone_count_summary == "1")[,"Buffa_signature_Sum_within"],
            subset(hypoxia_signatures_tum_data_single, Clone_count_summary == ">1")[,"Buffa_signature_Sum_within"])
```


```{r}
clone_barplot <- ggplot(subset(hypoxia_signatures_tum_data_single, !is.na(Clone_count)), aes(fill = as.factor(Clone_count), x = Buffa_signature_Sum_within_quartile_one_sample_per_donor))+
  geom_bar(stat = "count", position = "fill")+ 
  geom_text(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], label=percent(..count../tapply(..count.., ..x.. ,sum)[..x..]) ),
              stat="count", position=position_fill(0.5), vjust=0.5, colour = "white")+
  theme_light()+
  xlab("Hypoxia Score Quartile")+
  ylab("")+
  scale_fill_manual(values = clone_colours, name = "Number of Subclones")+
  theme(text = element_text(size = 16),
        legend.position = "bottom")
clone_barplot
```

```{r}
clone_sum_barplot <- ggplot(subset(hypoxia_signatures_tum_data_single, !is.na(Clone_count_summary)), aes(fill = as.factor(Clone_count_summary), x = Buffa_signature_Sum_within_quartile_one_sample_per_donor))+
  geom_bar(stat = "count", position = "fill")+ 
  geom_text(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], label=percent(..count../tapply(..count.., ..x.. ,sum)[..x..]) ),
              stat="count", position=position_fill(0.5), vjust=0.5, colour = "white")+
  theme_light()+
  xlab("Hypoxia Score Quartile")+
  ylab("")+
  scale_fill_manual(values = clone_colours[c(1,3)], name = "Number of Subclones")+
  theme(text = element_text(size = 16),
        legend.position = "bottom")
clone_sum_barplot
```

```{r}
fisher.test(table(subset(hypoxia_signatures_tum_data_single, Buffa_signature_Sum_within_quartile_one_sample_per_donor %in% c("Q1","Q4"))$Clone_count, 
          subset(hypoxia_signatures_tum_data_single, Buffa_signature_Sum_within_quartile_one_sample_per_donor %in% c("Q1","Q4"))$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
```

```{r}
fisher.test(table(subset(hypoxia_signatures_tum_data_single, Buffa_signature_Sum_within_quartile_one_sample_per_donor %in% c("Q1","Q4"))$Clone_count_summary, 
          subset(hypoxia_signatures_tum_data_single, Buffa_signature_Sum_within_quartile_one_sample_per_donor %in% c("Q1","Q4"))$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
```

```{r}
table(hypoxia_signatures_tum_data_single$Clone_count, hypoxia_signatures_tum_data_single$Buffa_signature_Sum_within_quartile_one_sample_per_donor)
```

```{r}
table(hypoxia_signatures_tum_data_single$Clone_count_summary, hypoxia_signatures_tum_data_single$Buffa_signature_Sum_within_quartile_one_sample_per_donor)
```

```{r}
# fisher.test(table(hypoxia_signatures_tum_data_single$Clone_count, 
#           hypoxia_signatures_tum_data_single$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
# 
# fisher.test(table(subset(hypoxia_signatures_tum_data_single, Buffa_signature_Sum_within_quartile_one_sample_per_donor %in% c("Q1","Q4"))$Clone_count, 
#           subset(hypoxia_signatures_tum_data_single, Buffa_signature_Sum_within_quartile_one_sample_per_donor %in% c("Q1","Q4"))$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
```

# Trajectory and Clone plot
```{r fig.height=11, fig.width=11}
grid.arrange(traj_score_diff+labs(tag = "A"), clone_boxplot+labs(tag = "B"),
             traj_quart_bar+labs(tag = "C"), clone_barplot+labs(tag = "D"),
             nrow = 2, ncol = 3,
             layout_matrix = rbind(c(1,2),
                                   c(3,4)))
```



# Germline variants
## Tier 1 variants
Sample vs gene list
```{r}
hypoxia_tier_1_lists <- merge(hypoxia_signatures_tum_data_single,
                               germline_tier_1_sample_list,
                               by.x = "PPCG_Donor_ID",
                               by.y = "Sample")
dim(hypoxia_tier_1_lists)
```

Individual genes vs samples
```{r}
germline_tier_1_sample_gene_sub <- subset(germline_tier_1_sample_gene, Genes %in% unique((c(drg_genes_36, drg_genes_128, hallmark_hypoxia))))
rownames(germline_tier_1_sample_gene_sub) <- germline_tier_1_sample_gene_sub$Genes
germline_tier_1_sample_gene_sub <- as.data.frame(t(as.matrix(germline_tier_1_sample_gene_sub[,3:ncol(germline_tier_1_sample_gene_sub)])))
germline_tier_1_sample_gene_sub$Sample <- rownames(germline_tier_1_sample_gene_sub)
```

```{r}
hypoxia_tier_1_genes <- merge(hypoxia_signatures_tum_data_single,
                               germline_tier_1_sample_gene_sub,
                               by.x = "PPCG_Donor_ID",
                               by.y = "Sample")
dim(hypoxia_tier_1_genes)
```

### Hallmark hypoxia
#### Any gene in list
##### Quartiles
```{r}
hypoxia_tier_1_lists_quart <- subset(hypoxia_tier_1_lists, Buffa_signature_Sum_within_quartile_one_sample_per_donor %in% c("Q1","Q4"))
```

```{r}
fisher.test(table(hypoxia_tier_1_lists_quart$HALLMARK_HYPOXIA, hypoxia_tier_1_lists_quart$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
```
```{r}
table(hypoxia_tier_1_lists_quart$HALLMARK_HYPOXIA, hypoxia_tier_1_lists_quart$Buffa_signature_Sum_within_quartile_one_sample_per_donor)
```


##### Scores
```{r}
wilcox.test(subset(hypoxia_tier_1_lists, HALLMARK_HYPOXIA == "1")[,"Buffa_signature_Sum_within"], 
            subset(hypoxia_tier_1_lists, HALLMARK_HYPOXIA == "0")[,"Buffa_signature_Sum_within"])
```


#### Individual genes
##### Quartiles
```{r}
hypoxia_tier_1_genes_quart <- subset(hypoxia_tier_1_genes, Buffa_signature_Sum_within_quartile_one_sample_per_donor %in% c("Q1","Q4"))
```

Make a table for Fisher exact test results
```{r}
germline_hypoxia_enrichment_t1 <- as.data.frame(matrix(nrow = length(hallmark_hypoxia), ncol = 8))
colnames(germline_hypoxia_enrichment_t1) <- c("Gene","n_q4_mutated","n_q1_mutated","n_q4_wt","n_q1_wt","fisher_p_val","fisher_p_adj","odds_ratio")
germline_hypoxia_enrichment_t1$Gene <- hallmark_hypoxia

for(i in 1:nrow(germline_hypoxia_enrichment_t1)){
  gene <- germline_hypoxia_enrichment_t1$Gene[i]

  
  test_data <- hypoxia_tier_1_genes_quart[,c("Buffa_signature_Sum_within_quartile_one_sample_per_donor",gene)]
  
  table_data <- as.data.frame(table(test_data[,gene], test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
  
  if(nrow(table_data) > 2){
    germline_hypoxia_enrichment_t1$fisher_p_val[i] <- fisher.test(table(test_data[,gene], test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))[[1]]
    germline_hypoxia_enrichment_t1$odds_ratio[i] <- fisher.test(table(test_data[,gene], test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))[[3]]
  
  
  
    germline_hypoxia_enrichment_t1$n_q4_mutated[i] <- subset(table_data, (Var1 == "1") & (Var2 == "Q4"))[,"Freq"]
    germline_hypoxia_enrichment_t1$n_q1_mutated[i] <- subset(table_data, (Var1 == "1") & (Var2 == "Q1"))[,"Freq"]
    germline_hypoxia_enrichment_t1$n_q4_wt[i] <- subset(table_data, (Var1 == "0") & (Var2 == "Q4"))[,"Freq"]
    germline_hypoxia_enrichment_t1$n_q1_wt[i] <- subset(table_data, (Var1 == "0") & (Var2 == "Q1"))[,"Freq"]
  }
  
}
# germline_hypoxia_enrichment_t1 <- subset(germline_hypoxia_enrichment_t1, !is.na(fisher_p_val))
germline_hypoxia_enrichment_t1$fisher_p_adj <- p.adjust(germline_hypoxia_enrichment_t1$fisher_p_val, method = "BH")
```

```{r}
germline_hypoxia_enrichment_t1
```

###### Binomial model
```{r}
hallmark_hypoxia_quartile_coefficients
```


##### Scores
Make a table for Wilcox test results
```{r}
germline_hypoxia_wilcox_t1 <- as.data.frame(matrix(nrow = length(hallmark_hypoxia), ncol = 5))
colnames(germline_hypoxia_wilcox_t1) <- c("Gene","n_mutated","n_wt","wilcox_p_val","wilcox_p_adj")
germline_hypoxia_wilcox_t1$Gene <- hallmark_hypoxia

for(i in 1:nrow(germline_hypoxia_wilcox_t1)){
  gene <- germline_hypoxia_wilcox_t1$Gene[i]

  
  test_data <- hypoxia_tier_1_genes[,c("Buffa_signature_Sum_within",gene)]
  
  data_mut <- test_data[test_data[,2] == "1",]
  data_wt <- test_data[test_data[,2] == "0",]
  
  if(nrow(data_mut) > 3){
    germline_hypoxia_wilcox_t1$wilcox_p_val[i] <- wilcox.test(data_mut$Buffa_signature_Sum_within,
                                                              data_wt$Buffa_signature_Sum_within)[[3]]

    germline_hypoxia_wilcox_t1$n_mutated[i] <- nrow(data_mut)
    germline_hypoxia_wilcox_t1$n_wt[i] <- nrow(data_wt)
  }
  
}
# germline_hypoxia_wilcox_t1 <- subset(germline_hypoxia_wilcox_t1, !is.na(wilcox_p_val))
germline_hypoxia_wilcox_t1$wilcox_p_adj <- p.adjust(germline_hypoxia_wilcox_t1$wilcox_p_val, method = "BH")
```

```{r}
germline_hypoxia_wilcox_t1
```


```{r}
dim(germline_hypoxia_wilcox_t1)
```

###### Binomial model
```{r}
hallmark_hypoxia_continuous_coefficients
```


### DRG 36
#### Any gene in list
##### Quartiles
```{r}
fisher.test(table(hypoxia_tier_1_lists_quart$DRG.Clinical..36.,
                  hypoxia_tier_1_lists_quart$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
```
```{r}
table(hypoxia_tier_1_lists_quart$DRG.Clinical..36.,
      hypoxia_tier_1_lists_quart$Buffa_signature_Sum_within_quartile_one_sample_per_donor)
```


##### Scores
```{r}
wilcox.test(subset(hypoxia_tier_1_lists, DRG.Clinical..36. == "1")[,"Buffa_signature_Sum_within"], 
            subset(hypoxia_tier_1_lists, DRG.Clinical..36. == "0")[,"Buffa_signature_Sum_within"])
```


#### Individual genes
##### Quartiles
Make a table for Fisher exact test results
```{r}
germline_drg36_enrichment_t1 <- as.data.frame(matrix(nrow = length(drg_genes_36), ncol = 8))
colnames(germline_drg36_enrichment_t1) <- c("Gene","n_q4_mutated","n_q1_mutated","n_q4_wt","n_q1_wt","fisher_p_val","fisher_p_adj","odds_ratio")
germline_drg36_enrichment_t1$Gene <- drg_genes_36

for(i in 1:nrow(germline_drg36_enrichment_t1)){
  gene <- germline_drg36_enrichment_t1$Gene[i]

  
  test_data <- hypoxia_tier_1_genes_quart[,c("Buffa_signature_Sum_within_quartile_one_sample_per_donor",gene)]
  
  table_data <- as.data.frame(table(test_data[,gene], test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
  
  if(nrow(table_data) > 2){
    germline_drg36_enrichment_t1$fisher_p_val[i] <- fisher.test(table(test_data[,gene], test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))[[1]]
    germline_drg36_enrichment_t1$odds_ratio[i] <- fisher.test(table(test_data[,gene], test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))[[3]]
  
  
  
    germline_drg36_enrichment_t1$n_q4_mutated[i] <- subset(table_data, (Var1 == "1") & (Var2 == "Q4"))[,"Freq"]
    germline_drg36_enrichment_t1$n_q1_mutated[i] <- subset(table_data, (Var1 == "1") & (Var2 == "Q1"))[,"Freq"]
    germline_drg36_enrichment_t1$n_q4_wt[i] <- subset(table_data, (Var1 == "0") & (Var2 == "Q4"))[,"Freq"]
    germline_drg36_enrichment_t1$n_q1_wt[i] <- subset(table_data, (Var1 == "0") & (Var2 == "Q1"))[,"Freq"]
  }
  
}
# germline_drg36_enrichment_t1 <- subset(germline_drg36_enrichment_t1, !is.na(fisher_p_val))
germline_drg36_enrichment_t1$fisher_p_adj <- p.adjust(germline_drg36_enrichment_t1$fisher_p_val, method = "BH")
```

```{r}
germline_drg36_enrichment_t1
```

###### Binomial model
```{r}
drg36_quartile_coefficients
```


##### Scores
Make a table for Wilcox test results
```{r}
germline_drg36_wilcox_t1 <- as.data.frame(matrix(nrow = length(drg_genes_36), ncol = 5))
colnames(germline_drg36_wilcox_t1) <- c("Gene","n_mutated","n_wt","wilcox_p_val","wilcox_p_adj")
germline_drg36_wilcox_t1$Gene <- drg_genes_36

for(i in 1:nrow(germline_drg36_wilcox_t1)){
  gene <- germline_drg36_wilcox_t1$Gene[i]

  
  test_data <- hypoxia_tier_1_genes[,c("Buffa_signature_Sum_within",gene)]
  
  data_mut <- test_data[test_data[,2] == "1",]
  data_wt <- test_data[test_data[,2] == "0",]
  
  if(nrow(data_mut) > 3){
    germline_drg36_wilcox_t1$wilcox_p_val[i] <- wilcox.test(data_mut$Buffa_signature_Sum_within,
                                                            data_wt$Buffa_signature_Sum_within)[[3]]

    germline_drg36_wilcox_t1$n_mutated[i] <- nrow(data_mut)
    germline_drg36_wilcox_t1$n_wt[i] <- nrow(data_wt)
  }
  
}
# germline_drg36_wilcox_t1 <- subset(germline_drg36_wilcox_t1, !is.na(wilcox_p_val))
germline_drg36_wilcox_t1$wilcox_p_adj <- p.adjust(germline_drg36_wilcox_t1$wilcox_p_val, method = "BH")
```

```{r}
germline_drg36_wilcox_t1
```


```{r}
dim(germline_drg36_wilcox_t1)
```

####### Binomial model
```{r}
drg36_continuous_coefficients
```

### DRG 128
#### Any gene in list
##### Quartiles
```{r}
fisher.test(table(hypoxia_tier_1_lists_quart$DRG..128., hypoxia_tier_1_lists_quart$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
```
```{r}
table(hypoxia_tier_1_lists_quart$DRG..128., 
    hypoxia_tier_1_lists_quart$Buffa_signature_Sum_within_quartile_one_sample_per_donor)
```

##### Scores
```{r}
wilcox.test(subset(hypoxia_tier_1_lists, DRG..128. == "1")[,"Buffa_signature_Sum_within"], 
            subset(hypoxia_tier_1_lists, DRG..128. == "0")[,"Buffa_signature_Sum_within"])
```


#### Individual genes
##### Quartiles
Make a table for Fisher exact test results
```{r}
germline_drg128_enrichment_t1 <- as.data.frame(matrix(nrow = length(drg_genes_128), ncol = 8))
colnames(germline_drg128_enrichment_t1) <- c("Gene","n_q4_mutated","n_q1_mutated","n_q4_wt","n_q1_wt","fisher_p_val","fisher_p_adj","odds_ratio")
germline_drg128_enrichment_t1$Gene <- drg_genes_128

for(i in 1:nrow(germline_drg128_enrichment_t1)){
  gene <- germline_drg128_enrichment_t1$Gene[i]

  
  test_data <- hypoxia_tier_1_genes_quart[,c("Buffa_signature_Sum_within_quartile_one_sample_per_donor",gene)]
  
  table_data <- as.data.frame(table(test_data[,gene], test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
  
  if(nrow(table_data) > 2){
    germline_drg128_enrichment_t1$fisher_p_val[i] <- fisher.test(table(test_data[,gene], test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))[[1]]
    germline_drg128_enrichment_t1$odds_ratio[i] <- fisher.test(table(test_data[,gene], test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))[[3]]
  
  
  
    germline_drg128_enrichment_t1$n_q4_mutated[i] <- subset(table_data, (Var1 == "1") & (Var2 == "Q4"))[,"Freq"]
    germline_drg128_enrichment_t1$n_q1_mutated[i] <- subset(table_data, (Var1 == "1") & (Var2 == "Q1"))[,"Freq"]
    germline_drg128_enrichment_t1$n_q4_wt[i] <- subset(table_data, (Var1 == "0") & (Var2 == "Q4"))[,"Freq"]
    germline_drg128_enrichment_t1$n_q1_wt[i] <- subset(table_data, (Var1 == "0") & (Var2 == "Q1"))[,"Freq"]
  }
  
}
# germline_drg128_enrichment_t1 <- subset(germline_drg128_enrichment_t1, !is.na(fisher_p_val))
germline_drg128_enrichment_t1$fisher_p_adj <- p.adjust(germline_drg128_enrichment_t1$fisher_p_val, method = "BH")
```

```{r}
germline_drg128_enrichment_t1
```


###### Binomial model
```{r}
drg128_quartile_coefficients
```



##### Scores
Make a table for Wilcox test results
```{r}
germline_drg128_wilcox_t1 <- as.data.frame(matrix(nrow = length(drg_genes_128), ncol = 5))
colnames(germline_drg128_wilcox_t1) <- c("Gene","n_mutated","n_wt","wilcox_p_val","wilcox_p_adj")
germline_drg128_wilcox_t1$Gene <- drg_genes_128

for(i in 1:nrow(germline_drg128_wilcox_t1)){
  gene <- germline_drg128_wilcox_t1$Gene[i]

  
  test_data <- hypoxia_tier_1_genes[,c("Buffa_signature_Sum_within",gene)]
  
  data_mut <- test_data[test_data[,2] == "1",]
  data_wt <- test_data[test_data[,2] == "0",]
  
  if(nrow(data_mut) > 3){
    germline_drg128_wilcox_t1$wilcox_p_val[i] <- wilcox.test(data_mut$Buffa_signature_Sum_within, data_wt$Buffa_signature_Sum_within)[[3]]

    germline_drg128_wilcox_t1$n_mutated[i] <- nrow(data_mut)
    germline_drg128_wilcox_t1$n_wt[i] <- nrow(data_wt)
  }
  
}
# germline_drg128_wilcox_t1 <- subset(germline_drg128_wilcox_t1, !is.na(wilcox_p_val))
germline_drg128_wilcox_t1$wilcox_p_adj <- p.adjust(germline_drg128_wilcox_t1$wilcox_p_val, method = "BH")
```

```{r}
germline_drg128_wilcox_t1
```


```{r}
dim(germline_drg128_wilcox_t1)
```


####### Binomial model
```{r}
drg128_continuous_coefficients
```


## Tier 1 and 2 variants
Sample vs gene list
```{r}
hypoxia_tier_1_2_lists <- merge(hypoxia_signatures_tum_data_single,
                               germline_tier_1_2_sample_list,
                               by.x = "PPCG_Donor_ID",
                               by.y = "Column1")
dim(hypoxia_tier_1_2_lists)
```

Individual genes vs samples
```{r}
germline_tier_1_2_sample_gene_sub <- subset(germline_tier_1_2_sample_gene, Genes %in% unique((c(drg_genes_36, drg_genes_128, hallmark_hypoxia))))
rownames(germline_tier_1_2_sample_gene_sub) <- germline_tier_1_2_sample_gene_sub$Genes
germline_tier_1_2_sample_gene_sub <- as.data.frame(t(as.matrix(germline_tier_1_2_sample_gene_sub[,3:ncol(germline_tier_1_2_sample_gene_sub)])))
germline_tier_1_2_sample_gene_sub$Sample <- rownames(germline_tier_1_2_sample_gene_sub)
```

```{r}
hypoxia_tier_1_2_genes <- merge(hypoxia_signatures_tum_data_single,
                               germline_tier_1_2_sample_gene_sub,
                               by.x = "PPCG_Donor_ID",
                               by.y = "Sample")
dim(hypoxia_tier_1_2_genes)
```

### Hallmark hypoxia
#### Any gene in list
```{r}
table(hypoxia_tier_1_2_lists$HALLMARK_HYPOXIA)
```

##### Quartiles

```{r}
hypoxia_tier_1_2_lists_quart <- subset(hypoxia_tier_1_2_lists, Buffa_signature_Sum_within_quartile_one_sample_per_donor %in% c("Q1","Q4"))
```

```{r}
fisher.test(table(hypoxia_tier_1_2_lists_quart$HALLMARK_HYPOXIA, hypoxia_tier_1_2_lists_quart$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
```
```{r}
table(hypoxia_tier_1_2_lists_quart$HALLMARK_HYPOXIA, hypoxia_tier_1_2_lists_quart$Buffa_signature_Sum_within_quartile_one_sample_per_donor)
```


##### Scores
```{r}
wilcox.test(subset(hypoxia_tier_1_2_lists, HALLMARK_HYPOXIA == "1")[,"Buffa_signature_Sum_within"], 
            subset(hypoxia_tier_1_2_lists, HALLMARK_HYPOXIA == "0")[,"Buffa_signature_Sum_within"])
```


#### Individual genes
##### Quartiles
```{r}
hypoxia_tier_1_2_genes_quart <- subset(hypoxia_tier_1_2_genes, Buffa_signature_Sum_within_quartile_one_sample_per_donor %in% c("Q1","Q4"))
```

Make a table for Fisher exact test results
```{r}
germline_hypoxia_enrichment_t1_2 <- as.data.frame(matrix(nrow = length(hallmark_hypoxia), ncol = 8))
colnames(germline_hypoxia_enrichment_t1_2) <- c("Gene","n_q4_mutated","n_q1_mutated","n_q4_wt","n_q1_wt","fisher_p_val","fisher_p_adj","odds_ratio")
germline_hypoxia_enrichment_t1_2$Gene <- hallmark_hypoxia

for(i in 1:nrow(germline_hypoxia_enrichment_t1_2)){
  gene <- germline_hypoxia_enrichment_t1_2$Gene[i]

  
  test_data <- hypoxia_tier_1_2_genes_quart[,c("Buffa_signature_Sum_within_quartile_one_sample_per_donor",gene)]
  
  table_data <- as.data.frame(table(test_data[,gene], test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
  
  if(nrow(table_data) > 2){
    germline_hypoxia_enrichment_t1_2$fisher_p_val[i] <- fisher.test(table(test_data[,gene], test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))[[1]]
    germline_hypoxia_enrichment_t1_2$odds_ratio[i] <- fisher.test(table(test_data[,gene], test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))[[3]]
  
  
  
    germline_hypoxia_enrichment_t1_2$n_q4_mutated[i] <- subset(table_data, (Var1 == "1") & (Var2 == "Q4"))[,"Freq"]
    germline_hypoxia_enrichment_t1_2$n_q1_mutated[i] <- subset(table_data, (Var1 == "1") & (Var2 == "Q1"))[,"Freq"]
    germline_hypoxia_enrichment_t1_2$n_q4_wt[i] <- subset(table_data, (Var1 == "0") & (Var2 == "Q4"))[,"Freq"]
    germline_hypoxia_enrichment_t1_2$n_q1_wt[i] <- subset(table_data, (Var1 == "0") & (Var2 == "Q1"))[,"Freq"]
  }
  
}
# germline_hypoxia_enrichment_t1_2 <- subset(germline_hypoxia_enrichment_t1_2, !is.na(fisher_p_val))
germline_hypoxia_enrichment_t1_2$fisher_p_adj <- p.adjust(germline_hypoxia_enrichment_t1_2$fisher_p_val, method = "BH")
```

```{r}
germline_hypoxia_enrichment_t1_2
```

##### Scores
Make a table for Wilcox test results
```{r}
germline_hypoxia_wilcox_t1_2 <- as.data.frame(matrix(nrow = length(hallmark_hypoxia), ncol = 5))
colnames(germline_hypoxia_wilcox_t1_2) <- c("Gene","n_mutated","n_wt","wilcox_p_val","wilcox_p_adj")
germline_hypoxia_wilcox_t1_2$Gene <- hallmark_hypoxia

for(i in 1:nrow(germline_hypoxia_wilcox_t1_2)){
  gene <- germline_hypoxia_wilcox_t1_2$Gene[i]

  
  test_data <- hypoxia_tier_1_2_genes[,c("Buffa_signature_Sum_within",gene)]
  
  data_mut <- test_data[test_data[,2] == "1",]
  data_wt <- test_data[test_data[,2] == "0",]
  
  if(nrow(data_mut) > 3){
    germline_hypoxia_wilcox_t1_2$wilcox_p_val[i] <- wilcox.test(data_mut$Buffa_signature_Sum_within, data_wt$Buffa_signature_Sum_within)[[3]]

    germline_hypoxia_wilcox_t1_2$n_mutated[i] <- nrow(data_mut)
    germline_hypoxia_wilcox_t1_2$n_wt[i] <- nrow(data_wt)
  }
  
}
# germline_hypoxia_wilcox_t1_2 <- subset(germline_hypoxia_wilcox_t1_2, !is.na(wilcox_p_val))
germline_hypoxia_wilcox_t1_2$wilcox_p_adj <- p.adjust(germline_hypoxia_wilcox_t1_2$wilcox_p_val, method = "BH")
```

```{r}
germline_hypoxia_wilcox_t1_2
```


```{r}
dim(germline_hypoxia_wilcox_t1_2)
```

### DRG 36
```{r}
table(hypoxia_tier_1_2_lists$DRG.Clinical..36.)
```
#### Any gene in list
##### Quartiles
```{r}
fisher.test(table(hypoxia_tier_1_2_lists_quart$DRG.Clinical..36.,
                  hypoxia_tier_1_2_lists_quart$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
```
```{r}
table(hypoxia_tier_1_2_lists_quart$DRG.Clinical..36.,
      hypoxia_tier_1_2_lists_quart$Buffa_signature_Sum_within_quartile_one_sample_per_donor)
```


##### Scores
```{r}
wilcox.test(subset(hypoxia_tier_1_2_lists, DRG.Clinical..36. == "1")[,"Buffa_signature_Sum_within"], 
            subset(hypoxia_tier_1_2_lists, DRG.Clinical..36. == "0")[,"Buffa_signature_Sum_within"])
```


#### Individual genes
##### Quartiles
Make a table for Fisher exact test results
```{r}
germline_drg36_enrichment_t1_2 <- as.data.frame(matrix(nrow = length(drg_genes_36), ncol = 8))
colnames(germline_drg36_enrichment_t1_2) <- c("Gene","n_q4_mutated","n_q1_mutated","n_q4_wt","n_q1_wt","fisher_p_val","fisher_p_adj","odds_ratio")
germline_drg36_enrichment_t1_2$Gene <- drg_genes_36

for(i in 1:nrow(germline_drg36_enrichment_t1_2)){
  gene <- germline_drg36_enrichment_t1_2$Gene[i]

  
  test_data <- hypoxia_tier_1_2_genes_quart[,c("Buffa_signature_Sum_within_quartile_one_sample_per_donor",gene)]
  
  table_data <- as.data.frame(table(test_data[,gene], test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
  
  if(nrow(table_data) > 2){
    germline_drg36_enrichment_t1_2$fisher_p_val[i] <- fisher.test(table(test_data[,gene], test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))[[1]]
    germline_drg36_enrichment_t1_2$odds_ratio[i] <- fisher.test(table(test_data[,gene], test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))[[3]]
  
  
  
    germline_drg36_enrichment_t1_2$n_q4_mutated[i] <- subset(table_data, (Var1 == "1") & (Var2 == "Q4"))[,"Freq"]
    germline_drg36_enrichment_t1_2$n_q1_mutated[i] <- subset(table_data, (Var1 == "1") & (Var2 == "Q1"))[,"Freq"]
    germline_drg36_enrichment_t1_2$n_q4_wt[i] <- subset(table_data, (Var1 == "0") & (Var2 == "Q4"))[,"Freq"]
    germline_drg36_enrichment_t1_2$n_q1_wt[i] <- subset(table_data, (Var1 == "0") & (Var2 == "Q1"))[,"Freq"]
  }
  
}
# germline_drg36_enrichment_t1_2 <- subset(germline_drg36_enrichment_t1_2, !is.na(fisher_p_val))
germline_drg36_enrichment_t1_2$fisher_p_adj <- p.adjust(germline_drg36_enrichment_t1_2$fisher_p_val, method = "BH")
```

```{r}
germline_drg36_enrichment_t1_2
```

##### Scores
Make a table for Wilcox test results
```{r}
germline_drg36_wilcox_t1_2 <- as.data.frame(matrix(nrow = length(drg_genes_36), ncol = 5))
colnames(germline_drg36_wilcox_t1_2) <- c("Gene","n_mutated","n_wt","wilcox_p_val","wilcox_p_adj")
germline_drg36_wilcox_t1_2$Gene <- drg_genes_36

for(i in 1:nrow(germline_drg36_wilcox_t1_2)){
  gene <- germline_drg36_wilcox_t1_2$Gene[i]

  
  test_data <- hypoxia_tier_1_2_genes[,c("Buffa_signature_Sum_within",gene)]
  
  data_mut <- test_data[test_data[,2] == "1",]
  data_wt <- test_data[test_data[,2] == "0",]
  
  if(nrow(data_mut) > 3){
    germline_drg36_wilcox_t1_2$wilcox_p_val[i] <- wilcox.test(data_mut$Buffa_signature_Sum_within,
                                                              data_wt$Buffa_signature_Sum_within)[[3]]

    germline_drg36_wilcox_t1_2$n_mutated[i] <- nrow(data_mut)
    germline_drg36_wilcox_t1_2$n_wt[i] <- nrow(data_wt)
  }
  
}
# germline_drg36_wilcox_t1_2 <- subset(germline_drg36_wilcox_t1_2, !is.na(wilcox_p_val))
germline_drg36_wilcox_t1_2$wilcox_p_adj <- p.adjust(germline_drg36_wilcox_t1_2$wilcox_p_val, method = "BH")
```

```{r}
germline_drg36_wilcox_t1_2
```


```{r}
dim(germline_drg36_wilcox_t1_2)
```

### DRG 128
#### Any gene in list
```{r}
table(hypoxia_tier_1_2_lists$DRG..128.)
```
##### Quartiles
```{r}
fisher.test(table(hypoxia_tier_1_2_lists_quart$DRG..128., hypoxia_tier_1_2_lists_quart$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
```
```{r}
table(hypoxia_tier_1_2_lists_quart$DRG..128., 
      hypoxia_tier_1_2_lists_quart$Buffa_signature_Sum_within_quartile_one_sample_per_donor)
```


##### Scores
```{r}
wilcox.test(subset(hypoxia_tier_1_2_lists, DRG..128. == "1")[,"Buffa_signature_Sum_within"], 
            subset(hypoxia_tier_1_2_lists, DRG..128. == "0")[,"Buffa_signature_Sum_within"])
```


#### Individual genes
##### Quartiles
Make a table for Fisher exact test results
```{r}
germline_drg128_enrichment_t1_2 <- as.data.frame(matrix(nrow = length(drg_genes_128), ncol = 8))
colnames(germline_drg128_enrichment_t1_2) <- c("Gene","n_q4_mutated","n_q1_mutated","n_q4_wt","n_q1_wt","fisher_p_val","fisher_p_adj","odds_ratio")
germline_drg128_enrichment_t1_2$Gene <- drg_genes_128

for(i in 1:nrow(germline_drg128_enrichment_t1_2)){
  gene <- germline_drg128_enrichment_t1_2$Gene[i]

  
  test_data <- hypoxia_tier_1_2_genes_quart[,c("Buffa_signature_Sum_within_quartile_one_sample_per_donor",gene)]
  
  table_data <- as.data.frame(table(test_data[,gene], test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
  
  if(nrow(table_data) > 2){
    germline_drg128_enrichment_t1_2$fisher_p_val[i] <- fisher.test(table(test_data[,gene], test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))[[1]]
    germline_drg128_enrichment_t1_2$odds_ratio[i] <- fisher.test(table(test_data[,gene], test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))[[3]]
  
  
  
    germline_drg128_enrichment_t1_2$n_q4_mutated[i] <- subset(table_data, (Var1 == "1") & (Var2 == "Q4"))[,"Freq"]
    germline_drg128_enrichment_t1_2$n_q1_mutated[i] <- subset(table_data, (Var1 == "1") & (Var2 == "Q1"))[,"Freq"]
    germline_drg128_enrichment_t1_2$n_q4_wt[i] <- subset(table_data, (Var1 == "0") & (Var2 == "Q4"))[,"Freq"]
    germline_drg128_enrichment_t1_2$n_q1_wt[i] <- subset(table_data, (Var1 == "0") & (Var2 == "Q1"))[,"Freq"]
  }
  
}
# germline_drg128_enrichment_t1_2 <- subset(germline_drg128_enrichment_t1_2, !is.na(fisher_p_val))
germline_drg128_enrichment_t1_2$fisher_p_adj <- p.adjust(germline_drg128_enrichment_t1_2$fisher_p_val, method = "BH")
```

```{r}
germline_drg128_enrichment_t1_2
```

##### Scores
Make a table for Wilcox test results
```{r}
germline_drg128_wilcox_t1_2 <- as.data.frame(matrix(nrow = length(drg_genes_128), ncol = 5))
colnames(germline_drg128_wilcox_t1_2) <- c("Gene","n_mutated","n_wt","wilcox_p_val","wilcox_p_adj")
germline_drg128_wilcox_t1_2$Gene <- drg_genes_128

for(i in 1:nrow(germline_drg128_wilcox_t1_2)){
  gene <- germline_drg128_wilcox_t1_2$Gene[i]

  
  test_data <- hypoxia_tier_1_2_genes[,c("Buffa_signature_Sum_within",gene)]
  
  data_mut <- test_data[test_data[,2] == "1",]
  data_wt <- test_data[test_data[,2] == "0",]
  
  if(nrow(data_mut) > 3){
    germline_drg128_wilcox_t1_2$wilcox_p_val[i] <- wilcox.test(data_mut$Buffa_signature_Sum_within,
                                                               data_wt$Buffa_signature_Sum_within)[[3]]

    germline_drg128_wilcox_t1_2$n_mutated[i] <- nrow(data_mut)
    germline_drg128_wilcox_t1_2$n_wt[i] <- nrow(data_wt)
  }
  
}
# germline_drg128_wilcox_t1_2 <- subset(germline_drg128_wilcox_t1_2, !is.na(wilcox_p_val))
germline_drg128_wilcox_t1_2$wilcox_p_adj <- p.adjust(germline_drg128_wilcox_t1_2$wilcox_p_val, method = "BH")
```

```{r}
germline_drg128_wilcox_t1_2
```


```{r}
dim(germline_drg128_wilcox_t1_2)
```

## Follow up on individual genes
### Hallmark hypoxia
#### SLC2A3 Tier 1 and 2
```{r}
ggplot(hypoxia_tier_1_2_genes[,c("Buffa_signature_Sum_within","SLC2A3")], aes(x = as.factor(SLC2A3), y = Buffa_signature_Sum_within))+
  geom_boxplot(outlier.colour = NA)+
  geom_jitter(alpha = 0.3, height = 0)+
  theme_light()+
  geom_signif(map_signif_level = TRUE, tip_length = 0, comparisons = list(c("0","1")), test = "wilcox.test")+
  ylab("Hypoxia Score")+
  xlab("SLC2A3 Germline Mutation Status")
```



# Somatic variants in hypoxia genes

Hallmark hypoxia genes with somatic mutations identified by both callers or a tier 1 indel (filtered data)
```{r}
hallmark_hypoxia[hallmark_hypoxia %in% rownames(driver_data)]
```

DRG genes that do not have somatic mutations identified by both callers or a tier 1 indel (filtered data)
```{r}
hallmark_hypoxia[hallmark_hypoxia %!in% rownames(driver_data)]
```

```{r}
somatic_hallmark_hypoxia_mat <- as.matrix(driver_data[rownames(driver_data) %in% hallmark_hypoxia,])
somatic_hallmark_hypoxia_mat = somatic_hallmark_hypoxia_mat[, -c(1,2)]
somatic_hallmark_hypoxia_mat <- gsub("\\.","", somatic_hallmark_hypoxia_mat)
```

```{r}
dim(somatic_hallmark_hypoxia_mat)
```

Reorder signatures for oncoprint.  Only plot samples for which we have data from WGS/drivers. 
```{r}
# order driver samples by hypoxia scores
somatic_hallmark_hypoxia_mat <- somatic_hallmark_hypoxia_mat[,hypoxia_signatures_tum_data_single$WGS_AssayID[hypoxia_signatures_tum_data_single$WGS_AssayID %in% colnames(somatic_hallmark_hypoxia_mat)]]

```

```{r}
dim(somatic_hallmark_hypoxia_mat)
```

```{r}
table(somatic_hallmark_hypoxia_mat)
```

```{r}
somatic_hallmark_hypoxia_mat[somatic_hallmark_hypoxia_mat == "nsSNV"] <- "Mutation"
somatic_hallmark_hypoxia_mat[somatic_hallmark_hypoxia_mat == "splicing"] <- "Mutation"
somatic_hallmark_hypoxia_mat[somatic_hallmark_hypoxia_mat == "stopgain"] <- "Mutation"
somatic_hallmark_hypoxia_mat[somatic_hallmark_hypoxia_mat == "stoploss"] <- "Mutation"
somatic_hallmark_hypoxia_mat[somatic_hallmark_hypoxia_mat == "startloss"] <- "Mutation"
somatic_hallmark_hypoxia_mat[somatic_hallmark_hypoxia_mat == "stopgain+nsSNV"] <- "Mutation"
somatic_hallmark_hypoxia_mat[somatic_hallmark_hypoxia_mat == "nsSNV+splicing"] <- "Mutation"
somatic_hallmark_hypoxia_mat[somatic_hallmark_hypoxia_mat == "nsSNV+stopgain"] <- "Mutation"
somatic_hallmark_hypoxia_mat[somatic_hallmark_hypoxia_mat == "splicing+nsSNV"] <- "Mutation"
somatic_hallmark_hypoxia_mat[somatic_hallmark_hypoxia_mat == "startloss+nsSNV"] <- "Mutation"

somatic_hallmark_hypoxia_mat[somatic_hallmark_hypoxia_mat == ""] <- "No Mutation"
```

```{r}
table(somatic_hallmark_hypoxia_mat)
```

## Test for differences in scores
```{r}
somatic_hallmark_hypoxia_mat_hypoxia_scores <- as.data.frame(matrix(nrow = nrow(somatic_hallmark_hypoxia_mat), 
                                                                    ncol = 5))
colnames(somatic_hallmark_hypoxia_mat_hypoxia_scores) <- c("Gene","n_mutated","n_wt","wilcox_p_val","wilcox_p_adj")
somatic_hallmark_hypoxia_mat_hypoxia_scores$Gene <- rownames(somatic_hallmark_hypoxia_mat)

for(i in 1:nrow(somatic_hallmark_hypoxia_mat_hypoxia_scores)){
  gene <- somatic_hallmark_hypoxia_mat_hypoxia_scores$Gene[i]
  test_gene <- as.data.frame(t(somatic_hallmark_hypoxia_mat)[,gene])
  colnames(test_gene) <-"Test_gene"
  test_gene$sample <- rownames(test_gene)
  
  test_data <- merge(test_gene, hypoxia_signatures_tum_data_single[,c("Buffa_signature_Sum_within","WGS_AssayID")],
                     by.x = "sample", by.y = "WGS_AssayID")
  
  table_data <- as.data.frame(table(test_data$Test_gene))
  if(nrow(table_data) > 1){
    somatic_hallmark_hypoxia_mat_hypoxia_scores$wilcox_p_val[i] <- wilcox.test(subset(test_data, Test_gene == "Mutation")[,"Buffa_signature_Sum_within"],
                                                         subset(test_data, Test_gene == "No Mutation")[,"Buffa_signature_Sum_within"])[[3]]
  
    somatic_hallmark_hypoxia_mat_hypoxia_scores$n_mutated[i] <- nrow(subset(test_data, Test_gene == "Mutation"))
    somatic_hallmark_hypoxia_mat_hypoxia_scores$n_wt[i] <- nrow(subset(test_data, Test_gene == "No Mutation"))
  }
  
}

somatic_hallmark_hypoxia_mat_hypoxia_scores$wilcox_p_adj <- p.adjust(somatic_hallmark_hypoxia_mat_hypoxia_scores$wilcox_p_val, method = "BH")
```

```{r}
somatic_hallmark_hypoxia_mat_hypoxia_scores$Significance <- ifelse(somatic_hallmark_hypoxia_mat_hypoxia_scores$wilcox_p_adj < 0.05, yes = "Significant", no = "N.S.")
```

```{r}
rownames(somatic_hallmark_hypoxia_mat_hypoxia_scores) <- somatic_hallmark_hypoxia_mat_hypoxia_scores$Gene
```

Order data by p-value
```{r}
somatic_hallmark_hypoxia_mat_hypoxia_scores <- somatic_hallmark_hypoxia_mat_hypoxia_scores[order(somatic_hallmark_hypoxia_mat_hypoxia_scores$wilcox_p_val),]
rownames(somatic_hallmark_hypoxia_mat_hypoxia_scores) <- somatic_hallmark_hypoxia_mat_hypoxia_scores$Gene
```

```{r}
(somatic_hallmark_hypoxia_mat_hypoxia_scores)
```

## Q1 Q4 difference
```{r}
somatic_hallmark_hypoxia_mat_hypoxia_enrichment_quart <- as.data.frame(matrix(nrow = nrow(somatic_hallmark_hypoxia_mat), ncol = 8))
colnames(somatic_hallmark_hypoxia_mat_hypoxia_enrichment_quart) <- c("Gene","n_q4_mutated","n_q1_mutated","n_q4_wt","n_q1_wt","fisher_p_val","fisher_p_adj","odds_ratio")
somatic_hallmark_hypoxia_mat_hypoxia_enrichment_quart$Gene <- rownames(somatic_hallmark_hypoxia_mat)

for(i in 1:nrow(somatic_hallmark_hypoxia_mat_hypoxia_enrichment_quart)){
  gene <- somatic_hallmark_hypoxia_mat_hypoxia_enrichment_quart$Gene[i]
  test_gene <- as.data.frame(t(somatic_hallmark_hypoxia_mat)[,gene])
  colnames(test_gene) <-"Test_gene"
  test_gene$sample <- rownames(test_gene)
  
  test_data <- merge(test_gene, hypoxia_signatures_tum_data_single[,c("Buffa_signature_Sum_within_quartile_one_sample_per_donor","WGS_AssayID")],
                     by.x = "sample", by.y = "WGS_AssayID")
  test_data <- subset(test_data, Buffa_signature_Sum_within_quartile_one_sample_per_donor %in% c("Q1","Q4"))
  
  table_data <- as.data.frame(table(test_data$Test_gene, test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
  if(nrow(table_data) > 2){
    somatic_hallmark_hypoxia_mat_hypoxia_enrichment_quart$fisher_p_val[i] <- fisher.test(table(test_data$Test_gene, test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))[[1]]
    somatic_hallmark_hypoxia_mat_hypoxia_enrichment_quart$odds_ratio[i] <- fisher.test(table(test_data$Test_gene, test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))[[3]]
  
  
  
    somatic_hallmark_hypoxia_mat_hypoxia_enrichment_quart$n_q4_mutated[i] <- subset(table_data, (Var1 == "Mutation") & (Var2 == "Q4"))[,"Freq"]
    somatic_hallmark_hypoxia_mat_hypoxia_enrichment_quart$n_q1_mutated[i] <- subset(table_data, (Var1 == "Mutation") & (Var2 == "Q1"))[,"Freq"]
    somatic_hallmark_hypoxia_mat_hypoxia_enrichment_quart$n_q4_wt[i] <- subset(table_data, (Var1 == "No Mutation") & (Var2 == "Q4"))[,"Freq"]
    somatic_hallmark_hypoxia_mat_hypoxia_enrichment_quart$n_q1_wt[i] <- subset(table_data, (Var1 == "No Mutation") & (Var2 == "Q1"))[,"Freq"]
  }
  
}

somatic_hallmark_hypoxia_mat_hypoxia_enrichment_quart$fisher_p_adj <- p.adjust(somatic_hallmark_hypoxia_mat_hypoxia_enrichment_quart$fisher_p_val, method = "BH")
```

```{r}
# subset(somatic_hallmark_hypoxia_mat_hypoxia_enrichment_quart, fisher_p_adj < 0.05)
```

```{r}
somatic_hallmark_hypoxia_mat_hypoxia_enrichment_quart$Significance <- ifelse(somatic_hallmark_hypoxia_mat_hypoxia_enrichment_quart$fisher_p_adj < 0.05, yes = "Significant", no = "N.S.")
```

```{r}
rownames(somatic_hallmark_hypoxia_mat_hypoxia_enrichment_quart) <- somatic_hallmark_hypoxia_mat_hypoxia_enrichment_quart$Gene
```

```{r}
somatic_hallmark_hypoxia_mat_hypoxia_enrichment_quart
```

## Check mutation in group vs not
Check if mutation in any associated gene has differences in hypoxia score
```{r}
somatic_hallmark_hypoxia_mat_bin <- somatic_hallmark_hypoxia_mat
somatic_hallmark_hypoxia_mat_bin[somatic_hallmark_hypoxia_mat_bin == "Mutation"] <- as.numeric(1)
somatic_hallmark_hypoxia_mat_bin[somatic_hallmark_hypoxia_mat_bin == "No Mutation"] <- as.numeric(0)
```

```{r}
class(somatic_hallmark_hypoxia_mat_bin) <- "numeric"
```

```{r}
summary_somatic_hallmark_hypoxia <- as.data.frame(colSums(somatic_hallmark_hypoxia_mat_bin))
colnames(summary_somatic_hallmark_hypoxia) <- c("Muts_in_hallmark_hypoxia")
summary_somatic_hallmark_hypoxia$WGS_AssayID <- rownames(summary_somatic_hallmark_hypoxia)
summary_somatic_hallmark_hypoxia$bin_muts_in_hallmark_hypoxia <- ifelse(summary_somatic_hallmark_hypoxia$Muts_in_hallmark_hypoxia == 0, yes = "No Mutation", no = "Mutation")
```

```{r}
summary_somatic_hallmark_hypoxia <- merge(summary_somatic_hallmark_hypoxia, 
                               hypoxia_signatures_tum_data_single[,c("Buffa_signature_Sum_within","WGS_AssayID",
                                             "Buffa_signature_Sum_within_quartile_one_sample_per_donor")],
                               by = "WGS_AssayID")
```

```{r}
summary_somatic_hallmark_hypoxia_quart <- subset(summary_somatic_hallmark_hypoxia, Buffa_signature_Sum_within_quartile_one_sample_per_donor %in% c("Q1","Q4"))
```

```{r}
fisher.test(table(summary_somatic_hallmark_hypoxia_quart$bin_muts_in_hallmark_hypoxia,
                  summary_somatic_hallmark_hypoxia_quart$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
```
```{r}
table(summary_somatic_hallmark_hypoxia_quart$bin_muts_in_hallmark_hypoxia, 
      summary_somatic_hallmark_hypoxia_quart$Buffa_signature_Sum_within_quartile_one_sample_per_donor)
```

```{r}
wilcox.test(subset(summary_somatic_hallmark_hypoxia, bin_muts_in_hallmark_hypoxia == "Mutation")[,"Buffa_signature_Sum_within"], 
            subset(summary_somatic_hallmark_hypoxia, bin_muts_in_hallmark_hypoxia == "No Mutation")[,"Buffa_signature_Sum_within"])
```

```{r}
ggplot(summary_somatic_hallmark_hypoxia, aes(x = bin_muts_in_hallmark_hypoxia, y = Buffa_signature_Sum_within))+
  geom_boxplot()+
  geom_jitter(alpha = 0.5, height = 0)+
  ylab("Hypoxia Score")+
  xlab("Somatic Mutations in Hallmark Hypoxia Genes")+
  geom_signif(map_signif_level = TRUE, tip_length = 0, test = "wilcox.test", comparisons = list(c("Mutation","No Mutation")))
```


# Heatmap of hallmark hypoxia genes
```{r}
rownames(hypoxia_tier_1_2_genes) <- hypoxia_tier_1_2_genes$WGS_AssayID
hypoxia_tier_1_2_genes_mat <- t(as.matrix(hypoxia_tier_1_2_genes[,colnames(hypoxia_tier_1_2_genes) %in% hallmark_hypoxia]))
```
```{r}
hypoxia_tier_1_2_genes_mat[hypoxia_tier_1_2_genes_mat == "1"] <- "Mutation"

hypoxia_tier_1_2_genes_mat[hypoxia_tier_1_2_genes_mat == "0"] <- "No Mutation"
```

```{r}
hypoxia_tier_1_2_genes_mat[1:4,1:4]
```

```{r}
somatic_hallmark_hypoxia_mat[1:4,1:4]
```

# Combine somatic and germline hallmark hypoxia variants
Combine somatic and germline variants so we can just plot as one heatmap then check any variant?
```{r}
germ_som_hallmark_hypoxia_mat <- matrix(nrow = length(hallmark_hypoxia), ncol = ncol(somatic_hallmark_hypoxia_mat))
rownames(germ_som_hallmark_hypoxia_mat) <- hallmark_hypoxia
colnames(germ_som_hallmark_hypoxia_mat) <- colnames(somatic_hallmark_hypoxia_mat)
```

```{r}
for(i in 1:length(hallmark_hypoxia)){
  
  for(j in 1:ncol(germ_som_hallmark_hypoxia_mat)){
  somatic_gene <- "NA"
  somatic_sample <- "NA"
  somatic_status <- "NA"
  
  germline_gene <- "NA"
  germline_sample <- "NA"
  germline_status <- "NA"
  
  combined_label <- "NA"
    
    if(hallmark_hypoxia[i] %in% rownames(somatic_hallmark_hypoxia_mat)){
      somatic_gene <- hallmark_hypoxia[i]
      if(colnames(germ_som_hallmark_hypoxia_mat)[j] %in% colnames(somatic_hallmark_hypoxia_mat)){
        somatic_sample <- colnames(germ_som_hallmark_hypoxia_mat)[j]
        somatic_status <- somatic_hallmark_hypoxia_mat[somatic_gene,somatic_sample]
      }
    }
    if(hallmark_hypoxia[i] %in% rownames(hypoxia_tier_1_2_genes_mat)){
      germline_gene <- hallmark_hypoxia[i]
      if(colnames(germ_som_hallmark_hypoxia_mat)[j] %in% colnames(hypoxia_tier_1_2_genes_mat)){
        germline_sample <- colnames(germ_som_hallmark_hypoxia_mat)[j]
        germline_status <- hypoxia_tier_1_2_genes_mat[germline_gene,germline_sample]
      }
    }
    
    # Make class
    ## Not found in either
    if(all(c(somatic_status == "NA", germline_status == "NA"))){
      combined_label <- "NA"
    }else if(all(c(somatic_status == "Mutation", germline_status == "NA"))){## Somatic mutation not found in germline
      combined_label <- "Somatic mutation only"
    }else if(all(c(somatic_status == "NA", germline_status == "Mutation"))){## Germline mutation not found in somatic
      combined_label <- "Germline mutation only"
    }else if(all(c(somatic_status == "Mutation", germline_status == "Mutation"))){## Mutation in both
      combined_label <- "Both"
    }else if(all(c(somatic_status == "No Mutation", germline_status == "No Mutation"))){## Mutation in neither
      combined_label <- "NA"
    }else if(all(c(somatic_status == "Mutation", germline_status == "No Mutation"))){## Somatic mutation not germline
      combined_label <- "Somatic mutation only"
    }else if(all(c(somatic_status == "No Mutation", germline_status == "Mutation"))){## Germline mutation not somatic
      combined_label <- "Germline mutation only"
    }
    
    
    # Add label to new matrix
    germ_som_hallmark_hypoxia_mat[i,j] <- combined_label
  }
  
}
```

```{r}
hypoxia_op <- subset(hypoxia_signatures_tum_data_single, !is.na(WGS_AssayID) &
                       WGS_AssayID %!in% c("ADD",""))

rownames(hypoxia_op) <- hypoxia_op$WGS_AssayID

# order data by hypoxia scores
hypoxia_op <- hypoxia_op[rownames(hypoxia_op) %in% colnames(germ_som_hallmark_hypoxia_mat),]
hypoxia_op <- hypoxia_op[order(hypoxia_op$Buffa_signature_Sum_within, decreasing = FALSE),]

# order driver samples by hypoxia scores
germ_som_hallmark_hypoxia_mat <- germ_som_hallmark_hypoxia_mat[,rownames(hypoxia_op)[rownames(hypoxia_op) %in% colnames(germ_som_hallmark_hypoxia_mat)]]

```

Remove rows that are all NA. (only include rows where sum of is.na is not equal to the number of columns)
```{r}
germ_som_hallmark_hypoxia_mat <- germ_som_hallmark_hypoxia_mat[(rowSums((germ_som_hallmark_hypoxia_mat == "NA")) != ncol(germ_som_hallmark_hypoxia_mat)),]
```

```{r}
table(germ_som_hallmark_hypoxia_mat)
```


## Prepare heatmap
## Heatmap annotations
### Annotation colours
```{r}
cellularity_colours <- colorRamp2(c(0,max(hypoxia_op$Cellularity, na.rm = TRUE)),c("white","navyblue"))

age_colours <- c("< 40" = "white","40-50" = "grey75","50-65" = "grey50","65-70" = "grey25","≥ 70" = "black")
psa_colours <- c("0-9.9" = "navajowhite", "10-19.9" = "orange", "≥20" = "indianred")
gleason_colours <- c("3+3" = "white", "3+4" = "yellow", "4+3" = "orange", "4+4" = "red", "3+5" = "maroon3", "4+5" = "orangered4", "5+4" = "firebrick4", "5+5" = "black")
clin_t_colours <- c("T1" = "olivedrab2", "T2" = "green3", "T3" = "green4", "T4" = "darkgreen")

germ_som_colours <- c("Both" = "slateblue1",
                      "Neither" = "grey",
                      "NA" = "grey90",
                      "Somatic mutation only" = "red4",
                      "Germline mutation only" = "royalblue",
                      "Somatic mutation; NA" = "tomato",
                      "Germline mutation; NA" = "lightskyblue"
                      )
mut_colours <- c("Mutation" = "black",
                 "No Mutation" = "grey")
```

### Top annotation

```{r}
ha_row_annotation_hypoxia <- HeatmapAnnotation(Hypoxia = anno_barplot(hypoxia_op$Buffa_signature_Sum_within))
ha_row_annotation_top <- HeatmapAnnotation(Hypoxia = anno_barplot(hypoxia_op$Buffa_signature_Sum_within),
                                           Gleason = hypoxia_op$gleason_total,
                                           ClinicalT = hypoxia_op$clin_t_bracket,
                                           PSA = hypoxia_op$psa_bracket,
                                           Age = hypoxia_op$age_bracket,
                                           Cellularity = hypoxia_op$Cellularity,
                                           col = list(Gleason = gleason_colours,
                                                      ClinicalT = clin_t_colours,
                                                      PSA = psa_colours,
                                                      Age = age_colours,
                                                      Cellularity = cellularity_colours
                                           ))

```

## Hallmark Hypoxia Mutation heatmap
```{r fig.height=20, fig.width=17}
hallmark_hypoxia_mut_heatmap <- Heatmap(germ_som_hallmark_hypoxia_mat,
                                        name = "Hallmark Hypoxia Mutations",
                                        top_annotation = ha_row_annotation_top,
                                        show_column_names = FALSE,
                                        column_split = hypoxia_op$Buffa_signature_Sum_within_quartile_one_sample_per_donor,
                                        col = germ_som_colours,
                                        cluster_rows = TRUE)
```

```{r fig.height=20, fig.width=17}
draw(hallmark_hypoxia_mut_heatmap)
```

```{r fig.height=15, fig.width=17}
draw(Heatmap(somatic_hallmark_hypoxia_mat[,rownames(hypoxia_op)[rownames(hypoxia_op) %in% colnames(somatic_hallmark_hypoxia_mat)]],
     name = "Somatic mutations",
     top_annotation = ha_row_annotation_top,
     show_column_names = FALSE,
     column_split = hypoxia_op$Buffa_signature_Sum_within_quartile_one_sample_per_donor,
     col = mut_colours,
     cluster_rows = TRUE))
```

```{r fig.height=15, fig.width=17}
draw(Heatmap(hypoxia_tier_1_2_genes_mat[,rownames(hypoxia_op)[rownames(hypoxia_op) %in% colnames(hypoxia_tier_1_2_genes_mat)]],
     name = "Germline mutations",
     # top_annotation = ha_row_annotation_top,
     show_column_names = FALSE,
     # column_split = hypoxia_op[rownames(hypoxia_op)[rownames(hypoxia_op) %in% colnames(hypoxia_tier_1_2_genes_mat)],"Buffa_signature_Sum_within_quartile_one_sample_per_donor"],
     col = mut_colours,
     cluster_rows = TRUE))
```

# Somatic/Germline combined stats
## Update matrix
Save a copy of the matrix for later plots
```{r}
germ_som_hallmark_hypoxia_mat_2 <- germ_som_hallmark_hypoxia_mat
```
```{r}
germ_som_hallmark_hypoxia_mat[germ_som_hallmark_hypoxia_mat %in% c("Somatic mutation only","Germline mutation only")] <- "Mutation"
```


## Test for differences in scores
```{r}
germ_som_hallmark_hypoxia_mat_hypoxia_scores <- as.data.frame(matrix(nrow = nrow(germ_som_hallmark_hypoxia_mat), 
                                                                    ncol = 5))
colnames(germ_som_hallmark_hypoxia_mat_hypoxia_scores) <- c("Gene","n_mutated","n_wt","wilcox_p_val","wilcox_p_adj")
germ_som_hallmark_hypoxia_mat_hypoxia_scores$Gene <- rownames(germ_som_hallmark_hypoxia_mat)

for(i in 1:nrow(germ_som_hallmark_hypoxia_mat_hypoxia_scores)){
  gene <- germ_som_hallmark_hypoxia_mat_hypoxia_scores$Gene[i]
  test_gene <- as.data.frame(t(germ_som_hallmark_hypoxia_mat)[,gene])
  colnames(test_gene) <-"Test_gene"
  test_gene$sample <- rownames(test_gene)
  
  test_data <- merge(test_gene, hypoxia_signatures_tum_data_single[,c("Buffa_signature_Sum_within","WGS_AssayID")],
                     by.x = "sample", by.y = "WGS_AssayID")
  
  table_data <- as.data.frame(table(test_data$Test_gene))
  if(nrow(table_data) > 1){
    germ_som_hallmark_hypoxia_mat_hypoxia_scores$wilcox_p_val[i] <- wilcox.test(subset(test_data, Test_gene == "Mutation")[,"Buffa_signature_Sum_within"],
                                                         subset(test_data, Test_gene == "NA")[,"Buffa_signature_Sum_within"])[[3]]
  
    germ_som_hallmark_hypoxia_mat_hypoxia_scores$n_mutated[i] <- nrow(subset(test_data, Test_gene == "Mutation"))
    germ_som_hallmark_hypoxia_mat_hypoxia_scores$n_wt[i] <- nrow(subset(test_data, Test_gene == "NA"))
  }
  
}

germ_som_hallmark_hypoxia_mat_hypoxia_scores$wilcox_p_adj <- p.adjust(germ_som_hallmark_hypoxia_mat_hypoxia_scores$wilcox_p_val, method = "BH")
```

```{r}
germ_som_hallmark_hypoxia_mat_hypoxia_scores$Significance <- ifelse(germ_som_hallmark_hypoxia_mat_hypoxia_scores$wilcox_p_adj < 0.05, yes = "Significant", no = "N.S.")
```

```{r}
rownames(germ_som_hallmark_hypoxia_mat_hypoxia_scores) <- germ_som_hallmark_hypoxia_mat_hypoxia_scores$Gene
```

Order data by p-value
```{r}
germ_som_hallmark_hypoxia_mat_hypoxia_scores <- germ_som_hallmark_hypoxia_mat_hypoxia_scores[order(germ_som_hallmark_hypoxia_mat_hypoxia_scores$wilcox_p_val),]
rownames(germ_som_hallmark_hypoxia_mat_hypoxia_scores) <- germ_som_hallmark_hypoxia_mat_hypoxia_scores$Gene
```
```{r}
(germ_som_hallmark_hypoxia_mat_hypoxia_scores)
```

## Q1 Q4 difference
```{r}
germ_som_hallmark_hypoxia_mat_hypoxia_enrichment_quart <- as.data.frame(matrix(nrow = nrow(germ_som_hallmark_hypoxia_mat), ncol = 8))
colnames(germ_som_hallmark_hypoxia_mat_hypoxia_enrichment_quart) <- c("Gene","n_q4_mutated","n_q1_mutated","n_q4_wt","n_q1_wt","fisher_p_val","fisher_p_adj","odds_ratio")
germ_som_hallmark_hypoxia_mat_hypoxia_enrichment_quart$Gene <- rownames(germ_som_hallmark_hypoxia_mat)

for(i in 1:nrow(germ_som_hallmark_hypoxia_mat_hypoxia_enrichment_quart)){
  gene <- germ_som_hallmark_hypoxia_mat_hypoxia_enrichment_quart$Gene[i]
  test_gene <- as.data.frame(t(germ_som_hallmark_hypoxia_mat)[,gene])
  colnames(test_gene) <-"Test_gene"
  test_gene$sample <- rownames(test_gene)
  
  test_data <- merge(test_gene, hypoxia_signatures_tum_data_single[,c("Buffa_signature_Sum_within_quartile_one_sample_per_donor","WGS_AssayID")],
                     by.x = "sample", by.y = "WGS_AssayID")
  test_data <- subset(test_data, Buffa_signature_Sum_within_quartile_one_sample_per_donor %in% c("Q1","Q4"))
  
  table_data <- as.data.frame(table(test_data$Test_gene, test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
  if(nrow(table_data) > 2){
    germ_som_hallmark_hypoxia_mat_hypoxia_enrichment_quart$fisher_p_val[i] <- fisher.test(table(test_data$Test_gene, test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))[[1]]
    germ_som_hallmark_hypoxia_mat_hypoxia_enrichment_quart$odds_ratio[i] <- fisher.test(table(test_data$Test_gene, test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))[[3]]
  
  
  
    germ_som_hallmark_hypoxia_mat_hypoxia_enrichment_quart$n_q4_mutated[i] <- subset(table_data, (Var1 == "Mutation") & (Var2 == "Q4"))[,"Freq"]
    germ_som_hallmark_hypoxia_mat_hypoxia_enrichment_quart$n_q1_mutated[i] <- subset(table_data, (Var1 == "Mutation") & (Var2 == "Q1"))[,"Freq"]
    germ_som_hallmark_hypoxia_mat_hypoxia_enrichment_quart$n_q4_wt[i] <- subset(table_data, (Var1 == "NA") & (Var2 == "Q4"))[,"Freq"]
    germ_som_hallmark_hypoxia_mat_hypoxia_enrichment_quart$n_q1_wt[i] <- subset(table_data, (Var1 == "NA") & (Var2 == "Q1"))[,"Freq"]
  }
  
}

germ_som_hallmark_hypoxia_mat_hypoxia_enrichment_quart$fisher_p_adj <- p.adjust(germ_som_hallmark_hypoxia_mat_hypoxia_enrichment_quart$fisher_p_val, method = "BH")
```

```{r}
# subset(germ_som_hallmark_hypoxia_mat_hypoxia_enrichment_quart, fisher_p_adj < 0.05)
```

```{r}
germ_som_hallmark_hypoxia_mat_hypoxia_enrichment_quart$Significance <- ifelse(germ_som_hallmark_hypoxia_mat_hypoxia_enrichment_quart$fisher_p_adj < 0.05, yes = "Significant", no = "N.S.")
```

Order data by p-value
```{r}
germ_som_hallmark_hypoxia_mat_hypoxia_enrichment_quart <- germ_som_hallmark_hypoxia_mat_hypoxia_enrichment_quart[order(germ_som_hallmark_hypoxia_mat_hypoxia_enrichment_quart$fisher_p_val),]
rownames(germ_som_hallmark_hypoxia_mat_hypoxia_enrichment_quart) <- germ_som_hallmark_hypoxia_mat_hypoxia_enrichment_quart$Gene
```
```{r}
(germ_som_hallmark_hypoxia_mat_hypoxia_enrichment_quart)
```

## Check mutation in group vs not
Check if mutation in any associated gene has differences in hypoxia score
```{r}
germ_som_hallmark_hypoxia_mat_bin <- germ_som_hallmark_hypoxia_mat
germ_som_hallmark_hypoxia_mat_bin[germ_som_hallmark_hypoxia_mat_bin == "Mutation"] <- as.numeric(1)
germ_som_hallmark_hypoxia_mat_bin[germ_som_hallmark_hypoxia_mat_bin == "NA"] <- as.numeric(0)
```

```{r}
class(germ_som_hallmark_hypoxia_mat_bin) <- "numeric"
```

```{r}
summary_germ_som_hallmark_hypoxia <- as.data.frame(colSums(germ_som_hallmark_hypoxia_mat_bin))
colnames(summary_germ_som_hallmark_hypoxia) <- c("Muts_in_hallmark_hypoxia")
summary_germ_som_hallmark_hypoxia$WGS_AssayID <- rownames(summary_germ_som_hallmark_hypoxia)
summary_germ_som_hallmark_hypoxia$bin_muts_in_hallmark_hypoxia <- ifelse(summary_germ_som_hallmark_hypoxia$Muts_in_hallmark_hypoxia == 0, yes = "NA", no = "Mutation")
```

```{r}
summary_germ_som_hallmark_hypoxia <- merge(summary_germ_som_hallmark_hypoxia, 
                               hypoxia_signatures_tum_data_single[,c("Buffa_signature_Sum_within","WGS_AssayID",
                                             "Buffa_signature_Sum_within_quartile_one_sample_per_donor")],
                               by = "WGS_AssayID")
```

```{r}
summary_germ_som_hallmark_hypoxia_quart <- subset(summary_germ_som_hallmark_hypoxia, Buffa_signature_Sum_within_quartile_one_sample_per_donor %in% c("Q1","Q4"))
```

```{r}
fisher.test(table(summary_germ_som_hallmark_hypoxia_quart$bin_muts_in_hallmark_hypoxia,
                  summary_germ_som_hallmark_hypoxia_quart$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
```
```{r}
table(summary_germ_som_hallmark_hypoxia_quart$bin_muts_in_hallmark_hypoxia, 
      summary_germ_som_hallmark_hypoxia_quart$Buffa_signature_Sum_within_quartile_one_sample_per_donor)
```

```{r}
wilcox.test(subset(summary_germ_som_hallmark_hypoxia, bin_muts_in_hallmark_hypoxia == "Mutation")[,"Buffa_signature_Sum_within"], 
            subset(summary_germ_som_hallmark_hypoxia, bin_muts_in_hallmark_hypoxia == "NA")[,"Buffa_signature_Sum_within"])
```

## any germline or any somatic in set
```{r}
som_hallmark_hypoxia_muts <- as.data.frame(colSums(somatic_hallmark_hypoxia_mat_bin))
colnames(som_hallmark_hypoxia_muts) <- c("Somatic_mutation")
som_hallmark_hypoxia_muts$WGS_AssayID <- rownames(som_hallmark_hypoxia_muts)

ger_hallmark_hypoxia_muts <- hypoxia_tier_1_2_lists[,c("WGS_AssayID","HALLMARK_HYPOXIA")]

hallmark_hypoxia_muts <- merge(som_hallmark_hypoxia_muts,
                               ger_hallmark_hypoxia_muts,
                               by = "WGS_AssayID", all = TRUE)

rownames(hallmark_hypoxia_muts) <- hallmark_hypoxia_muts$WGS_AssayID

summary(hallmark_hypoxia_muts)

hallmark_hypoxia_muts$Somatic_mutation <- ifelse(hallmark_hypoxia_muts$Somatic_mutation > 0, yes = "Mutation", no = "No Mutation")
hallmark_hypoxia_muts$HALLMARK_HYPOXIA[hallmark_hypoxia_muts$HALLMARK_HYPOXIA == 1] <- "Mutation"
hallmark_hypoxia_muts$HALLMARK_HYPOXIA[hallmark_hypoxia_muts$HALLMARK_HYPOXIA == 0] <- "No Mutation"

colnames(hallmark_hypoxia_muts) <- c( "WGS_AssayID", "Somatic_mutation", "Germline_mutation")
```

```{r}
dim(hallmark_hypoxia_muts)
```

```{r}
hallmark_hypoxia_muts_t <- t(hallmark_hypoxia_muts)
# arrange by hypoxia score
hallmark_hypoxia_muts_t <- hallmark_hypoxia_muts_t[,rownames(hypoxia_op)[rownames(hypoxia_op) %in% colnames(hallmark_hypoxia_muts_t)]]
```


# Combine somatic and germline DRG36 variants
## Somatic DRG36 variants
```{r}
somatic_drg36_mat <- as.matrix(driver_data[rownames(driver_data) %in% drg_genes_36,])
somatic_drg36_mat = somatic_drg36_mat[, -c(1,2)]
somatic_drg36_mat <- gsub("\\.","", somatic_drg36_mat)
```

```{r}
dim(somatic_drg36_mat)
```

```{r}
somatic_drg36_mat[somatic_drg36_mat == "nsSNV"] <- "Mutation"
somatic_drg36_mat[somatic_drg36_mat == "splicing"] <- "Mutation"
somatic_drg36_mat[somatic_drg36_mat == "stopgain"] <- "Mutation"
somatic_drg36_mat[somatic_drg36_mat == "stoploss"] <- "Mutation"
somatic_drg36_mat[somatic_drg36_mat == "startloss"] <- "Mutation"
somatic_drg36_mat[somatic_drg36_mat == "stopgain+nsSNV"] <- "Mutation"
somatic_drg36_mat[somatic_drg36_mat == "nsSNV+splicing"] <- "Mutation"
somatic_drg36_mat[somatic_drg36_mat == "nsSNV+stopgain"] <- "Mutation"
somatic_drg36_mat[somatic_drg36_mat == "splicing+nsSNV"] <- "Mutation"
somatic_drg36_mat[somatic_drg36_mat == "startloss+nsSNV"] <- "Mutation"

somatic_drg36_mat[somatic_drg36_mat == ""] <- "No Mutation"
```

```{r}
dim(somatic_drg36_mat)
```

```{r}
table(somatic_drg36_mat)
```

## Germline DRG36 variants
```{r}
drg36_tier_1_2_genes_mat <- t(as.matrix(hypoxia_tier_1_2_genes[,colnames(hypoxia_tier_1_2_genes) %in% drg_genes_36]))
```
```{r}
drg36_tier_1_2_genes_mat[drg36_tier_1_2_genes_mat == "1"] <- "Mutation"

drg36_tier_1_2_genes_mat[drg36_tier_1_2_genes_mat == "0"] <- "No Mutation"
```

## Combine
```{r}
drg36_tier_1_2_genes_mat[1:4,1:4]
```

```{r}
somatic_drg36_mat[1:4,1:4]
```

Combine somatic and germline variants so we can just plot as one heatmap then check any variant?
```{r}
germ_som_drg36_mat <- matrix(nrow = length(drg_genes_36), ncol = ncol(somatic_drg36_mat))
rownames(germ_som_drg36_mat) <- drg_genes_36
colnames(germ_som_drg36_mat) <- colnames(somatic_drg36_mat)
```

```{r}
for(i in 1:length(drg_genes_36)){
  
  for(j in 1:ncol(somatic_drg36_mat)){
  somatic_gene <- "NA"
  somatic_sample <- "NA"
  somatic_status <- "NA"
  
  germline_gene <- "NA"
  germline_sample <- "NA"
  germline_status <- "NA"
  
  combined_label <- "NA"
    
    if(drg_genes_36[i] %in% rownames(somatic_drg36_mat)){
      somatic_gene <- drg_genes_36[i]
      if(colnames(germ_som_drg36_mat)[j] %in% colnames(somatic_drg36_mat)){
        somatic_sample <- colnames(germ_som_drg36_mat)[j]
        somatic_status <- somatic_drg36_mat[somatic_gene,somatic_sample]
      }
    }
    if(drg_genes_36[i] %in% rownames(drg36_tier_1_2_genes_mat)){
      germline_gene <- drg_genes_36[i]
      if(colnames(germ_som_drg36_mat)[j] %in% colnames(drg36_tier_1_2_genes_mat)){
        germline_sample <- colnames(germ_som_drg36_mat)[j]
        germline_status <- drg36_tier_1_2_genes_mat[germline_gene,germline_sample]
      }
    }
    
    # Make class
    ## Not found in either
    if(all(c(somatic_status == "NA", germline_status == "NA"))){
      combined_label <- "NA"
    }else if(all(c(somatic_status == "Mutation", germline_status == "NA"))){## Somatic mutation not found in germline
      combined_label <- "Somatic mutation only"
    }else if(all(c(somatic_status == "NA", germline_status == "Mutation"))){## Germline mutation not found in somatic
      combined_label <- "Germline mutation only"
    }else if(all(c(somatic_status == "Mutation", germline_status == "Mutation"))){## Mutation in both
      combined_label <- "Both"
    }else if(all(c(somatic_status == "No Mutation", germline_status == "No Mutation"))){## Mutation in neither
      combined_label <- "NA"
    }else if(all(c(somatic_status == "Mutation", germline_status == "No Mutation"))){## Somatic mutation not germline
      combined_label <- "Somatic mutation only"
    }else if(all(c(somatic_status == "No Mutation", germline_status == "Mutation"))){## Germline mutation not somatic
      combined_label <- "Germline mutation only"
    }
    
    
    # Add label to new matrix
    germ_som_drg36_mat[i,j] <- combined_label
  }
  
}
```

```{r}
# order driver samples by hypoxia scores
germ_som_drg36_mat <- germ_som_drg36_mat[,rownames(hypoxia_op)[rownames(hypoxia_op) %in% colnames(germ_som_drg36_mat)]]

```

Remove rows that are all NA. (only include rows where sum of is.na is not equal to the number of columns)
```{r}
germ_som_drg36_mat <- germ_som_drg36_mat[(rowSums((germ_som_drg36_mat == "NA")) != ncol(germ_som_drg36_mat)),]
```

## Heatmap
```{r fig.height=20, fig.width=17}
drg36_mut_heatmap <- Heatmap(germ_som_drg36_mat,
                                        name = "DRG36 Mutations",
                                        top_annotation = ha_row_annotation_top,
                                        show_column_names = FALSE,
                                        column_split = hypoxia_op$Buffa_signature_Sum_within_quartile_one_sample_per_donor,
                                        col = germ_som_colours,
                                        cluster_rows = TRUE)
```

```{r fig.height=20, fig.width=17}
draw(drg36_mut_heatmap)
```

## Stats
### Update matrix
```{r}
germ_som_drg36_mat[germ_som_drg36_mat %in% c("Somatic mutation only","Germline mutation only","Both")] <- "Mutation"
```


### Test for differences in scores
```{r}
germ_som_drg36_mat_hypoxia_scores <- as.data.frame(matrix(nrow = nrow(germ_som_drg36_mat), 
                                                                    ncol = 5))
colnames(germ_som_drg36_mat_hypoxia_scores) <- c("Gene","n_mutated","n_wt","wilcox_p_val","wilcox_p_adj")
germ_som_drg36_mat_hypoxia_scores$Gene <- rownames(germ_som_drg36_mat)

for(i in 1:nrow(germ_som_drg36_mat_hypoxia_scores)){
  gene <- germ_som_drg36_mat_hypoxia_scores$Gene[i]
  test_gene <- as.data.frame(t(germ_som_drg36_mat)[,gene])
  colnames(test_gene) <-"Test_gene"
  test_gene$sample <- rownames(test_gene)
  
  test_data <- merge(test_gene, hypoxia_signatures_tum_data_single[,c("Buffa_signature_Sum_within","WGS_AssayID")],
                     by.x = "sample", by.y = "WGS_AssayID")
  
  table_data <- as.data.frame(table(test_data$Test_gene))
  if(nrow(table_data) > 1){
    germ_som_drg36_mat_hypoxia_scores$wilcox_p_val[i] <- wilcox.test(subset(test_data, Test_gene == "Mutation")[,"Buffa_signature_Sum_within"],
                                                         subset(test_data, Test_gene == "NA")[,"Buffa_signature_Sum_within"])[[3]]
  
    germ_som_drg36_mat_hypoxia_scores$n_mutated[i] <- nrow(subset(test_data, Test_gene == "Mutation"))
    germ_som_drg36_mat_hypoxia_scores$n_wt[i] <- nrow(subset(test_data, Test_gene == "NA"))
  }
  
}

germ_som_drg36_mat_hypoxia_scores$wilcox_p_adj <- p.adjust(germ_som_drg36_mat_hypoxia_scores$wilcox_p_val, method = "BH")
```

```{r}
germ_som_drg36_mat_hypoxia_scores$Significance <- ifelse(germ_som_drg36_mat_hypoxia_scores$wilcox_p_adj < 0.05, yes = "Significant", no = "N.S.")
```

```{r}
rownames(germ_som_drg36_mat_hypoxia_scores) <- germ_som_drg36_mat_hypoxia_scores$Gene
```

Order data by p-value
```{r}
germ_som_drg36_mat_hypoxia_scores <- germ_som_drg36_mat_hypoxia_scores[order(germ_som_drg36_mat_hypoxia_scores$wilcox_p_val),]
rownames(germ_som_drg36_mat_hypoxia_scores) <- germ_som_drg36_mat_hypoxia_scores$Gene
```
```{r}
(germ_som_drg36_mat_hypoxia_scores)
```

### Q1 Q4 difference
```{r}
germ_som_drg36_mat_hypoxia_enrichment_quart <- as.data.frame(matrix(nrow = nrow(germ_som_drg36_mat), ncol = 8))
colnames(germ_som_drg36_mat_hypoxia_enrichment_quart) <- c("Gene","n_q4_mutated","n_q1_mutated","n_q4_wt","n_q1_wt","fisher_p_val","fisher_p_adj","odds_ratio")
germ_som_drg36_mat_hypoxia_enrichment_quart$Gene <- rownames(germ_som_drg36_mat)

for(i in 1:nrow(germ_som_drg36_mat_hypoxia_enrichment_quart)){
  gene <- germ_som_drg36_mat_hypoxia_enrichment_quart$Gene[i]
  test_gene <- as.data.frame(t(germ_som_drg36_mat)[,gene])
  colnames(test_gene) <-"Test_gene"
  test_gene$sample <- rownames(test_gene)
  
  test_data <- merge(test_gene, hypoxia_signatures_tum_data_single[,c("Buffa_signature_Sum_within_quartile_one_sample_per_donor","WGS_AssayID")],
                     by.x = "sample", by.y = "WGS_AssayID")
  test_data <- subset(test_data, Buffa_signature_Sum_within_quartile_one_sample_per_donor %in% c("Q1","Q4"))
  
  table_data <- as.data.frame(table(test_data$Test_gene, test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
  if(nrow(table_data) > 2){
    germ_som_drg36_mat_hypoxia_enrichment_quart$fisher_p_val[i] <- fisher.test(table(test_data$Test_gene, test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))[[1]]
    germ_som_drg36_mat_hypoxia_enrichment_quart$odds_ratio[i] <- fisher.test(table(test_data$Test_gene, test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))[[3]]
  
  
  
    germ_som_drg36_mat_hypoxia_enrichment_quart$n_q4_mutated[i] <- subset(table_data, (Var1 == "Mutation") & (Var2 == "Q4"))[,"Freq"]
    germ_som_drg36_mat_hypoxia_enrichment_quart$n_q1_mutated[i] <- subset(table_data, (Var1 == "Mutation") & (Var2 == "Q1"))[,"Freq"]
    germ_som_drg36_mat_hypoxia_enrichment_quart$n_q4_wt[i] <- subset(table_data, (Var1 == "NA") & (Var2 == "Q4"))[,"Freq"]
    germ_som_drg36_mat_hypoxia_enrichment_quart$n_q1_wt[i] <- subset(table_data, (Var1 == "NA") & (Var2 == "Q1"))[,"Freq"]
  }
  
}

germ_som_drg36_mat_hypoxia_enrichment_quart$fisher_p_adj <- p.adjust(germ_som_drg36_mat_hypoxia_enrichment_quart$fisher_p_val, method = "BH")
```

```{r}
# subset(germ_som_drg36_mat_hypoxia_enrichment_quart, fisher_p_adj < 0.05)
```

```{r}
germ_som_drg36_mat_hypoxia_enrichment_quart$Significance <- ifelse(germ_som_drg36_mat_hypoxia_enrichment_quart$fisher_p_adj < 0.05, yes = "Significant", no = "N.S.")
```

Order data by p-value
```{r}
germ_som_drg36_mat_hypoxia_enrichment_quart <- germ_som_drg36_mat_hypoxia_enrichment_quart[order(germ_som_drg36_mat_hypoxia_enrichment_quart$fisher_p_val),]
rownames(germ_som_drg36_mat_hypoxia_enrichment_quart) <- germ_som_drg36_mat_hypoxia_enrichment_quart$Gene
```
```{r}
(germ_som_drg36_mat_hypoxia_enrichment_quart)
```

### Check mutation in group vs not
Check if mutation in any associated gene has differences in hypoxia score
```{r}
germ_som_drg36_mat_bin <- germ_som_drg36_mat
germ_som_drg36_mat_bin[germ_som_drg36_mat_bin == "Mutation"] <- as.numeric(1)
germ_som_drg36_mat_bin[germ_som_drg36_mat_bin == "NA"] <- as.numeric(0)
```

```{r}
class(germ_som_drg36_mat_bin) <- "numeric"
```

```{r}
summary_germ_som_drg36 <- as.data.frame(colSums(germ_som_drg36_mat_bin))
colnames(summary_germ_som_drg36) <- c("Muts_in_drg36")
summary_germ_som_drg36$WGS_AssayID <- rownames(summary_germ_som_drg36)
summary_germ_som_drg36$bin_muts_in_drg36 <- ifelse(summary_germ_som_drg36$Muts_in_drg36 == 0, yes = "NA", no = "Mutation")
```

```{r}
summary_germ_som_drg36 <- merge(summary_germ_som_drg36, 
                               hypoxia_signatures_tum_data_single[,c("Buffa_signature_Sum_within","WGS_AssayID",
                                             "Buffa_signature_Sum_within_quartile_one_sample_per_donor","PPCG_Donor_ID")],
                               by = "WGS_AssayID")
```

```{r}
summary_germ_som_drg36_quart <- subset(summary_germ_som_drg36, Buffa_signature_Sum_within_quartile_one_sample_per_donor %in% c("Q1","Q4"))
```

```{r}
fisher.test(table(summary_germ_som_drg36_quart$bin_muts_in_drg36,
                  summary_germ_som_drg36_quart$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
```
```{r}
table(summary_germ_som_drg36_quart$bin_muts_in_drg36, 
      summary_germ_som_drg36_quart$Buffa_signature_Sum_within_quartile_one_sample_per_donor)
```

```{r}
wilcox.test(subset(summary_germ_som_drg36_quart, bin_muts_in_drg36 == "Mutation")[,"Buffa_signature_Sum_within"], 
            subset(summary_germ_som_drg36_quart, bin_muts_in_drg36 == "NA")[,"Buffa_signature_Sum_within"])
```

# Combine somatic and germline DRG128 variants
## Somatic DRG128 variants
```{r}
somatic_drg128_mat <- as.matrix(driver_data[rownames(driver_data) %in% drg_genes_128,])
somatic_drg128_mat = somatic_drg128_mat[, -c(1,2)]
somatic_drg128_mat <- gsub("\\.","", somatic_drg128_mat)
```

```{r}
dim(somatic_drg128_mat)
```

```{r}
somatic_drg128_mat[somatic_drg128_mat == "nsSNV"] <- "Mutation"
somatic_drg128_mat[somatic_drg128_mat == "splicing"] <- "Mutation"
somatic_drg128_mat[somatic_drg128_mat == "stopgain"] <- "Mutation"
somatic_drg128_mat[somatic_drg128_mat == "stoploss"] <- "Mutation"
somatic_drg128_mat[somatic_drg128_mat == "startloss"] <- "Mutation"
somatic_drg128_mat[somatic_drg128_mat == "stopgain+nsSNV"] <- "Mutation"
somatic_drg128_mat[somatic_drg128_mat == "nsSNV+splicing"] <- "Mutation"
somatic_drg128_mat[somatic_drg128_mat == "nsSNV+stopgain"] <- "Mutation"
somatic_drg128_mat[somatic_drg128_mat == "splicing+nsSNV"] <- "Mutation"
somatic_drg128_mat[somatic_drg128_mat == "startloss+nsSNV"] <- "Mutation"

somatic_drg128_mat[somatic_drg128_mat == ""] <- "No Mutation"
```

```{r}
dim(somatic_drg128_mat)
```

```{r}
table(somatic_drg128_mat)
```

## Germline drg128 variants
```{r}
drg128_tier_1_2_genes_mat <- t(as.matrix(hypoxia_tier_1_2_genes[,colnames(hypoxia_tier_1_2_genes) %in% drg_genes_128]))
```
```{r}
drg128_tier_1_2_genes_mat[drg128_tier_1_2_genes_mat == "1"] <- "Mutation"

drg128_tier_1_2_genes_mat[drg128_tier_1_2_genes_mat == "0"] <- "No Mutation"
```

## Combine
```{r}
drg128_tier_1_2_genes_mat[1:4,1:4]
```

```{r}
somatic_drg128_mat[1:4,1:4]
```

Combine somatic and germline variants so we can just plot as one heatmap then check any variant?
```{r}
germ_som_drg128_mat <- matrix(nrow = length(drg_genes_128), ncol = ncol(somatic_drg128_mat))
rownames(germ_som_drg128_mat) <- drg_genes_128
colnames(germ_som_drg128_mat) <- colnames(somatic_drg128_mat)
```

```{r}
for(i in 1:length(drg_genes_128)){
  
  for(j in 1:ncol(somatic_drg128_mat)){
  somatic_gene <- "NA"
  somatic_sample <- "NA"
  somatic_status <- "NA"
  
  germline_gene <- "NA"
  germline_sample <- "NA"
  germline_status <- "NA"
  
  combined_label <- "NA"
    
    if(drg_genes_128[i] %in% rownames(somatic_drg128_mat)){
      somatic_gene <- drg_genes_128[i]
      if(colnames(germ_som_drg128_mat)[j] %in% colnames(somatic_drg128_mat)){
        somatic_sample <- colnames(germ_som_drg128_mat)[j]
        somatic_status <- somatic_drg128_mat[somatic_gene,somatic_sample]
      }
    }
    if(drg_genes_128[i] %in% rownames(drg128_tier_1_2_genes_mat)){
      germline_gene <- drg_genes_128[i]
      if(colnames(germ_som_drg128_mat)[j] %in% colnames(drg128_tier_1_2_genes_mat)){
        germline_sample <- colnames(germ_som_drg128_mat)[j]
        germline_status <- drg128_tier_1_2_genes_mat[germline_gene,germline_sample]
      }
    }
    
    # Make class
    ## Not found in either
    if(all(c(somatic_status == "NA", germline_status == "NA"))){
      combined_label <- "NA"
    }else if(all(c(somatic_status == "Mutation", germline_status == "NA"))){## Somatic mutation not found in germline
      combined_label <- "Somatic mutation only"
    }else if(all(c(somatic_status == "NA", germline_status == "Mutation"))){## Germline mutation not found in somatic
      combined_label <- "Germline mutation only"
    }else if(all(c(somatic_status == "Mutation", germline_status == "Mutation"))){## Mutation in both
      combined_label <- "Both"
    }else if(all(c(somatic_status == "No Mutation", germline_status == "No Mutation"))){## Mutation in neither
      combined_label <- "NA"
    }else if(all(c(somatic_status == "Mutation", germline_status == "No Mutation"))){## Somatic mutation not germline
      combined_label <- "Somatic mutation only"
    }else if(all(c(somatic_status == "No Mutation", germline_status == "Mutation"))){## Germline mutation not somatic
      combined_label <- "Germline mutation only"
    }
    
    
    # Add label to new matrix
    germ_som_drg128_mat[i,j] <- combined_label
  }
  
}
```

```{r}
# order driver samples by hypoxia scores
germ_som_drg128_mat <- germ_som_drg128_mat[,rownames(hypoxia_op)[rownames(hypoxia_op) %in% colnames(germ_som_drg128_mat)]]

```

Remove rows that are all NA. (only include rows where sum of is.na is not equal to the number of columns)
```{r}
germ_som_drg128_mat <- germ_som_drg128_mat[(rowSums((germ_som_drg128_mat == "NA")) != ncol(germ_som_drg128_mat)),]
```

## Heatmap
```{r fig.height=20, fig.width=17}
drg128_mut_heatmap <- Heatmap(germ_som_drg128_mat,
                                        name = "DRG128 Mutations",
                                        top_annotation = ha_row_annotation_top,
                                        show_column_names = FALSE,
                                        column_split = hypoxia_op$Buffa_signature_Sum_within_quartile_one_sample_per_donor,
                                        col = germ_som_colours,
                                        cluster_rows = TRUE)
```

```{r fig.height=20, fig.width=17}
draw(drg128_mut_heatmap)
```

## Stats
### Update matrix
Save copy of matrix for later plots
```{r}
germ_som_drg128_mat_2 <- germ_som_drg128_mat
```
```{r}
germ_som_drg128_mat[germ_som_drg128_mat %in% c("Somatic mutation only","Germline mutation only","Both")] <- "Mutation"
```


### Test for differences in scores
```{r}
germ_som_drg128_mat_hypoxia_scores <- as.data.frame(matrix(nrow = nrow(germ_som_drg128_mat), 
                                                                    ncol = 5))
colnames(germ_som_drg128_mat_hypoxia_scores) <- c("Gene","n_mutated","n_wt","wilcox_p_val","wilcox_p_adj")
germ_som_drg128_mat_hypoxia_scores$Gene <- rownames(germ_som_drg128_mat)

for(i in 1:nrow(germ_som_drg128_mat_hypoxia_scores)){
  gene <- germ_som_drg128_mat_hypoxia_scores$Gene[i]
  test_gene <- as.data.frame(t(germ_som_drg128_mat)[,gene])
  colnames(test_gene) <-"Test_gene"
  test_gene$sample <- rownames(test_gene)
  
  test_data <- merge(test_gene, hypoxia_signatures_tum_data_single[,c("Buffa_signature_Sum_within","WGS_AssayID")],
                     by.x = "sample", by.y = "WGS_AssayID")
  
  table_data <- as.data.frame(table(test_data$Test_gene))
  if(nrow(table_data) > 1){
    germ_som_drg128_mat_hypoxia_scores$wilcox_p_val[i] <- wilcox.test(subset(test_data, Test_gene == "Mutation")[,"Buffa_signature_Sum_within"],
                                                         subset(test_data, Test_gene == "NA")[,"Buffa_signature_Sum_within"])[[3]]
  
    germ_som_drg128_mat_hypoxia_scores$n_mutated[i] <- nrow(subset(test_data, Test_gene == "Mutation"))
    germ_som_drg128_mat_hypoxia_scores$n_wt[i] <- nrow(subset(test_data, Test_gene == "NA"))
  }
  
}

germ_som_drg128_mat_hypoxia_scores$wilcox_p_adj <- p.adjust(germ_som_drg128_mat_hypoxia_scores$wilcox_p_val, method = "BH")
```

```{r}
germ_som_drg128_mat_hypoxia_scores$Significance <- ifelse(germ_som_drg128_mat_hypoxia_scores$wilcox_p_adj < 0.05, yes = "Significant", no = "N.S.")
```

```{r}
rownames(germ_som_drg128_mat_hypoxia_scores) <- germ_som_drg128_mat_hypoxia_scores$Gene
```

Order data by p-value
```{r}
germ_som_drg128_mat_hypoxia_scores <- germ_som_drg128_mat_hypoxia_scores[order(germ_som_drg128_mat_hypoxia_scores$wilcox_p_val),]
rownames(germ_som_drg128_mat_hypoxia_scores) <- germ_som_drg128_mat_hypoxia_scores$Gene
```
```{r}
(germ_som_drg128_mat_hypoxia_scores)
```

### Q1 Q4 difference
```{r}
germ_som_drg128_mat_hypoxia_enrichment_quart <- as.data.frame(matrix(nrow = nrow(germ_som_drg128_mat), ncol = 8))
colnames(germ_som_drg128_mat_hypoxia_enrichment_quart) <- c("Gene","n_q4_mutated","n_q1_mutated","n_q4_wt","n_q1_wt","fisher_p_val","fisher_p_adj","odds_ratio")
germ_som_drg128_mat_hypoxia_enrichment_quart$Gene <- rownames(germ_som_drg128_mat)

for(i in 1:nrow(germ_som_drg128_mat_hypoxia_enrichment_quart)){
  gene <- germ_som_drg128_mat_hypoxia_enrichment_quart$Gene[i]
  test_gene <- as.data.frame(t(germ_som_drg128_mat)[,gene])
  colnames(test_gene) <-"Test_gene"
  test_gene$sample <- rownames(test_gene)
  
  test_data <- merge(test_gene, hypoxia_signatures_tum_data_single[,c("Buffa_signature_Sum_within_quartile_one_sample_per_donor",
                                                                      "WGS_AssayID")],
                     by.x = "sample", by.y = "WGS_AssayID")
  test_data <- subset(test_data, Buffa_signature_Sum_within_quartile_one_sample_per_donor %in% c("Q1","Q4"))
  
  table_data <- as.data.frame(table(test_data$Test_gene, test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
  if(nrow(table_data) > 2){
    germ_som_drg128_mat_hypoxia_enrichment_quart$fisher_p_val[i] <- fisher.test(table(test_data$Test_gene, test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))[[1]]
    germ_som_drg128_mat_hypoxia_enrichment_quart$odds_ratio[i] <- fisher.test(table(test_data$Test_gene, test_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))[[3]]
  
  
  
    germ_som_drg128_mat_hypoxia_enrichment_quart$n_q4_mutated[i] <- subset(table_data, (Var1 == "Mutation") & (Var2 == "Q4"))[,"Freq"]
    germ_som_drg128_mat_hypoxia_enrichment_quart$n_q1_mutated[i] <- subset(table_data, (Var1 == "Mutation") & (Var2 == "Q1"))[,"Freq"]
    germ_som_drg128_mat_hypoxia_enrichment_quart$n_q4_wt[i] <- subset(table_data, (Var1 == "NA") & (Var2 == "Q4"))[,"Freq"]
    germ_som_drg128_mat_hypoxia_enrichment_quart$n_q1_wt[i] <- subset(table_data, (Var1 == "NA") & (Var2 == "Q1"))[,"Freq"]
  }
  
}

germ_som_drg128_mat_hypoxia_enrichment_quart$fisher_p_adj <- p.adjust(germ_som_drg128_mat_hypoxia_enrichment_quart$fisher_p_val, method = "BH")
```

```{r}
# subset(germ_som_drg128_mat_hypoxia_enrichment_quart, fisher_p_adj < 0.05)
```

```{r}
germ_som_drg128_mat_hypoxia_enrichment_quart$Significance <- ifelse(germ_som_drg128_mat_hypoxia_enrichment_quart$fisher_p_adj < 0.05, yes = "Significant", no = "N.S.")
```

Order data by p-value
```{r}
germ_som_drg128_mat_hypoxia_enrichment_quart <- germ_som_drg128_mat_hypoxia_enrichment_quart[order(germ_som_drg128_mat_hypoxia_enrichment_quart$fisher_p_val),]
rownames(germ_som_drg128_mat_hypoxia_enrichment_quart) <- germ_som_drg128_mat_hypoxia_enrichment_quart$Gene
```
```{r}
(germ_som_drg128_mat_hypoxia_enrichment_quart)
```

### Check mutation in group vs not
Check if mutation in any associated gene has differences in hypoxia score
```{r}
germ_som_drg128_mat_bin <- germ_som_drg128_mat
germ_som_drg128_mat_bin[germ_som_drg128_mat_bin == "Mutation"] <- as.numeric(1)
germ_som_drg128_mat_bin[germ_som_drg128_mat_bin == "NA"] <- as.numeric(0)
```

```{r}
class(germ_som_drg128_mat_bin) <- "numeric"
```

```{r}
summary_germ_som_drg128 <- as.data.frame(colSums(germ_som_drg128_mat_bin))
colnames(summary_germ_som_drg128) <- c("Muts_in_drg128")
summary_germ_som_drg128$WGS_AssayID <- rownames(summary_germ_som_drg128)
summary_germ_som_drg128$bin_muts_in_drg128 <- ifelse(summary_germ_som_drg128$Muts_in_drg128 == 0, yes = "NA", no = "Mutation")
```

```{r}
summary_germ_som_drg128 <- merge(summary_germ_som_drg128, 
                               hypoxia_signatures_tum_data_single[,c("Buffa_signature_Sum_within","WGS_AssayID",
                                             "Buffa_signature_Sum_within_quartile_one_sample_per_donor")],
                               by = "WGS_AssayID")
```

```{r}
summary_germ_som_drg128_quart <- subset(summary_germ_som_drg128, Buffa_signature_Sum_within_quartile_one_sample_per_donor %in% c("Q1","Q4"))
```

```{r}
fisher.test(table(summary_germ_som_drg128_quart$bin_muts_in_drg128,
                  summary_germ_som_drg128_quart$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
```
```{r}
table(summary_germ_som_drg128_quart$bin_muts_in_drg128, 
      summary_germ_som_drg128_quart$Buffa_signature_Sum_within_quartile_one_sample_per_donor)
```

```{r}
wilcox.test(subset(summary_germ_som_drg128_quart, bin_muts_in_drg128 == "Mutation")[,"Buffa_signature_Sum_within"], 
            subset(summary_germ_som_drg128_quart, bin_muts_in_drg128 == "NA")[,"Buffa_signature_Sum_within"])
```


# Heatmaps with stats
## DRG 128
```{r}
gene_group <- as.data.frame(drg_genes_128)
colnames(gene_group) <- c("Gene")
gene_group$Gene_list_group <- ifelse(gene_group$Gene %in% drg_genes_36, yes = "DRG36 + DRG128", no = "DRG128 Only")
rownames(gene_group) <- gene_group$Gene
```

Colours
```{r}
drg_group_colours <- c("DRG36 + DRG128" = "palegreen4", "DRG128 Only" = "mediumaquamarine")
```

Make a table with relevant stats
```{r}
colnames(germ_som_drg128_mat_hypoxia_enrichment_quart) <- paste0(colnames(germ_som_drg128_mat_hypoxia_enrichment_quart),"_g_s_drg128")
colnames(germline_drg128_enrichment_t1_2) <- paste0(colnames(germline_drg128_enrichment_t1_2),"_g_drg128")
# colnames(somatic_drg128_mat_hypoxia_enrichment_quart) <- paste0(colnames(somatic_drg128_mat_hypoxia_enrichment_quart),"_s_drg128")
drg_heatmap_stats <- merge(germ_som_drg128_mat_hypoxia_enrichment_quart, 
                          germline_drg128_enrichment_t1_2, 
                          by.x = "Gene_g_s_drg128", 
                          by.y = "Gene_g_drg128")
drg_heatmap_stats <- merge(drg_heatmap_stats,
                           germline_drg128_wilcox_t1_2,
                           by.x = "Gene_g_s_drg128",
                           by.y = "Gene")

drg_heatmap_stats$g_s_drg128_significance <- ifelse(drg_heatmap_stats$fisher_p_adj_g_s_drg128 < 0.05, yes = "Significant", no = "N.S.")
drg_heatmap_stats$g_drg128_significance <- ifelse(drg_heatmap_stats$fisher_p_adj_g_drg128 < 0.05, yes = "Significant", no = "N.S.")
drg_heatmap_stats$g_drg128_wilcox_significance <- ifelse(drg_heatmap_stats$wilcox_p_adj < 0.05, yes = "Significant", no = "N.S.")

rownames(drg_heatmap_stats) <- drg_heatmap_stats$Gene_g_s_drg128
```

```{r}
coefficient_colours_drg128_cont <- colorRamp2(c(min(c(drg128_continuous_coefficients$Hypoxia_coefficient,
                                                  hallmark_hypoxia_continuous_coefficients$Hypoxia_coefficient), na.rm = TRUE),
                                    0,
                                    max(c(drg128_continuous_coefficients$Hypoxia_coefficient,
                                          hallmark_hypoxia_continuous_coefficients$Hypoxia_coefficient), na.rm = TRUE)),
                                  c("grey20","white","firebrick"))
coefficient_colours_drg128_quart <- colorRamp2(c(min(c(drg128_quartile_coefficients$Hypoxia_coefficient,
                                                  hallmark_hypoxia_quartile_coefficients$Hypoxia_coefficient), na.rm = TRUE),
                                    0,
                                    max(c(drg128_quartile_coefficients$Hypoxia_coefficient,
                                          hallmark_hypoxia_quartile_coefficients$Hypoxia_coefficient), na.rm = TRUE)),
                                  c("grey20","white","firebrick"))
```

```{r}
drg128_quartile_coefficients_merged <- merge(drg128_quartile_coefficients,
                                             subset(gene_group, Gene %in% rownames(germ_som_drg128_mat)),
                                             by = "Gene",
                                             all.y = TRUE)
drg128_continuous_coefficients_merged <- merge(drg128_continuous_coefficients,
                                             subset(gene_group, Gene %in% rownames(germ_som_drg128_mat)),
                                             by = "Gene",
                                             all.y = TRUE)
```

```{r}
ha_drg128_significance_annotation <- rowAnnotation("Q1 v Q4 - Germline" = drg_heatmap_stats[,"g_drg128_significance"],
                                                   "Mut v Not - Germline" = drg_heatmap_stats[,"g_drg128_wilcox_significance"],
                                                   "Score Coef - Somatic" = drg128_continuous_coefficients_merged$Hypoxia_coefficient,
                                                   "Score Signif - Somatic" = drg128_continuous_coefficients_merged$Significance,
                                                   "Quartile Coef - Somatic" = drg128_quartile_coefficients_merged$Hypoxia_coefficient,
                                                   "Quartile Signif - Somatic" = drg128_quartile_coefficients_merged$Significance,
                                                   "DRG Group" = subset(gene_group, Gene %in% rownames(germ_som_drg128_mat))$Gene_list_group,
                                                   col = list("Q1 v Q4 - Germline" = c("Significant" = "firebrick", "N.S." = "grey90"),
                                                              "Mut v Not - Germline" = c("Significant" = "orangered", "N.S." = "grey90"),
                                                              "Score Coef - Somatic" = coefficient_colours_drg128_cont,
                                                              "Score Signif - Somatic" = c("Significant" = "goldenrod", "N.S." = "grey90"),
                                                              "Quartile Coef - Somatic" = coefficient_colours_drg128_quart,
                                                              "Quartile Signif - Somatic" = c("Significant" = "goldenrod", "N.S." = "grey90"),
                                                              "DRG Group" = drg_group_colours),
                                         annotation_name_rot = 45)
```

```{r}
drg128_mut_heatmap_2 <- Heatmap(germ_som_drg128_mat_2,
                                        name = "DRG128 Mutations",
                                        top_annotation = ha_row_annotation_top,
                                        right_annotation = ha_drg128_significance_annotation,
                                        show_column_names = FALSE,
                                        column_split = hypoxia_op$Buffa_signature_Sum_within_quartile_one_sample_per_donor,
                                        col = germ_som_colours,
                                        cluster_rows = TRUE)
```

```{r fig.height=20, fig.width=17}
draw(drg128_mut_heatmap_2)
```


## Hallmark hypoxia
```{r}
hypoxia_heatmap_stats <- as.data.frame(rownames(germ_som_hallmark_hypoxia_mat))
colnames(hypoxia_heatmap_stats) <- c("Gene")

colnames(germ_som_hallmark_hypoxia_mat_hypoxia_enrichment_quart) <- paste0(colnames(germ_som_hallmark_hypoxia_mat_hypoxia_enrichment_quart),"_g_s_hypoxia")
colnames(germline_hypoxia_enrichment_t1_2) <- paste0(colnames(germline_hypoxia_enrichment_t1_2),"_g_hypoxia")
colnames(somatic_hallmark_hypoxia_mat_hypoxia_enrichment_quart) <- paste0(colnames(somatic_hallmark_hypoxia_mat_hypoxia_enrichment_quart),"_s_hypoxia")

hypoxia_heatmap_stats <- merge(hypoxia_heatmap_stats, 
                          germ_som_hallmark_hypoxia_mat_hypoxia_enrichment_quart, 
                          by.x = "Gene", 
                          by.y = "Gene_g_s_hypoxia", all = TRUE)
hypoxia_heatmap_stats <- merge(hypoxia_heatmap_stats, 
                          germline_hypoxia_enrichment_t1_2, 
                          by.x = "Gene", 
                          by.y = "Gene_g_hypoxia", all = TRUE)
hypoxia_heatmap_stats <- merge(hypoxia_heatmap_stats, 
                          somatic_hallmark_hypoxia_mat_hypoxia_enrichment_quart, 
                          by.x = "Gene", 
                          by.y = "Gene_s_hypoxia", all = TRUE)
hypoxia_heatmap_stats <- merge(hypoxia_heatmap_stats,
                               germline_hypoxia_wilcox_t1_2,
                               by.x = "Gene",
                               by.y = "Gene")

hypoxia_heatmap_stats$g_s_hypoxia_significance <- ifelse(hypoxia_heatmap_stats$fisher_p_adj_g_s_hypoxia < 0.05, yes = "Significant", no = "N.S.")
hypoxia_heatmap_stats$s_hypoxia_significance <- ifelse(hypoxia_heatmap_stats$fisher_p_adj_s_hypoxia < 0.05, yes = "Significant", no = "N.S.")
hypoxia_heatmap_stats$g_hypoxia_significance <- ifelse(hypoxia_heatmap_stats$fisher_p_adj_g_hypoxia < 0.05, yes = "Significant", no = "N.S.")
hypoxia_heatmap_stats$g_hypoxia_wilcox_significance <- ifelse(hypoxia_heatmap_stats$wilcox_p_adj < 0.05, yes = "Significant", no = "N.S.")

rownames(hypoxia_heatmap_stats) <- hypoxia_heatmap_stats$Gene
hypoxia_heatmap_stats_sub <- subset(hypoxia_heatmap_stats, Gene %in% rownames(germ_som_hallmark_hypoxia_mat))
```

```{r}
hallmark_hypoxia_continuous_coefficients_merged <- merge(hallmark_hypoxia_continuous_coefficients,
                                                         hypoxia_heatmap_stats_sub,
                                                         by = "Gene",
                                                         all.y = TRUE)
hallmark_hypoxia_quartile_coefficients_merged <- merge(hallmark_hypoxia_quartile_coefficients,
                                                         hypoxia_heatmap_stats_sub,
                                                         by = "Gene",
                                                         all.y = TRUE)
```

```{r}
ha_hypoxia_significance_annotation <- rowAnnotation("Q1 v Q4 - Germline" = hypoxia_heatmap_stats_sub[,"g_hypoxia_significance"],
                                                    "Mut v Not - Germline" = hypoxia_heatmap_stats_sub[,"g_hypoxia_wilcox_significance"],
                                                    "Score Coef - Somatic" = hallmark_hypoxia_continuous_coefficients_merged$Hypoxia_coefficient,
                                                    "Score Signif - Somatic" = hallmark_hypoxia_continuous_coefficients_merged$Significance,
                                                    "Quartile Coef - Somatic" = hallmark_hypoxia_quartile_coefficients_merged$Hypoxia_coefficient,
                                                    "Quartile Signif - Somatic" = hallmark_hypoxia_quartile_coefficients_merged$Significance,
                                                    col = list("Q1 v Q4 - Germline" = c("Significant" = "firebrick", "N.S." = "grey90"),
                                                               "Mut v Not - Germline" = c("Significant" = "coral", "N.S." = "grey90"),
                                                               "Q1 v Q4 Combined" = c("Significant" = "orangered", "N.S." = "grey90"),
                                                               "Score Coef - Somatic" = coefficient_colours_drg128_cont,
                                                               "Score Signif - Somatic" = c("Significant" = "goldenrod", "N.S." = "grey90"),
                                                               "Quartile Coef - Somatic" = coefficient_colours_drg128_quart,
                                                               "Quartile Signif - Somatic" = c("Significant" = "goldenrod", "N.S." = "grey90")),
                                         annotation_name_rot = 45)
```

```{r}
mut_colours <- c("Mutation" = "forestgreen", "No Mutation" = "grey")
```


```{r}
hallmark_hypoxia_mut_heatmap_2 <- Heatmap(germ_som_hallmark_hypoxia_mat_2,
                                        name = "Hallmark Hypoxia Mutations",
                                        right_annotation = ha_hypoxia_significance_annotation,
                                        show_column_names = FALSE,
                                        column_split = hypoxia_op$Buffa_signature_Sum_within_quartile_one_sample_per_donor,
                                        col = germ_som_colours,
                                        cluster_rows = TRUE)
```

```{r fig.height=20, fig.width=17}
draw(hallmark_hypoxia_mut_heatmap_2)
```


```{r}
cumulative_heatmap <- Heatmap(hallmark_hypoxia_muts_t[c("Somatic_mutation","Germline_mutation"),],
                              name = "Any Hallmark Hypoxia Mutation",
                                        top_annotation = ha_row_annotation_top,
                              show_column_names = FALSE,
                              column_split = hypoxia_op$Buffa_signature_Sum_within_quartile_one_sample_per_donor,
                              col = mut_colours,
                              cluster_rows = TRUE)
```

```{r}
draw(cumulative_heatmap)
```

# Combine heatmaps
```{r}
ht_list = cumulative_heatmap %v%  hallmark_hypoxia_mut_heatmap_2
```

```{r fig.height=20, fig.width=16}
draw(ht_list, merge_legend = TRUE, ht_gap = unit(c(3), "mm"))
```

# Session Info
```{r}
sessionInfo()
```

