---
title: "Modelling driver mutations accounting for mutation burden without studies"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2:
    toc: true 
    toc_float: true
    toc_depth: 6
---
# Libraries used
```{r message=FALSE, warning=FALSE}
library(circlize)
library(ComplexHeatmap)
library(data.table)
library(dplyr)
library(ggplot2)
library(ggsci)
library(ggsignif)
library(ggthemes)
library(gridExtra)
library(RColorBrewer)
library(scales)
library(xtable)
library(lme4)
library(lmerTest)
library(DHARMa)

options(bitmapType='cairo')
'%!in%' <- function(x,y)!('%in%'(x,y))
```

Heatmap parameters
```{r}
ht_opt$heatmap_row_names_gp = gpar(fontsize = 12)
```

# Data
```{r}
outdir <- "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_paper_revisions/outputs/driver_models/"
model_residuals_dir <- "residual_plots"
model_objects_dir <- "model_objects"
subscript <- "revisions"
```

Read in data

```{r}
params <- list(ppcg_colourlist = "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/GitHub/PPCG_colours/PPCG_colourlist.RData",
               combined_data = "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_data_private/combined_PPCG_WGS_RNA_summary_data_clinical_v7_2024-12-06.txt",
               drivers = "/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_Drivers/Somatic_drivers/DRIVER_ANNOTATIONS__2023-09-25_release4/DRIVER_ANNOTATIONS__2023-09-25_release4__SNV+indel__filteredVariants__allGenes.txt")
```

```{r}
load(params$ppcg_colourlist)
```
```{r}
hypoxia_colours_q <- list("Q1" = "#14556e",
                          "Q2" = "#67889a",
                          "Q3" = "#c06842",
                          "Q4" = "#9c2f15")
```

Parameters, here for simplicity, otherwise kept in submission script. 
```{r}
print(params)
```

## Combined data

```{r}
combined_data <- read.delim(params$combined_data, header = TRUE, sep = "\t")
```

## Driver data
Data release version: 4

Date: 2023-09-25

Source: Juri Reimand - FTP box

/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_Drivers/Somatic_drivers/DRIVER_ANNOTATIONS__2023-09-25_release4/DRIVER_ANNOTATIONS__2023-09-25_release4__SNV+indel__filteredVariants__allGenes.txt
```{r}
driver_data <- read.delim(params$drivers, row.names = 1, sep = "\t", header = TRUE)
```


## Other data
### Gene lists

```{r}
drg_genes_36 <- c("ATM","BAP1","BLM","BRCA1","BRCA2",
               "BRIP1","CDH1","CDK4","CDKN2A","CHEK2",
               "FANCA","FANCD2","FANCI","FANCL","FANCM",
               "MLH1","MSH2","MSH6","NBN","NTHL1",
               "PALB2","PMS2","POLD1","POLE","POT1",
               "PTCH1","PTEN","RAD51C","RAD51D","RB1",
               "SMAD4","SMARCA4","STK11","TP53","XPA",
               "XPC")
```

```{r}
drg_genes_128 <- c("AKT1","ALKBH3","APEX1","APLF","ATM",
                   "ATR","ATRIP","BAP1","BARD1","BLM",
                   "BRCA1","BRCA2","BRIP1","CCNH","CDC25A",
                   "CDC25C","CDH1","CDK1","CDK2","CDK4",
                   "CDKN2A","CHEK1","CHEK2","CLK2","DCLRE1A",
                   "DDB1","EME1","EME2","EPCAM","ERCC2",
                   "ERCC3","ERCC5","ERCC6","ESR2","EXO1",
                   "FAM175A","FANCA","FANCD2","FANCI","FANCL",
                   "FANCM","GADD45A","GEN1","GTF2H2","GTF2H3",
                   "GTF2H4","H2AFX","HUS1","LIG1","LIG3",
                   "LIG4","MLH1","MLH3","MMS19","MNAT1",
                   "MPG","MRE11","MRE11A","MSH2","MSH3",
                   "MSH5","MSH6","MSR1","NABP2","NBN",
                   "NEIL1","NEIL2","NHEJ1","NTHL1","OGG1",
                   "PALB2","PARP2","PER1","PER2","PMS1",
                   "PMS2","PNKP","POLD1","POLE","POLH",
                   "POLK","POLL","POLM","POLN","POLQ",
                   "POT1","PRSS1","PTCH1","PTEN","RAD1",
                   "RAD50","RAD51","RAD51B","RAD51C","RAD51D",
                   "RAD52","RAD54B","RAD54L","RB1","RDM1",
                   "RECQL","RECQL4","RECQL5","RINT1","RNF8",
                   "RPA1","RPA3","SETMAR","SLX4","SMAD4",
                   "SMARCA4","SMUG1","STK11","TDG","TDP1",
                   "TOP2A","TOP2B","TOP3A","TP53","TP53BP1",
                   "WRN","XAB2","XPA","XPC","XRCC1",
                   "XRCC2","XRCC4","XRCC5")
```

```{r}
hallmark_hypoxia <- c("ADM",      "ADORA2B",  "AK4",      "AKAP12",   "ALDOA",    "ALDOB",   
                                     "ALDOC",    "AMPD3",    "ANGPTL4",  "ANKZF1",   "ANXA2",    "ATF3",
                                     "ATP7A",    "B3GALT6",  "B4GALNT2", "BCAN",     "BCL2",     "BGN",
                                     "BHLHE40",  "BNIP3L",   "BRS3",     "BTG1",     "CA12",     "CASP6",
                                     "CAV1",     "CCNG2",    "CCRN4L",   "CDKN1A",   "CDKN1B",   "CDKN1C",
                                     "CHST2",    "CHST3",    "CITED2",   "COL5A1",   "CP",       "CSRP2",
                                     "CTGF",     "CXCR4",    "CXCR7",    "CYR61",    "DCN",      "DDIT3",
                                     "DDIT4",    "DPYSL4",   "DTNA",     "DUSP1",    "EDN2",     "EFNA1",
                                     "EFNA3",    "EGFR",     "ENO1",     "ENO2",     "ENO3",     "ERO1L",
                                     "ERRFI1",   "ETS1",     "EXT1",     "F3",       "FAM162A",  "FBP1",
                                     "FOS",      "FOSL2",    "FOXO3",    "GAA",      "GALK1",    "GAPDH",
                                     "GAPDHS",   "GBE1",     "GCK",      "GCNT2",    "GLRX",     "GPC1",
                                     "GPC3",     "GPC4",     "GPI",      "GRHPR",    "GYS1",     "HAS1",
                                     "HDLBP",    "HEXA",     "HK1",      "HK2",      "HMOX1",    "HOXB9",
                                     "HS3ST1",   "HSPA5",    "IDS",      "IER3",     "IGFBP1",   "IGFBP3",
                                     "IL6",      "ILVBL",    "INHA",     "IRS2",     "ISG20",    "JMJD6",
                                     "JUN",      "KDELR3",   "KDM3A",    "KIF5A",    "KLF6",     "KLF7",
                                    "KLHL24",   "LALBA",    "LARGE",    "LDHA",     "LDHC",     "LOX",
                                    "LXN",      "MAFF",     "MAP3K1",   "MIF",      "MT1E",     "MT2A",
                                    "MXI1",     "MYH9",     "NAGK",     "NCAN",     "NDRG1",    "NDST1",
                                    "NDST2",    "NEDD4L",   "NFIL3",    "NR3C1",    "P4HA1",    "P4HA2",
                                    "PAM",      "PCK1",     "PDGFB",    "PDK1",     "PDK3",     "PFKFB3",
                                    "PFKL",     "PFKP",     "PGAM2",    "PGF",      "PGK1",     "PGM1",
                                    "PGM2",     "PHKG1",    "PIM1",     "PKLR",     "PKP1",     "PLAC8",
                                    "PLAUR",    "PLIN2",    "PNRC1",    "PPARGC1A", "PPFIA4",   "PPP1R15A",
                                    "PPP1R3C",  "PRDX5",    "PRKCA",    "PRKCDBP",  "PTRF",     "PYGM",
                                    "RBPJ",     "RORA",     "RRAGD",    "S100A4",   "SAP30",    "SCARB1",
                                    "SDC2",     "SDC3",     "SDC4",     "SELENBP1", "SERPINE1", "SIAH2",
                                    "SLC25A1",  "SLC2A1",   "SLC2A3",   "SLC2A5",   "SLC37A4",  "SLC6A6",
                                    "SRPX",     "STBD1",    "STC1",     "STC2",     "SULT2B1",  "TES",
                                    "TGFB3",    "TGFBI",    "TGM2",     "TIPARP",   "TKTL1",    "TMEM45A",
                                    "TNFAIP3",  "TPBG",     "TPD52",    "TPI1",     "TPST2",    "UGP2",
                                    "VEGFA",    "VHL",      "VLDLR",    "WISP2",    "WSB1",     "XPNPEP1",
                                    "ZFP36",    "ZNF292")
```


# Merge data and select samples
Use Tier 1 Buffa signature sum.

## Selecting tumour only and one sample per patient
Select only tumour samples. Then only one sample per individual donor. 

```{r}
hypoxia_signatures_tum_data_single <- subset(combined_data, !is.na(Buffa_signature_Sum_within_quartile_one_sample_per_donor))
```

```{r}
dim(hypoxia_signatures_tum_data_single)
```

# Drivers

Preparing analysis and plotting of driver data as a heatmap/oncoprint. 

```{r}
driver_mat <- as.matrix(subset(driver_data, gene_source != "."))
driver_mat = driver_mat[, -c(1,2)]
driver_mat <- gsub("\\.","", driver_mat)
```

```{r}
# subset drivers by sample used
driver_mat <- driver_mat[,hypoxia_signatures_tum_data_single$WGS_AssayID]
driver_mat <- as.data.frame(t(driver_mat))
driver_mat$WGS_AssayID <- rownames(driver_mat)
```

```{r}
driver_mat[driver_mat == "nsSNV"] <- "1"
driver_mat[driver_mat == "splicing"] <- "1"
driver_mat[driver_mat == "stopgain"] <- "1"
driver_mat[driver_mat == "stoploss"] <- "1"
driver_mat[driver_mat == "startloss"] <- "1"
driver_mat[driver_mat == "stopgain+nsSNV"] <- "1"
driver_mat[driver_mat == "nsSNV+splicing"] <- "1"
driver_mat[driver_mat == "nsSNV+stopgain"] <- "1"
driver_mat[driver_mat == "splicing+nsSNV"] <- "1"

driver_mat[driver_mat == ""] <- "0"
```

Try modelling and test for differences in scores
```{r}
hypoxia_signatures_tum_data_single_driver <- merge(hypoxia_signatures_tum_data_single[,c("Buffa_signature_Sum_within",
                                                                                         "Buffa_signature_Sum_within_quartile_one_sample_per_donor",
                                                           "Studies","num.SNVs",
                                                           "WGS_AssayID")],
                                                   driver_mat,
                                                   by = "WGS_AssayID")
```
```{r}
hypoxia_signatures_tum_data_single_driver$Studies <- as.factor(hypoxia_signatures_tum_data_single_driver$Studies)
```

## Continuous hypoxia
Set up table for outputs
```{r}
driver_hypoxia_scores <- as.data.frame(matrix(nrow = ncol(driver_mat)-1, ncol = 5))
colnames(driver_hypoxia_scores) <- c("Gene","n_mutated","n_wt",
                                     "Hypoxia_coefficient","Hypoxia_p_value")
driver_hypoxia_scores$Gene <- colnames(driver_mat)[1:(ncol(driver_mat)-1)]
```

```{r}
for(i in 1:nrow(driver_hypoxia_scores)){
  gene <- driver_hypoxia_scores$Gene[i]
  sub_data <- hypoxia_signatures_tum_data_single_driver[,c("Buffa_signature_Sum_within",
                                                           "Studies","num.SNVs",
                                                           gene,"WGS_AssayID")]
  colnames(sub_data) <- c("Buffa_signature_Sum_within",
                          "Studies","num.SNVs",
                          "gene","WGS_AssayID")
  
  driver_hypoxia_scores$n_mutated[i] <- sum(sub_data$gene == 1)
  driver_hypoxia_scores$n_wt[i] <- sum(sub_data$gene == 0)
  
  if(sum(sub_data$gene == 1) > 2){
    test_model <- glm(as.numeric(gene) ~ Buffa_signature_Sum_within + log(num.SNVs), 
                        family = binomial, data = sub_data)
    test_model_summary <- summary(test_model)
    hypoxia_coefficient <- test_model_summary$coefficients[,1][[2]]
    hypoxia_p <- test_model_summary$coefficients[,4][[2]]
    
    # save model in case we need to go back to it
    saveRDS(test_model, 
            file = paste0(outdir,model_objects_dir,"/",gene,"_drivers","_continuous_hypoxia","_binomial_glm","_",subscript,".rds"))
    
    # save residual image
    png(filename = paste0(outdir,model_residuals_dir,"/",gene,"_drivers","_continuous_hypoxia","_residual_plots","_",subscript,".png"), 
        width = 20, height = 10, res = 200, units = "cm")
    plot(test_model)
    dev.off()
    
    driver_hypoxia_scores$Hypoxia_coefficient[i] <- hypoxia_coefficient
    driver_hypoxia_scores$Hypoxia_p_value[i] <- hypoxia_p

  }
}
driver_hypoxia_scores$Hypoxia_p_adj <- p.adjust(driver_hypoxia_scores$Hypoxia_p_value, method = "BH")

```

```{r}
driver_hypoxia_scores
```
```{r}
write.table(driver_hypoxia_scores, file = paste0(outdir,"drivers","_continuous_hypoxia","_binomial_glm_summary","_",subscript,".txt"), sep = "\t", row.names = FALSE)
```


## Q1 Q4 difference
```{r}
driver_hypoxia_enrichment_quart <- as.data.frame(matrix(nrow = ncol(driver_mat)-1, ncol = 7))
colnames(driver_hypoxia_enrichment_quart) <- c("Gene","n_q4_mutated","n_q1_mutated","n_q4_wt","n_q1_wt",
                                     "Hypoxia_coefficient","Hypoxia_p_value")
driver_hypoxia_enrichment_quart$Gene <- colnames(driver_mat)[1:(ncol(driver_mat)-1)]

for(i in 1:nrow(driver_hypoxia_enrichment_quart)){
  gene <- driver_hypoxia_enrichment_quart$Gene[i]
  sub_data <- hypoxia_signatures_tum_data_single_driver[,c("Buffa_signature_Sum_within_quartile_one_sample_per_donor",
                                                           "Studies","num.SNVs",
                                                           gene,"WGS_AssayID")]
  colnames(sub_data) <- c("Buffa_signature_Sum_within_quartile_one_sample_per_donor",
                          "Studies","num.SNVs",
                          "gene","WGS_AssayID")
  
  table_data <- as.data.frame(table(sub_data$gene, 
                                    sub_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
  
  if(sum(sub_data$gene == 1) > 2){
    test_model <- glm(as.numeric(gene) ~ Buffa_signature_Sum_within_quartile_one_sample_per_donor + log(num.SNVs), 
                        family = binomial, data = sub_data)
    test_model_summary <- summary(test_model)
    hypoxia_coefficient <- test_model_summary$coefficients[,1][[4]]
    hypoxia_p <- test_model_summary$coefficients[,4][[4]]
    
    # save model in case we need to go back to it
    saveRDS(test_model, 
            file = paste0(outdir,model_objects_dir,"/",gene,"_drivers","_quartile_hypoxia","_binomial_glm","_",subscript,".rds"))
    
    # save residual image
    png(filename = paste0(outdir,model_residuals_dir,"/",gene,"_drivers","_quartile_hypoxia","_residual_plots","_",subscript,".png"), 
        width = 20, height = 10, res = 200, units = "cm")
    plot(test_model)
    dev.off()
    
    driver_hypoxia_enrichment_quart$Hypoxia_coefficient[i] <- hypoxia_coefficient
    driver_hypoxia_enrichment_quart$Hypoxia_p_value[i] <- hypoxia_p

  }
  if(sum(sub_data$gene == 1) > 0){
    driver_hypoxia_enrichment_quart$n_q4_mutated[i] <- subset(table_data, (Var1 == "1") & (Var2 == "Q4"))[,"Freq"]
    driver_hypoxia_enrichment_quart$n_q1_mutated[i] <- subset(table_data, (Var1 == "1") & (Var2 == "Q1"))[,"Freq"]
  }
    driver_hypoxia_enrichment_quart$n_q4_wt[i] <- subset(table_data, (Var1 == "0") & (Var2 == "Q4"))[,"Freq"]
    driver_hypoxia_enrichment_quart$n_q1_wt[i] <- subset(table_data, (Var1 == "0") & (Var2 == "Q1"))[,"Freq"]
  
}

driver_hypoxia_enrichment_quart$Hypoxia_p_adj <- p.adjust(driver_hypoxia_enrichment_quart$Hypoxia_p_value, method = "BH")
```

```{r}
write.table(driver_hypoxia_enrichment_quart, file = paste0(outdir,"drivers","_quartile_hypoxia","_binomial_glm_summary","_",subscript,".txt"), sep = "\t", row.names = FALSE)
```

## Check mutation in group vs not
Check if mutation in any associated gene has differences in hypoxia score
```{r}
driver_mat$WGS_AssayID <- as.factor(driver_mat$WGS_AssayID)
driver_mat <- driver_mat %>% mutate_if(is.character, as.numeric)
driver_mat$WGS_AssayID <- as.character(driver_mat$WGS_AssayID)
# driver_mat[,1:ncol(driver_mat)-1] <- as.numeric(driver_mat[,1:ncol(driver_mat)-1])
driver_mat$any_driver_sum <- rowSums(driver_mat[,1:ncol(driver_mat)-1], )
driver_mat$any_driver_bin <- ifelse(driver_mat$any_driver_sum > 0, yes = 1, no = 0)
```

```{r}
hypoxia_signatures_tum_data_single_driver_bin <- merge(hypoxia_signatures_tum_data_single[,c("Buffa_signature_Sum_within",
                                                                                         "Buffa_signature_Sum_within_quartile_one_sample_per_donor",
                                                           "Studies","num.SNVs",
                                                           "WGS_AssayID")],
                                                   driver_mat[,c("WGS_AssayID","any_driver_sum","any_driver_bin")],
                                                   by = "WGS_AssayID")
```


Check if mutation in any driver and continuous are related. 
```{r}
model_driv_continuous <- glm(any_driver_bin ~ Buffa_signature_Sum_within + log(num.SNVs), 
                        family = binomial, data = hypoxia_signatures_tum_data_single_driver_bin)
summary(model_driv_continuous)
```
```{r}
plot(model_driv_continuous)
```

Check if mutation in any driver and quartile are related.
```{r}
model_driv_quart <- glm(any_driver_bin ~ Buffa_signature_Sum_within_quartile_one_sample_per_donor + log(num.SNVs), 
                        family = binomial, data = hypoxia_signatures_tum_data_single_driver_bin)
summary(model_driv_quart)
```
```{r}
plot(model_driv_quart)
```

Check if number of mutations in driver and continuous are related.
```{r}
model_driv_continuous_cont <- glm(any_driver_sum ~ Buffa_signature_Sum_within + log(num.SNVs), 
                        family = poisson, data = hypoxia_signatures_tum_data_single_driver_bin)
summary(model_driv_continuous_cont)
```
```{r}
plot(model_driv_continuous_cont)
```

Check if number of mutations in driver and quartile are related.
```{r}
model_driv_quart_cont <- glm(any_driver_sum ~ Buffa_signature_Sum_within_quartile_one_sample_per_donor + log(num.SNVs), 
                        family = poisson, data = hypoxia_signatures_tum_data_single_driver_bin)
summary(model_driv_quart_cont)
```
```{r}
plot(model_driv_quart_cont)
```

------------------------------

# DRG 36 Somatic
DRG genes with somatic mutations identified by both callers or a tier 1 indel (filtered data)
```{r}
drg_genes_36[drg_genes_36 %in% rownames(driver_data)]
```

DRG genes that do not have somatic mutations identified by both callers or a tier 1 indel (filtered data)
```{r}
drg_genes_36[drg_genes_36 %!in% rownames(driver_data)]
```

```{r}
drg36_mat <- as.matrix(driver_data[rownames(driver_data) %in% drg_genes_36,])
drg36_mat = drg36_mat[, -c(1,2)]
drg36_mat <- gsub("\\.","", drg36_mat)
```

```{r}
# subset drivers by sample used
drg36_mat <- drg36_mat[,hypoxia_signatures_tum_data_single$WGS_AssayID]
drg36_mat <- as.data.frame(t(drg36_mat))
drg36_mat$WGS_AssayID <- rownames(drg36_mat)
```

```{r}
drg36_mat[drg36_mat == "nsSNV"] <- "1"
drg36_mat[drg36_mat == "splicing"] <- "1"
drg36_mat[drg36_mat == "stopgain"] <- "1"
drg36_mat[drg36_mat == "stoploss"] <- "1"
drg36_mat[drg36_mat == "startloss"] <- "1"
drg36_mat[drg36_mat == "stopgain+nsSNV"] <- "1"
drg36_mat[drg36_mat == "nsSNV+splicing"] <- "1"
drg36_mat[drg36_mat == "nsSNV+stopgain"] <- "1"
drg36_mat[drg36_mat == "splicing+nsSNV"] <- "1"

drg36_mat[drg36_mat == ""] <- "0"
```

```{r}
hypoxia_signatures_tum_data_single_drg36 <- merge(hypoxia_signatures_tum_data_single[,c("Buffa_signature_Sum_within",
                                                                                         "Buffa_signature_Sum_within_quartile_one_sample_per_donor",
                                                           "Studies","num.SNVs",
                                                           "WGS_AssayID")],
                                                   drg36_mat,
                                                   by = "WGS_AssayID")
```
```{r}
hypoxia_signatures_tum_data_single_drg36$Studies <- as.factor(hypoxia_signatures_tum_data_single_drg36$Studies)
```

## Continuous hypoxia
Set up table for outputs
```{r}
drg36_hypoxia_scores <- as.data.frame(matrix(nrow = ncol(drg36_mat)-1, ncol = 5))
colnames(drg36_hypoxia_scores) <- c("Gene","n_mutated","n_wt",
                                     "Hypoxia_coefficient","Hypoxia_p_value")
drg36_hypoxia_scores$Gene <- colnames(drg36_mat)[1:(ncol(drg36_mat)-1)]
```

```{r}
for(i in 1:nrow(drg36_hypoxia_scores)){
  gene <- drg36_hypoxia_scores$Gene[i]
  sub_data <- hypoxia_signatures_tum_data_single_drg36[,c("Buffa_signature_Sum_within",
                                                           "Studies","num.SNVs",
                                                           gene,"WGS_AssayID")]
  colnames(sub_data) <- c("Buffa_signature_Sum_within",
                          "Studies","num.SNVs",
                          "gene","WGS_AssayID")
  
  drg36_hypoxia_scores$n_mutated[i] <- sum(sub_data$gene == 1)
  drg36_hypoxia_scores$n_wt[i] <- sum(sub_data$gene == 0)
  
  if(sum(sub_data$gene == 1) > 2){
    test_model <- glm(as.numeric(gene) ~ Buffa_signature_Sum_within + log(num.SNVs), 
                        family = binomial, data = sub_data)
    test_model_summary <- summary(test_model)
    hypoxia_coefficient <- test_model_summary$coefficients[,1][[2]]
    hypoxia_p <- test_model_summary$coefficients[,4][[2]]
    
    # save model in case we need to go back to it
    saveRDS(test_model, 
            file = paste0(outdir,model_objects_dir,"/",gene,"_drg36s","_continuous_hypoxia","_binomial_glm","_",subscript,".rds"))
    
    # save residual image
    png(filename = paste0(outdir,model_residuals_dir,"/",gene,"_drg36s","_continuous_hypoxia","_residual_plots","_",subscript,".png"), 
        width = 20, height = 10, res = 200, units = "cm")
    plot(test_model)
    dev.off()
    
    drg36_hypoxia_scores$Hypoxia_coefficient[i] <- hypoxia_coefficient
    drg36_hypoxia_scores$Hypoxia_p_value[i] <- hypoxia_p

  }
}
drg36_hypoxia_scores$Hypoxia_p_adj <- p.adjust(drg36_hypoxia_scores$Hypoxia_p_value, method = "BH")

```

```{r}
drg36_hypoxia_scores
```
```{r}
write.table(drg36_hypoxia_scores, file = paste0(outdir,"drg36s","_continuous_hypoxia","_binomial_glm_summary","_",subscript,".txt"), sep = "\t", row.names = FALSE)
```


## Q1 Q4 difference
```{r}
drg36_hypoxia_enrichment_quart <- as.data.frame(matrix(nrow = ncol(drg36_mat)-1, ncol = 7))
colnames(drg36_hypoxia_enrichment_quart) <- c("Gene","n_q4_mutated","n_q1_mutated","n_q4_wt","n_q1_wt",
                                     "Hypoxia_coefficient","Hypoxia_p_value")
drg36_hypoxia_enrichment_quart$Gene <- colnames(drg36_mat)[1:(ncol(drg36_mat)-1)]

for(i in 1:nrow(drg36_hypoxia_enrichment_quart)){
  gene <- drg36_hypoxia_enrichment_quart$Gene[i]
  sub_data <- hypoxia_signatures_tum_data_single_drg36[,c("Buffa_signature_Sum_within_quartile_one_sample_per_donor",
                                                           "Studies","num.SNVs",
                                                           gene,"WGS_AssayID")]
  colnames(sub_data) <- c("Buffa_signature_Sum_within_quartile_one_sample_per_donor",
                          "Studies","num.SNVs",
                          "gene","WGS_AssayID")
  
  table_data <- as.data.frame(table(sub_data$gene, 
                                    sub_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
  
  if(sum(sub_data$gene == 1) > 2){
    test_model <- glm(as.numeric(gene) ~ Buffa_signature_Sum_within_quartile_one_sample_per_donor + log(num.SNVs), 
                        family = binomial, data = sub_data)
    test_model_summary <- summary(test_model)
    hypoxia_coefficient <- test_model_summary$coefficients[,1][[4]]
    hypoxia_p <- test_model_summary$coefficients[,4][[4]]
    
    # save model in case we need to go back to it
    saveRDS(test_model, 
            file = paste0(outdir,model_objects_dir,"/",gene,"_drg36s","_quartile_hypoxia","_binomial_glm","_",subscript,".rds"))
    
    # save residual image
    png(filename = paste0(outdir,model_residuals_dir,"/",gene,"_drg36s","_quartile_hypoxia","_residual_plots","_",subscript,".png"), 
        width = 20, height = 10, res = 200, units = "cm")
    plot(test_model)
    dev.off()
    
    drg36_hypoxia_enrichment_quart$Hypoxia_coefficient[i] <- hypoxia_coefficient
    drg36_hypoxia_enrichment_quart$Hypoxia_p_value[i] <- hypoxia_p

  }
  if(sum(sub_data$gene == 1) > 0){
    drg36_hypoxia_enrichment_quart$n_q4_mutated[i] <- subset(table_data, (Var1 == "1") & (Var2 == "Q4"))[,"Freq"]
    drg36_hypoxia_enrichment_quart$n_q1_mutated[i] <- subset(table_data, (Var1 == "1") & (Var2 == "Q1"))[,"Freq"]
  }
    drg36_hypoxia_enrichment_quart$n_q4_wt[i] <- subset(table_data, (Var1 == "0") & (Var2 == "Q4"))[,"Freq"]
    drg36_hypoxia_enrichment_quart$n_q1_wt[i] <- subset(table_data, (Var1 == "0") & (Var2 == "Q1"))[,"Freq"]
  
}

drg36_hypoxia_enrichment_quart$Hypoxia_p_adj <- p.adjust(drg36_hypoxia_enrichment_quart$Hypoxia_p_value, method = "BH")
```

```{r}
write.table(drg36_hypoxia_enrichment_quart, file = paste0(outdir,"drg36s","_quartile_hypoxia","_binomial_glm_summary","_",subscript,".txt"), sep = "\t", row.names = FALSE)
```

## Check mutation in group vs not
Check if mutation in any associated gene has differences in hypoxia score
```{r}
drg36_mat$WGS_AssayID <- as.factor(drg36_mat$WGS_AssayID)
drg36_mat <- drg36_mat %>% mutate_if(is.character, as.numeric)
drg36_mat$WGS_AssayID <- as.character(drg36_mat$WGS_AssayID)
# drg36_mat[,1:ncol(drg36_mat)-1] <- as.numeric(drg36_mat[,1:ncol(drg36_mat)-1])
drg36_mat$any_drg36_sum <- rowSums(drg36_mat[,1:ncol(drg36_mat)-1], )
drg36_mat$any_drg36_bin <- ifelse(drg36_mat$any_drg36_sum > 0, yes = 1, no = 0)
```

```{r}
hypoxia_signatures_tum_data_single_drg36_bin <- merge(hypoxia_signatures_tum_data_single[,c("Buffa_signature_Sum_within",
                                                                                         "Buffa_signature_Sum_within_quartile_one_sample_per_donor",
                                                           "Studies","num.SNVs",
                                                           "WGS_AssayID")],
                                                   drg36_mat[,c("WGS_AssayID","any_drg36_sum","any_drg36_bin")],
                                                   by = "WGS_AssayID")
```


Check if mutation in any drg36 and continuous are related. 
```{r}
model_drg36_continuous <- glm(any_drg36_bin ~ Buffa_signature_Sum_within + log(num.SNVs), 
                        family = binomial, data = hypoxia_signatures_tum_data_single_drg36_bin)
summary(model_drg36_continuous)
```
```{r}
plot(model_drg36_continuous)
```

Check if mutation in any drg36 and quartile are related.
```{r}
model_drg36_quart <- glm(any_drg36_bin ~ Buffa_signature_Sum_within_quartile_one_sample_per_donor + log(num.SNVs), 
                        family = binomial, data = hypoxia_signatures_tum_data_single_drg36_bin)
summary(model_drg36_quart)
```
```{r}
plot(model_drg36_quart)
```

Check if number of mutations in drg36 and continuous are related.
```{r}
model_drg36_continuous_cont <- glm(any_drg36_sum ~ Buffa_signature_Sum_within + log(num.SNVs), 
                        family = poisson, data = hypoxia_signatures_tum_data_single_drg36_bin)
summary(model_drg36_continuous_cont)
```
```{r}
plot(model_drg36_continuous_cont)
```

Check if number of mutations in drg36 and quartile are related.
```{r}
model_drg36_quart_cont <- glm(any_drg36_sum ~ Buffa_signature_Sum_within_quartile_one_sample_per_donor + log(num.SNVs), 
                        family = poisson, data = hypoxia_signatures_tum_data_single_drg36_bin)
summary(model_drg36_quart_cont)
```
```{r}
plot(model_drg36_quart_cont)
```


# DRG 128 Somatic
DRG genes with somatic mutations identified by both callers or a tier 1 indel (filtered data)
```{r}
drg_genes_128[drg_genes_128 %in% rownames(driver_data)]
```

DRG genes that do not have somatic mutations identified by both callers or a tier 1 indel (filtered data)
```{r}
drg_genes_128[drg_genes_128 %!in% rownames(driver_data)]
```

```{r}
drg128_mat <- as.matrix(driver_data[rownames(driver_data) %in% drg_genes_128,])
drg128_mat = drg128_mat[, -c(1,2)]
drg128_mat <- gsub("\\.","", drg128_mat)
```

```{r}
# subset drivers by sample used
drg128_mat <- drg128_mat[,hypoxia_signatures_tum_data_single$WGS_AssayID]
drg128_mat <- as.data.frame(t(drg128_mat))
drg128_mat$WGS_AssayID <- rownames(drg128_mat)
```

```{r}
drg128_mat[drg128_mat == "nsSNV"] <- "1"
drg128_mat[drg128_mat == "splicing"] <- "1"
drg128_mat[drg128_mat == "stopgain"] <- "1"
drg128_mat[drg128_mat == "stoploss"] <- "1"
drg128_mat[drg128_mat == "startloss"] <- "1"
drg128_mat[drg128_mat == "stopgain+nsSNV"] <- "1"
drg128_mat[drg128_mat == "nsSNV+splicing"] <- "1"
drg128_mat[drg128_mat == "nsSNV+stopgain"] <- "1"
drg128_mat[drg128_mat == "splicing+nsSNV"] <- "1"

drg128_mat[drg128_mat == ""] <- "0"
```

```{r}
hypoxia_signatures_tum_data_single_drg128 <- merge(hypoxia_signatures_tum_data_single[,c("Buffa_signature_Sum_within",
                                                                                         "Buffa_signature_Sum_within_quartile_one_sample_per_donor",
                                                           "Studies","num.SNVs",
                                                           "WGS_AssayID")],
                                                   drg128_mat,
                                                   by = "WGS_AssayID")
```
```{r}
hypoxia_signatures_tum_data_single_drg128$Studies <- as.factor(hypoxia_signatures_tum_data_single_drg128$Studies)
```

## Continuous hypoxia
Set up table for outputs
```{r}
drg128_hypoxia_scores <- as.data.frame(matrix(nrow = ncol(drg128_mat)-1, ncol = 5))
colnames(drg128_hypoxia_scores) <- c("Gene","n_mutated","n_wt",
                                     "Hypoxia_coefficient","Hypoxia_p_value")
drg128_hypoxia_scores$Gene <- colnames(drg128_mat)[1:(ncol(drg128_mat)-1)]
```

```{r}
for(i in 1:nrow(drg128_hypoxia_scores)){
  gene <- drg128_hypoxia_scores$Gene[i]
  sub_data <- hypoxia_signatures_tum_data_single_drg128[,c("Buffa_signature_Sum_within",
                                                           "Studies","num.SNVs",
                                                           gene,"WGS_AssayID")]
  colnames(sub_data) <- c("Buffa_signature_Sum_within",
                          "Studies","num.SNVs",
                          "gene","WGS_AssayID")
  
  drg128_hypoxia_scores$n_mutated[i] <- sum(sub_data$gene == 1)
  drg128_hypoxia_scores$n_wt[i] <- sum(sub_data$gene == 0)
  
  if(sum(sub_data$gene == 1) > 2){
    test_model <- glm(as.numeric(gene) ~ Buffa_signature_Sum_within + log(num.SNVs), 
                        family = binomial, data = sub_data)
    test_model_summary <- summary(test_model)
    hypoxia_coefficient <- test_model_summary$coefficients[,1][[2]]
    hypoxia_p <- test_model_summary$coefficients[,4][[2]]
    
    # save model in case we need to go back to it
    saveRDS(test_model, 
            file = paste0(outdir,model_objects_dir,"/",gene,"_drg128s","_continuous_hypoxia","_binomial_glm","_",subscript,".rds"))
    
    # save residual image
    png(filename = paste0(outdir,model_residuals_dir,"/",gene,"_drg128s","_continuous_hypoxia","_residual_plots","_",subscript,".png"), 
        width = 20, height = 10, res = 200, units = "cm")
    plot(test_model)
    dev.off()
    
    drg128_hypoxia_scores$Hypoxia_coefficient[i] <- hypoxia_coefficient
    drg128_hypoxia_scores$Hypoxia_p_value[i] <- hypoxia_p

  }
}
drg128_hypoxia_scores$Hypoxia_p_adj <- p.adjust(drg128_hypoxia_scores$Hypoxia_p_value, method = "BH")

```

```{r}
drg128_hypoxia_scores
```
```{r}
write.table(drg128_hypoxia_scores, file = paste0(outdir,"drg128s","_continuous_hypoxia","_binomial_glm_summary","_",subscript,".txt"), sep = "\t", row.names = FALSE)
```


## Q1 Q4 difference
```{r}
drg128_hypoxia_enrichment_quart <- as.data.frame(matrix(nrow = ncol(drg128_mat)-1, ncol = 7))
colnames(drg128_hypoxia_enrichment_quart) <- c("Gene","n_q4_mutated","n_q1_mutated","n_q4_wt","n_q1_wt",
                                     "Hypoxia_coefficient","Hypoxia_p_value")
drg128_hypoxia_enrichment_quart$Gene <- colnames(drg128_mat)[1:(ncol(drg128_mat)-1)]

for(i in 1:nrow(drg128_hypoxia_enrichment_quart)){
  gene <- drg128_hypoxia_enrichment_quart$Gene[i]
  sub_data <- hypoxia_signatures_tum_data_single_drg128[,c("Buffa_signature_Sum_within_quartile_one_sample_per_donor",
                                                           "Studies","num.SNVs",
                                                           gene,"WGS_AssayID")]
  colnames(sub_data) <- c("Buffa_signature_Sum_within_quartile_one_sample_per_donor",
                          "Studies","num.SNVs",
                          "gene","WGS_AssayID")
  
  table_data <- as.data.frame(table(sub_data$gene, 
                                    sub_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
  
  if(sum(sub_data$gene == 1) > 2){
    test_model <- glm(as.numeric(gene) ~ Buffa_signature_Sum_within_quartile_one_sample_per_donor + log(num.SNVs), 
                        family = binomial, data = sub_data)
    test_model_summary <- summary(test_model)
    hypoxia_coefficient <- test_model_summary$coefficients[,1][[4]]
    hypoxia_p <- test_model_summary$coefficients[,4][[4]]
    
    # save model in case we need to go back to it
    saveRDS(test_model, 
            file = paste0(outdir,model_objects_dir,"/",gene,"_drg128s","_quartile_hypoxia","_binomial_glm","_",subscript,".rds"))
    
    # save residual image
    png(filename = paste0(outdir,model_residuals_dir,"/",gene,"_drg128s","_quartile_hypoxia","_residual_plots","_",subscript,".png"), 
        width = 20, height = 10, res = 200, units = "cm")
    plot(test_model)
    dev.off()
    
    drg128_hypoxia_enrichment_quart$Hypoxia_coefficient[i] <- hypoxia_coefficient
    drg128_hypoxia_enrichment_quart$Hypoxia_p_value[i] <- hypoxia_p
  }
  if(sum(sub_data$gene == 1) > 0){
    drg128_hypoxia_enrichment_quart$n_q4_mutated[i] <- subset(table_data, (Var1 == "1") & (Var2 == "Q4"))[,"Freq"]
    drg128_hypoxia_enrichment_quart$n_q1_mutated[i] <- subset(table_data, (Var1 == "1") & (Var2 == "Q1"))[,"Freq"]
  }
    drg128_hypoxia_enrichment_quart$n_q4_wt[i] <- subset(table_data, (Var1 == "0") & (Var2 == "Q4"))[,"Freq"]
    drg128_hypoxia_enrichment_quart$n_q1_wt[i] <- subset(table_data, (Var1 == "0") & (Var2 == "Q1"))[,"Freq"]
  
}

drg128_hypoxia_enrichment_quart$Hypoxia_p_adj <- p.adjust(drg128_hypoxia_enrichment_quart$Hypoxia_p_value, method = "BH")
```

```{r}
write.table(drg128_hypoxia_enrichment_quart, file = paste0(outdir,"drg128s","_quartile_hypoxia","_binomial_glm_summary","_",subscript,".txt"), sep = "\t", row.names = FALSE)
```

## Check mutation in group vs not
Check if mutation in any associated gene has differences in hypoxia score
```{r}
drg128_mat$WGS_AssayID <- as.factor(drg128_mat$WGS_AssayID)
drg128_mat <- drg128_mat %>% mutate_if(is.character, as.numeric)
drg128_mat$WGS_AssayID <- as.character(drg128_mat$WGS_AssayID)
# drg128_mat[,1:ncol(drg128_mat)-1] <- as.numeric(drg128_mat[,1:ncol(drg128_mat)-1])
drg128_mat$any_drg128_sum <- rowSums(drg128_mat[,1:ncol(drg128_mat)-1], )
drg128_mat$any_drg128_bin <- ifelse(drg128_mat$any_drg128_sum > 0, yes = 1, no = 0)
```

```{r}
hypoxia_signatures_tum_data_single_drg128_bin <- merge(hypoxia_signatures_tum_data_single[,c("Buffa_signature_Sum_within",
                                                                                         "Buffa_signature_Sum_within_quartile_one_sample_per_donor",
                                                           "Studies","num.SNVs",
                                                           "WGS_AssayID")],
                                                   drg128_mat[,c("WGS_AssayID","any_drg128_sum","any_drg128_bin")],
                                                   by = "WGS_AssayID")
```


Check if mutation in any drg128 and continuous are related. 
```{r}
model_drg128_continuous <- glm(any_drg128_bin ~ Buffa_signature_Sum_within + log(num.SNVs), 
                        family = binomial, data = hypoxia_signatures_tum_data_single_drg128_bin)
summary(model_drg128_continuous)
```
```{r}
plot(model_drg128_continuous)
```

Check if mutation in any drg128 and quartile are related.
```{r}
model_drg128_quart <- glm(any_drg128_bin ~ Buffa_signature_Sum_within_quartile_one_sample_per_donor + log(num.SNVs), 
                        family = binomial, data = hypoxia_signatures_tum_data_single_drg128_bin)
summary(model_drg128_quart)
```
```{r}
plot(model_drg128_quart)
```

Check if number of mutations in drg128 and continuous are related.
```{r}
model_drg128_continuous_cont <- glm(any_drg128_sum ~ Buffa_signature_Sum_within + log(num.SNVs), 
                        family = poisson, data = hypoxia_signatures_tum_data_single_drg128_bin)
summary(model_drg128_continuous_cont)
```
```{r}
plot(model_drg128_continuous_cont)
```

Check if number of mutations in drg128 and quartile are related.
```{r}
model_drg128_quart_cont <- glm(any_drg128_sum ~ Buffa_signature_Sum_within_quartile_one_sample_per_donor + log(num.SNVs), 
                        family = poisson, data = hypoxia_signatures_tum_data_single_drg128_bin)
summary(model_drg128_quart_cont)
```
```{r}
plot(model_drg128_quart_cont)
```

# Hallmark Hypoxia Somatic
DRG genes with somatic mutations identified by both callers or a tier 1 indel (filtered data)
```{r}
hallmark_hypoxia[hallmark_hypoxia %in% rownames(driver_data)]
```

DRG genes that do not have somatic mutations identified by both callers or a tier 1 indel (filtered data)
```{r}
hallmark_hypoxia[hallmark_hypoxia %!in% rownames(driver_data)]
```

```{r}
hallmark_hypoxia_mat <- as.matrix(driver_data[rownames(driver_data) %in% hallmark_hypoxia,])
hallmark_hypoxia_mat = hallmark_hypoxia_mat[, -c(1,2)]
hallmark_hypoxia_mat <- gsub("\\.","", hallmark_hypoxia_mat)
```

```{r}
# subset drivers by sample used
hallmark_hypoxia_mat <- hallmark_hypoxia_mat[,hypoxia_signatures_tum_data_single$WGS_AssayID]
hallmark_hypoxia_mat <- as.data.frame(t(hallmark_hypoxia_mat))
hallmark_hypoxia_mat$WGS_AssayID <- rownames(hallmark_hypoxia_mat)
```

```{r}
hallmark_hypoxia_mat[hallmark_hypoxia_mat == "nsSNV"] <- "1"
hallmark_hypoxia_mat[hallmark_hypoxia_mat == "splicing"] <- "1"
hallmark_hypoxia_mat[hallmark_hypoxia_mat == "stopgain"] <- "1"
hallmark_hypoxia_mat[hallmark_hypoxia_mat == "stoploss"] <- "1"
hallmark_hypoxia_mat[hallmark_hypoxia_mat == "startloss"] <- "1"
hallmark_hypoxia_mat[hallmark_hypoxia_mat == "stopgain+nsSNV"] <- "1"
hallmark_hypoxia_mat[hallmark_hypoxia_mat == "nsSNV+splicing"] <- "1"
hallmark_hypoxia_mat[hallmark_hypoxia_mat == "nsSNV+stopgain"] <- "1"
hallmark_hypoxia_mat[hallmark_hypoxia_mat == "splicing+nsSNV"] <- "1"

hallmark_hypoxia_mat[hallmark_hypoxia_mat == ""] <- "0"
```

```{r}
hypoxia_signatures_tum_data_single_hallmark_hypoxia <- merge(hypoxia_signatures_tum_data_single[,c("Buffa_signature_Sum_within",
                                                                                         "Buffa_signature_Sum_within_quartile_one_sample_per_donor",
                                                           "Studies","num.SNVs",
                                                           "WGS_AssayID")],
                                                   hallmark_hypoxia_mat,
                                                   by = "WGS_AssayID")
```
```{r}
hypoxia_signatures_tum_data_single_hallmark_hypoxia$Studies <- as.factor(hypoxia_signatures_tum_data_single_hallmark_hypoxia$Studies)
```

## Continuous hypoxia
Set up table for outputs
```{r}
hallmark_hypoxia_hypoxia_scores <- as.data.frame(matrix(nrow = ncol(hallmark_hypoxia_mat)-1, ncol = 5))
colnames(hallmark_hypoxia_hypoxia_scores) <- c("Gene","n_mutated","n_wt",
                                     "Hypoxia_coefficient","Hypoxia_p_value")
hallmark_hypoxia_hypoxia_scores$Gene <- colnames(hallmark_hypoxia_mat)[1:(ncol(hallmark_hypoxia_mat)-1)]
```

```{r}
for(i in 1:nrow(hallmark_hypoxia_hypoxia_scores)){
  gene <- hallmark_hypoxia_hypoxia_scores$Gene[i]
  sub_data <- hypoxia_signatures_tum_data_single_hallmark_hypoxia[,c("Buffa_signature_Sum_within",
                                                           "Studies","num.SNVs",
                                                           gene,"WGS_AssayID")]
  colnames(sub_data) <- c("Buffa_signature_Sum_within",
                          "Studies","num.SNVs",
                          "gene","WGS_AssayID")
  
  hallmark_hypoxia_hypoxia_scores$n_mutated[i] <- sum(sub_data$gene == 1)
  hallmark_hypoxia_hypoxia_scores$n_wt[i] <- sum(sub_data$gene == 0)
  
  if(sum(sub_data$gene == 1) > 2){
    test_model <- glm(as.numeric(gene) ~ Buffa_signature_Sum_within + log(num.SNVs), 
                        family = binomial, data = sub_data)
    test_model_summary <- summary(test_model)
    hypoxia_coefficient <- test_model_summary$coefficients[,1][[2]]
    hypoxia_p <- test_model_summary$coefficients[,4][[2]]
    
    # save model in case we need to go back to it
    saveRDS(test_model, 
            file = paste0(outdir,model_objects_dir,"/",gene,"_hallmark_hypoxia","_continuous_hypoxia","_binomial_glm","_",subscript,".rds"))
    
    # save residual image
    png(filename = paste0(outdir,model_residuals_dir,"/",gene,"_hallmark_hypoxia","_continuous_hypoxia","_residual_plots","_",subscript,".png"), 
        width = 20, height = 10, res = 200, units = "cm")
    plot(test_model)
    dev.off()
    
    hallmark_hypoxia_hypoxia_scores$Hypoxia_coefficient[i] <- hypoxia_coefficient
    hallmark_hypoxia_hypoxia_scores$Hypoxia_p_value[i] <- hypoxia_p

  }
}
hallmark_hypoxia_hypoxia_scores$Hypoxia_p_adj <- p.adjust(hallmark_hypoxia_hypoxia_scores$Hypoxia_p_value, method = "BH")

```

```{r}
hallmark_hypoxia_hypoxia_scores
```
```{r}
write.table(hallmark_hypoxia_hypoxia_scores, file = paste0(outdir,"hallmark_hypoxia","_continuous_hypoxia","_binomial_glm_summary","_",subscript,".txt"), sep = "\t", row.names = FALSE)
```


## Q1 Q4 difference
```{r}
hallmark_hypoxia_hypoxia_enrichment_quart <- as.data.frame(matrix(nrow = ncol(hallmark_hypoxia_mat)-1, ncol = 7))
colnames(hallmark_hypoxia_hypoxia_enrichment_quart) <- c("Gene","n_q4_mutated","n_q1_mutated","n_q4_wt","n_q1_wt",
                                     "Hypoxia_coefficient","Hypoxia_p_value")
hallmark_hypoxia_hypoxia_enrichment_quart$Gene <- colnames(hallmark_hypoxia_mat)[1:(ncol(hallmark_hypoxia_mat)-1)]

for(i in 1:nrow(hallmark_hypoxia_hypoxia_enrichment_quart)){
  gene <- hallmark_hypoxia_hypoxia_enrichment_quart$Gene[i]
  sub_data <- hypoxia_signatures_tum_data_single_hallmark_hypoxia[,c("Buffa_signature_Sum_within_quartile_one_sample_per_donor",
                                                           "Studies","num.SNVs",
                                                           gene,"WGS_AssayID")]
  colnames(sub_data) <- c("Buffa_signature_Sum_within_quartile_one_sample_per_donor",
                          "Studies","num.SNVs",
                          "gene","WGS_AssayID")
  
  table_data <- as.data.frame(table(sub_data$gene, 
                                    sub_data$Buffa_signature_Sum_within_quartile_one_sample_per_donor))
  
  if(sum(sub_data$gene == 1) > 2){
    test_model <- glm(as.numeric(gene) ~ Buffa_signature_Sum_within_quartile_one_sample_per_donor + log(num.SNVs), 
                        family = binomial, data = sub_data)
    test_model_summary <- summary(test_model)
    hypoxia_coefficient <- test_model_summary$coefficients[,1][[4]]
    hypoxia_p <- test_model_summary$coefficients[,4][[4]]
    
    # save model in case we need to go back to it
    saveRDS(test_model, 
            file = paste0(outdir,model_objects_dir,"/",gene,"_hallmark_hypoxia","_quartile_hypoxia","_binomial_glm","_",subscript,".rds"))
    
    # save residual image
    png(filename = paste0(outdir,model_residuals_dir,"/",gene,"_hallmark_hypoxia","_quartile_hypoxia","_residual_plots","_",subscript,".png"), 
        width = 20, height = 10, res = 200, units = "cm")
    plot(test_model)
    dev.off()
    
    hallmark_hypoxia_hypoxia_enrichment_quart$Hypoxia_coefficient[i] <- hypoxia_coefficient
    hallmark_hypoxia_hypoxia_enrichment_quart$Hypoxia_p_value[i] <- hypoxia_p
  }
  if(sum(sub_data$gene == 1) > 0){
    hallmark_hypoxia_hypoxia_enrichment_quart$n_q4_mutated[i] <- subset(table_data, (Var1 == "1") & (Var2 == "Q4"))[,"Freq"]
    hallmark_hypoxia_hypoxia_enrichment_quart$n_q1_mutated[i] <- subset(table_data, (Var1 == "1") & (Var2 == "Q1"))[,"Freq"]
  }
    hallmark_hypoxia_hypoxia_enrichment_quart$n_q4_wt[i] <- subset(table_data, (Var1 == "0") & (Var2 == "Q4"))[,"Freq"]
    hallmark_hypoxia_hypoxia_enrichment_quart$n_q1_wt[i] <- subset(table_data, (Var1 == "0") & (Var2 == "Q1"))[,"Freq"]
  
}

hallmark_hypoxia_hypoxia_enrichment_quart$Hypoxia_p_adj <- p.adjust(hallmark_hypoxia_hypoxia_enrichment_quart$Hypoxia_p_value, method = "BH")
```

```{r}
write.table(hallmark_hypoxia_hypoxia_enrichment_quart, file = paste0(outdir,"hallmark_hypoxia","_quartile_hypoxia","_binomial_glm_summary","_",subscript,".txt"), sep = "\t", row.names = FALSE)
```

## Check mutation in group vs not
Check if mutation in any associated gene has differences in hypoxia score
```{r}
hallmark_hypoxia_mat$WGS_AssayID <- as.factor(hallmark_hypoxia_mat$WGS_AssayID)
hallmark_hypoxia_mat <- hallmark_hypoxia_mat %>% mutate_if(is.character, as.numeric)
hallmark_hypoxia_mat$WGS_AssayID <- as.character(hallmark_hypoxia_mat$WGS_AssayID)
# hallmark_hypoxia_mat[,1:ncol(hallmark_hypoxia_mat)-1] <- as.numeric(hallmark_hypoxia_mat[,1:ncol(hallmark_hypoxia_mat)-1])
hallmark_hypoxia_mat$any_hallmark_hypoxia_sum <- rowSums(hallmark_hypoxia_mat[,1:ncol(hallmark_hypoxia_mat)-1], )
hallmark_hypoxia_mat$any_hallmark_hypoxia_bin <- ifelse(hallmark_hypoxia_mat$any_hallmark_hypoxia_sum > 0, yes = 1, no = 0)
```

```{r}
hypoxia_signatures_tum_data_single_hallmark_hypoxia_bin <- merge(hypoxia_signatures_tum_data_single[,c("Buffa_signature_Sum_within",
                                                                                         "Buffa_signature_Sum_within_quartile_one_sample_per_donor",
                                                           "Studies","num.SNVs",
                                                           "WGS_AssayID")],
                                                   hallmark_hypoxia_mat[,c("WGS_AssayID","any_hallmark_hypoxia_sum","any_hallmark_hypoxia_bin")],
                                                   by = "WGS_AssayID")
```


Check if mutation in any hallmark_hypoxia and continuous are related. 
```{r}
model_hallmark_hypoxia_continuous <- glm(any_hallmark_hypoxia_bin ~ Buffa_signature_Sum_within + log(num.SNVs), 
                        family = binomial, data = hypoxia_signatures_tum_data_single_hallmark_hypoxia_bin)
summary(model_hallmark_hypoxia_continuous)
```
```{r}
plot(model_hallmark_hypoxia_continuous)
```

Check if mutation in any hallmark_hypoxia and quartile are related.
```{r}
model_hallmark_hypoxia_quart <- glm(any_hallmark_hypoxia_bin ~ Buffa_signature_Sum_within_quartile_one_sample_per_donor + log(num.SNVs), 
                        family = binomial, data = hypoxia_signatures_tum_data_single_hallmark_hypoxia_bin)
summary(model_hallmark_hypoxia_quart)
```
```{r}
plot(model_hallmark_hypoxia_quart)
```

Check if number of mutations in hallmark_hypoxia and continuous are related.
```{r}
model_hallmark_hypoxia_continuous_cont <- glm(any_hallmark_hypoxia_sum ~ Buffa_signature_Sum_within + log(num.SNVs), 
                        family = poisson, data = hypoxia_signatures_tum_data_single_hallmark_hypoxia_bin)
summary(model_hallmark_hypoxia_continuous_cont)
```
```{r}
plot(model_hallmark_hypoxia_continuous_cont)
```

Check if number of mutations in hallmark_hypoxia and quartile are related.
```{r}
model_hallmark_hypoxia_quart_cont <- glm(any_hallmark_hypoxia_sum ~ Buffa_signature_Sum_within_quartile_one_sample_per_donor + log(num.SNVs), 
                        family = poisson, data = hypoxia_signatures_tum_data_single_hallmark_hypoxia_bin)
summary(model_hallmark_hypoxia_quart_cont)
```
```{r}
plot(model_hallmark_hypoxia_quart_cont)
```


# Output of gene set models
```{r}
mut_sets_annotation <- as.data.frame(c("Mutations in Drivers","Mutations in DRG36","Mutations in DRG128", "Mutations in Hallmark Hypoxia"))
colnames(mut_sets_annotation) <- c("mutation_set")
rownames(mut_sets_annotation) <- mut_sets_annotation$mutation_set
mut_sets_annotation$Score_coef <- NA
mut_sets_annotation$Score_p <- NA
mut_sets_annotation$Quart_coef <- NA
mut_sets_annotation$Quart_p <- NA
```

```{r}
driv_model_summary <- summary(model_driv_continuous)
driv_model_summary_q <- summary(model_driv_quart)

mut_sets_annotation$Score_coef[1] <- driv_model_summary$coefficients[,1][[2]]
mut_sets_annotation$Score_p[1] <- driv_model_summary$coefficients[,4][[2]]
mut_sets_annotation$Quart_coef[1] <- driv_model_summary_q$coefficients[,1][[4]]
mut_sets_annotation$Quart_p[1] <- driv_model_summary_q$coefficients[,4][[4]]
```

```{r}
drg36_model_summary <- summary(model_drg36_continuous)
drg36_model_summary_q <- summary(model_drg36_quart)

mut_sets_annotation$Score_coef[2] <- drg36_model_summary$coefficients[,1][[2]]
mut_sets_annotation$Score_p[2] <- drg36_model_summary$coefficients[,4][[2]]
mut_sets_annotation$Quart_coef[2] <- drg36_model_summary_q$coefficients[,1][[4]]
mut_sets_annotation$Quart_p[2] <- drg36_model_summary_q$coefficients[,4][[4]]
```

```{r}
drg128_model_summary <- summary(model_drg128_continuous)
drg128_model_summary_q <- summary(model_drg128_quart)

mut_sets_annotation$Score_coef[3] <- drg128_model_summary$coefficients[,1][[2]]
mut_sets_annotation$Score_p[3] <- drg128_model_summary$coefficients[,4][[2]]
mut_sets_annotation$Quart_coef[3] <- drg128_model_summary_q$coefficients[,1][[4]]
mut_sets_annotation$Quart_p[3] <- drg128_model_summary_q$coefficients[,4][[4]]
```

```{r}
hallmark_hypoxia_model_summary <- summary(model_hallmark_hypoxia_continuous)
hallmark_hypoxia_model_summary_q <- summary(model_hallmark_hypoxia_quart)

mut_sets_annotation$Score_coef[4] <- hallmark_hypoxia_model_summary$coefficients[,1][[2]]
mut_sets_annotation$Score_p[4] <- hallmark_hypoxia_model_summary$coefficients[,4][[2]]
mut_sets_annotation$Quart_coef[4] <- hallmark_hypoxia_model_summary_q$coefficients[,1][[4]]
mut_sets_annotation$Quart_p[4] <- hallmark_hypoxia_model_summary_q$coefficients[,4][[4]]
```

```{r}
write.table(mut_sets_annotation, file = paste0(outdir,"gene_set","_hypoxia","_binomial_glm_summary","_",subscript,".txt"), sep = "\t", row.names = FALSE)
```


# Session info
```{r}
sessionInfo()
```

