---
title: "Figure 2. Hypoxia and ssGSEA results - release 1.4 - TCGA"
author: Maria Jakobsdottir (Maria.Jakobsdottir@manchester.ac.uk)
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2:
    toc: true 
    toc_float: true
    toc_depth: 6
---

# Libraries
```{r}
library(circlize)
library(ComplexHeatmap)
library(ggplot2)
library(gridExtra)
library(kableExtra)
library(robustbase)
library(qs)
library(scales)
library(SummarizedExperiment)
library(viridis)
```

# Data
```{r}
load("/mnt/bmh01-rds/UoOxford_David_W/b05055gj/GitHub/PPCG_colours/PPCG_colourlist.RData")
```


```{r}
outdir <- "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_paper_final_results/"
```

```{r}
fig_outdir_supp <- "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_paper_final_results/Supplementary_Figures/"
```

```{r}
supp_fig_data_outdir <- "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_paper_final_results/Supplementary_Figure_Data/"
```

```{r}
supp_tab_data_outdir <- "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_paper_final_results/Supplementary_Table_Data/"
```

## Combined data

```{r}
combined_data <- read.delim("/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_data_private/combined_PPCG_WGS_RNA_summary_data_clinical_v9_2026-05-31.txt", header = TRUE, sep = "\t")
```

```{r}
table(combined_data$Studies,combined_data$group, useNA = "always")
```


## Selecting tumour only and one sample per patient
```{r}
hypoxia_signatures_tum_data_single <- subset(combined_data, !is.na(Buffa_signature_Sum_within_quartile_extended))
```

```{r}
hypoxia_signatures_tum_data_single <- hypoxia_signatures_tum_data_single[grep("TCGA",hypoxia_signatures_tum_data_single$RNA_Local_Assay_ID),]
```


```{r}
dim(hypoxia_signatures_tum_data_single)
```

Select one sample per patient by making sure it's tumour
```{r}
hypoxia_signatures_tum_data_single <- subset(hypoxia_signatures_tum_data_single, RNA_Sample_Type == "TUM")
```

```{r}
length(unique(hypoxia_signatures_tum_data_single$RNA_Local_Assay_ID))
```

```{r}
dim(hypoxia_signatures_tum_data_single)
```


## Tidy data levels and labels

```{r}
hypoxia_signatures_tum_data_single$gleason_total <- as.factor(hypoxia_signatures_tum_data_single$gleason_total)
hypoxia_signatures_tum_data_single$gleason_total <- factor(hypoxia_signatures_tum_data_single$gleason_total, levels = c("3+3","3+4","4+3","4+4","3+5","4+5","5+4","5+5"))
```

```{r}
hypoxia_signatures_tum_data_single$psa_bracket <- as.factor(hypoxia_signatures_tum_data_single$psa_bracket)
hypoxia_signatures_tum_data_single$psa_bracket <- factor(hypoxia_signatures_tum_data_single$psa_bracket, levels = c("0-9.9","10-19.9","≥20"))
```

```{r}
hypoxia_signatures_tum_data_single$age_bracket <- as.factor(hypoxia_signatures_tum_data_single$age_bracket)
hypoxia_signatures_tum_data_single$age_bracket <- factor(hypoxia_signatures_tum_data_single$age_bracket, levels = c("< 40","40-50","50-65","65-70","≥ 70"))
```


## ssGSEA

```{r}
ssgsea_hallmark <- read.delim("/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_paper_final_results/outputs/ssgsea/revised_ssgsea_hallmark_gs_2026-05-31.txt")
```

```{r}
colnames(ssgsea_hallmark) <- gsub("HALLMARK_","", colnames(ssgsea_hallmark))
```

## Expression data
```{r}
rna_dat_supervised = qread('/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_RNA/DEPRECATED_Release_1.3/PPCG_RNAseq_RUVIIIPRPS_Supervised.qs')
rna_dat_k10 = qread('/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_RNA/K10_unsupervised/PPCG_RNAseq_RUVIII_K10.qs')
```

## Gene lists
Gene lists from Skingen et al., 2023.
```{r}
peak1_gene_list <- structure(list(SYMBOL = c("BIRC5", "CDC20", "CDK4", "CDKN2A", "CENPF", #
                                             "CYC1", "E2F2", "STMN1", "MCM2", "PA2G4", #
                                             "PMS2", "PSMA2", "TK1", "UBE2E1", "RAD54L", 
                                             "PRC1", "EXO1", "PTTG1", "TRIP13", "MELK", 
                                             "DDX39A", "SMC2", "SPAG5", "POLQ", "UBE2C", # DDX39 changed to DDX39A
                                             "CBX3", "TPX2", "KIF4A", "PTTG3P", "PABPC1", 
                                             "UBE2T", "PSMC3IP", "NUSAP1", "ASF1B", "SPC25", 
                                             "CDCA3"),
                                  ENSG = c("ENSG00000089685", "ENSG00000117399", "ENSG00000135446", "ENSG00000147889", "ENSG00000117724",
                                           "ENSG00000179091", "ENSG00000007968", "ENSG00000117632", "ENSG00000073111", "ENSG00000170515",
                                           "ENSG00000122512", "ENSG00000106588", "ENSG00000167900", "ENSG00000170142", "ENSG00000085999", 
                                           "ENSG00000198901", "ENSG00000174371", "ENSG00000164611", "ENSG00000071539", "ENSG00000165304", 
                                           "ENSG00000123136", "ENSG00000136824", "ENSG00000076382", "ENSG00000051341", "ENSG00000175063", 
                                           "ENSG00000122565", "ENSG00000088325", "ENSG00000090889", "ENSG00000213005", "ENSG00000070756", 
                                           "ENSG00000077152", "ENSG00000131470", "ENSG00000137804", "ENSG00000105011", "ENSG00000152253",
                                           "ENSG00000111665")),
                             class = "data.frame", 
                             row.names = c(NA,-36L))
```

```{r}
peak2_gene_list <- structure(list(SYMBOL = c("AMD1", "CHEK1", "DR1", "EIF1AX", "EIF4G2", #
                                             "ETF1", "HDAC2", "HIF1A", "HNRNPA2B1", "HNRNPC", #HNRPA2B1 changed to HNRNPA2B1, HNRPC changed to HNRNPC
                                             "HNRNPD", "KPNB1", "MARCKS", "MNAT1", "NDUFAB1", #
                                             "PCNA", "PURA", "RAD23B", "RPA1", "TRA2B", #SRSF10 changed to TRA2B maybe
                                             "SS18", "TCP1", "USP1", "DEK", "EED", #
                                             "PRPF4B", "CCP110", "PSMD14", "STAG1", "POLD3", #CP110 changed to CCP110
                                             "PTGES3", "CBX1", "POLS", "PSIP1", "MTF2", #POLS changed to TENT4A
                                             "PDS5B", "FAM120A", "LUC7L3"),# CROP changed to LUC7L3
                                  ENSG = c("ENSG00000123505", "ENSG00000149554", "ENSG00000117505", "ENSG00000173674", "ENSG00000110321",
                                           "ENSG00000120705", "ENSG00000196591", "ENSG00000100644", "ENSG00000122566", "ENSG00000092199",
                                           "ENSG00000138668", "ENSG00000108424", "ENSG00000277443", "ENSG00000020426", "ENSG00000004779",
                                           "ENSG00000132646", "ENSG00000185129", "ENSG00000119318", "ENSG00000132383", "ENSG00000136527",
                                           "ENSG00000141380", "ENSG00000120438", "ENSG00000162607", "ENSG00000124795", "ENSG00000074266",
                                           "ENSG00000112739", "ENSG00000103540", "ENSG00000115233", "ENSG00000118007", "ENSG00000077514",
                                           "ENSG00000110958", "ENSG00000108468", "ENSG00000112941", "ENSG00000164985", "ENSG00000143033",
                                           "ENSG00000083642", "ENSG00000048828", "ENSG00000108848")),
                             class = "data.frame", 
                             row.names = c(NA,-38L))
```

## Peak 1 
Check status of genes in PRPS data
```{r}
peak1_gene_list <- peak1_gene_list

peak1_gene_list$removed.pseudogenes <- NA
peak1_gene_list$removed.ig <- NA
peak1_gene_list$removed.noIds <- NA
peak1_gene_list$removed.NotDetected <- NA

for(i in 1:nrow(peak1_gene_list)){
  if(peak1_gene_list$ENSG[i] %in% rna_dat_supervised@elementMetadata$all.ppcg.trans){
    gene_index <- which(rna_dat_supervised@elementMetadata$all.ppcg.trans == peak1_gene_list$ENSG[i])
    
    peak1_gene_list$removed.pseudogenes[i] <- rna_dat_supervised@elementMetadata$removed.pseudogene[gene_index]
    peak1_gene_list$removed.ig[i] <- rna_dat_supervised@elementMetadata$removed.ig.mt.other.genes[gene_index]
    peak1_gene_list$removed.noIds[i] <- rna_dat_supervised@elementMetadata$removed.no.gene.ids[gene_index]
    peak1_gene_list$removed.NotDetected[i] <- rna_dat_supervised@elementMetadata$removed.no.expression[gene_index]
  }
}
```

```{r}
kable_paper(kbl(peak1_gene_list), 
            "hover", full_width = F)
```
Merge gene list with metadata to get correct row names when calculating signature.

```{r}
peak1_gene_list <- merge(peak1_gene_list, 
                      as.data.frame((rna_dat_supervised@elementMetadata)[which(rna_dat_supervised@elementMetadata$all.ppcg.trans %in% peak1_gene_list$ENSG),]), 
                      by.x = "ENSG", by.y = "all.ppcg.trans",
                      all.x = TRUE)

```

Make a column that matches expression matrix gene name
```{r}
peak1_gene_list$ID_SYMBOL <- peak1_gene_list$hgnc_symbol
```

```{r}
peak1_gene_list
```


## Peak 2 
Check status of genes in PRPS data
```{r}
peak2_gene_list <- peak2_gene_list

peak2_gene_list$removed.pseudogenes <- NA
peak2_gene_list$removed.ig <- NA
peak2_gene_list$removed.noIds <- NA
peak2_gene_list$removed.NotDetected <- NA

for(i in 1:nrow(peak2_gene_list)){
  if(peak2_gene_list$ENSG[i] %in% rna_dat_supervised@elementMetadata$all.ppcg.trans){
    gene_index <- which(rna_dat_supervised@elementMetadata$all.ppcg.trans == peak2_gene_list$ENSG[i])
    
    peak2_gene_list$removed.pseudogenes[i] <- rna_dat_supervised@elementMetadata$removed.pseudogene[gene_index]
    peak2_gene_list$removed.ig[i] <- rna_dat_supervised@elementMetadata$removed.ig.mt.other.genes[gene_index]
    peak2_gene_list$removed.noIds[i] <- rna_dat_supervised@elementMetadata$removed.no.gene.ids[gene_index]
    peak2_gene_list$removed.NotDetected[i] <- rna_dat_supervised@elementMetadata$removed.no.expression[gene_index]
  }
}
```

```{r}
kable_paper(kbl(peak2_gene_list), 
            "hover", full_width = F)
```
Merge gene list with metadata to get correct row names when calculating signature.

```{r}
peak2_gene_list <- merge(peak2_gene_list, 
                      as.data.frame((rna_dat_supervised@elementMetadata)[which(rna_dat_supervised@elementMetadata$all.ppcg.trans %in% peak2_gene_list$ENSG),]), 
                      by.x = "ENSG", by.y = "all.ppcg.trans",
                      all.x = TRUE)

```

Make a column that matches expression matrix gene name
```{r}
peak2_gene_list$ID_SYMBOL <- peak2_gene_list$hgnc_symbol
```

```{r}
peak2_gene_list
```


# Functions
```{r}
peak_score <- function(expression_matrix, annotation, gene_list, scaling, peak_number){
  # helper not in statement
  '%!in%' <- function(x,y)!('%in%'(x,y))
  
  # Stop statements
  if("matrix" %!in% class(expression_matrix)){
    stop("Input data must be an expression matrix")
  }
  
  if(ncol(expression_matrix) < 3){## Think about threshold here
    stop("Not enough samples to calculate signature")
  }
  
  if((annotation %in% c("ENSG","SYMBOL","ID_SYMBOL")) == FALSE){
    stop("Annotation must be one of 'ENSG' or 'SYMBOL' or 'ID_SYMBOL'.")
  }

  if((peak_number %in% c("Peak1","Peak2")) == FALSE){
    stop("peak_number must be one of 'Peak1' or 'Peak2'.")
  }
  
  # Set up output table
  output_data <- as.data.frame(colnames(expression_matrix))
  colnames(output_data) <- c("Sample")
  output_data$Genes_used_peak <- NA
  
  Signature_Genes_Present <- NA
  Signature_Genes_Absent <- NA
  
  ccp_genes <- gene_list
  
  # Warnings about missing signature genes
  Signature_Genes_Present <- ccp_genes[ccp_genes[,annotation] %in% rownames(expression_matrix), annotation]
  Signature_Genes_Absent <- ccp_genes[ccp_genes[,annotation] %!in% rownames(expression_matrix), annotation]
  
  if(length(Signature_Genes_Present) < nrow(ccp_genes)){
    warning(paste("Warning: ",length(Signature_Genes_Absent),"signature genes not present in input matrix; results may be incomplete or unreliable. Missing genes:",Signature_Genes_Absent, sep = " "))
  }
  
  output_data$Genes_used_peak <- toString(Signature_Genes_Present)
  
  expression_matrix_peak <- as.matrix(expression_matrix[ Signature_Genes_Present, ])
  
  # add column to output table
  output_data$peak_average <- NA
  output_data$peak_average_log2 <- NA
  
  if(scaling == TRUE){
    # Median centre and scale to unit variance per gene across patients
    expression_matrix_peak  <- expression_matrix_peak - rowMedians(expression_matrix_peak)
    expression_matrix_peak_scaled <- t(scale(t(expression_matrix_peak), center = F, scale = T))
    
    # Peak score is average expression
    # calculate signature
    output_data$peak_average <- colMeans(expression_matrix_peak_scaled)
    
  }else{
    # Median centre and scale to unit variance per gene across patients
    expression_matrix_peak  <- expression_matrix_peak - rowMedians(expression_matrix_peak)
    # calculate signature
    output_data$peak_average <- colMeans(expression_matrix_peak)
  }
  
  colnames(output_data) <- gsub("peak",peak_number,colnames(output_data))

  return(output_data)
}

```


```{r}
my_point_plot <- function(object){ggplot()+geom_point(aes(x = object$fitted.values, y = object$residuals, colour = object$rweights))+geom_hline(yintercept = 0)}

my_qq_plot <- function(object){ggplot()+geom_qq_line(aes(sample = object$residuals))+geom_qq(aes(sample = object$residuals))+geom_point(aes(y = object$residuals,colour =  object$rweights, x = -2.2), shape = 17)+ggtitle("triangles are just for showing weights\nand are not part of the plot")}
```

# Calculate peak scores
## Peak1
```{r}
peak1_unscaled <- peak_score(rna_dat_k10, annotation = "ID_SYMBOL", gene_list = peak1_gene_list, scaling = FALSE, peak_number = "Peak1")
```

```{r}
hist(peak1_unscaled$Peak1_average, breaks = 25)
```

### Calculate Peak1 without KIF4A
```{r}
peak1_unscaled_noKIF4A <- peak_score(rna_dat_k10, annotation = "ID_SYMBOL", gene_list = peak1_gene_list[-which(peak1_gene_list$SYMBOL == "KIF4A"),], scaling = FALSE, peak_number = "Peak1")
```

```{r}
hist(peak1_unscaled_noKIF4A$Peak1_average, breaks = 25)
```

Compare
```{r}
colnames(peak1_unscaled) <- paste0(colnames(peak1_unscaled),"_complete")
peak1_comparison <- merge(peak1_unscaled, peak1_unscaled_noKIF4A, by.x = "Sample_complete", by.y = "Sample")
```

```{r}
ggplot(peak1_comparison, aes(x = Peak1_average_complete, y = Peak1_average))+
  geom_point()+
  geom_smooth(method = "lm")+
  xlab("Peak 1 Complete Gene List")+
  ylab("Peak 1 Without KIF4A")+
  theme_light()
```

```{r}
ggplot(peak1_comparison, aes(x = Peak1_average_complete, y = Peak1_average))+
  geom_point()+
  geom_smooth(method = "lm")+
  xlab("Peak 1 Complete Gene List")+
  ylab("Peak 1 Without KIF4A")+
  theme_light()+
  annotate("text", label = paste0("Spearman's ρ = ",
                                 round(cor.test(peak1_comparison$Peak1_average_complete, 
                                                peak1_comparison$Peak1_average, method = "spearman")[[4]], digits = 3),
                                 "; p = ",
            format(cor.test(peak1_comparison$Peak1_average_complete, 
                            peak1_comparison$Peak1_average, method = "spearman")[[3]], scientific = TRUE, digits = 3)), 
           x = -1, y = 1.5, colour = "black")
```

```{r}
cor.test(peak1_comparison$Peak1_average_complete, peak1_comparison$Peak1_average, method = "spearman")
```


## Peak2
```{r}
peak2_unscaled <- peak_score(rna_dat_k10, annotation = "ID_SYMBOL", gene_list = peak2_gene_list, scaling = FALSE, peak_number = "Peak2")
```

```{r}
hist(peak2_unscaled$Peak2_average, breaks = 25)
```


# Combine ssGSEA and data
```{r}
ssgsea_hallmark_hypoxia <- merge(hypoxia_signatures_tum_data_single,
                                 ssgsea_hallmark,
                                 by.x = "PPCG_RNA_Assay_ID",
                                 by.y = "sample")
ssgsea_hallmark_hypoxia <- merge(ssgsea_hallmark_hypoxia,
                                 peak1_unscaled_noKIF4A,
                                 by.x = "PPCG_RNA_Assay_ID",
                                 by.y = "Sample")
ssgsea_hallmark_hypoxia <- merge(ssgsea_hallmark_hypoxia,
                                 peak2_unscaled,
                                 by.x = "PPCG_RNA_Assay_ID",
                                 by.y = "Sample")
# add RNA purity for modelling
ssgsea_hallmark_hypoxia <- merge(ssgsea_hallmark_hypoxia, 
                                     rna_dat_supervised@colData[,c("PPCG_RNA_Assay_ID",
                                                       "tumour.purity.estimate")],
                                     by.x = "PPCG_RNA_Assay_ID",
                                     by.y = "PPCG_RNA_Assay_ID",
                                     all.x = TRUE)
```

```{r}
as.data.frame(ssgsea_hallmark_hypoxia)
```

```{r}
rownames(ssgsea_hallmark_hypoxia) <- ssgsea_hallmark_hypoxia$RNA_Local_Assay_ID
```


```{r}
dim(ssgsea_hallmark_hypoxia)
```

# Plots and analysis
## Difference between groups
```{r}
ssgsea_wilcox_table <- as.data.frame(matrix(nrow = length(colnames(ssgsea_hallmark)[-51]), ncol = 7))
colnames(ssgsea_wilcox_table) <- c("Hallmark_pathway",
                                   "Direction_q1_q4","p_value_q1_q4","p_adj_q1_q4", "wilcox_w", "wilcox_L95", "wilcox_U95")
ssgsea_wilcox_table$Hallmark_pathway <- colnames(ssgsea_hallmark)[-51]
```

```{r}
for(i in 1:nrow(ssgsea_wilcox_table)){
  
  set <- ssgsea_wilcox_table$Hallmark_pathway[i]
  
  ssgsea_wilcox_table$p_value_q1_q4[i] <- wilcox.test(subset(ssgsea_hallmark_hypoxia, Buffa_signature_Sum_within_quartile_extended == "Q4")[,set],
                                                         subset(ssgsea_hallmark_hypoxia, Buffa_signature_Sum_within_quartile_extended == "Q1")[,set],
                                                      conf.int = TRUE)[[3]]
  
  ssgsea_wilcox_table$wilcox_w[i] <- wilcox.test(subset(ssgsea_hallmark_hypoxia, Buffa_signature_Sum_within_quartile_extended == "Q4")[,set],
                                                         subset(ssgsea_hallmark_hypoxia, Buffa_signature_Sum_within_quartile_extended == "Q1")[,set],
                                                      conf.int = TRUE)[[1]]
  ssgsea_wilcox_table$wilcox_L95[i] <- wilcox.test(subset(ssgsea_hallmark_hypoxia, Buffa_signature_Sum_within_quartile_extended == "Q4")[,set],
                                                         subset(ssgsea_hallmark_hypoxia, Buffa_signature_Sum_within_quartile_extended == "Q1")[,set],
                                                      conf.int = TRUE)[[8]][1]
  ssgsea_wilcox_table$wilcox_U95[i] <- wilcox.test(subset(ssgsea_hallmark_hypoxia, Buffa_signature_Sum_within_quartile_extended == "Q4")[,set],
                                                         subset(ssgsea_hallmark_hypoxia, Buffa_signature_Sum_within_quartile_extended == "Q1")[,set],
                                                      conf.int = TRUE)[[8]][2]
  
  ssgsea_wilcox_table$Direction_q1_q4[i] <- ifelse(median(subset(ssgsea_hallmark_hypoxia, Buffa_signature_Sum_within_quartile_extended == "Q4")[,set]) > median( subset(ssgsea_hallmark_hypoxia, Buffa_signature_Sum_within_quartile_extended == "Q1")[,set]), yes = "Higher in Q4", no = "Higher in Q1")
  
}

ssgsea_wilcox_table$p_adj_q1_q4 <- p.adjust(ssgsea_wilcox_table$p_value_q1_q4, method = "BH")
```

```{r}
ssgsea_wilcox_table
```

```{r}
ssgsea_wilcox_table$significance_q1_q4 <- ifelse(ssgsea_wilcox_table$p_adj_q1_q4 < 0.05, yes = "Significant", no = "N.S.")
```


## Heatmap data
```{r}
heatmap_mat <- as.matrix(t(scale(ssgsea_hallmark_hypoxia[,colnames(ssgsea_hallmark)[-51]])))
colnames(heatmap_mat) <- ssgsea_hallmark_hypoxia$RNA_Local_Assay_ID
# order data by hypoxia scores
ssgsea_hallmark_hypoxia <- ssgsea_hallmark_hypoxia[order(ssgsea_hallmark_hypoxia$Buffa_signature_Sum_within, decreasing = FALSE),]
# order driver samples by hypoxia scores
heatmap_mat <- heatmap_mat[,ssgsea_hallmark_hypoxia$RNA_Local_Assay_ID]
```

```{r}
ha_row_annotation_top <- HeatmapAnnotation(Hypoxia = anno_barplot(ssgsea_hallmark_hypoxia$Buffa_signature_Sum_within))
```

```{r}
ha_row_annotation_right <- rowAnnotation(HighvLow = ssgsea_wilcox_table$significance_2_groups,
                                         Q1vQ4 = ssgsea_wilcox_table$significance_q1_q4,
                                         col = list(HighvLow = c("Significant" = "coral", "N.S." = "grey"),
                                                    Q1vQ4 = c("Significant" = "firebrick", "N.S." = "grey")))
```

```{r fig.height=10, fig.width=15}
draw(Heatmap(heatmap_mat,
             name = "ssGSEA Score",
             top_annotation = ha_row_annotation_top,
             right_annotation = ha_row_annotation_right,
             column_split = ssgsea_hallmark_hypoxia$Buffa_signature_Sum_within_quartile_extended,
             show_column_names = FALSE,
             cluster_columns = FALSE)#,
     # heatmap_legend_side = "left"
)
```

# Robust linear mixed effects models
## Prepare model data
```{r}
ssgsea_hallmark_hypoxia_modeldat <- as.data.frame(subset(ssgsea_hallmark_hypoxia, !is.na(age_at_tumour_collection) &
                                              !is.na(ISUP_Grade) &
                                              !is.na(tumour.purity.estimate)))
```

```{r}
ssgsea_hallmark_hypoxia_modeldat$group <- as.factor(ssgsea_hallmark_hypoxia_modeldat$group)
```


## Build models
1
```{r}
lmrob_ADIPOGENESIS <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + ADIPOGENESIS, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_ADIPOGENESIS)
my_qq_plot(lmrob_ADIPOGENESIS)
```

2
```{r}
lmrob_ALLOGRAFT_REJECTION <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + ALLOGRAFT_REJECTION, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_ALLOGRAFT_REJECTION)
my_qq_plot(lmrob_ALLOGRAFT_REJECTION)
```

3
```{r}
lmrob_ANDROGEN_RESPONSE <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + ANDROGEN_RESPONSE, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_ANDROGEN_RESPONSE)
my_qq_plot(lmrob_ANDROGEN_RESPONSE)
```

4
```{r}
lmrob_ANGIOGENESIS <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + ANGIOGENESIS, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_ANGIOGENESIS)
my_qq_plot(lmrob_ANGIOGENESIS)
```

5
```{r}
lmrob_APICAL_JUNCTION <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + APICAL_JUNCTION, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_APICAL_JUNCTION)
my_qq_plot(lmrob_APICAL_JUNCTION)
```

6
```{r}
lmrob_APICAL_SURFACE <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + APICAL_SURFACE, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_APICAL_SURFACE)
my_qq_plot(lmrob_APICAL_SURFACE)
```

7
```{r}
lmrob_APOPTOSIS <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + APOPTOSIS, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_APOPTOSIS)
my_qq_plot(lmrob_APOPTOSIS)
```

8
```{r}
lmrob_BILE_ACID_METABOLISM <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + BILE_ACID_METABOLISM, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_BILE_ACID_METABOLISM)
my_qq_plot(lmrob_BILE_ACID_METABOLISM)
```

9
```{r}
lmrob_CHOLESTEROL_HOMEOSTASIS <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + CHOLESTEROL_HOMEOSTASIS, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_CHOLESTEROL_HOMEOSTASIS)
my_qq_plot(lmrob_CHOLESTEROL_HOMEOSTASIS)
```

10
```{r}
lmrob_COAGULATION <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + COAGULATION, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_COAGULATION)
my_qq_plot(lmrob_COAGULATION)
```

11
```{r}
lmrob_COMPLEMENT <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + COMPLEMENT, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_COMPLEMENT)
my_qq_plot(lmrob_COMPLEMENT)
```

12
```{r}
lmrob_DNA_REPAIR <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + DNA_REPAIR, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_DNA_REPAIR)
```

13
```{r}
lmrob_E2F_TARGETS <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + E2F_TARGETS, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_E2F_TARGETS)
my_qq_plot(lmrob_E2F_TARGETS)
```

14
```{r}
lmrob_EPITHELIAL_MESENCHYMAL_TRANSITION <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + EPITHELIAL_MESENCHYMAL_TRANSITION, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_EPITHELIAL_MESENCHYMAL_TRANSITION)
my_qq_plot(lmrob_EPITHELIAL_MESENCHYMAL_TRANSITION)
```

15
```{r}
lmrob_ESTROGEN_RESPONSE_EARLY <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + ESTROGEN_RESPONSE_EARLY, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_ESTROGEN_RESPONSE_EARLY)
my_qq_plot(lmrob_ESTROGEN_RESPONSE_EARLY)
```

16
```{r}
lmrob_ESTROGEN_RESPONSE_LATE <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + ESTROGEN_RESPONSE_LATE, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_ESTROGEN_RESPONSE_LATE)
my_qq_plot(lmrob_ESTROGEN_RESPONSE_LATE)
```

17
```{r}
lmrob_FATTY_ACID_METABOLISM <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + FATTY_ACID_METABOLISM, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_FATTY_ACID_METABOLISM)
my_qq_plot(lmrob_FATTY_ACID_METABOLISM)
```

18
```{r}
lmrob_G2M_CHECKPOINT <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + G2M_CHECKPOINT, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_G2M_CHECKPOINT)
my_qq_plot(lmrob_G2M_CHECKPOINT)
```

19
```{r}
lmrob_GLYCOLYSIS <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + GLYCOLYSIS, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_GLYCOLYSIS)
my_qq_plot(lmrob_GLYCOLYSIS)
```

20
```{r}
lmrob_HEDGEHOG_SIGNALING <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + HEDGEHOG_SIGNALING, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_HEDGEHOG_SIGNALING)
my_qq_plot(lmrob_HEDGEHOG_SIGNALING)
```

21
```{r}
lmrob_HEME_METABOLISM <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + HEME_METABOLISM, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_HEME_METABOLISM)
my_qq_plot(lmrob_HEME_METABOLISM)
```

22
```{r}
lmrob_HYPOXIA <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + HYPOXIA, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_HYPOXIA)
my_qq_plot(lmrob_HYPOXIA)
```

23
```{r}
lmrob_IL2_STAT5_SIGNALING <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + IL2_STAT5_SIGNALING, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_IL2_STAT5_SIGNALING)
my_qq_plot(lmrob_IL2_STAT5_SIGNALING)
```

24
```{r}
lmrob_IL6_JAK_STAT3_SIGNALING <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + IL6_JAK_STAT3_SIGNALING, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_IL6_JAK_STAT3_SIGNALING)
my_qq_plot(lmrob_IL6_JAK_STAT3_SIGNALING)
```

25
```{r}
lmrob_INFLAMMATORY_RESPONSE <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + INFLAMMATORY_RESPONSE, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_INFLAMMATORY_RESPONSE)
my_qq_plot(lmrob_INFLAMMATORY_RESPONSE)
```

26
```{r}
lmrob_INTERFERON_ALPHA_RESPONSE <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + INTERFERON_ALPHA_RESPONSE, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_INTERFERON_ALPHA_RESPONSE)
my_qq_plot(lmrob_INTERFERON_ALPHA_RESPONSE)
```

27
```{r}
lmrob_INTERFERON_GAMMA_RESPONSE <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + INTERFERON_GAMMA_RESPONSE, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_INTERFERON_GAMMA_RESPONSE)
my_qq_plot(lmrob_INTERFERON_GAMMA_RESPONSE)
```

28
```{r}
lmrob_KRAS_SIGNALING_DN <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + KRAS_SIGNALING_DN, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_KRAS_SIGNALING_DN)
my_qq_plot(lmrob_KRAS_SIGNALING_DN)
```

29
```{r}
lmrob_KRAS_SIGNALING_UP <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + KRAS_SIGNALING_UP, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_KRAS_SIGNALING_UP)
my_qq_plot(lmrob_KRAS_SIGNALING_UP)
```

30
```{r}
lmrob_MITOTIC_SPINDLE <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + MITOTIC_SPINDLE, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_MITOTIC_SPINDLE)
my_qq_plot(lmrob_MITOTIC_SPINDLE)
```

31
```{r}
lmrob_MTORC1_SIGNALING <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + MTORC1_SIGNALING, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_MTORC1_SIGNALING)
my_qq_plot(lmrob_MTORC1_SIGNALING)
```

32
```{r}
lmrob_MYC_TARGETS_V1 <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + MYC_TARGETS_V1, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_MYC_TARGETS_V1)
my_qq_plot(lmrob_MYC_TARGETS_V1)
```

33
```{r}
lmrob_MYC_TARGETS_V2 <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + MYC_TARGETS_V2, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_MYC_TARGETS_V2)
my_qq_plot(lmrob_MYC_TARGETS_V2)
```

34
```{r}
lmrob_MYOGENESIS <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + MYOGENESIS, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_MYOGENESIS)
my_qq_plot(lmrob_MYOGENESIS)
```

35
```{r}
lmrob_NOTCH_SIGNALING <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + NOTCH_SIGNALING, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_NOTCH_SIGNALING)
my_qq_plot(lmrob_NOTCH_SIGNALING)
```

36
```{r}
lmrob_OXIDATIVE_PHOSPHORYLATION <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + OXIDATIVE_PHOSPHORYLATION, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_OXIDATIVE_PHOSPHORYLATION)
my_qq_plot(lmrob_OXIDATIVE_PHOSPHORYLATION)
```

37
```{r}
lmrob_P53_PATHWAY <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + P53_PATHWAY, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_P53_PATHWAY)
my_qq_plot(lmrob_P53_PATHWAY)
```

38
```{r}
lmrob_PANCREAS_BETA_CELLS <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + PANCREAS_BETA_CELLS, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_PANCREAS_BETA_CELLS)
my_qq_plot(lmrob_PANCREAS_BETA_CELLS)
```

39
```{r}
lmrob_PEROXISOME <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + PEROXISOME, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_PEROXISOME)
my_qq_plot(lmrob_PEROXISOME)
```

40
```{r}
lmrob_PI3K_AKT_MTOR_SIGNALING <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + PI3K_AKT_MTOR_SIGNALING, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_PI3K_AKT_MTOR_SIGNALING)
my_qq_plot(lmrob_PI3K_AKT_MTOR_SIGNALING)
```

41
```{r}
lmrob_PROTEIN_SECRETION <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + PROTEIN_SECRETION, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_PROTEIN_SECRETION)
my_qq_plot(lmrob_PROTEIN_SECRETION)
```

42
```{r}
lmrob_REACTIVE_OXYGEN_SPECIES_PATHWAY <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + REACTIVE_OXYGEN_SPECIES_PATHWAY, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_REACTIVE_OXYGEN_SPECIES_PATHWAY)
my_qq_plot(lmrob_REACTIVE_OXYGEN_SPECIES_PATHWAY)
```

43
```{r}
lmrob_SPERMATOGENESIS <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + SPERMATOGENESIS, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_SPERMATOGENESIS)
my_qq_plot(lmrob_SPERMATOGENESIS)
```

44
```{r}
lmrob_TGF_BETA_SIGNALING <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + TGF_BETA_SIGNALING, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_TGF_BETA_SIGNALING)
my_qq_plot(lmrob_TGF_BETA_SIGNALING)
```

45
```{r}
lmrob_TNFA_SIGNALING_VIA_NFKB <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + TNFA_SIGNALING_VIA_NFKB, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_TNFA_SIGNALING_VIA_NFKB)
my_qq_plot(lmrob_TNFA_SIGNALING_VIA_NFKB)
```

46
```{r}
lmrob_UNFOLDED_PROTEIN_RESPONSE <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + UNFOLDED_PROTEIN_RESPONSE, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_UNFOLDED_PROTEIN_RESPONSE)
my_qq_plot(lmrob_UNFOLDED_PROTEIN_RESPONSE)
```

47
```{r}
lmrob_UV_RESPONSE_DN <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + UV_RESPONSE_DN, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_UV_RESPONSE_DN)
my_qq_plot(lmrob_UV_RESPONSE_DN)
```

48
```{r}
lmrob_UV_RESPONSE_UP <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + UV_RESPONSE_UP, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_UV_RESPONSE_UP)
my_qq_plot(lmrob_UV_RESPONSE_UP)
```

49
```{r}
lmrob_WNT_BETA_CATENIN_SIGNALING <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + WNT_BETA_CATENIN_SIGNALING, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_WNT_BETA_CATENIN_SIGNALING)
my_qq_plot(lmrob_WNT_BETA_CATENIN_SIGNALING)
```

50
```{r}
lmrob_XENOBIOTIC_METABOLISM <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + XENOBIOTIC_METABOLISM, data = ssgsea_hallmark_hypoxia_modeldat)
```

```{r}
my_point_plot(lmrob_XENOBIOTIC_METABOLISM)
my_qq_plot(lmrob_XENOBIOTIC_METABOLISM)
```


## Model output table
```{r}
lmrob_hallmark_outputs <- as.data.frame(matrix(nrow = 50, ncol = 11))
colnames(lmrob_hallmark_outputs) <- c("Model","Measure","Parameter","Coefficient","SE","CI","CI_low",
                                        "CI_high","t","df_error","p")
lmrob_hallmark_outputs$Model <- c("lmrob_ADIPOGENESIS", "lmrob_ALLOGRAFT_REJECTION", "lmrob_ANDROGEN_RESPONSE", "lmrob_ANGIOGENESIS", "lmrob_APICAL_JUNCTION",
                   "lmrob_APICAL_SURFACE", "lmrob_APOPTOSIS", "lmrob_BILE_ACID_METABOLISM", "lmrob_CHOLESTEROL_HOMEOSTASIS", "lmrob_COAGULATION", 
                   "lmrob_COMPLEMENT", "lmrob_DNA_REPAIR", "lmrob_E2F_TARGETS", "lmrob_EPITHELIAL_MESENCHYMAL_TRANSITION", "lmrob_ESTROGEN_RESPONSE_EARLY",
                   "lmrob_ESTROGEN_RESPONSE_LATE", "lmrob_FATTY_ACID_METABOLISM", "lmrob_G2M_CHECKPOINT", "lmrob_GLYCOLYSIS", "lmrob_HEDGEHOG_SIGNALING",
                   "lmrob_HEME_METABOLISM", "lmrob_HYPOXIA", "lmrob_IL2_STAT5_SIGNALING", "lmrob_IL6_JAK_STAT3_SIGNALING", "lmrob_INFLAMMATORY_RESPONSE",
                   "lmrob_INTERFERON_ALPHA_RESPONSE", "lmrob_INTERFERON_GAMMA_RESPONSE", "lmrob_KRAS_SIGNALING_DN", "lmrob_KRAS_SIGNALING_UP", "lmrob_MITOTIC_SPINDLE",
                   "lmrob_MTORC1_SIGNALING", "lmrob_MYC_TARGETS_V1", "lmrob_MYC_TARGETS_V2", "lmrob_MYOGENESIS", "lmrob_NOTCH_SIGNALING",
                   "lmrob_OXIDATIVE_PHOSPHORYLATION", "lmrob_P53_PATHWAY", "lmrob_PANCREAS_BETA_CELLS", "lmrob_PEROXISOME", "lmrob_PI3K_AKT_MTOR_SIGNALING",
                   "lmrob_PROTEIN_SECRETION", "lmrob_REACTIVE_OXYGEN_SPECIES_PATHWAY", "lmrob_SPERMATOGENESIS", "lmrob_TGF_BETA_SIGNALING", "lmrob_TNFA_SIGNALING_VIA_NFKB",
                   "lmrob_UNFOLDED_PROTEIN_RESPONSE", "lmrob_UV_RESPONSE_DN", "lmrob_UV_RESPONSE_UP", "lmrob_WNT_BETA_CATENIN_SIGNALING", "lmrob_XENOBIOTIC_METABOLISM")
model_list <- list(lmrob_ADIPOGENESIS, lmrob_ALLOGRAFT_REJECTION, lmrob_ANDROGEN_RESPONSE, lmrob_ANGIOGENESIS, lmrob_APICAL_JUNCTION,
                   lmrob_APICAL_SURFACE, lmrob_APOPTOSIS, lmrob_BILE_ACID_METABOLISM, lmrob_CHOLESTEROL_HOMEOSTASIS, lmrob_COAGULATION, 
                   lmrob_COMPLEMENT, lmrob_DNA_REPAIR, lmrob_E2F_TARGETS, lmrob_EPITHELIAL_MESENCHYMAL_TRANSITION, lmrob_ESTROGEN_RESPONSE_EARLY,
                   lmrob_ESTROGEN_RESPONSE_LATE, lmrob_FATTY_ACID_METABOLISM, lmrob_G2M_CHECKPOINT, lmrob_GLYCOLYSIS, lmrob_HEDGEHOG_SIGNALING,
                   lmrob_HEME_METABOLISM, lmrob_HYPOXIA, lmrob_IL2_STAT5_SIGNALING, lmrob_IL6_JAK_STAT3_SIGNALING, lmrob_INFLAMMATORY_RESPONSE,
                   lmrob_INTERFERON_ALPHA_RESPONSE, lmrob_INTERFERON_GAMMA_RESPONSE, lmrob_KRAS_SIGNALING_DN, lmrob_KRAS_SIGNALING_UP, lmrob_MITOTIC_SPINDLE,
                   lmrob_MTORC1_SIGNALING, lmrob_MYC_TARGETS_V1, lmrob_MYC_TARGETS_V2, lmrob_MYOGENESIS, lmrob_NOTCH_SIGNALING,
                   lmrob_OXIDATIVE_PHOSPHORYLATION, lmrob_P53_PATHWAY, lmrob_PANCREAS_BETA_CELLS, lmrob_PEROXISOME, lmrob_PI3K_AKT_MTOR_SIGNALING,
                   lmrob_PROTEIN_SECRETION, lmrob_REACTIVE_OXYGEN_SPECIES_PATHWAY, lmrob_SPERMATOGENESIS, lmrob_TGF_BETA_SIGNALING, lmrob_TNFA_SIGNALING_VIA_NFKB,
                   lmrob_UNFOLDED_PROTEIN_RESPONSE, lmrob_UV_RESPONSE_DN, lmrob_UV_RESPONSE_UP, lmrob_WNT_BETA_CATENIN_SIGNALING, lmrob_XENOBIOTIC_METABOLISM)
lmrob_hallmark_outputs$Measure <- gsub("lmrob_","",lmrob_hallmark_outputs$Model)
```

```{r}
for(i in 1:nrow(lmrob_hallmark_outputs)){

  model <- model_list[[i]]

  model_dat <- as.data.frame(parameters::parameters(model))
  
  lmrob_hallmark_outputs[i,3:11] <- model_dat[model_dat$Parameter == lmrob_hallmark_outputs$Measure[i],1:9]
}

lmrob_hallmark_outputs$measure_p_adj <- p.adjust(lmrob_hallmark_outputs$p, method = "BH")
```

```{r}
lmrob_hallmark_outputs
```

```{r}
lmrob_hallmark_outputs$significance <- ifelse(lmrob_hallmark_outputs$measure_p_adj < 0.05, yes = "Significant", no = "N.S.")
```

# New heatmap
```{r}
age_colours <- c("< 40" = "white","40-50" = "grey75","50-65" = "grey50","65-70" = "grey25","≥ 70" = "black")
psa_colours <- c("0-9.9" = "navajowhite", "10-19.9" = "orange", "≥20" = "indianred")
gleason_colours <- c("3+3" = "white", "3+4" = "yellow", "4+3" = "orange", "4+4" = "red", "3+5" = "maroon3", "4+5" = "orangered4", "5+4" = "firebrick4", "5+5" = "black")
clin_t_colours <- c("T1" = "olivedrab2", "T2" = "green3", "T3" = "green4", "T4" = "darkgreen")
cellularity_colours <- colorRamp2(c(0,max(ssgsea_hallmark_hypoxia$tumour.purity.estimate)),c("white","navyblue"))
isup_colours <- c("1"="#FFC1C9","2"="#E78A96","3"="#A84955","4"="#7F000F","5"="#4C191E")
erg_colours <- c("TRUE" = PPCG_colourlist$ERG_colourlist$`ERG positive`, "FALSE" = PPCG_colourlist$ERG_colourlist$`ERG negative`)
isup_colours <- c("1" = PPCG_colourlist$ISUP_colourlist$`1`, 
                  "2" = PPCG_colourlist$ISUP_colourlist$`2`, 
                  "3" = PPCG_colourlist$ISUP_colourlist$`3`, 
                  "4" = PPCG_colourlist$ISUP_colourlist$`4`, 
                  "5" = PPCG_colourlist$ISUP_colourlist$`5`)

ha_row_annotation_top <- HeatmapAnnotation(Hypoxia = anno_barplot(ssgsea_hallmark_hypoxia$Buffa_signature_Sum_within),
                                           Gleason = ssgsea_hallmark_hypoxia$gleason_total,
                                           "Clinical T" = ssgsea_hallmark_hypoxia$clin_t_bracket,
                                           "Gleason Grade Group (GG)" = ssgsea_hallmark_hypoxia$ISUP,
                                           PSA = ssgsea_hallmark_hypoxia$psa_bracket,
                                           Age = ssgsea_hallmark_hypoxia$age_bracket,
                                           Purity = ssgsea_hallmark_hypoxia$tumour.purity.estimate,
                                           "ETS Status" = ssgsea_hallmark_hypoxia$ETS,
                                           col = list(Gleason = gleason_colours,
                                                      "Clinical T" = clin_t_colours,
                                                      "Gleason Grade Group (GG)" = isup_colours,
                                                      PSA = psa_colours,
                                                      Age = age_colours,
                                                      Purity = cellularity_colours,
                                                      "ETS Status" = erg_colours))
```

```{r}
coefficient_colours <- colorRamp2(c(min(lmrob_hallmark_outputs$Coefficient, na.rm = TRUE),
                                    0,
                                    max(lmrob_hallmark_outputs$Coefficient, na.rm = TRUE)),
                                  c("purple1","white","orange1"))
ha_row_annotation_right <- rowAnnotation(Q1vQ4 = ssgsea_wilcox_table$significance_q1_q4,
                                         robustlm = lmrob_hallmark_outputs$significance,
                                         Coefficient = lmrob_hallmark_outputs$Coefficient,
                                         col = list(Q1vQ4 = c("Significant" = "firebrick", "N.S." = "grey"),
                                                    robustlm = c("Significant" = "indianred", "N.S." = "grey"),
                                                    Coefficient = coefficient_colours))
```

```{r}

ha_annotation_bottom <- HeatmapAnnotation(Gleason = ssgsea_hallmark_hypoxia$gleason_total,
                                              ClinicalT = ssgsea_hallmark_hypoxia$clin_t_bracket,
                                              ISUP = ssgsea_hallmark_hypoxia$ISUP,
                                              "PSA (ng/ml)" = ssgsea_hallmark_hypoxia$psa_bracket,
                                              "Age (Years)" = ssgsea_hallmark_hypoxia$age_bracket,
                                              Purity = ssgsea_hallmark_hypoxia$tumour.purity.estimate,
                                              col = list(Gleason = gleason_colours,
                                                         ClinicalT = clin_t_colours,
                                                         ISUP = isup_colours,
                                                         "PSA (ng/ml)" = psa_colours,
                                                         "Age (Years)" = age_colours,
                                                         Purity = cellularity_colours))
```

```{r}
ssgsea_heatmap <- Heatmap(heatmap_mat,
             name = "ssGSEA Score",
             top_annotation = ha_row_annotation_top,
             right_annotation = ha_row_annotation_right,
             # bottom_annotation = ha_annotation_bottom,
             column_split = ssgsea_hallmark_hypoxia$Buffa_signature_Sum_within_quartile_extended,
             show_column_names = FALSE,
             cluster_columns = FALSE)
```

```{r}
ht_opt$HEATMAP_LEGEND_PADDING = unit(3, "cm")
```

```{r fig.height=13.5, fig.width=15}
draw(ssgsea_heatmap#,
     # heatmap_legend_side = "left"
)
```

```{r}
ssgsea_heatmap_grob = grid.grabExpr(draw(ssgsea_heatmap#,
     # heatmap_legend_side = "left"
)) 
```

```{r}
pdf(file = paste0(fig_outdir_supp,"Supplementary_figure_12_heatmap_", Sys.Date(), ".pdf"), 
    width = 18, height = 13.5)
draw(ssgsea_heatmap)
dev.off()
```


## Save heatmap data
Top data
```{r}
# Hypoxia = anno_barplot(ssgsea_hallmark_hypoxia$Buffa_signature_Sum_within),
#                                            Gleason = ssgsea_hallmark_hypoxia$gleason_total,
#                                            "Clinical T" = ssgsea_hallmark_hypoxia$clin_t_bracket,
#                                            "Gleason Grade Group (GG)" = ssgsea_hallmark_hypoxia$ISUP,
#                                            PSA = ssgsea_hallmark_hypoxia$psa_bracket,
#                                            Age = ssgsea_hallmark_hypoxia$age_bracket,
#                                            Purity = ssgsea_hallmark_hypoxia$tumour.purity.estimate,
#                                            "ETS Status" = ssgsea_hallmark_hypoxia$ETS

write.csv(ssgsea_hallmark_hypoxia[,c("Buffa_signature_Sum_within",
                                     "Buffa_signature_Sum_within_quartile_extended",
                                     "gleason_total",
                                     "clin_t_bracket",
                                     "ISUP_Grade",
                                     "psa_bracket",
                                     "age_bracket",
                                     "tumour.purity.estimate",
                                     "ETS")],
          file = paste0(supp_fig_data_outdir, "Supplementary_figure_12_top_anno.csv"))
```


Matrix
```{r}
write.csv(heatmap_mat,
          file = paste0(supp_fig_data_outdir, "Supplementary_figure_12_heatmap_data.csv"))
```

Right annotation
```{r}
# Q1vQ4 = ssgsea_wilcox_table$significance_q1_q4,
#                                          robustlm = lmrob_hallmark_outputs$significance,
#                                          Coefficient = lmrob_hallmark_outputs$Coefficient
ssgsea_right_anno <- rbind("Q1vQ4" = ssgsea_wilcox_table$significance_q1_q4,
                                         "robustlm" = lmrob_hallmark_outputs$significance,
                                         "Coefficient" = lmrob_hallmark_outputs$Coefficient,
                           "robustlmm p" = lmrob_hallmark_outputs$p ,
                           "robustlmm padj" = lmrob_hallmark_outputs$measure_p_adj)
colnames(ssgsea_right_anno) <- lmrob_hallmark_outputs$Measure

write.csv(ssgsea_right_anno, 
          file = paste0(supp_fig_data_outdir, "Supplementary_figure_12_right_anno.csv"))
```

# Peak score plots
```{r}
cor.test(ssgsea_hallmark_hypoxia$Peak1_average, ssgsea_hallmark_hypoxia$Buffa_signature_Sum_within, method = "spearman")
```

```{r}
peak1_hypox_cor <- cor.test(ssgsea_hallmark_hypoxia$Peak1_average, ssgsea_hallmark_hypoxia$Buffa_signature_Sum_within, method = "spearman")
```

```{r}
peak1_hypox_cor$p.value
```

```{r}
cor.test(ssgsea_hallmark_hypoxia$Peak2_average, ssgsea_hallmark_hypoxia$Buffa_signature_Sum_within, method = "spearman")
```

```{r}
peak1_peak2_scatter <- ggplot(ssgsea_hallmark_hypoxia, aes(x = Peak1_average, y = Peak2_average, colour = Buffa_signature_Sum_within))+
  geom_point(alpha = 0.5)+
  theme_light()+
  scale_colour_gradient2(low = "#14556e",  mid = "white",  high = "#9c2f15",  midpoint = 0, name = "Hypoxia Score")+
  xlab("Peak1 Score")+
  ylab("Peak2 Score")+
  theme(legend.position = "bottom")
peak1_peak2_scatter
```

```{r}
hypoxia_peak1_scatter <- ggplot(ssgsea_hallmark_hypoxia, aes(x = Peak1_average, y = Buffa_signature_Sum_within))+
  geom_point(alpha = 0.5)+
  theme_light()+
  xlab("Peak1 Score")+
  ylab("Hypoxia Score")+
  geom_smooth(method = "lm", se = FALSE, colour = "grey")
hypoxia_peak1_scatter
```

```{r}
hypoxia_peak2_scatter <- ggplot(ssgsea_hallmark_hypoxia, aes(x = Peak2_average, y = Buffa_signature_Sum_within))+
  geom_point(alpha = 0.5)+
  theme_light()+
  xlab("Peak2 Score")+
  ylab("Hypoxia Score")+
  geom_smooth(method = "lm", se = FALSE, colour = "grey")
hypoxia_peak2_scatter
```

```{r}
hypoxia_peak1_box <- ggplot(ssgsea_hallmark_hypoxia, aes(y = Peak1_average, x = Buffa_signature_Sum_within_quartile_extended))+
  geom_boxplot(outlier.colour = NA)+
  geom_jitter(alpha = 0.3, height = 0, aes(colour = Buffa_signature_Sum_within_quartile_extended))+
  theme_light()+
  scale_colour_manual(values = c("Q1" = "#14556e",
                          "Q2" = "#67889a",
                          "Q3" = "#c06842",
                          "Q4" = "#9c2f15"))+
  ylab("Peak1 Score")+
  xlab("Hypoxia Quartile")+
  theme_light()+
  theme(legend.position = "none")
hypoxia_peak1_box
```

```{r}
hypoxia_peak2_box <- ggplot(ssgsea_hallmark_hypoxia, aes(y = Peak2_average, x = Buffa_signature_Sum_within_quartile_extended))+
  geom_boxplot(outlier.colour = NA)+
  geom_jitter(alpha = 0.3, height = 0, aes(colour = Buffa_signature_Sum_within_quartile_extended))+
  scale_colour_manual(values = c("Q1" = "#14556e",
                          "Q2" = "#67889a",
                          "Q3" = "#c06842",
                          "Q4" = "#9c2f15"))+
  ylab("Peak2 Score")+
  xlab("Hypoxia Quartile")+
  theme_light()+
  theme(legend.position = "none")
hypoxia_peak2_box
```


```{r}
ggplot(ssgsea_hallmark_hypoxia, aes(x = Peak1_average, y = Peak2_average, colour = Buffa_signature_Sum_within_quartile_extended))+
  geom_point()+
  theme_light()+
  scale_colour_manual(values = c("Q1" = "#14556e",
                          "Q2" = "#67889a",
                          "Q3" = "#c06842",
                          "Q4" = "#9c2f15"))+
  theme(legend.position = "bottom")+
  geom_hline(yintercept = median(ssgsea_hallmark_hypoxia$Peak2_average), linetype = 2, colour = "grey")+
  geom_vline(xintercept = median(ssgsea_hallmark_hypoxia$Peak1_average), linetype = 2, colour = "grey")
```

```{r}
cor.test(ssgsea_hallmark_hypoxia$Peak1_average, ssgsea_hallmark_hypoxia$Peak2_average, method = "spearman")
```

# Combined ssGSEA and peak score plot
combine in post...
```{r fig.height=18, fig.width=15}
grid.arrange(ggplot()+labs(tag="A"),
             hypoxia_peak1_scatter+labs(tag = "B"), hypoxia_peak2_scatter+labs(tag="C"),
             ncol = 2, nrow = 4,
             layout_matrix = rbind(c(1,1),
                                   c(1,1),
                                   c(1,1),
                                   c(2,3)))
```

```{r}
png(filename = paste0(fig_outdir_supp, "Supplementary_figure_12_peak_plots_base_",Sys.Date(),".png"), 
    width = 18, height = 18, units = "cm", res = 300)
grid.arrange(ggplot()+labs(tag="A"),
             hypoxia_peak1_scatter+labs(tag = "B"), hypoxia_peak2_scatter+labs(tag="C"),
             ncol = 2, nrow = 4,
             layout_matrix = rbind(c(1,1),
                                   c(1,1),
                                   c(1,1),
                                   c(2,3)))
dev.off()
```

# Saving peak score and ssgsea data
```{r}
write.csv(ssgsea_hallmark_hypoxia[,c("Buffa_signature_Sum_within",
                                     "Peak1_average",
                                     "Peak2_average")],
          file = paste0(supp_fig_data_outdir,"Supplementary_figure_12_peak_data.csv"), row.names = FALSE)
```




# Session info
```{r}
sessionInfo()
```

