---
title: "lmrob of hypoxia in RNA modelled for TCGA"
author: Maria Jakobsdottir (Maria.Jakobsdottir@manchester.ac.uk)
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2:
    toc: true 
    toc_float: true
    toc_depth: 6
---

Somewhat complete robust lmm analysis for publication. 

Hypoxia signatures from previous runs have been read in. In this case hypoxia scores were calculated by scaling the expression of the hypoxia genes within the sample rather than across. Cell line experiments suggest this works best. 

# Libraries used
```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(ggsci)
library(ggthemes)
library(gridExtra)
library(patchwork)
library(performance)
library(RColorBrewer)
library(robustbase)
library(sjPlot)
library(SummarizedExperiment)
library(xtable)
library(qs)

options(bitmapType='cairo')
'%!in%' <- function(x,y)!('%in%'(x,y))
```

# Functions
```{r}
my_point_plot <- function(object){ggplot()+geom_point(aes(x = object$fitted.values, y = object$residuals, colour = object$rweights))+geom_hline(yintercept = 0)}

my_qq_plot <- function(object){ggplot()+geom_qq_line(aes(sample = object$residuals))+geom_qq(aes(sample = object$residuals))+geom_point(aes(y = object$residuals,colour =  object$rweights, x = -2.2), shape = 17)+ggtitle("triangles are just for showing weights\nand are not part of the plot")}
```


# Data
Read in data

```{r}
params <- list(ppcg_colourlist = "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/GitHub/PPCG_colours/PPCG_colourlist.RData",
               combined_data = "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_data_private/combined_PPCG_WGS_RNA_summary_data_clinical_v7_2024-12-06.txt",
               TCGA_FGA = "/mnt/bmh01-rds/UoOxford_David_W/shared/TCGA-PRAD/TCGA_PRAD_cBioPortal_Fraction_Genome_Altered.txt",
               TCGA_CX = "/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_TCGA/TCGA/CINSignature_Activities_THRESH95_TCGA-PRAD-WGS_Jun2026.tsv",
               TCGA_SBS = "/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_TCGA/TCGA/PPCG_SBS96_assigned_to_TCGA_samples/Assignment_Solution/Activities/Assignment_Solution_Activities.txt",
               TCGA_ascat_CX = "/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_TCGA/TCGA/tcga_prad_wgs_ascat_segments.rds",
               tcga_snv_count = "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_paper_revisions/TCGA_analysis/TCGA_PRAD_num_SNVs_2025-10-12.txt")
```

```{r}
load(params$ppcg_colourlist)
```

Parameters, here for simplicity, otherwise kept in submission script. 
```{r}
print(params)
```

## RNA data
```{r}
rna_dat_supervised = qread('/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_RNA/DEPRECATED_Release_1.3/PPCG_RNAseq_RUVIIIPRPS_Supervised.qs')
```


## Combined data
Combined data from `../manuscript_dec_2024_code/Preparation_2_combining_PPCG_WGS_RNA_with_clinical_data_v7.Rmd` and `../manuscript_dec_2024_code/Preparation_1_combining_PPCG_WGS_RNA_data_v7.Rmd`. Check this file for version info. 

```{r}
combined_data <- read.delim(params$combined_data, header = TRUE, sep = "\t")
```

Make RNA local donor ID
```{r}
combined_data$RNA_Local_Donor_ID <- substr(combined_data$RNA_Local_Assay_ID,1,nchar(combined_data$RNA_Local_Assay_ID)-4)
```

## TCGA SNV count
```{r}
tcga_snv_count <- read.delim(params$tcga_snv_count)
```

```{r}
tcga_snv_count <- subset(tcga_snv_count, file_exists == "yes")
```

## TCGA ASCAT segs used for CX
```{r}
ascat_segs <-  readRDS(params$TCGA_ascat_CX)
```
```{r}
ascat_counts <- as.data.frame(table(ascat_segs$sample))
ascat_counts <- subset(ascat_counts, Freq > 0)
colnames(ascat_counts) <- c("tcga.id","num_cnas")
```

## TCGA CN signatures
```{r}
tcga_cn_signatures <- read.delim(params$TCGA_CX)
```
```{r}
tcga_cn_signatures <- merge(tcga_cn_signatures, ascat_counts, 
                            by.x = "sample",
                            by.y = "tcga.id")
```

## TCGA SBS
```{r}
tcga_sbs <- read.delim(params$TCGA_SBS)
```
```{r}
colnames(tcga_sbs) <- paste0(colnames(tcga_sbs),"_TCGA")
```


## TCGA FGA
```{r}
tcga_fga <- read.delim(params$TCGA_FGA)
```

## TCGA CN signature numbers
```{r}
cna_colnames <- c("CX1",  "CX2",  "CX3",  "CX4",  "CX5",  "CX7",  "CX8",  "CX11")
cna_colnames_counts <- paste0(cna_colnames,"_numbers")
```

### Multiply CNA signature proportions by number of CNAs
```{r}
tcga_cn_signatures[,cna_colnames_counts] <- tcga_cn_signatures[,cna_colnames]*tcga_cn_signatures$num_cnas
```

# Merge data and select samples
Use Buffa signature sum.

## Selecting tumour only and one sample per patient
Select only tumour samples. Then only one sample per individual donor. 
```{r}
hypoxia_signatures_tum_data_single <- subset(combined_data, !is.na(Buffa_signature_Sum_within_quartile_extended))
```


```{r}
hypoxia_signatures_tum_data_single <- hypoxia_signatures_tum_data_single[grep("TCGA",hypoxia_signatures_tum_data_single$RNA_Local_Assay_ID),]
```


```{r}
dim(hypoxia_signatures_tum_data_single)
```

Select one sample per patient by making sure it's tumour
```{r}
hypoxia_signatures_tum_data_single <- subset(hypoxia_signatures_tum_data_single, RNA_Sample_Type == "TUM")
```

```{r}
length(unique(hypoxia_signatures_tum_data_single$PPCG_Donor_ID))
```

```{r}
dim(hypoxia_signatures_tum_data_single)
```

Merge with RNA purity for modelling
```{r}
# add RNA purity for modelling
hypoxia_signatures_tum_data_single <- merge(hypoxia_signatures_tum_data_single, 
                                     rna_dat_supervised@colData[,c("PPCG_RNA_Assay_ID",
                                                       "tumour.purity.estimate")],
                                     by.x = "PPCG_RNA_Assay_ID",
                                     by.y = "PPCG_RNA_Assay_ID",
                                     all.x = TRUE)
```

```{r}
hypoxia_signatures_tum_data_single$RNA_Local_Sample_ID <- substr(hypoxia_signatures_tum_data_single$RNA_Local_Assay_ID,1,nchar(hypoxia_signatures_tum_data_single$RNA_Local_Assay_ID)-1)
```

Merge with TCGA CN signatures
```{r}
hypoxia_signatures_tum_data_single <- merge(hypoxia_signatures_tum_data_single,
                                            tcga_cn_signatures,
                                            by.x = "RNA_Local_Donor_ID",
                                            by.y = "sample",
                                            all.x = TRUE)
```

Merge with TCGA SBS signatures
```{r}
hypoxia_signatures_tum_data_single <- merge(hypoxia_signatures_tum_data_single,
                                            tcga_sbs,
                                            by.x = "RNA_Local_Assay_ID",
                                            by.y = "Samples_TCGA",
                                            all.x = TRUE)
```

Merge with TCGA FGA values
```{r}
hypoxia_signatures_tum_data_single <- merge(hypoxia_signatures_tum_data_single,
                                            tcga_fga[,c("Sample.ID","Fraction.Genome.Altered")],
                                            by.x = "RNA_Local_Sample_ID",
                                            by.y = "Sample.ID",
                                            all.x = TRUE)
```

Merge TCGA SNV count
```{r}
hypoxia_signatures_tum_data_single <- merge(hypoxia_signatures_tum_data_single,
                                            tcga_snv_count,
                                            by.x = "RNA_Local_Assay_ID",
                                            by.y = "tcga_id",
                                            all.x = TRUE)
```

## Check data
```{r}
dim(hypoxia_signatures_tum_data_single)
```

```{r}
length(unique(hypoxia_signatures_tum_data_single$RNA_Local_Assay_ID))
```

```{r}
length(unique(hypoxia_signatures_tum_data_single$PPCG_Donor_ID))
```

```{r}
summary(as.data.frame(hypoxia_signatures_tum_data_single))
```

# Subset data for modelling

## Genomic instability
### PGA Clonal
```{r}
lm_fga_clonal_data <- subset(hypoxia_signatures_tum_data_single, 
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate) &
                               !is.na(Fraction.Genome.Altered))
```
```{r}
dim(lm_fga_clonal_data)
```

### SNVs
```{r}
lm_snvs_data <- subset(hypoxia_signatures_tum_data_single, 
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate) &
                               !is.na(num.SNVs.TCGA))
```
```{r}
dim(lm_snvs_data)
```

### CNAs
```{r}
lm_cnas_data <- subset(hypoxia_signatures_tum_data_single, 
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate) &
                               !is.na(num_cnas))
```
```{r}
dim(lm_cnas_data)
```


# Robust linear models
robustlmm package

## PGA Clonal
```{r}
lmrob_fga_clonal <- lmrob(Fraction.Genome.Altered ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_fga_clonal_data)
```

```{r}
summary(lmrob_fga_clonal)
```


#### Check model diagnostics with tests

```{r}
check_collinearity(lmrob_fga_clonal)
# # check_normality(lmrob_pga_clonal)
check_autocorrelation(lmrob_fga_clonal)
check_heteroscedasticity(lmrob_fga_clonal)
# check_outliers(lmrob_pga_clonal)
```

Partial residual plot to detect poor specification

```{r}
my_point_plot(lmrob_fga_clonal)
my_qq_plot(lmrob_fga_clonal)
```

```{r}
car::qqPlot(residuals(lmrob_fga_clonal))
```


### model output table
```{r}
tab_model(lmrob_fga_clonal)
```


## SNVs
```{r}
lm_snvs_data$num.SNVs.TCGA_div100 <- lm_snvs_data$num.SNVs.TCGA/100
```

```{r}
lmrob_snvs <- lmrob(num.SNVs.TCGA_div100 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_snvs_data)
```

```{r}
summary(lmrob_snvs)
```


```{r}
my_point_plot(lmrob_snvs)
my_qq_plot(lmrob_snvs)
```


### model output table
```{r}
tab_model(lmrob_snvs)
```


## CNAs

```{r}
lmrob_cnas <- lmrob(num_cnas ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_cnas_data)
```

```{r}
summary(lmrob_cnas)
```

```{r}
my_point_plot(lmrob_cnas)
my_qq_plot(lmrob_cnas)
```


### model output table
```{r}
tab_model(lmrob_cnas)
```



## GI model p-value and coefficient table
Make a table with all the p-values from the robustlmm GI models with p-values and coefficients and adjust the p-values for plotting. 

Prepare output table
```{r}
lmrob_GI_outputs <- as.data.frame(matrix(nrow = 3, ncol = 12))
colnames(lmrob_GI_outputs) <- c("Model","Measure","Parameter","Coefficient","SE","CI","CI_low",
                                        "CI_high","t","df_error","p","Variable")
lmrob_GI_outputs$Model <- c("lmrob_fga_clonal", "lmrob_snvs", 
                   "lmrob_cnas")
model_list <- list(lmrob_fga_clonal, lmrob_snvs, 
                   lmrob_cnas)
lmrob_GI_outputs$Variable <- c("fga","num.SNVs.TCGA_div100","num_cnvs")
lmrob_GI_outputs$Measure <- c("Buffa_signature_Sum_within","Buffa_signature_Sum_within",
                              "Buffa_signature_Sum_within")

```

Make model output tables
```{r}
for(i in 1:nrow(lmrob_GI_outputs)){

  model <- model_list[[i]]

  model_dat <- as.data.frame(parameters::parameters(model))
  
  lmrob_GI_outputs[i,3:11] <- model_dat[model_dat$Parameter == lmrob_GI_outputs$Measure[i],1:9]
}

lmrob_GI_outputs$measure_p_adj <- p.adjust(lmrob_GI_outputs$p, method = "BH")
```

```{r}
model_list_GI <- model_list
```


```{r}
lmrob_GI_outputs
```

```{r}
dput(lmrob_GI_outputs)
```

Extract data and put in relevant slot

## Mutational signatures
### Data
```{r}
# SBS

lm_SBS1_data <- subset(hypoxia_signatures_tum_data_single, !is.na(SBS1_TCGA) &
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate))
lm_SBS1_data$SBS1_div10 <- lm_SBS1_data$SBS1_TCGA

lm_SBS2_data <- subset(hypoxia_signatures_tum_data_single, !is.na(SBS2_TCGA) &
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate))
lm_SBS2_data$SBS2_div10 <- lm_SBS2_data$SBS2_TCGA

lm_SBS3_data <- subset(hypoxia_signatures_tum_data_single, !is.na(SBS3_TCGA) &
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate))
lm_SBS3_data$SBS3_div10 <- lm_SBS3_data$SBS3_TCGA

lm_SBS5_data <- subset(hypoxia_signatures_tum_data_single, !is.na(SBS5_TCGA) &
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate))
lm_SBS5_data$SBS5_div10 <- lm_SBS5_data$SBS5_TCGA

lm_SBS8_data <- subset(hypoxia_signatures_tum_data_single, !is.na(SBS8_TCGA) &
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate))
lm_SBS8_data$SBS8_div10 <- lm_SBS8_data$SBS8_TCGA

lm_SBS13_data <- subset(hypoxia_signatures_tum_data_single, !is.na(SBS13_TCGA) &
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate))
lm_SBS13_data$SBS13_div10 <- lm_SBS13_data$SBS13_TCGA

lm_SBS15_data <- subset(hypoxia_signatures_tum_data_single, !is.na(SBS15_TCGA) &
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate))
lm_SBS15_data$SBS15_div10 <- lm_SBS15_data$SBS15_TCGA

lm_SBS17a_data <- subset(hypoxia_signatures_tum_data_single, !is.na(SBS17a_TCGA) &
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate))
lm_SBS17a_data$SBS17a_div10 <- lm_SBS17a_data$SBS17a_TCGA

lm_SBS17b_data <- subset(hypoxia_signatures_tum_data_single, !is.na(SBS17b_TCGA) &
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate) )
lm_SBS17b_data$SBS17b_div10 <- lm_SBS17b_data$SBS17b_TCGA

lm_SBS18_data <- subset(hypoxia_signatures_tum_data_single, !is.na(SBS18_TCGA) &
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate))
lm_SBS18_data$SBS18_div10 <- lm_SBS18_data$SBS18_TCGA

lm_SBS21_data <- subset(hypoxia_signatures_tum_data_single, !is.na(SBS21_TCGA) &
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate))
lm_SBS21_data$SBS21_div10 <- lm_SBS21_data$SBS21_TCGA

lm_SBS22b_data <- subset(hypoxia_signatures_tum_data_single, !is.na(SBS22b_TCGA) &
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate))
lm_SBS22b_data$SBS22b_div10 <- lm_SBS22b_data$SBS22b_TCGA

lm_SBS28_data <- subset(hypoxia_signatures_tum_data_single, !is.na(SBS28_TCGA) &
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate))
lm_SBS28_data$SBS28_div10 <- lm_SBS28_data$SBS28_TCGA

lm_SBS34_data <- subset(hypoxia_signatures_tum_data_single, !is.na(SBS34_TCGA) &
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate))
lm_SBS34_data$SBS34_div10 <- lm_SBS34_data$SBS34_TCGA

lm_SBS37_data <- subset(hypoxia_signatures_tum_data_single, !is.na(SBS37_TCGA) &
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate))
lm_SBS37_data$SBS37_div10 <- lm_SBS37_data$SBS37_TCGA

lm_SBS38_data <- subset(hypoxia_signatures_tum_data_single, !is.na(SBS38_TCGA) &
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate))
lm_SBS38_data$SBS38_div10 <- lm_SBS38_data$SBS38_TCGA

lm_SBS40a_data <- subset(hypoxia_signatures_tum_data_single, !is.na(SBS40a_TCGA) &
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate))
lm_SBS40a_data$SBS40a_div10 <- lm_SBS40a_data$SBS40a_TCGA

lm_SBS40b_data <- subset(hypoxia_signatures_tum_data_single, !is.na(SBS40b_TCGA) &
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate))
lm_SBS40b_data$SBS40b_div10 <- lm_SBS40b_data$SBS40b_TCGA

lm_SBS44_data <- subset(hypoxia_signatures_tum_data_single, !is.na(SBS44_TCGA) &
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate))
lm_SBS44_data$SBS44_div10 <- lm_SBS44_data$SBS44_TCGA

lm_SBS52_data <- subset(hypoxia_signatures_tum_data_single, !is.na(SBS52_TCGA) &
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate))
lm_SBS52_data$SBS52_div10 <- lm_SBS52_data$SBS52_TCGA

lm_SBS54_data <- subset(hypoxia_signatures_tum_data_single, !is.na(SBS54_TCGA) &
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate))
lm_SBS54_data$SBS54_div10 <- lm_SBS54_data$SBS54_TCGA

lm_SBS96D_data <- subset(hypoxia_signatures_tum_data_single, !is.na(SBS96D_TCGA) &
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate))
lm_SBS96D_data$SBS96D_div10 <- lm_SBS96D_data$SBS96D_TCGA


```


### Models
#### SBS
##### SBS1
```{r}
lmrob_SBS1 <- lmrob(SBS1_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_SBS1_data)
```

```{r}
summary(lmrob_SBS1)
```

```{r}
my_point_plot(lmrob_SBS1)
my_qq_plot(lmrob_SBS1)
```


```{r}
tab_model(lmrob_SBS1)
```

##### SBS2
```{r}
lmrob_SBS2 <- lmrob(SBS2_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_SBS2_data)
```

```{r}
summary(lmrob_SBS2)
```

```{r}
my_point_plot(lmrob_SBS2)
my_qq_plot(lmrob_SBS2)
```

```{r}
tab_model(lmrob_SBS2)
```

##### SBS3
```{r}
lmrob_SBS3 <- lmrob(SBS3_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_SBS3_data)
```

```{r}
summary(lmrob_SBS3)
```

```{r}
my_point_plot(lmrob_SBS3)
my_qq_plot(lmrob_SBS3)
```

```{r}
tab_model(lmrob_SBS3)
```

##### SBS5
```{r}
lmrob_SBS5 <- lmrob(SBS5_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_SBS5_data)
```

```{r}
summary(lmrob_SBS5)
```

```{r}
my_point_plot(lmrob_SBS5)
my_qq_plot(lmrob_SBS5)
```

```{r}
tab_model(lmrob_SBS5)
```

##### SBS8
```{r}
lmrob_SBS8 <- lmrob(SBS8_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_SBS8_data)
```

```{r}
summary(lmrob_SBS8)
```

```{r}
my_point_plot(lmrob_SBS8)
my_qq_plot(lmrob_SBS8)
```

```{r}
tab_model(lmrob_SBS8)
```

##### SBS13
```{r}
lmrob_SBS13 <- lmrob(SBS13_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_SBS13_data)
```

```{r}
summary(lmrob_SBS13)
```

```{r}
my_point_plot(lmrob_SBS13)
my_qq_plot(lmrob_SBS13)
```

```{r}
tab_model(lmrob_SBS13)
```

##### SBS15
```{r}
lmrob_SBS15 <- lmrob(SBS15_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_SBS15_data)
```

```{r}
summary(lmrob_SBS15)
```

```{r}
my_point_plot(lmrob_SBS15)
my_qq_plot(lmrob_SBS15)
```

```{r}
tab_model(lmrob_SBS15)
```

##### SBS17a
```{r}
lmrob_SBS17a <- lmrob(SBS17a_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_SBS17a_data)
```

```{r}
summary(lmrob_SBS17a)
```

```{r}
my_point_plot(lmrob_SBS17a)
my_qq_plot(lmrob_SBS17a)
```

```{r}
tab_model(lmrob_SBS17a)
```

##### SBS17b
```{r}
lmrob_SBS17b <- lmrob(SBS17b_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_SBS17b_data)
```

```{r}
summary(lmrob_SBS17b)
```

```{r}
my_point_plot(lmrob_SBS17b)
my_qq_plot(lmrob_SBS17b)
```

```{r}
tab_model(lmrob_SBS17b)
```

##### SBS18
```{r}
lmrob_SBS18 <- lmrob(SBS18_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_SBS18_data)
```

```{r}
summary(lmrob_SBS18)
```

```{r}
my_point_plot(lmrob_SBS18)
my_qq_plot(lmrob_SBS18)
```

```{r}
tab_model(lmrob_SBS18)
```

##### SBS21
```{r}
lmrob_SBS21 <- lmrob(SBS21_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_SBS21_data)
```

```{r}
summary(lmrob_SBS21)
```

```{r}
my_point_plot(lmrob_SBS21)
my_qq_plot(lmrob_SBS21)
```

```{r}
tab_model(lmrob_SBS21)
```

##### SBS22b
```{r}
dim(lm_SBS22b_data)
```

```{r}
# lmrob_SBS22b <- lmrob(SBS22b_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_SBS22b_data)
```

```{r}
# summary(lmrob_SBS22b)
```

```{r}
# my_point_plot(lmrob_SBS22b)
# my_qq_plot(lmrob_SBS22b)
```

```{r}
# tab_model(lmrob_SBS22b)
```

##### SBS28
```{r}
lmrob_SBS28 <- lmrob(SBS28_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_SBS28_data)
```

```{r}
summary(lmrob_SBS28)
```

```{r}
my_point_plot(lmrob_SBS28)
my_qq_plot(lmrob_SBS28)
```

```{r}
tab_model(lmrob_SBS28)
```

##### SBS34
```{r}
lm_SBS34_data$SBS34_div10
```

```{r}
# lmrob_SBS34 <- lmrob(SBS34_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_SBS34_data)
```

```{r}
# summary(lmrob_SBS34)
```

```{r}
# my_point_plot(lmrob_SBS34)
# my_qq_plot(lmrob_SBS34)
```

```{r}
# tab_model(lmrob_SBS34)
```

##### SBS37
```{r}
lmrob_SBS37 <- lmrob(SBS37_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_SBS37_data)
```

```{r}
summary(lmrob_SBS37)
```

```{r}
my_point_plot(lmrob_SBS37)
my_qq_plot(lmrob_SBS37)
```

```{r}
tab_model(lmrob_SBS37)
```

##### SBS38
```{r}
lmrob_SBS38 <- lmrob(SBS38_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_SBS38_data)
```

```{r}
summary(lmrob_SBS38)
```

```{r}
my_point_plot(lmrob_SBS38)
my_qq_plot(lmrob_SBS38)
```

```{r}
tab_model(lmrob_SBS38)
```

##### SBS40a
```{r}
lmrob_SBS40a <- lmrob(SBS40a_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_SBS40a_data)
```

```{r}
summary(lmrob_SBS40a)
```

```{r}
my_point_plot(lmrob_SBS40a)
my_qq_plot(lmrob_SBS40a)
```

```{r}
tab_model(lmrob_SBS40a)
```

##### SBS40b
```{r}
lmrob_SBS40b <- lmrob(SBS40b_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_SBS40b_data)
```

```{r}
summary(lmrob_SBS40b)
```

```{r}
my_point_plot(lmrob_SBS40b)
my_qq_plot(lmrob_SBS40b)
```

```{r}
tab_model(lmrob_SBS40b)
```

##### SBS44
```{r}
lmrob_SBS44 <- lmrob(SBS44_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_SBS44_data)
```

```{r}
summary(lmrob_SBS44)
```

```{r}
my_point_plot(lmrob_SBS44)
my_qq_plot(lmrob_SBS44)
```

```{r}
tab_model(lmrob_SBS44)
```

##### SBS52
```{r}
lmrob_SBS52 <- lmrob(SBS52_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_SBS52_data)
```

```{r}
summary(lmrob_SBS52)
```

```{r}
my_point_plot(lmrob_SBS52)
my_qq_plot(lmrob_SBS52)
```

```{r}
tab_model(lmrob_SBS52)
```

##### SBS54
```{r}
lmrob_SBS54 <- lmrob(SBS54_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_SBS54_data)
```

```{r}
summary(lmrob_SBS54)
```

```{r}
my_point_plot(lmrob_SBS54)
my_qq_plot(lmrob_SBS54)
```

```{r}
tab_model(lmrob_SBS54)
```

##### SBS96D
```{r}
lmrob_SBS96D <- lmrob(SBS96D_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_SBS96D_data)
```

```{r}
summary(lmrob_SBS96D)
```

```{r}
my_point_plot(lmrob_SBS96D)
my_qq_plot(lmrob_SBS96D)
```

```{r}
tab_model(lmrob_SBS96D)
```

#### SBS model p-value and coefficient table
Make a table with all the p-values from the robustlmm SBS models with p-values and coefficients and adjust the p-values for plotting. 

Prepare output table
```{r}
lmrob_SBS_outputs <- as.data.frame(matrix(nrow = 20, ncol = 12))
colnames(lmrob_SBS_outputs) <- c("Model","Measure","Parameter","Coefficient","SE","CI","CI_low",
                                        "CI_high","t","df_error","p","Variable")
lmrob_SBS_outputs$Model <- c("lmrob_SBS1",	"lmrob_SBS2",	"lmrob_SBS3",	"lmrob_SBS5",	"lmrob_SBS8",	
                               "lmrob_SBS13",	"lmrob_SBS15",	"lmrob_SBS17a",	"lmrob_SBS17b",	"lmrob_SBS18",
                               "lmrob_SBS21",	#"lmrob_SBS22b",	
                             "lmrob_SBS28",	#"lmrob_SBS34",	
                             "lmrob_SBS37",
                               "lmrob_SBS38",	"lmrob_SBS40a",	"lmrob_SBS40b",	"lmrob_SBS44",	"lmrob_SBS52",
                               "lmrob_SBS54",	"lmrob_SBS96D")
model_list <- list(lmrob_SBS1,	lmrob_SBS2,	lmrob_SBS3,	lmrob_SBS5,	lmrob_SBS8,	
                               lmrob_SBS13,	lmrob_SBS15,	lmrob_SBS17a,	lmrob_SBS17b,	lmrob_SBS18,
                               lmrob_SBS21,	#lmrob_SBS22b,	
                   lmrob_SBS28,	#lmrob_SBS34,	
                   lmrob_SBS37,
                               lmrob_SBS38,	lmrob_SBS40a,	lmrob_SBS40b,	lmrob_SBS44,	lmrob_SBS52,
                               lmrob_SBS54,	lmrob_SBS96D)
lmrob_SBS_outputs$Variable <- c("SBS1_div10",	"SBS2_div10",	"SBS3_div10",	"SBS5_div10",	"SBS8_div10",	
                               "SBS13_div10",	"SBS15_div10",	"SBS17a_div10",	"SBS17b_div10",	"SBS18_div10",
                               "SBS21_div10",	#"SBS22b_div10",	
                               "SBS28_div10",	#"SBS34_div10",	
                               "SBS37_div10",
                               "SBS38_div10",	"SBS40a_div10",	"SBS40b_div10",	"SBS44_div10",	"SBS52_div10",
                               "SBS54_div10",	"SBS96D_div10")
lmrob_SBS_outputs$Measure <- c("Buffa_signature_Sum_within",	"Buffa_signature_Sum_within",	"Buffa_signature_Sum_within",
                               "Buffa_signature_Sum_within",	"Buffa_signature_Sum_within",	
                               "Buffa_signature_Sum_within",	"Buffa_signature_Sum_within",	"Buffa_signature_Sum_within",
                               "Buffa_signature_Sum_within",	"Buffa_signature_Sum_within",
                               "Buffa_signature_Sum_within",	"Buffa_signature_Sum_within",	"Buffa_signature_Sum_within",
                               "Buffa_signature_Sum_within",	"Buffa_signature_Sum_within",
                               "Buffa_signature_Sum_within",	"Buffa_signature_Sum_within",	"Buffa_signature_Sum_within",
                               #"Buffa_signature_Sum_within",	#"Buffa_signature_Sum_within",
                               "Buffa_signature_Sum_within",	"Buffa_signature_Sum_within")
```

Make model output tables
```{r}
for(i in 1:nrow(lmrob_SBS_outputs)){

  model <- model_list[[i]]

  model_dat <- as.data.frame(parameters::parameters(model))
  
  lmrob_SBS_outputs[i,3:11] <- model_dat[model_dat$Parameter == lmrob_SBS_outputs$Measure[i],1:9]
}

lmrob_SBS_outputs$measure_p_adj <- p.adjust(lmrob_SBS_outputs$p, method = "BH")
```

```{r}
lmrob_SBS_outputs
```

```{r}
dput(lmrob_SBS_outputs)
```

```{r}
model_list_SBS <- model_list
```


#### CN Signatures
##### Data
```{r}
lm_CX1_data <- subset(hypoxia_signatures_tum_data_single, !is.na(CX1_numbers) &
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate))
lm_CX2_data <- subset(hypoxia_signatures_tum_data_single, !is.na(CX2_numbers) &
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate))
lm_CX3_data <- subset(hypoxia_signatures_tum_data_single, !is.na(CX3_numbers) &
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate))
lm_CX4_data <- subset(hypoxia_signatures_tum_data_single, !is.na(CX4_numbers) &
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate))
lm_CX5_data <- subset(hypoxia_signatures_tum_data_single, !is.na(CX5_numbers) &
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate))
lm_CX7_data <- subset(hypoxia_signatures_tum_data_single, !is.na(CX7_numbers) &
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate))
lm_CX8_data <- subset(hypoxia_signatures_tum_data_single, !is.na(CX8_numbers) &
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate))
lm_CX11_data <- subset(hypoxia_signatures_tum_data_single, !is.na(CX11_numbers) &
                               !is.na(age_at_tumour_collection) &
                               !is.na(ISUP_Grade) &
                               !is.na(tumour.purity.estimate))
```

##### Models
###### CX1
```{r}
lmrob_CX1 <- lmrob(CX1_numbers ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_CX1_data)
```

```{r}
summary(lmrob_CX1)
```

```{r}
my_point_plot(lmrob_CX1)
my_qq_plot(lmrob_CX1)
```

```{r}
tab_model(lmrob_CX1)
```

###### CX2
```{r}
lmrob_CX2 <- lmrob(CX2_numbers ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_CX2_data)
```

```{r}
summary(lmrob_CX2)
```

```{r}
my_point_plot(lmrob_CX2)
my_qq_plot(lmrob_CX2)
```

```{r}
tab_model(lmrob_CX2)
```

###### CX3
```{r}
lmrob_CX3 <- lmrob(CX3_numbers ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_CX3_data)
```

```{r}
summary(lmrob_CX3)
```

```{r}
my_point_plot(lmrob_CX3)
my_qq_plot(lmrob_CX3)
```

```{r}
tab_model(lmrob_CX3)
```

###### CX4
```{r}
lmrob_CX4 <- lmrob(CX4_numbers ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_CX4_data)
```

```{r}
summary(lmrob_CX4)
```

```{r}
my_point_plot(lmrob_CX4)
my_qq_plot(lmrob_CX4)
```

```{r}
tab_model(lmrob_CX4)
```

###### CX5
```{r}
lmrob_CX5 <- lmrob(CX5_numbers ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_CX5_data)
```

```{r}
summary(lmrob_CX5)
```

```{r}
my_point_plot(lmrob_CX5)
my_qq_plot(lmrob_CX5)
```

```{r}
tab_model(lmrob_CX5)
```

###### CX7
```{r}
lmrob_CX7 <- lmrob(CX7_numbers ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_CX7_data)
```

```{r}
summary(lmrob_CX7)
```

```{r}
my_point_plot(lmrob_CX7)
my_qq_plot(lmrob_CX7)
```

```{r}
tab_model(lmrob_CX7)
```

###### CX8
```{r}
lmrob_CX8 <- lmrob(CX8_numbers ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_CX8_data)
```

```{r}
summary(lmrob_CX8)
```

```{r}
my_point_plot(lmrob_CX8)
my_qq_plot(lmrob_CX8)
```

```{r}
tab_model(lmrob_CX8)
```

###### CX11
```{r}
lmrob_CX11 <- lmrob(CX11_numbers ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = lm_CX11_data)
```

```{r}
summary(lmrob_CX11)
```

```{r}
my_point_plot(lmrob_CX11)
my_qq_plot(lmrob_CX11)
```

```{r}
tab_model(lmrob_CX11)
```

##### Output table
Make a table with all the p-values from the robustlmm CX models with p-values and coefficients and adjust the p-values for plotting. 

Prepare output table
```{r}
lmrob_CX_outputs <- as.data.frame(matrix(nrow = 8, ncol = 12))
colnames(lmrob_CX_outputs) <- c("Model","Measure","Parameter","Coefficient","SE","CI","CI_low",
                                        "CI_high","t","df_error","p","Variable")
lmrob_CX_outputs$Model <- c("lmrob_CX1","lmrob_CX2","lmrob_CX3","lmrob_CX4","lmrob_CX5",
                             "lmrob_CX7","lmrob_CX8","lmrob_CX11")
model_list <- list(lmrob_CX1,lmrob_CX2,lmrob_CX3,lmrob_CX4,lmrob_CX5,
                             lmrob_CX7,lmrob_CX8,lmrob_CX11)
lmrob_CX_outputs$Variable <- c("CX1_numbers","CX2_numbers","CX3_numbers","CX4_numbers","CX5_numbers",
                             "CX7_numbers","CX8_numbers","CX11_numbers")
lmrob_CX_outputs$Measure <- c("Buffa_signature_Sum_within","Buffa_signature_Sum_within","Buffa_signature_Sum_within","Buffa_signature_Sum_within","Buffa_signature_Sum_within",
                             "Buffa_signature_Sum_within","Buffa_signature_Sum_within","Buffa_signature_Sum_within")
```

Make model output tables
```{r}
for(i in 1:nrow(lmrob_CX_outputs)){

  model <- model_list[[i]]

  model_dat <- as.data.frame(parameters::parameters(model))
  
  lmrob_CX_outputs[i,3:11] <- model_dat[model_dat$Parameter == lmrob_CX_outputs$Measure[i],1:9]
}

lmrob_CX_outputs$measure_p_adj <- p.adjust(lmrob_CX_outputs$p, method = "BH")
```

```{r}
lmrob_CX_outputs
```

```{r}
dput(lmrob_CX_outputs)
```

```{r}
model_list_CX <- model_list
```

# Non-Zero Robust mixed effects models
robustlmm package

## Mutational signatures
### Models
#### SBS
##### SBS1
```{r}
lmrob_SBS1_nonzero <- lmrob(SBS1_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = subset(lm_SBS1_data, SBS1_div10 > 0))
```

```{r}
summary(lmrob_SBS1_nonzero)
```

```{r}
my_point_plot(lmrob_SBS1_nonzero)
my_qq_plot(lmrob_SBS1_nonzero)
```


```{r}
tab_model(lmrob_SBS1_nonzero)
```

##### SBS2
```{r}
nrow(subset(lm_SBS2_data, SBS2_div10 > 0))
```

```{r}
lmrob_SBS2_nonzero <- lmrob(SBS2_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = subset(lm_SBS2_data, SBS2_div10 > 0))
```

```{r}
summary(lmrob_SBS2_nonzero)
```

```{r}
my_point_plot(lmrob_SBS2_nonzero)
my_qq_plot(lmrob_SBS2_nonzero)
```

```{r}
tab_model(lmrob_SBS2_nonzero)
```

##### SBS3
```{r}
lmrob_SBS3_nonzero <- lmrob(SBS3_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = subset(lm_SBS3_data, SBS3_div10 > 0))
```

```{r}
summary(lmrob_SBS3_nonzero)
```

```{r}
my_point_plot(lmrob_SBS3_nonzero)
my_qq_plot(lmrob_SBS3_nonzero)
```

```{r}
tab_model(lmrob_SBS3_nonzero)
```

##### SBS5
```{r}
lmrob_SBS5_nonzero <- lmrob(SBS5_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = subset(lm_SBS5_data, SBS5_div10 > 0))
```

```{r}
summary(lmrob_SBS5_nonzero)
```

```{r}
my_point_plot(lmrob_SBS5_nonzero)
my_qq_plot(lmrob_SBS5_nonzero)
```

```{r}
tab_model(lmrob_SBS5_nonzero)
```

##### SBS8
```{r}
nrow(subset(lm_SBS8_data, SBS8_div10 > 0))
```

```{r}
# lmrob_SBS8_nonzero <- lmrob(SBS8_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = subset(lm_SBS8_data, SBS8_div10 > 0))
```

```{r}
# summary(lmrob_SBS8_nonzero)
```

```{r}
# my_point_plot(lmrob_SBS8_nonzero)
# my_qq_plot(lmrob_SBS8_nonzero)
```

```{r}
# tab_model(lmrob_SBS8_nonzero)
```

##### SBS13
```{r}
nrow(subset(lm_SBS13_data, SBS13_div10 > 0))
```

```{r}
lmrob_SBS13_nonzero <- lmrob(SBS13_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = subset(lm_SBS13_data, SBS13_div10 > 0))
```

```{r}
summary(lmrob_SBS13_nonzero)
```

```{r}
my_point_plot(lmrob_SBS13_nonzero)
my_qq_plot(lmrob_SBS13_nonzero)
```

```{r}
tab_model(lmrob_SBS13_nonzero)
```

##### SBS15
```{r}
nrow(subset(lm_SBS15_data, SBS15_div10 > 0))
```

```{r}
lmrob_SBS15_nonzero <- lmrob(SBS15_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = subset(lm_SBS15_data, SBS15_div10 > 0))
```

```{r}
summary(lmrob_SBS15_nonzero)
```

```{r}
my_point_plot(lmrob_SBS15_nonzero)
my_qq_plot(lmrob_SBS15_nonzero)
```

```{r}
tab_model(lmrob_SBS15_nonzero)
```

##### SBS17a
```{r}
nrow(subset(lm_SBS17a_data, SBS17a_div10 > 0))
```

```{r}
nrow(subset(lm_SBS17a_data, SBS17a_div10 > 0))
```

```{r}
# lmrob_SBS17a_nonzero <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + SBS17a_div10, data = subset(lm_SBS17a_data, SBS17a_div10 > 0))
```

```{r}
# summary(lmrob_SBS17a_nonzero)
```

```{r}
# my_point_plot(lmrob_SBS17a_nonzero)
# my_qq_plot(lmrob_SBS17a_nonzero)
```

```{r}
# tab_model(lmrob_SBS17a_nonzero)
```

##### SBS17b
```{r}
# summary(subset(lm_SBS17b_data, SBS17b_div10 > 0))
```
```{r}
nrow(subset(lm_SBS17b_data, SBS17b_div10 > 0))
```


```{r}
# lmrob_SBS17b_nonzero <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + SBS17b_div10, data = subset(lm_SBS17b_data, SBS17b_div10 > 0))
```

```{r}
# summary(lmrob_SBS17b_nonzero)
```

```{r}
# my_point_plot(lmrob_SBS17b_nonzero)
# my_qq_plot(lmrob_SBS17b_nonzero)
```

```{r}
# tab_model(lmrob_SBS17b_nonzero)
```

##### SBS18
```{r}
nrow(subset(lm_SBS18_data, SBS18_div10 > 0))
```

Issues with modelling - exact fit and not converging
```{r}
# lmrob_SBS18_nonzero <- lmrob(SBS18_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = subset(lm_SBS18_data, SBS18_div10 > 0))
```

```{r}
# summary(lmrob_SBS18_nonzero)
```

```{r}
# my_point_plot(lmrob_SBS18_nonzero)
# my_qq_plot(lmrob_SBS18_nonzero)
```

```{r}
# tab_model(lmrob_SBS18_nonzero)
```

##### SBS21
```{r}
nrow(subset(lm_SBS21_data, SBS21_div10 > 0))
```

```{r}
# lmrob_SBS21_nonzero <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + SBS21_div10, data = subset(lm_SBS21_data, SBS21_div10 > 0))
```

```{r}
# summary(lmrob_SBS21_nonzero)
```

```{r}
# my_point_plot(lmrob_SBS21_nonzero)
# my_qq_plot(lmrob_SBS21_nonzero)
```

```{r}
# tab_model(lmrob_SBS21_nonzero)
```

##### SBS22b
```{r}
nrow(subset(lm_SBS22b_data, SBS22b_div10 > 0))
```

```{r}
# lmrob_SBS22b_nonzero <- lmrob(SBS22b_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = subset(lm_SBS22b_data, SBS22b_div10 > 0))
```

```{r}
# summary(lmrob_SBS22b_nonzero)
```

```{r}
# my_point_plot(lmrob_SBS22b_nonzero)
# my_qq_plot(lmrob_SBS22b_nonzero)
```

```{r}
# tab_model(lmrob_SBS22b_nonzero)
```

##### SBS28
```{r}
nrow(subset(lm_SBS28_data, SBS28_div10 > 0))
```

```{r}
# lmrob_SBS28_nonzero <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + SBS28_div10, data = subset(lm_SBS28_data, SBS28_div10 > 0))
```

```{r}
# summary(lmrob_SBS28_nonzero)
```

```{r}
# my_point_plot(lmrob_SBS28_nonzero)
# my_qq_plot(lmrob_SBS28_nonzero)
```

```{r}
# tab_model(lmrob_SBS28_nonzero)
```

##### SBS34
```{r}
nrow(subset(lm_SBS34_data, SBS34_div10 > 0))
```

```{r}
# lmrob_SBS34_nonzero <- lmrob(SBS34_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = subset(lm_SBS34_data, SBS34_div10 > 0))
```

```{r}
# summary(lmrob_SBS34_nonzero)
```

```{r}
# my_point_plot(lmrob_SBS34_nonzero)
# my_qq_plot(lmrob_SBS34_nonzero)
```

```{r}
# tab_model(lmrob_SBS34_nonzero)
```

##### SBS37
```{r}
nrow(subset(lm_SBS37_data, SBS37_div10 > 0))
```


```{r}
# lmrob_SBS37_nonzero <- lmrob(SBS37_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = subset(lm_SBS37_data, SBS37_div10 > 0))
```

```{r}
# summary(lmrob_SBS37_nonzero)
```

```{r}
# my_point_plot(lmrob_SBS37_nonzero)
# my_qq_plot(lmrob_SBS37_nonzero)
```

```{r}
# tab_model(lmrob_SBS37_nonzero)
```

##### SBS38
```{r}
nrow(subset(lm_SBS38_data, SBS38_div10 > 0))
```


```{r}
lmrob_SBS38_nonzero <- lmrob(SBS38_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = subset(lm_SBS38_data, SBS38_div10 > 0))
```

```{r}
summary(lmrob_SBS38_nonzero)
```

```{r}
my_point_plot(lmrob_SBS38_nonzero)
my_qq_plot(lmrob_SBS38_nonzero)
```

```{r}
tab_model(lmrob_SBS38_nonzero)
```

##### SBS40a
```{r}
nrow(subset(lm_SBS40a_data, SBS40a_div10 > 0))
```

```{r}
lmrob_SBS40a_nonzero <- lmrob(SBS40a_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = subset(lm_SBS40a_data, SBS40a_div10 > 0))
```

```{r}
summary(lmrob_SBS40a_nonzero)
```

```{r}
my_point_plot(lmrob_SBS40a_nonzero)
my_qq_plot(lmrob_SBS40a_nonzero)
```

```{r}
tab_model(lmrob_SBS40a_nonzero)
```

##### SBS40b
```{r}
nrow(subset(lm_SBS40b_data, SBS40b_div10 > 0))
```

```{r}
# lmrob_SBS40b_nonzero <- lmrob(SBS40b_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = subset(lm_SBS40b_data, SBS40b_div10 > 0))
```

```{r}
# summary(lmrob_SBS40b_nonzero)
```

```{r}
# my_point_plot(lmrob_SBS40b_nonzero)
# my_qq_plot(lmrob_SBS40b_nonzero)
```

```{r}
# tab_model(lmrob_SBS40b_nonzero)
```

##### SBS44
```{r}
nrow(subset(lm_SBS44_data, SBS44_div10 > 0))
```

```{r}
lmrob_SBS44_nonzero <- lmrob(SBS44_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = subset(lm_SBS44_data, SBS44_div10 > 0))
```

```{r}
summary(lmrob_SBS44_nonzero)
```

```{r}
my_point_plot(lmrob_SBS44_nonzero)
my_qq_plot(lmrob_SBS44_nonzero)
```

```{r}
tab_model(lmrob_SBS44_nonzero)
```

##### SBS52
```{r}
nrow(subset(lm_SBS52_data, SBS52_div10 > 0))
```

Issues with modelling - exact fit and not converging

```{r}
# lmrob_SBS52_nonzero <- lmrob(SBS52_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = subset(lm_SBS52_data, SBS52_div10 > 0))
```

```{r}
# summary(lmrob_SBS52_nonzero)
```

```{r}
# my_point_plot(lmrob_SBS52_nonzero)
# my_qq_plot(lmrob_SBS52_nonzero)
```

```{r}
# tab_model(lmrob_SBS52_nonzero)
```

##### SBS54
```{r}
nrow(subset(lm_SBS54_data, SBS54_div10 > 0))
```

```{r}
# lmrob_SBS54_nonzero <- lmrob(SBS54_div10 ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = subset(lm_SBS54_data, SBS54_div10 > 0))
```

```{r}
# summary(lmrob_SBS54_nonzero)
```

```{r}
# my_point_plot(lmrob_SBS54_nonzero)
# my_qq_plot(lmrob_SBS54_nonzero)
```

```{r}
# tab_model(lmrob_SBS54_nonzero)
```

##### SBS96D
```{r}
nrow(subset(lm_SBS96D_data, SBS96D_div10 > 0))
```

```{r}
# lmrob_SBS96D_nonzero <- lmrob(Buffa_signature_Sum_within ~ age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate + SBS96D_div10, data = subset(lm_SBS96D_data, SBS96D_div10 > 0))
```

```{r}
# summary(lmrob_SBS96D_nonzero)
```

```{r}
# my_point_plot(lmrob_SBS96D_nonzero)
# my_qq_plot(lmrob_SBS96D_nonzero)
```

```{r}
# tab_model(lmrob_SBS96D_nonzero)
```

#### SBS model p-value and coefficient table
Make a table with all the p-values from the robustlmm SBS models with p-values and coefficients and adjust the p-values for plotting. 

Prepare output table
```{r}
lmrob_SBS_outputs_nonzero <- as.data.frame(matrix(nrow = 9, ncol = 12))
colnames(lmrob_SBS_outputs_nonzero) <- c("Model","Measure","Parameter","Coefficient","SE","CI","CI_low",
                                        "CI_high","t","df_error","p","Variable")
lmrob_SBS_outputs_nonzero$Model <- c("lmrob_SBS1_nonzero",	
                                     "lmrob_SBS2_nonzero",
                                     "lmrob_SBS3_nonzero",	"lmrob_SBS5_nonzero",	#"lmrob_SBS8_nonzero",	
                               "lmrob_SBS13_nonzero",	"lmrob_SBS15_nonzero",	#"lmrob_SBS17a_nonzero",
                               #"lmrob_SBS17b_nonzero",
                               # "lmrob_SBS18_nonzero",
                               # "lmrob_SBS21_nonzero",
                               # "lmrob_SBS22b_nonzero",
                               # "lmrob_SBS28_nonzero",
                               # "lmrob_SBS34_nonzero",	#"lmrob_SBS37_nonzero",
                               "lmrob_SBS38_nonzero",
                               "lmrob_SBS40a_nonzero",	#"lmrob_SBS40b_nonzero",	
                               "lmrob_SBS44_nonzero"#,
                               # "lmrob_SBS52_nonzero"#,
                               # "lmrob_SBS54_nonzero"	,
                               # "lmrob_SBS96D_nonzero"
                               )
model_list_nonzero <- list(lmrob_SBS1_nonzero,	lmrob_SBS2_nonzero,
                           lmrob_SBS3_nonzero,	lmrob_SBS5_nonzero,	#lmrob_SBS8_nonzero,	
                               lmrob_SBS13_nonzero,	lmrob_SBS15_nonzero,	#lmrob_SBS17a_nonzero,
                           # lmrob_SBS17b_nonzero,
                           # lmrob_SBS18_nonzero,
                               # lmrob_SBS21_nonzero,
                           # lmrob_SBS22b_nonzero,
                           # lmrob_SBS28_nonzero,
                           # lmrob_SBS34_nonzero,	
                           # lmrob_SBS37_nonzero,
                               lmrob_SBS38_nonzero,
                           lmrob_SBS40a_nonzero,	#lmrob_SBS40b_nonzero,	
                           lmrob_SBS44_nonzero#,
                           # lmrob_SBS52_nonzero#,
                               # lmrob_SBS54_nonzero,	
                           # lmrob_SBS96D_nonzero
                           )
lmrob_SBS_outputs_nonzero$Variable <- c("SBS1_div10",	
                                       "SBS2_div10",
                                       "SBS3_div10",	"SBS5_div10",	#"SBS8_div10",	
                               "SBS13_div10",	"SBS15_div10",	#"SBS17a_div10",	
                               # "SBS17b_div10",
                               # "SBS18_div10",
                               # "SBS21_div10",
                               # "SBS22b_div10",
                               # "SBS28_div10",
                               # "SBS34_div10",	
                               # "SBS37_div10",
                               "SBS38_div10",
                               "SBS40a_div10",	#"SBS40b_div10",	
                               "SBS44_div10"#,
                               # "SBS52_div10"#,
                               # "SBS54_div10",	
                               # "SBS96D_div10"
                               )
lmrob_SBS_outputs_nonzero$Measure <- c("Buffa_signature_Sum_within",	
                                       "Buffa_signature_Sum_within",
                                       "Buffa_signature_Sum_within",	"Buffa_signature_Sum_within",	#"Buffa_signature_Sum_within",	
                               "Buffa_signature_Sum_within",	"Buffa_signature_Sum_within",#	"Buffa_signature_Sum_within",	#"Buffa_signature_Sum_within",
                               # "Buffa_signature_Sum_within",
                               # "Buffa_signature_Sum_within",
                               # "Buffa_signature_Sum_within",
                               # "Buffa_signature_Sum_within",
                               # "Buffa_signature_Sum_within",	#"Buffa_signature_Sum_within",
                               # "Buffa_signature_Sum_within",
                               # "Buffa_signature_Sum_within",	#"Buffa_signature_Sum_within",	
                               "Buffa_signature_Sum_within",
                               # "Buffa_signature_Sum_within",
                               "Buffa_signature_Sum_within",	"Buffa_signature_Sum_within"
                               )
```

Make model output tables
```{r}
for(i in 1:nrow(lmrob_SBS_outputs_nonzero)){
print(i)
  model <- model_list_nonzero[[i]]

  model_dat <- as.data.frame(parameters::parameters(model))

  lmrob_SBS_outputs_nonzero[i,3:11] <- model_dat[model_dat$Parameter == lmrob_SBS_outputs_nonzero$Measure[i],1:9]
}

lmrob_SBS_outputs_nonzero$measure_p_adj <- p.adjust(lmrob_SBS_outputs_nonzero$p, method = "BH")
```

```{r}
lmrob_SBS_outputs_nonzero
```

```{r}
dput(lmrob_SBS_outputs_nonzero)
```

```{r}
model_list_nonzero_SBS <- model_list_nonzero
```

#### CN Signatures
##### Models
###### CX1
```{r}
lmrob_CX1_nonzero <- lmrob(CX1_numbers ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = subset(lm_CX1_data, CX1_numbers > 0))
```

```{r}
summary(lmrob_CX1_nonzero)
```

```{r}
my_point_plot(lmrob_CX1_nonzero)
my_qq_plot(lmrob_CX1_nonzero)
```

```{r}
tab_model(lmrob_CX1_nonzero)
```

###### CX2
```{r}
lmrob_CX2_nonzero <- lmrob(CX2_numbers ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = subset(lm_CX2_data, CX2_numbers > 0))
```

```{r}
summary(lmrob_CX2_nonzero)
```

```{r}
my_point_plot(lmrob_CX2_nonzero)
my_qq_plot(lmrob_CX2_nonzero)
```

```{r}
tab_model(lmrob_CX2_nonzero)
```

###### CX3
```{r}
lmrob_CX3_nonzero <- lmrob(CX3_numbers ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = subset(lm_CX3_data, CX3_numbers > 0))
```

```{r}
summary(lmrob_CX3_nonzero)
```

```{r}
my_point_plot(lmrob_CX3_nonzero)
my_qq_plot(lmrob_CX3_nonzero)
```

```{r}
tab_model(lmrob_CX3_nonzero)
```

###### CX4
```{r}
lmrob_CX4_nonzero <- lmrob(CX4_numbers ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = subset(lm_CX4_data, CX4_numbers > 0))
```

```{r}
summary(lmrob_CX4_nonzero)
```

```{r}
my_point_plot(lmrob_CX4_nonzero)
my_qq_plot(lmrob_CX4_nonzero)
```

```{r}
tab_model(lmrob_CX4_nonzero)
```

###### CX5
```{r}
lmrob_CX5_nonzero <- lmrob(CX5_numbers ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = subset(lm_CX5_data, CX5_numbers > 0))
```

```{r}
summary(lmrob_CX5_nonzero)
```

```{r}
my_point_plot(lmrob_CX5_nonzero)
my_qq_plot(lmrob_CX5_nonzero)
```

```{r}
tab_model(lmrob_CX5_nonzero)
```

###### CX7
```{r}
lmrob_CX7_nonzero <- lmrob(CX7_numbers ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = subset(lm_CX7_data, CX7_numbers > 0))
```

```{r}
summary(lmrob_CX7_nonzero)
```

```{r}
my_point_plot(lmrob_CX7_nonzero)
my_qq_plot(lmrob_CX7_nonzero)
```

```{r}
tab_model(lmrob_CX7_nonzero)
```

###### CX8
```{r}
lmrob_CX8_nonzero <- lmrob(CX8_numbers ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = subset(lm_CX8_data, CX8_numbers > 0))
```

```{r}
summary(lmrob_CX8_nonzero)
```

```{r}
my_point_plot(lmrob_CX8_nonzero)
my_qq_plot(lmrob_CX8_nonzero)
```

```{r}
tab_model(lmrob_CX8_nonzero)
```

###### CX11
```{r}
lmrob_CX11_nonzero <- lmrob(CX11_numbers ~ Buffa_signature_Sum_within + age_at_tumour_collection + ISUP_Grade + tumour.purity.estimate, data = subset(lm_CX11_data, CX11_numbers > 0))
```

```{r}
summary(lmrob_CX11_nonzero)
```

```{r}
my_point_plot(lmrob_CX11_nonzero)
my_qq_plot(lmrob_CX11_nonzero)
```

```{r}
tab_model(lmrob_CX11_nonzero)
```

##### Output table
Make a table with all the p-values from the robustlmm CX models with p-values and coefficients and adjust the p-values for plotting. 

Prepare output table
```{r}
lmrob_CX_outputs_nonzero <- as.data.frame(matrix(nrow = 8, ncol = 12))
colnames(lmrob_CX_outputs_nonzero) <- c("Model","Measure","Parameter","Coefficient","SE","CI","CI_low",
                                        "CI_high","t","df_error","p","Variable")
lmrob_CX_outputs_nonzero$Model <- c("lmrob_CX1_nonzero","lmrob_CX2_nonzero","lmrob_CX3_nonzero","lmrob_CX4_nonzero","lmrob_CX5_nonzero",
                             "lmrob_CX7_nonzero","lmrob_CX8_nonzero","lmrob_CX11_nonzero")
model_list_nonzero <- list(lmrob_CX1_nonzero,lmrob_CX2_nonzero,lmrob_CX3_nonzero,lmrob_CX4_nonzero,lmrob_CX5_nonzero,
                             lmrob_CX7_nonzero,lmrob_CX8_nonzero,lmrob_CX11_nonzero)
lmrob_CX_outputs_nonzero$Variable <- c("CX1_numbers","CX2_numbers","CX3_numbers","CX4_numbers","CX5_numbers",
                             "CX7_numbers","CX8_numbers","CX11_numbers")
lmrob_CX_outputs_nonzero$Measure <- c("Buffa_signature_Sum_within","Buffa_signature_Sum_within","Buffa_signature_Sum_within",
                                      "Buffa_signature_Sum_within","Buffa_signature_Sum_within",
                             "Buffa_signature_Sum_within","Buffa_signature_Sum_within","Buffa_signature_Sum_within")
```

Make model output tables
```{r}
for(i in 1:nrow(lmrob_CX_outputs_nonzero)){

  model <- model_list_nonzero[[i]]

  model_dat <- as.data.frame(parameters::parameters(model))
  
  lmrob_CX_outputs_nonzero[i,3:11] <- model_dat[model_dat$Parameter == lmrob_CX_outputs_nonzero$Measure[i],1:9]
}

lmrob_CX_outputs_nonzero$measure_p_adj <- p.adjust(lmrob_CX_outputs_nonzero$p, method = "BH")
```

```{r}
lmrob_CX_outputs_nonzero
```

```{r}
dput(lmrob_CX_outputs_nonzero)
```

```{r}
model_list_nonzero_CX <- model_list_nonzero
```

# Save model outputs
## Saving coefficient tables
```{r}
save(lmrob_GI_outputs,
     lmrob_SBS_outputs,
     lmrob_CX_outputs,
     file = "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_paper_revisions/TCGA_analysis/lmrob_outputs/total_coefficient_tables_k10_rearranged_20251018.RData")
```

```{r}
save(lmrob_SBS_outputs_nonzero,
     lmrob_CX_outputs_nonzero,
     file = "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_paper_revisions/TCGA_analysis/lmrob_outputs/nonzero_coefficient_tables_k10_rearranged_20251018.RData")
```

## Saving models
```{r}
save(model_list_GI,
     model_list_SBS,
     model_list_CX,
     file = "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_paper_revisions/TCGA_analysis/lmrob_outputs/total_models_k10_rearranged_20251018.RData")
```

```{r}
save(model_list_nonzero_SBS,
     model_list_nonzero_CX,
     file = "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_paper_revisions/TCGA_analysis/lmrob_outputs/nonzero_models_k10_rearranged_20251018.RData")
```

# Session info
```{r}
sessionInfo()
```

