---
title: "Creating PPCG combined table - V9"
author: Maria Jakobsdottir (Maria.Jakobsdottir@manchester.ac.uk)
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2:
    toc: true 
    toc_float: true
    toc_depth: 3
---

The purpose of this R notebook is to harmonise the merging of different PPCG data into one big table. 

# Libraries used
```{r message=FALSE, warning=FALSE}
options(bitmapType='cairo')
'%!in%' <- function(x,y)!('%in%'(x,y))
```


# Data
Read in data

## Original files used
```{r}
files = list(
    SBS_signatures = '/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_Signatures/MUTATIONAL_SIGNATURES_v0.16_DRAFT_VERSION/SBS96_12_Signatures/Decompose_Solution/Activities/Decompose_Solution_Activities.txt',
    ID_signatures = '/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_Signatures/MUTATIONAL_SIGNATURES_v0.16_DRAFT_VERSION/COSMIC_ID83_Decomposed_Solution/Activities/COSMIC_ID83_Activities.txt',
    CNA_signatures = '/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_CN_signatures/CINSignature_Activities_THRESH95_PPCG_ASCAT_Dec2023.tsv',
    tracking_wgs = '/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_other/Main_Tracking_Sheet_ICR_WGS_v2.2.tsv',
    tracking_rna = '/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_other/MTS_new.csv',
    exclude_list = '/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_other/Main_Tracking_Sheet_ICR_Exclusion_list_v1.0.tsv',
    greylist = '/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_other/Main_Tracking_Sheet_ICR_greylist_v1.0.tsv',
    pga = '/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_analyses/2026-04-05pcawgsummary_pgaSCNA_with_Brass_Delly_SVs_20260212_2026-04-05.txt',
    snvcounts = '/mnt/bmh01-rds/UoOxford_David_W/b05055gj/preliminary_ICGC_UK_RNAseq/PPCG_20April2021_SNV_number_2021-11-29.txt',
    indelcounts = '/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_other/Somatic_variants/InDels/PPCG_indels_12_Dec_2023/Tiers/release_summary.tsv',
    cnacounts = '/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_analyses/SCNA_with_Brass_Delly_SVs_20260212_num_cnas_2026-04-05.txt',
    svcounts = '/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_analyses/PPCG_SV_Data_ReleaseV2_04_October_2023_vcf_sv_number_2024-04-09.txt',
    kataegiscounts = '/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_Data_Release/Somatic_variants/Kataegis/Kataegis_02_June_2021/PPCG_Kataegis_Calls_02_June_2021.tsv',
    mito_cn = '/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_Mitochondria/PPCG_MT_Data_Release_2021_Oct_07/PPCG_MT_Data_Release_2021_Oct_07_copynumber.csv',
    hypoxia_signatures = '/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_analyses/manuscript_v2_outputs/hypoxia_scores/alternative_within_sample_scaled_hypoxia_scores_RNAv1.k10_2024-08-23.txt',
    cellularity = '/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_paper_revisions/2026_BB_update/cellularity_feb2026_2026-05-31.csv',
    ETS_status = '/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_other/Main_Tracking_Sheet_ICR_ETS.csv',
    subclone_number = '/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_paper_revisions/2026_BB_update/clone_count_2026-04-05.csv',
    evo_qc = '/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_Copy_Number/PPCG_Feb2026_NRPCC.txt'
    )
```

## Signature data
Data release version: v0.16

Date: 2024-04-21 (SBS/ID)

Source: Andreas Gruber - Globus

SBS: /mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_Signatures/MUTATIONAL_SIGNATURES_v0.16_DRAFT_VERSION/SBS96_12_Signatures/Decompose_Solution/Activities/Decompose_Solution_Activities.txt
```{r}
sbs_signatures <- read.delim(files$SBS_signatures, sep = "\t", header = T)
```
```{r}
str(sbs_signatures)
```

INDEL: /mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_Signatures/MUTATIONAL_SIGNATURES_v0.16_DRAFT_VERSION/COSMIC_ID83_Decomposed_Solution/Activities/COSMIC_ID83_Activities.txt
```{r}
id_signatures <- read.delim(files$ID_signatures, sep = "\t", header = T)
```
```{r}
str(id_signatures)
```

Date: 2023-12

Source: Barbara Hernando - FTP box

CN: /mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_CN_signatures/CINSignature_Activities_THRESH95_PPCG_ASCAT_Dec2023.tsv
```{r}
cna_signatures <- read.delim(files$CNA_signatures, sep = "\t", header = T)
```
```{r}
str(cna_signatures)
```

## Mitochondrial data
Data release version: 2021_Oct_07

Date: 2022-12-02

Source: Chol Hee Jung - FTP box

Copy number
/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_Mitochondria/PPCG_MT_Data_Release_2021_Oct_07/PPCG_MT_Data_Release_2021_Oct_07_copynumber.csv
```{r}
mito_cn <- read.csv(files$mito_cn, header = TRUE)
```
```{r}
str(mito_cn)
```

## Other data

### Master tracking sheet WGS
Data release version: WGS v2.2

Date: 2024-12-04

/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_other/Main_Tracking_Sheet_ICR_WGS_v2.2.tsv

```{r}
tracking_wgs <- read.delim(files$tracking_wgs, sep = "\t", header = T)
```
```{r}
str(tracking_wgs)
```

#### Master tracking sheet - Exclusion list
Data release version: Exclusion v1.0

Date: 2024-12-04

/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_other/Main_Tracking_Sheet_ICR_Exclusion_list_v1.0.tsv

```{r}
exclude_list <- read.delim(files$exclude_list, sep = "\t", header = T)
```
```{r}
str(exclude_list)
```

#### Master tracking sheet - Grey list
Data release version: Grey list v1.0

Date: 2024-12-04

/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_other/Main_Tracking_Sheet_ICR_greylist_v1.0.tsv

```{r}
greylist <- read.delim(files$greylist, sep = "\t", header = T)
```
```{r}
str(greylist)
```

### Master tracking sheet RNA
Data release version: 

Date: 2024-11-01
Shared by Eddie on slack

/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_other/MTS_new.csv

```{r}
tracking_rna <- read.csv(files$tracking_rna, header = T)
```
```{r}
str(tracking_rna)
```

### PGA
Data release: PPCG_Data_Release_01_June_2020 (uploaded by Atef to csf3)

Date: 03-08-2023 

Calculated by Maria using code originally from Bhandari.

/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_analyses/2023-08-03_pgappcgsummary_pgaPPCG_UK_2023-08-03.txt
```{r}
pga_values <- read.delim(files$pga, sep = "\t", header = T)
```
```{r}
str(pga_values)
```

### Count data
## SNV
Data release: SNV release 20 April 2021
File: /mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_Data_Release/Somatic_variants/SNVs/Filtered_SNV_VCFs_20_April_2021/
Date: Apr 2021-04-20

## INDEL
Data release: 12 Dec 2023
File: /mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_other/Somatic_variants/InDels/PPCG_indels_12_Dec_2023/Tiers/release_summary.tsv
Date: Dec 2023-12-12

## CNA + PGA
Data release:03_05_2023
File: /mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_Copy_Number/SCNA_with_Brass_Delly_SVs_20260212
Date: May 2026-04-05

## SV
Data release: V2
File: /mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_other/Somatic_variants/SVs/PPCG_SV_Data_ReleaseV2_04102023/SVs/PPCG_SV_Data_ReleaseV2_04_October_2023_vcf/
Date: Oct 2023-10-04

Summary data calculated by Maria.  (/mnt/bmh01-rds/UoOxford_David_W/b05055gj/preliminary_ICGC_UK_RNAseq/ppcg_uk_rna_preliminary/20211129_calculate_GI_muts_indels_etc.R - ignore kataegis counts)

```{r}
snv_counts <- read.delim(files$snvcounts, sep = "\t", header = T)
indel_counts <- read.delim(files$indelcounts, sep = "\t", header = T)
cna_counts <- read.delim(files$cnacounts, sep = "\t", header = T)
sv_counts <- read.delim(files$svcounts, sep = "\t", header = T)
```
```{r}
str(snv_counts)
```
```{r}
str(indel_counts)
```
```{r}
str(cna_counts)
```
```{r}
str(sv_counts)
```


### Kataegis
Data release: 02-06-2021

Date: Jun 30-06-2021 (uploaded by Atef to csf3)

/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_Data_Release/Somatic_variants/Kataegis/Kataegis_02_June_2021/PPCG_Kataegis_Calls_02_June_2021.tsv
```{r}
kataegis_counts <- read.delim(files$kataegiscounts, sep = "\t", header = TRUE)
```
```{r}
str(kataegis_counts)
```

### Cellularity/purity
Data release: Feb 2026

Date: Feb 2026

```{r}
cellularity <- read.csv(files$cellularity, header = TRUE)
```
```{r}
str(cellularity)
```


## Hypoxia signatures
Calculated from data release version 1.4/k10 - extra columns are from RNA metadata (1.3). Signatures were calculated using gene expression values for signature genes centred around the median expression of the gene in the data set and then the values of the signature genes were scaled within each sample. Signatures were then calculated using either the sum, binary, or ranked method. 

Recommended column to use: Buffa_signature_Sum_within
```{r}
hypoxia_signatures <- read.delim(files$hypoxia_signatures, sep = "\t", header = TRUE)
```
```{r}
str(hypoxia_signatures)
```

## ETS calls
ETS status is from the main tracking sheet
```{r}
ets_status <- read.csv(files$ETS_status, header = TRUE)
```

## Subclone-numbers
```{r}
subclone_numbers <- read.csv(files$subclone_number, header = TRUE)
```
```{r}
evo_qc <- read.delim(files$evo_qc, header = TRUE, sep = "\t")
```


# Data cleaning and summarising

```{r}
snv_counts$file_dna <- gsub("_vs_./*.*","",snv_counts$filename_snv)
sv_counts$file_dna <- gsub("_ppcg_consensus_annotated.somatic.sv.vcf.gz","",sv_counts$filename_sv)
cna_counts$file_dna <- gsub("_vs_./*.*","",cna_counts$filename_cna)
pga_values$file_dna <- gsub("_vs_./*.*","",pga_values$patient_id)
```

Changing sample names for signature files. Removing blood comparison.
```{r}
sbs_signatures$Samples_DNA <- gsub("_DNA/*.*","_DNA",sbs_signatures$Samples)
id_signatures$Samples_DNA <- gsub("_DNA/*.*","_DNA",id_signatures$Samples)
cna_signatures$Samples_DNA <- paste0(cna_signatures$sample,"_DNA")
```

Summarise Kataegis results
```{r}
kataegis_summary <- as.data.frame(table(kataegis_counts$PPCG_Sample_ID))
colnames(kataegis_summary) <- c("Samples_DNA","Kataegis_count")
```
```{r}
str(kataegis_summary)
```

Combine kataegis and SNV data so that samples without kataegis can be assigned 0
```{r}
kataegis_snv_summary <- merge(snv_counts,
                              kataegis_summary,
                              by.x = "file_dna",
                              by.y = "Samples_DNA",
                              all = TRUE)
kataegis_snv_summary$Kataegis_count[is.na(kataegis_snv_summary$Kataegis_count)] <- 0
```
```{r}
str(kataegis_snv_summary)
```


Combine grey and exclusion lists. If a sample is in both lists only keep the exclusion list entry. 
```{r}
greylist_unique <- subset(greylist, WGS_AssayID %!in% exclude_list$WGS_AssayID)
grey_exclude_list <- rbind(exclude_list, greylist_unique)
```

## Remove subclone data for samples that don't pass NRPCC > 10
Feb 2026 version of copy number data only has samples that passed
```{r}
evo_qc$PPCG_Sample_ID <- gsub("_vs_.*","", evo_qc$Barcode)
evo_qc$pass_nrpcc_qc <- ifelse(evo_qc$NRPCC > 10, yes = TRUE, no = FALSE)
```

```{r}

# cna_qc_pass <- evo_qc$PPCG_Sample_ID[evo_qc$pass_cna_qc %!in% greylist[grepl("Unsatisfactory",greylist$Reason_CV),"WGS_AssayID"]]
cna_qc_nrpcc_pass <- evo_qc$PPCG_Sample_ID[evo_qc$pass_nrpcc_qc == TRUE]
```

```{r}
# cna_signatures <- subset(cna_signatures, Samples_DNA %in% cna_qc_pass)
```

```{r}
# pga_values <- subset(pga_values, file_dna %in% cna_qc_pass)
```

```{r}
# 2026 version only contains files for samples with satisfacroty CNA calls
# cna_counts <- subset(cna_counts, file_dna %in% cna_qc_pass)
```

```{r}
subclone_numbers <- subset(subclone_numbers, PPCG_Sample_ID %in% cna_qc_nrpcc_pass)
```

## RNA
### Remove hypoxia scores for samples removed from RNA
```{r}
hypoxia_signatures <- subset(hypoxia_signatures, Sample %in% tracking_rna[is.na(tracking_rna$removed),"PPCG_RNA_Assay_ID"])
```

### Rename duplicated/mislabelled RNA samples so that they can be matched to WGS by PPCG ID
4 RNA samples had PPCG Donor IDs that were different in the RNA and the WGS/Methylation data. Re-naming the donor IDs so that the data can be matched accurately.  Keeping the RNA Assay ID for the same reason. Update matching WGS sample as well (if it exists). 

```{r}

tracking_rna$PPCG_Donor_ID[tracking_rna$PPCG_Donor_ID == "PPCG2136"] <- "PPCG1970"

tracking_rna$PPCG_Donor_ID[tracking_rna$PPCG_Donor_ID == "PPCG0145"] <- "PPCG0138"
tracking_rna$Matching_WGS_Sample[tracking_rna$PPCG_Donor_ID == "PPCG0138"] <- "PPCG0138a_DNA"

tracking_rna$PPCG_Donor_ID[tracking_rna$PPCG_Donor_ID == "PPCG2132"] <- "PPCG1968"

tracking_rna$PPCG_Donor_ID[tracking_rna$PPCG_Donor_ID == "PPCG2135"] <- "PPCG1969"
```


# Creating combined table

Merge RNA metadata with WGS Master tracking sheet. 

```{r}
dim(tracking_wgs)
dim(tracking_rna)
```


```{r}
PPCG_data <- merge(subset(tracking_wgs, SNV == TRUE),
                   subset(tracking_rna, is.na(removed)),
                   by.x = c("WGS_AssayID","PPCG_Donor_ID"),
                   by.y = c("Matching_WGS_Sample","PPCG_Donor_ID"),
                   all = TRUE,
                   suffixes = c(".WGS",".RNA"))
```
```{r}
dim(PPCG_data)
```
2194

```{r}
PPCG_data <- merge(PPCG_data,
                   subset(pga_values, file_dna %in% subset(tracking_wgs, SNV == TRUE)$WGS_AssayID),
                   by.x = "WGS_AssayID",
                   by.y = "file_dna",
                   all = TRUE)
dim(PPCG_data)
```
2194
```{r}
PPCG_data <- merge(PPCG_data,
                   subset(kataegis_snv_summary, file_dna %in% subset(tracking_wgs, SNV == TRUE)$WGS_AssayID),
                   by.x = "WGS_AssayID",
                   by.y = "file_dna",
                   all = TRUE)
dim(PPCG_data)
```
2194
```{r}
PPCG_data <- merge(PPCG_data,
                   subset(indel_counts, WGS_AssayID %in% subset(tracking_wgs, SNV == TRUE)$WGS_AssayID),
                   by.x = "WGS_AssayID",
                   by.y = "WGS_AssayID",
                   all = TRUE)
dim(PPCG_data)
```
2194

```{r}
PPCG_data <- merge(PPCG_data,
                   subset(cna_counts, file_dna %in% subset(tracking_wgs, SNV == TRUE)$WGS_AssayID),
                   by.x = "WGS_AssayID",
                   by.y = "file_dna",
                   all = TRUE)
dim(PPCG_data)
```
2194

```{r}
PPCG_data <- merge(PPCG_data,
                   subset(sv_counts, file_dna %in% subset(tracking_wgs, SNV == TRUE)$WGS_AssayID),
                   by.x = "WGS_AssayID",
                   by.y = "file_dna",
                   all = TRUE)
dim(PPCG_data)
```
2194

```{r}
PPCG_data <- merge(PPCG_data,
                   subset(cellularity, PPCG_Sample_ID %in% subset(tracking_wgs, SNV == TRUE)$WGS_AssayID),
                   by.x = "WGS_AssayID",
                   by.y = "PPCG_Sample_ID",
                   all = TRUE)
dim(PPCG_data)
```
2194

```{r}
PPCG_data <- merge(PPCG_data,
                   subset(mito_cn, PPCG_ID %in% subset(tracking_wgs, SNV == TRUE)$WGS_AssayID)[,c("PPCG_ID","Per.base.depth","WGS.depth","mtCN")],
                   by.x = "WGS_AssayID",
                   by.y = "PPCG_ID",
                   all = TRUE)
dim(PPCG_data)
```
2194

```{r}
PPCG_data <- merge(PPCG_data,
                   subset(sbs_signatures, Samples_DNA %in% subset(tracking_wgs, SNV == TRUE)$WGS_AssayID)[,-1],
                   by.x = "WGS_AssayID",
                   by.y = "Samples_DNA",
                   all = TRUE)
dim(PPCG_data)
```
2194

```{r}
PPCG_data <- merge(PPCG_data,
                   subset(id_signatures, Samples_DNA %in% subset(tracking_wgs, SNV == TRUE)$WGS_AssayID)[,-1],
                   by.x = "WGS_AssayID",
                   by.y = "Samples_DNA",
                   all = TRUE)
dim(PPCG_data)
```
2194

```{r}
PPCG_data <- merge(PPCG_data,
                   subset(cna_signatures, Samples_DNA %in% subset(tracking_wgs, SNV == TRUE)$WGS_AssayID),
                   by.x = "WGS_AssayID",
                   by.y = "Samples_DNA",
                   all = TRUE)
dim(PPCG_data)
```
2194

```{r}
PPCG_data <- merge(PPCG_data,
                   grey_exclude_list[,c("WGS_AssayID","Listed_category","Datatype","Reason_CV",
                                     "Reason_text","Person_reported","Date_reported","PPTECH_reviewed")],
                   by.x = "WGS_AssayID",
                   by.y = "WGS_AssayID",
                   all = TRUE)
dim(PPCG_data)
```
2213

```{r}
PPCG_data <- merge(PPCG_data,
                   ets_status[,-3],
                   by.x = c("WGS_AssayID","PPCG_Donor_ID"),
                   by.y = c("WGS_AssayID","PPCG_Donor_ID"),
                   all = TRUE)
dim(PPCG_data)
```
2215

```{r}
PPCG_data <- merge(PPCG_data,
                   subset(subclone_numbers, PPCG_Sample_ID %in% subset(tracking_wgs, SNV == TRUE)$WGS_AssayID),
                   by.x = "WGS_AssayID",
                   by.y = "PPCG_Sample_ID",
                   all = TRUE)
dim(PPCG_data)
```
2215

```{r}
PPCG_data <- merge(PPCG_data,
                   hypoxia_signatures[,c("Sample","Buffa_signature_Sum_within","Genes_used_Buffa_within")],
                   by.x = "PPCG_RNA_Assay_ID",
                   by.y = "Sample",
                   all = TRUE)
dim(PPCG_data)
```
2215

Remove exclusion list samples
```{r}
PPCG_data$Listed_category[is.na(PPCG_data$Listed_category)] <- "none"
```

```{r}
PPCG_data <- subset(PPCG_data, Listed_category != "black")
```
```{r}
dim(PPCG_data)
```
2215

Selecting single tumour sample per patient to stratify hypoxia into high/low, and merge back in. Subset for RNA tumour samples. 
```{r}
PPCG_data_tumour_only <- subset(PPCG_data, RNA_Sample_Type == "TUM")
```
```{r}
dim(PPCG_data_tumour_only)
```
1274

Subset for selected WGS tumour sample
```{r}
PPCG_data_tumour_only_single <- subset(PPCG_data_tumour_only, selected_one_sample_per_donor.WGS == TRUE)
```
```{r}
dim(PPCG_data_tumour_only_single)
```
508

Randomly pick one non-duplicated RNA sample
```{r}
rna_duplicated_ids <- PPCG_data_tumour_only_single$PPCG_Donor_ID[duplicated(PPCG_data_tumour_only_single$PPCG_Donor_ID)]
non_dup_rna <- subset(PPCG_data_tumour_only_single, PPCG_Donor_ID %!in% rna_duplicated_ids)

# keep selected sample for duplicated samples
dup_rna <- subset(PPCG_data_tumour_only_single, PPCG_Donor_ID %in% rna_duplicated_ids)
dup_rna <- dup_rna[dup_rna$PPCG_RNA_Assay_ID == dup_rna$selected_one_sample_per_donor.RNA,]

PPCG_data_tumour_only_single <- rbind(non_dup_rna, dup_rna)
```
```{r}
dim(PPCG_data_tumour_only_single)
```
506

Check how many don't have hypoxia scores
```{r}
summary(PPCG_data_tumour_only_single$Buffa_signature_Sum_within)
```

Subset for samples that have signature
```{r}
PPCG_data_tumour_only_single <- subset(PPCG_data_tumour_only_single, !is.na(Buffa_signature_Sum_within))
```
```{r}
dim(PPCG_data_tumour_only_single)
```
505

```{r}
PPCG_data_tumour_only_single$Buffa_signature_Sum_within_bin_one_sample_per_donor[PPCG_data_tumour_only_single$Buffa_signature_Sum_within >= median(PPCG_data_tumour_only_single$Buffa_signature_Sum_within)] <- "High"
PPCG_data_tumour_only_single$Buffa_signature_Sum_within_bin_one_sample_per_donor[PPCG_data_tumour_only_single$Buffa_signature_Sum_within < median(PPCG_data_tumour_only_single$Buffa_signature_Sum_within)] <- "Low"
```

```{r}
summary(PPCG_data_tumour_only_single)
```

Also split by quantiles
```{r}
hypoxia_quartiles <- quantile(PPCG_data_tumour_only_single$Buffa_signature_Sum_within)
hypoxia_quartiles <- as.data.frame(hypoxia_quartiles)
hypoxia_quartiles$quartile <- rownames(hypoxia_quartiles)
```

```{r}
hypoxia_quartiles
```

```{r}
PPCG_data_tumour_only_single$Buffa_signature_Sum_within_quartile_one_sample_per_donor[PPCG_data_tumour_only_single$Buffa_signature_Sum_within < hypoxia_quartiles$hypoxia_quartiles[hypoxia_quartiles$quartile == "25%"]] <- "Q1"

PPCG_data_tumour_only_single$Buffa_signature_Sum_within_quartile_one_sample_per_donor[PPCG_data_tumour_only_single$Buffa_signature_Sum_within >= hypoxia_quartiles$hypoxia_quartiles[hypoxia_quartiles$quartile == "25%"] & 
                                                                         PPCG_data_tumour_only_single$Buffa_signature_Sum_within < hypoxia_quartiles$hypoxia_quartiles[hypoxia_quartiles$quartile == "50%"]] <- "Q2"

PPCG_data_tumour_only_single$Buffa_signature_Sum_within_quartile_one_sample_per_donor[PPCG_data_tumour_only_single$Buffa_signature_Sum_within >= hypoxia_quartiles$hypoxia_quartiles[hypoxia_quartiles$quartile == "50%"] & 
                                                                         PPCG_data_tumour_only_single$Buffa_signature_Sum_within < hypoxia_quartiles$hypoxia_quartiles[hypoxia_quartiles$quartile == "75%"]] <- "Q3"

PPCG_data_tumour_only_single$Buffa_signature_Sum_within_quartile_one_sample_per_donor[PPCG_data_tumour_only_single$Buffa_signature_Sum_within >= hypoxia_quartiles$hypoxia_quartiles[hypoxia_quartiles$quartile == "75%"]] <- "Q4"
```

```{r}
PPCG_data <- merge(PPCG_data,
                   PPCG_data_tumour_only_single[,c("PPCG_RNA_Assay_ID","Buffa_signature_Sum_within_bin_one_sample_per_donor",
                                                   "Buffa_signature_Sum_within_quartile_one_sample_per_donor")],
                   by.x = "PPCG_RNA_Assay_ID",
                   by.y = "PPCG_RNA_Assay_ID",
                   all = TRUE)
dim(PPCG_data)
```
2215



```{r}
summary(PPCG_data)
```

## Extend quartiles for Barbara
```{r}
PPCG_data$Buffa_signature_Sum_within_quartile_extended[PPCG_data$Buffa_signature_Sum_within < hypoxia_quartiles$hypoxia_quartiles[hypoxia_quartiles$quartile == "25%"]] <- "Q1"

PPCG_data$Buffa_signature_Sum_within_quartile_extended[PPCG_data$Buffa_signature_Sum_within >= hypoxia_quartiles$hypoxia_quartiles[hypoxia_quartiles$quartile == "25%"] & 
                                                                         PPCG_data$Buffa_signature_Sum_within < hypoxia_quartiles$hypoxia_quartiles[hypoxia_quartiles$quartile == "50%"]] <- "Q2"

PPCG_data$Buffa_signature_Sum_within_quartile_extended[PPCG_data$Buffa_signature_Sum_within >= hypoxia_quartiles$hypoxia_quartiles[hypoxia_quartiles$quartile == "50%"] & 
                                                                         PPCG_data$Buffa_signature_Sum_within < hypoxia_quartiles$hypoxia_quartiles[hypoxia_quartiles$quartile == "75%"]] <- "Q3"

PPCG_data$Buffa_signature_Sum_within_quartile_extended[PPCG_data$Buffa_signature_Sum_within >= hypoxia_quartiles$hypoxia_quartiles[hypoxia_quartiles$quartile == "75%"]] <- "Q4"
```

# Saving data

```{r}
write.table(PPCG_data, paste0("/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_other/",
                              "combined_PPCG_WGS_RNA_summary_data_v9_",
                              Sys.Date(),
                              ".txt"),
            sep = "\t", row.names = FALSE)
```
```{r}
saveRDS(PPCG_data, paste0("/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_other/",
                              "combined_PPCG_WGS_RNA_summary_data_v9_",
                              Sys.Date(),
                              ".rds"))
```

# Session info

```{r}
sessionInfo()
```

