---
title: "Creating PPCG combined table with clinical data - V9"
author: Maria Jakobsdottir (Maria.Jakobsdottir@manchester.ac.uk)
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2:
    toc: true 
    toc_float: true
    toc_depth: 3
---

The purpose of this R notebook is to harmonise the merging of different PPCG data into one big table with clinical information. 

# Libraries used
```{r message=FALSE, warning=FALSE}
options(bitmapType='cairo')
'%!in%' <- function(x,y)!('%in%'(x,y))
```


# Data
Read in data

## Original files used

**Change these values on your system**
```{r}
files = list(
    combined_data = '/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_other/combined_PPCG_WGS_RNA_summary_data_v9_2026-05-31.txt',
    clinical_data = '/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_data_private/PPCGCoreDatav282024D_DATA_2024-12-02_1508.csv',
    path_review = '/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_data_private/PPCGPathReviewV10202_DATA_2024-09-27_1059.csv',
    outdir = '/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_data_private/'
    )
```

## Combined data
Date: 2026-05-31

Source: Maria Jakobsdottir
```{r}
PPCG_data <- read.delim(files$combined_data, header = TRUE, sep = "\t")
```
```{r}
dim(PPCG_data)
```
Should be 2215 rows, 150 columns


## Clinical data
Data release version: 

Date: 2024-12-02

Source: RedCap

/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_data_private/PPCGCoreDatav282024D_DATA_2024-12-02_1508.csv
```{r}
clinical_data <- read.csv(files$clinical_data)
```
```{r}
dim(clinical_data)
```
2098 rows, 28 columns

## Path review
Data release version: 

Date: 2024-09-27

Source: RedCap

/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_data_private/PPCGPathReviewV10202_DATA_2024-09-27_1059.csv
```{r}
path_data <- read.csv(files$path_review)
```
```{r}
dim(path_data)
```
476 rows, 18 columns


# Standardising clinical data
Re-formatting clinical data.  

Gleason. 
```{r}
# Make Gleason score total
clinical_data$gleason_total <- paste0(clinical_data$gleason_predominantly,"+",clinical_data$gleason_less_predominantly)

clinical_data$gleason_total[(clinical_data$gleason_total == "missing+missing")] <- NA
clinical_data$gleason_total[clinical_data$gleason_total == "not applicable+not applicable"] <- NA

clinical_data$gleason_total <- as.factor(clinical_data$gleason_total)
clinical_data$gleason_total <- factor(clinical_data$gleason_total, levels = c("3+3","3+4","4+3","4+4","3+5","4+5","5+4","5+5"))
```

PSA brackets
```{r}
clinical_data$psa_at_tumour_collection[(clinical_data$psa_at_tumour_collection == "missing")] <- NA
clinical_data$psa_at_tumour_collection[(clinical_data$psa_at_tumour_collection == "not applicable")] <- NA

clinical_data$psa_at_tumour_collection <- as.numeric(clinical_data$psa_at_tumour_collection)

clinical_data$psa_bracket[clinical_data$psa_at_tumour_collection < 10] <- "0-9.9"
clinical_data$psa_bracket[(clinical_data$psa_at_tumour_collection >= 10) & (clinical_data$psa_at_tumour_collection < 20)] <- "10-19.9"
clinical_data$psa_bracket[clinical_data$psa_at_tumour_collection >= 20] <- "≥20"

clinical_data$psa_bracket <- as.factor(clinical_data$psa_bracket)
clinical_data$psa_bracket <- factor(clinical_data$psa_bracket, levels = c("0-9.9","10-19.9","≥20"))
```

Age brackets
```{r}
clinical_data$age_at_tumour_collection[(clinical_data$age_at_tumour_collection == "missing")] <- NA
clinical_data$age_at_tumour_collection[(clinical_data$age_at_tumour_collection == "not applicable")] <- NA

clinical_data$age_at_tumour_collection <- as.numeric(clinical_data$age_at_tumour_collection)

clinical_data$age_bracket[clinical_data$age_at_tumour_collection < 40] <- "< 40"
clinical_data$age_bracket[(clinical_data$age_at_tumour_collection >= 40) & (clinical_data$age_at_tumour_collection < 50)] <- "40-50"
clinical_data$age_bracket[(clinical_data$age_at_tumour_collection >= 50) & (clinical_data$age_at_tumour_collection < 65)] <- "50-65"
clinical_data$age_bracket[(clinical_data$age_at_tumour_collection >= 65) & (clinical_data$age_at_tumour_collection < 70)] <- "65-70"
clinical_data$age_bracket[clinical_data$age_at_tumour_collection >= 70] <- "≥ 70"

clinical_data$age_bracket <- as.factor(clinical_data$age_bracket)
clinical_data$age_bracket <- factor(clinical_data$age_bracket, levels = c("< 40","40-50","50-65","65-70","≥ 70"))
```

Clinical T bracket
```{r}
clinical_data$clin_t_stage[clinical_data$clin_t_stage == "missing"] <- NA
clinical_data$clin_t_stage[clinical_data$clin_t_stage == "not applicable"] <- NA

clinical_data$clin_t_bracket <- substr(clinical_data$clin_t_stage, start = 0, stop = 2)
clinical_data$clin_t_bracket <- as.factor(clinical_data$clin_t_bracket)
```

ISUP grade group
```{r}
clinical_data$ISUP_Grade <- clinical_data$gleason_grade_group
```

```{r}
dim(clinical_data)
```
2098 33

IDC calls
Harmonising capitalisation of IDC calls
```{r}
path_data$Cribriform_Intraductal_Y_N <- toupper(path_data$cribriform__intraductal_yn)
```


# Merging data

Two samples from the blacklist is present in the clinical data. We'll merge keeping only all.x to avoid reintroducing those two samples. 

```{r}
clinical_data$ppcg_donor_id[clinical_data$ppcg_donor_id %!in% PPCG_data$PPCG_Donor_ID]
```


```{r}
PPCG_data <- merge(PPCG_data,
                   clinical_data,
                   by.x = "PPCG_Donor_ID",
                   by.y = "record_id",
                   all.x = TRUE)
dim(PPCG_data)
```
2215 182

```{r}
PPCG_data <- merge(PPCG_data,
                   path_data,
                   by.x = "PPCG_Donor_ID",
                   by.y = "record_id",
                   all.x = TRUE)
dim(PPCG_data)
```
2215 200

# Saving data

```{r}
write.table(PPCG_data, paste0(files$outdir,
                              "combined_PPCG_WGS_RNA_summary_data_clinical_v9_",
                              Sys.Date(),
                              ".txt"),
            sep = "\t", row.names = FALSE)
```
```{r}
saveRDS(PPCG_data, paste0(files$outdir,
                              "combined_PPCG_WGS_RNA_summary_data_clinical_v9_",
                              Sys.Date(),
                              ".rds"))
```

# Session info

```{r}
sessionInfo()
```

