---
title: "ssGSEA for PPCG RNAseq - v1.4k10prerelease"
output: html_notebook
---

# Libraries
```{r}
library(dplyr)
library(GSVA)
library(msigdbr)
library(org.Hs.eg.db)
library(qs)
library(SummarizedExperiment)
library(xtable)
options(bitmapType='cairo')
'%!in%' <- function(x,y)!('%in%'(x,y))
```

# Data
Release 1.3 is just used for the metadata associated with it. The was not provided with the 1.4k10 release but will have been the same. 
```{r}
files = list(
    rna_data_sup = '/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_RNA/K10_unsupervised/PPCG_RNAseq_RUVIII_K10.qs',
    rna_metadata = '/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_RNA/DEPRECATED_Release_1.3/PPCG_RNAseq_RUVIIIPRPS_Supervised.qs')
```

## RNA data

Data release version: 1.4/k10

Date: 2024-08-23

Source: Ramyar Molania - FTP box

rna_dat = qread("/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_RNA/K10_unsupervised/PPCG_RNAseq_RUVIII_K10.qs")
```{r}
rna_dat_unsupervised = qread(files$rna_data_sup)
```
```{r}
str(rna_dat_unsupervised)
```

```{r}
rna_dat_metadata = qread(files$rna_metadata)
```



# Gene sets

## Hallmark genes
"This package was generated with MSigDB v7.5.1 (released January 2022)"
```{r}
h_gene_sets <- msigdbr(species = "Homo sapiens", category = "H")
```

```{r}
table(h_gene_sets$gs_name)
```


## Match gene sets to names
Match genes to names
```{r}
h_gene_sets_partial <- merge(h_gene_sets, 
                     as.data.frame((rna_dat_metadata@elementMetadata)[,c("all.ppcg.trans","hgnc_symbol")]),
                     by.x = "ensembl_gene",
                     by.y = "all.ppcg.trans")
```
```{r}
table(h_gene_sets_partial$gs_name)
```


```{r}
h_gene_sets <- merge(h_gene_sets, 
                     as.data.frame((rna_dat_metadata@elementMetadata)[,c("all.ppcg.trans","hgnc_symbol")]),
                     by.x = "ensembl_gene",
                     by.y = "all.ppcg.trans",
                     all.x = TRUE)
```

```{r}
sum(is.na(h_gene_sets$hgnc_symbol))
```

## Summary of genes used and not
```{r}
original_gene_numbers <- as.data.frame(table(h_gene_sets$gs_name))
release_gene_numbers <- as.data.frame(table(h_gene_sets_partial$gs_name))

colnames(original_gene_numbers) <- c("Gene_set","Total_genes_in_set")
colnames(release_gene_numbers) <- c("Gene_set","Set_genes_in_tier_1")

hallmark_gene_numbers <- merge(original_gene_numbers,release_gene_numbers,
                               by = "Gene_set")
```

```{r}
xtable(hallmark_gene_numbers)
```

```{r}
missing_genes <- anti_join(h_gene_sets, h_gene_sets_partial)
```

```{r}
write.table(missing_genes,file = paste0("/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_data/ssgsea_hallmark_gs_missing_from_supervised_1.3_",Sys.Date(),".txt"), sep = "\t", row.names = FALSE)
```


## Gene set list
```{r}
msigdbr_list = split(x = h_gene_sets$hgnc_symbol, f = h_gene_sets$gs_name)
```

# ssGSEA

```{r}
ssgsea_parameters <- ssgseaParam(expr = rna_dat_unsupervised,
                      geneSets = msigdbr_list)

ssgseaResults <- gsva(ssgsea_parameters)
```

```{r}
ssgseaResults <- as.data.frame(t(ssgseaResults))
ssgseaResults$sample <- rownames(ssgseaResults)
```

```{r}
write.table(ssgseaResults,
            file = paste0("/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_data/ssgsea_hallmark_gs_",Sys.Date(),".txt"), sep = "\t", row.names = FALSE)
```

```{r}
sessionInfo()
```

