---
title: "ssGSEA for PPCG RNAseq - v1.4k10prerelease"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2:
    toc: true 
    toc_float: true
    toc_depth: 6
---

# Libraries
```{r}
library(dplyr)
library(GSVA)
library(msigdbr)
library(org.Hs.eg.db)
library(qs)
library(SummarizedExperiment)
library(xtable)
options(bitmapType='cairo')
'%!in%' <- function(x,y)!('%in%'(x,y))
```

```{r}
outdir <- "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_paper_final_results/outputs/"
```

```{r}
supp_tab_data_outdir <- "/mnt/bmh01-rds/UoOxford_David_W/b05055gj/PPCG_hypoxia_paper_final_results/Supplementary_Table_Data/"
```


# Data
Release 1.3 is just used for the metadata associated with it. The was not provided with the 1.4k10 release but will have been the same. 
```{r}
files = list(
    rna_data_sup = '/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_RNA/K10_unsupervised/PPCG_RNAseq_RUVIII_K10.qs',
    rna_metadata = '/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_RNA/DEPRECATED_Release_1.3/PPCG_RNAseq_RUVIIIPRPS_Supervised.qs')
```

## Buffa signature
```{r}
buffa_gene_list <- structure(list(SYMBOL = c("VEGFA", "SLC2A1", "PGAM1", "ENO1", "LDHA",
                                             "TPI1", "P4HA1", "MRPS17", "CDKN3", "ADM",
                                             "NDRG1", "TUBB6", "ALDOA", "MIF", "ACOT7",
                                             "MCTS1", "PSRC1", "PSMA7", "ANLN", "TUBA1B",
                                             "MAD2L2", "GPI", "TUBA1C", "MAP7D1", "DDIT4",
                                             "BNIP3", "MRGBP", "HILPDA", "GAPDH", "MRPL13",
                                             "CHCHD2", "YKT6", "PNP", "CORO1C", "SEC61G",
                                             "ANKRD37", "ESRP1", "PFKP", "SHCBP1", "CTSL2",
                                             "KIF20A", "SLC25A32", "UTP11", "SLC16A1", "MRPL15",
                                             "KIF4A", "LRRC42", "PGK1", "HK2", "AK3",
                                             "CA9"),
                                  ENSG = c("ENSG00000112715", "ENSG00000117394", "ENSG00000171314", "ENSG00000074800", "ENSG00000134333",
                                           "ENSG00000111669", "ENSG00000122884", "ENSG00000239789", "ENSG00000100526", "ENSG00000148926",
                                           "ENSG00000104419", "ENSG00000176014", "ENSG00000149925", "ENSG00000240972", "ENSG00000097021",
                                           "ENSG00000232119", "ENSG00000134222", "ENSG00000101182", "ENSG00000011426", "ENSG00000123416",
                                           "ENSG00000116670", "ENSG00000105220", "ENSG00000167553", "ENSG00000116871", "ENSG00000168209",
                                           "ENSG00000176171", "ENSG00000101189", "ENSG00000135245", "ENSG00000111640", "ENSG00000172172",
                                           "ENSG00000106153", "ENSG00000106636", "ENSG00000198805", "ENSG00000110880", "ENSG00000132432",
                                           "ENSG00000186352", "ENSG00000104413", "ENSG00000067057", "ENSG00000171241", "ENSG00000136943",
                                           "ENSG00000112984", "ENSG00000164933", "ENSG00000183520", "ENSG00000155380", "ENSG00000137547",
                                           "ENSG00000090889", "ENSG00000116212", "ENSG00000102144", "ENSG00000159399", "ENSG00000147853",
                                           "ENSG00000107159")),
                             class = "data.frame", 
                             row.names = c(NA,-51L))
```

## RNA data

Data release version: 1.4/k10

Date: 2024-08-23

Source: Ramyar Molania - FTP box

rna_dat = qread("/mnt/bmh01-rds/UoOxford_David_W/shared/PPCG_RNA/K10_unsupervised/PPCG_RNAseq_RUVIII_K10.qs")
```{r}
rna_dat_unsupervised = qread(files$rna_data_sup)
```
```{r}
str(rna_dat_unsupervised)
```

```{r}
rna_dat_metadata = qread(files$rna_metadata)
```



# Gene sets

## Hallmark genes
"This package was generated with MSigDB v7.5.1 (released January 2022)"
```{r}
h_gene_sets <- msigdbr(species = "Homo sapiens", collection = "H")
```

```{r}
table(h_gene_sets$gs_name)
```


## Match gene sets to names
Match genes to names
```{r}
h_gene_sets_partial <- merge(h_gene_sets, 
                     as.data.frame((rna_dat_metadata@elementMetadata)[,c("all.ppcg.trans","hgnc_symbol")]),
                     by.x = "ensembl_gene",
                     by.y = "all.ppcg.trans")
```
```{r}
table(h_gene_sets_partial$gs_name)
```


```{r}
h_gene_sets <- merge(h_gene_sets, 
                     as.data.frame((rna_dat_metadata@elementMetadata)[,c("all.ppcg.trans","hgnc_symbol")]),
                     by.x = "ensembl_gene",
                     by.y = "all.ppcg.trans",
                     all.x = TRUE)
```

```{r}
sum(is.na(h_gene_sets$hgnc_symbol))
```

Match gene sets to genes in Buffa signature
```{r}
h_gene_sets_buffa <- merge(h_gene_sets, 
                     buffa_gene_list,
                     by.x = "ensembl_gene",
                     by.y = "ENSG")
```

## Summary of genes used and not
```{r}
original_gene_numbers <- as.data.frame(table(h_gene_sets$gs_name))
release_gene_numbers <- as.data.frame(table(h_gene_sets_partial$gs_name))
buffa_gene_numbers <- as.data.frame(table(h_gene_sets_buffa$gs_name))

colnames(original_gene_numbers) <- c("Gene_set","Total_genes_in_set")
colnames(release_gene_numbers) <- c("Gene_set","Set_genes_in_tier_1")
colnames(buffa_gene_numbers) <- c("Gene_set","Set_genes_in_Buffa")

hallmark_gene_numbers <- merge(original_gene_numbers,release_gene_numbers,
                               by = "Gene_set",
                               all = TRUE)
hallmark_gene_numbers <- merge(hallmark_gene_numbers,buffa_gene_numbers,
                               by = "Gene_set",
                               all = TRUE)
```

```{r}
xtable(hallmark_gene_numbers)
```

```{r}
missing_genes <- anti_join(h_gene_sets, h_gene_sets_partial)
```

Add missing genes in column to `hallmark_gene_numbers` table
```{r}
missing_genes_names <- as.data.frame(unique(missing_genes$gs_name))
colnames(missing_genes_names) <- "gs_name"
missing_genes_names$Missing_genes_in_set <- NA

for(i in 1:nrow(missing_genes_names)){
  sub_gs <- subset(missing_genes, gs_name == missing_genes_names$gs_name[i])
  missing_genes_names$Missing_genes_in_set[i] <- paste(sub_gs$gene_symbol, collapse = ",")
}
```

Add buffa genes in column to `hallmark_gene_numbers` table
```{r}
buffa_genes <- semi_join(h_gene_sets, h_gene_sets_buffa)
```
```{r}
buffa_genes_names <- as.data.frame(unique(buffa_genes$gs_name))
colnames(buffa_genes_names) <- "gs_name"
buffa_genes_names$Buffa_genes_in_set <- NA

for(i in 1:nrow(buffa_genes_names)){
  sub_gs <- subset(buffa_genes, gs_name == buffa_genes_names$gs_name[i])
  buffa_genes_names$Buffa_genes_in_set[i] <- paste(sub_gs$gene_symbol, collapse = ",")
}
```

Add missing and buffa genes to set and final number
```{r}
hallmark_gene_numbers <- merge(hallmark_gene_numbers,missing_genes_names,
                               by.x = "Gene_set",
                               by.y = "gs_name",
                               all = TRUE)
hallmark_gene_numbers <- merge(hallmark_gene_numbers,buffa_genes_names,
                               by.x = "Gene_set",
                               by.y = "gs_name",
                               all = TRUE)
hallmark_gene_numbers$Set_genes_in_Buffa[is.na(hallmark_gene_numbers$Set_genes_in_Buffa)] <- 0
hallmark_gene_numbers$Total_Genes_Used_for_signature <- hallmark_gene_numbers$Set_genes_in_tier_1 - hallmark_gene_numbers$Set_genes_in_Buffa
```


```{r}
xtable(hallmark_gene_numbers)
```


```{r}
latex_summary <- print(xtable(hallmark_gene_numbers))
```


```{r}
used_genes <- semi_join(h_gene_sets, h_gene_sets_partial)
used_genes <- anti_join(used_genes, h_gene_sets_buffa)
```

```{r}
# missing genes and used genes
write.table(missing_genes,
            file = paste0(outdir,"/ssgsea/ssgsea_hallmark_gs_missing_from_1.4k10_",Sys.Date(),".txt"), 
            sep = "\t", row.names = FALSE)
write.table(used_genes,
            file = paste0(outdir,"/ssgsea/ssgsea_hallmark_gs_used_from_1.4k10_",Sys.Date(),".txt"), 
            sep = "\t", row.names = FALSE)

# summary table
write.table(hallmark_gene_numbers,
            file = paste0(outdir,"/ssgsea/text_summary_ssgsea_hallmark_gs_used_from_1.4k10_",Sys.Date(),".txt"), 
            sep = "\t", row.names = FALSE)
# summary table as latex
output_tex_file <- paste0(outdir,"/ssgsea/latex_summary_ssgsea_hallmark_gs_used_from_1.4k10_",Sys.Date(),".txt")
writeLines(text = latex_summary,
           con = output_tex_file)
```

```{r}
write.csv(hallmark_gene_numbers,
            file = paste0(supp_tab_data_outdir,"supp_tab_12_data",".csv"))
```


## Gene set list
```{r}
msigdbr_list = split(x = used_genes$hgnc_symbol, f = used_genes$gs_name)
```

# ssGSEA

```{r}
ssgsea_parameters <- ssgseaParam(expr = rna_dat_unsupervised,
                      geneSets = msigdbr_list)

ssgseaResults <- gsva(ssgsea_parameters)
```

```{r}
ssgseaResults <- as.data.frame(t(ssgseaResults))
ssgseaResults$sample <- rownames(ssgseaResults)
```

```{r}
write.table(ssgseaResults,
            file = paste0(outdir,"/ssgsea/revised_ssgsea_hallmark_gs_",Sys.Date(),".txt"), 
            sep = "\t", row.names = FALSE)
```

```{r}
sessionInfo()
```

